---
title: "Modeling Flight Response Summary File"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
rm(list=ls())

# set working directory
dir = "~/Desktop/git_repositories/SBB-dispersal/avbernat_working_on/"
setwd(dir)

knitr::opts_chunk$set(echo = TRUE)
```

# Data Cleaning And Exploration

## Read Libraries

```{r message=FALSE, warning=FALSE}
library(lme4) # fit regressions 
library(rethinking) # Bayesian data analysis and plotting
library(popbio) # logistic regression plotting
library(binom) # binomial confidence intervals
```

## Read Source Files

```{r message=FALSE, warning=FALSE}
source_path = paste0(dir,"/Rsrc/")

script_names = c("center_flight_data.R", # 1 function: center_data()
                 "clean_flight_data.R", # 1 function: clean_flight_data()
                 "unique_flight_data.R", # 1 function: create_delta_data()
                 "compare_models.R", # 1 function: model_comparisonsAIC()
                 "AICprobabilities.R", # 1 function: AICprobs() 
                 "get_warnings.R") # 1 function: withWarnings() 

for (script in script_names) { 
  path = paste0(source_path, script)
  source(path) 
}
```

## Read the Data

```{r message=FALSE, warning=FALSE}
data_path = paste0(dir,"/Dispersal/Winter_2020/stats/data/all_flight_data-Winter2020.csv")

data = read_flight_data(data_path)
data_all = data[[1]] 
data_tested = data[[2]] # subset of data already centered via read_flight_data

# keep bugs only tested once then re-center:
d <- create_delta_data(data_tested, remove_bugs_tested_once = FALSE) 
dc <- center_data(d, is_not_unique_data = FALSE)
```

## Data Definitions

* A **full dataset** is a dataset where each row has a unique ID and trial type. For example,

```{r}
data_tested[c(1:2,400:401), c("ID", "trial_type")]
```

* A **unique dataset** is a dataset where each row has a unique ID because each trial has been grouped by ID. For example,

```{r results='hold'}
dc[c(1:2,295:296), c("ID", "trial_type")]
```

## Data Transformations

* **var_name_b** means that the data are binary (either 0s or 1s)
* **var_name_c** means that the data had been centered 
* **var_name_s** means that the data had been standardized
* **avg_var_name** means that the data had been averaged across trial 1 (T1) and trial 2 (T2)
* **var_name_diff** means that the data are a difference between T1 and T2 (T2-T1)
* **var_name_per** means that the data a percent change between T1 and T2 (T2-T1)/T1 * 100
* **var_name_logsqrt** means that the data has been more normalized using a log-square-root transformation
* **var_name_logsqrt_i** means that the data has been more normalized using a log-square-root transformation but its sign is the inverse of the **var_name**. 

## Repeating Plot Parameters & Functions

```{r}
# scale/magnifications
c1 = 1.3*2 # size of points
c2 = 1.2*2 # size of large text
c3 = 2 # size of smaller text
c4 = 2*2 # size of title 

# compute 95% confidence interval
get_CI = function(x,y,m) {
  x.seq = seq(min(x) - sd(x), max(x) + sd(x), length.out=100)
  prd = data.frame(x=x.seq)
  err = predict(m, newdata = prd, se.fit = TRUE)
  prd$lci = err$fit - 1.96 * err$se.fit
  prd$fit = err$fit
  prd$uci = err$fit + 1.96 * err$se.fit
  mu_ci = t(matrix(c(prd$lci,prd$uci), ncol=2))
  return(list(mu_ci, prd))
}

# tailoring variables for plotting
d$mass_block=round(d$avg_mass/0.005)*0.005 # 0.005 g blocks 
d$wing2body_block=round(d$wing2body, digits=2) # 0.01 blocks
d$days_block=round(d$avg_days, digits=0) # integer blocks
```

# Across-Trial Flight Response

## Experimental Effects

```{r}
binary_counts = table(data_tested$flew_b, data_tested$trial_type)[,2:3] # flew yes or no per trial
dd = aggregate(flew_b ~ days_from_start, data=data_tested, FUN=mean)
dt = data_tested[,c("hr_start", "recording_duration")]
```

```{r echo=FALSE, fig.height=2.7*3, fig.width=6*3}
# day tested: flight prob vs. days from start

par(mfrow=c(1,2), mar = c(4.3, 7, 3.5, 0.1), cex.axis=c3, cex.lab=c2) 

# plot A (barplot)
barplot(binary_counts, 
        ylim=c(0,250), beside=TRUE,
        xlab="Trial", col=c("#b57783","#820331"),
        ylab="Frequency")
legend("topright",
       legend = c("no flew", "yes flew"),
       col=c("#b57783","#820331"),
       pch=15,
       cex = c2)
mtext("A", side=3, adj=0, line=0.5, cex=c4, font=1)
  
# plot B (linear regression)
x=dd$days_from_start
y=dd$flew_b
m = glm(y ~ x, data=dd, family="gaussian")
shading = get_CI(x,y,m)
mu_ci=shading[[1]]
prd=shading[[2]]

plot(dd$flew_b~dd$days_from_start, 
     ylab="Flight Probability", xlab="Days From Start", 
     pch=19, col=1, 
     cex=c1,
     ylim=c(0.4,1))
abline(m, lty=2)
shade(mu_ci, lim = prd$x)
pval = paste0("p = ", round(summary(m)$coefficients[8],3), "*")
text(11, 0.45, pval, cex=c2)
mtext("B", side=3, adj=0, line=0.5, cex=c4, font=1)

# inner plot (logistic regression)
par(fig=c(.70,0.97,0.5,0.99), new=TRUE, mar=c(1,1,1,1), cex.axis=c3, cex.lab=c3)

m = glm(flew_b ~ days_from_start, data=data_tested, family = "binomial" )
pvalue = paste0("p = ", round(summary(m)$coefficients[8],3), "*")

logi.hist.plot(data_tested$days_from_start, 
               data_tested$flew_b,boxp=FALSE,
               type="hist",col="indianred", ylabel="", ylabel2 = "", xlabel="")
text(7, 0.38, pvalue, cex=c3)
mtext("Frequency", side=4, cex=c3, line=1)
```

```{r echo=FALSE, fig.height=2.7*2.7, fig.width=6*3}
# time tested: recording duration (HR) vs. trial start time (HR)

par(mfrow=c(1,2), mar = c(4.3, 7, 3.5, 0.1), cex.axis=c3, cex.lab=c2) 

# plot C (all bugs)
x=dt$hr_start
y=dt$recording_duration/60/60
m = glm(y ~ x, data=dt, family="gaussian")
shading = get_CI(x,y,m)
mu_ci=shading[[1]]
prd=shading[[2]]

plot(dt$hr_start, dt$recording_duration/60/60,
     xlab="Trial Start Time (HR)", 
     ylab="Recording Duration (HR)",
     cex=c1)
abline(m, lty=2)
shade(mu_ci, lim = prd$x)
pval = paste0("p = ", round(summary(m)$coefficients[8],3)) # marginal effect
text(15, 11, pval, cex=c2) 
mtext("C", side=3, adj=0, line=0.5, cex=c4, font=1)


# plot D (filter out bugs that didn't fly)
dt=data_tested %>%
  filter(flew_b !=0)
x=dt$hr_start
y=dt$recording_duration/60/60
m = glm(y ~ x, data=dt, family="gaussian") 
shading = get_CI(x,y,m)
mu_ci=shading[[1]]
prd=shading[[2]]

plot(dt$hr_start, dt$recording_duration/60/60,
     xlab="Trial Start Time (HR)", 
     ylab="Recording Duration (HR)",
     cex=c1)
abline(m, lty=2)
shade(mu_ci, lim = prd$x)
pval = paste0("p = ", round(summary(m)$coefficients[8],3), "*")
text(15, 11, pval, cex=c2)
mtext("D", side=3, adj=0, line=0.5, cex=c4, font=1)
```

**A & B.** There was a negative effect of day a bug was tested (since the start of trials), but there was a significant effect only when the full dataset is considered (not the delta dataset, which is not shown here).
**C & D.** There was a negative effect of the trial start time but only after removing bugs that didn't fly (D).

## Binomial Modeling

We aimed to model the probability of a flight response with sex, host plant, and distance from the sympatric zone. Additionally, to understand how experimental factors affected flight response, we computed the average days since start and the average mass (log-square-root transformed). Finally, we used aggregated datasets for single-variate modeling and the unique dataset for multi-variate modeling.

### Average Days Since Start Histogram

```{r}
dc$num_trials <- as.factor(rowSums(dc[,c("num_flew", "num_notflew")]))
```

```{r echo=FALSE, fig.width=4.7, fig.height=3.4}
h <- ggplot(dc, aes(x=avg_days, color=num_trials)) +
  labs(title=" ", color="Number of Trials", 
       x="Average Days Since Start", y = "Frequency") +
  geom_histogram(binwidth=1, fill="white") + 
  theme(legend.position="top") +
  theme_classic() +
  scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9")) +
  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9"))  
h
```

The advantage of this computed variable is that it controls for the fact that some bugs were tested once late, and some by chance had been tested twice early. This allows the multi-variate models, which control for ID, to converge, but average days does not show the same strong, significant effect as days from start:

```{r}
days_model<-glm(cbind(num_flew,num_notflew)~avg_days_c, data=d, family=binomial)
summary(days_model)
```

## Single-Variate Effects

```{r}
# aggregate data for plotting
dt=aggregate(flew_prob~sex, data=d, FUN=mean)
dt$trials=c(sum(d$num_flew[d$sex=="F"]+d$num_notflew[d$sex=="F"]),
            sum(d$num_flew[d$sex=="M"]+d$num_notflew[d$sex=="M"]))

ds=aggregate(flew_prob~sex*wing2body_block, data=d, FUN=mean)
ds$n=aggregate(flew_prob~sex*wing2body_block, data=d, FUN=length)$flew_prob

dm=aggregate(flew_prob~sex*mass_block, data=d, FUN=mean)
dm$n=aggregate(flew_prob~sex*mass_block, data=d, FUN=length)$flew_prob

# calculate binomial confidence interval
dt$successes = c(sum(d$num_flew[d$sex=="F"]), 
                 sum(d$num_flew[d$sex=="M"]))
dt$CI = binom.confint(dt$successes, dt$trials, methods="exact")
```

```{r echo=FALSE, fig.height=2.7*2*2.7, fig.width=6*3}

par(mfrow=c(2,2), mar = c(4.3, 7, 3.8, 0.1), cex.axis=c3, cex.lab=c2) 

# plot A (flight probability grouped by sex)
sex_colors = c("red", "blue")
nfem = nrow(d[d$sex=="F",])
nfem = paste0("N=", nfem)
nmale = nrow(d[d$sex=="M",])
nmale = paste0("N=", nmale)

plot(dt$CI$mean~c(1,2), xaxt='n', 
     ylab="Flight probability", 
     xlab="Sex", ylim=c(0,1), 
     xlim=c(0.5,2.5), 
     cex=c1, pch=19, col=sex_colors)
lines(x=xy.coords(x=c(1,1), y=c(dt$CI$lower[1], dt$CI$upper[1])), lwd=3, col=sex_colors[1])
lines(x=xy.coords(x=c(2,2), y=c(dt$CI$lower[2], dt$CI$upper[2])), lwd=3, col=sex_colors[2])
axis(side=1, at=c(1,2), labels=c("female", "male"))
mtext(text=c(nfem,nmale), at=c(1,2), side=1, line=-3, cex=c2)
mtext("A", side=3, adj=0, line=0.5, cex=c4, font=1)

# plot B (wing2body ratio vs. flew_prob grouped by sex)

d1 = ds[ds$sex=="F",]
d2 = ds[ds$sex=="M",]

# females
x=d1[,"wing2body_block"]
y=d1[,"flew_prob"] 
m1 = lm(y ~ x, data=d1)
shading=get_CI(x,y,m1)
mu_ci1=shading[[1]]
prd1=shading[[2]]

# males
x=d2[,"wing2body_block"]
y=d2[,"flew_prob"] 
m2 = lm(y ~ x, data=d2)
shading=get_CI(x,y,m2)
mu_ci2=shading[[1]]
prd2=shading[[2]]

plot(ds$flew_prob~ds$wing2body_block, 
     ylab="Flight probability", xlab="wing-to-body ratio", 
     pch=19, col=c("red","blue")[as.factor(ds$sex)], 
     cex=c1, ylim=c(0,1))
abline(m1, lty=2, col="indianred1")
abline(m2, lty=2, col="darkblue")
shade(mu_ci1, lim = prd1$x, col=col.alpha("indianred1"))
shade(mu_ci2, lim = prd2$x) 
pval1 = paste0("p = ", round(summary(m1)$coefficients[8],2))
pval2 = paste0("p = ", round(summary(m2)$coefficients[8],5), "*")
text(0.72, 0.05, pval1, cex=c2, col="darkred")
text(0.67, 0.6, pval2, cex=c2) 
legend(0.64, 1, 
       legend=c("male", "female"), 
       col=c("blue", "red"),
       pch=19, y.intersp=1.3, 
       pt.cex=c1, cex = c2)
mtext("B", side=3, adj=0, line=0.5, cex=c4, font=1)

# plot C (mass vs. flew_prob grouped by sex)
dm=aggregate(flew_prob~sex*mass_block, data=d, FUN=mean)
dm$n=aggregate(flew_prob~sex*mass_block, data=d, FUN=length)$flew_prob

d1 = dm[dm$sex=="F",]
d2 = dm[dm$sex=="M",]

# females
x=d1[,"mass_block"]
y=d1[,"flew_prob"] 
m1 = lm(y ~ x, data=d1)
shading=get_CI(x,y,m1)
mu_ci1=shading[[1]]
prd1=shading[[2]]

# males
x=d2[,"mass_block"]
y=d2[,"flew_prob"] 
m2 = lm(y ~ x, data=d2)
shading=get_CI(x,y,m2)
mu_ci2=shading[[1]]
prd2=shading[[2]]

plot(dm$flew_prob~dm$mass_block, 
     ylab="Flight probability", 
     xlab="Mass (g)", 
     pch=19, 
     col=sex_colors[as.factor(dm$sex)], 
     cex=c1,
     ylim=c(0,1), xlim=c(0.02, 0.18))
legend(0.14, 1, 
       legend=c("male", "female"), 
       col=c("blue", "red"),
       pch=19, 
       y.intersp=1.3, pt.cex=c1, cex = c2)
abline(m1, lty=2)
abline(m2, lty=2, col="darkblue")
shade(mu_ci1, lim = prd1$x)
shade(mu_ci2, lim = prd2$x, col=col.alpha("darkblue"))
pval1 = paste0("p = ", round(summary(m1)$coefficients[8],3), "*")
pval2 = paste0("p = ", round(summary(m2)$coefficients[8],2))
text(0.11, 0.5, pval1, cex=c2)
text(0.1, 0.9, pval2, cex=c2, col="darkblue")
mtext("C", side=3, adj=0, line=0.5, cex=c4, font=1)
```

## Multi-Variate Modeling

```{r results='hold'}
data<-data.frame(R1 = dc$num_flew, 
                 R2 = dc$num_notflew, 
                 A = dc$host_c,
                 B = dc$sex_c,
                 C = dc$sym_dist,
                 D = dc$avg_mass_logsqrt,
                 E = dc$avg_days_c)

model_script = paste0(source_path,"generic models-binomial glm 2R ~ 4-FF + E.R")
errors = withWarnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

&nbsp;

```{r results='hold'}
anova(m63, m85, test="Chisq") # Adding B*D does not improve fit
anova(m63, m36, test="Chisq") # Adding C*D does improve fit 
```

### Best Fit 

```{r}
M1 <- glm(cbind(num_flew, num_notflew) ~ host_c * avg_mass_logsqrt 
          + sym_dist_s * avg_mass_logsqrt + sex_c + avg_days_c, data=d, family=binomial)
summary(M1)
```

&nbsp;

### By Sex

### Females

```{r}
data_fem <- dc[dc$sex=="F",]
data_fem <- center_data(data_fem, is_not_unique_data = FALSE)
```

```{r results='hold'}
data<-data.frame(R1 = data_fem$num_flew,
                 R2 = data_fem$num_notflew,
                 A = data_fem$host_c,
                 B = data_fem$sym_dist, 
                 C = data_fem$avg_mass_logsqrt, 
                 D = data_fem$wing2body_logsqrt_i, 
                 E = data_fem$avg_days_c)

model_script = paste0(source_path,"generic models-binomial glm 2R ~ 4-FF + E.R")
errors = withWarnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

&nbsp;

```{r results='hold'}
anova(m25, m45, test='Chisq') #adding A*D does not improve fit 
anova(m25, m13, test='Chisq') #adding A*C improves fit
anova(m25, m17, test="Chisq") #adding D improves fit
anova(m25, m45, test="Chisq") #adding D improves fit
```

### Best Fit 

```{r}
M2 <- glm(cbind(num_flew, num_notflew) ~ host_c * avg_mass_logsqrt +  wing2body_logsqrt_i + 
            avg_days_c, data=data_fem, family=binomial) 
summary(M2)
```

&nbsp;

### Males

```{r}
data_male <- dc[dc$sex=="M",]
data_male <- center_data(data_male, is_not_unique_data = FALSE)
```

```{r results='hold'}
data<-data.frame(R1 = data_male$num_flew,
                 R2 = data_male$num_notflew,
                 A = data_male$host_c,
                 B = data_male$sym_dist, 
                 C = data_male$avg_mass_logsqrt, 
                 D = data_male$wing2body_logsqrt_i, 
                 E = data_male$avg_days_c)

model_script = paste0(source_path,"generic models-binomial glm 2R ~ 4-FF + E.R")
errors = withWarnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

&nbsp;

```{r results='hold'}
anova(m83, m105, test="Chisq") # adding C*D marginally improves fit
anova(m83, m62, test="Chisq") # adding B*C marginally improves fit
anova(m50, m62, test="Chisq") # adding C does not improve fit 
```

### Best Fit 

```{r}
M3<-glm(cbind(num_flew, num_notflew)~host_c*wing2body_logsqrt_i + sym_dist*wing2body_logsqrt_i 
        + avg_days_c, family=binomial, data=data_male) 
summary(M3)
```

&nbsp;

# Between-Trial Flight Response (T1 vs. T2)

## Read Libraries

```{r warning=FALSE, message=FALSE}
library(dplyr) # data manipulation 
library(zoo) # data manipulation
library(nnet) # multinomial modeling 
library(kableExtra) # table formatting
library(plot.matrix) # enables matrix/heatmap plotting
```

## Read Source Files

```{r}
script_names = c( "multinom_functions.R") # 4 relevant functions:
                                          # calculate_P2(), calculate_P3(),
                                          # get_significant_models(),
                                          # get_significant_modelsf(),

for (script in script_names) { 
  path = paste0(source_path, script)
  source(path) 
}
```

## Read the Data

```{r}
d = create_delta_data(data_tested, remove_bugs_tested_once=TRUE) # remove bugs tested only once
```

## Encodings & Signs

We aimed to model the probability of different delta flight response cases with sex, host plant, percent changes in mass, and percent changes in egg-laying response as predictors. Since the outcomes (or response variables) were no longer binomial, we used multi-categorical logit models. Below are the categorical encodings and/or signs used. See the Appendix for additional explanations and examples of computing multi-categorical logit models.

```{r echo=FALSE}
Event = c("flew in both trials", "flew in T2 only", " flew in neither trials", "flew in T1 only")
Encoding = c(2, 1,0,-1)
key = cbind(Event, Encoding)

kable(key) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))  %>%
  kable_classic(html_font = "Cambria") %>%
  add_header_above(c("Delta Flight Response Key" = 2 )) 
```

```{r echo=FALSE}
Event = c("gained % mass from T1 to T2", "no % mass change between trails", "lost % mass from T1 to T2")
Sign = c("+","0","-")
key = cbind(Event, Sign)

kable(key) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))  %>%
  kable_classic(html_font = "Cambria") %>%
  add_header_above(c("Delta Percent Mass Key (%)" = 2 )) 
```

```{r echo=FALSE, warning=FALSE}
Host = c("Golden Rain Tree (GRT)", "Balloon Vine (BV)")
Encoding = c(1,-1)
key = cbind(Host, Encoding)

kable(key)  %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))  %>%
  kable_classic(html_font = "Cambria") %>%
  add_header_above(c("Host Plant Key" = 2 )) 
```
```{r echo=FALSE, warning=FALSE}
Sex = c("Female", "Male")
Encoding = c(1,-1)
key = cbind(Sex, Encoding)

kable(key)  %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))  %>%
  kable_classic(html_font = "Cambria") %>%
  add_header_above(c("Sex Key" = 2 )) 
```

## Multinomial Modeling

### Baseline

```{r}
# remove any missing values for flight case or mass percent change between trials
df = d[with(d,!is.na(flight_case) & !is.na(mass_per)),]

# order the dataset by ascending mass percent change values
df = df[with(df, order(mass_per)),]

# relevel the flight case factors so as to set 0 as the first level.
df$flight_case = relevel(as.factor(df$flight_case), ref = "0")
```

### Null Model

```{r}
null = multinom(flight_case ~ 1, data = df) 
```

### Compare Models - predictors: % mass, sex, host

```{r warning=FALSE}
data = data.frame(R = df$flight_case, 
         A = df$mass_per,
         B = df$sex_c,
         C = df$host_c)
model_script = paste0(source_path,"generic multinomial models- multinom 1RF + 3 FF.R")
model_comparisonsAIC(model_script)
```

&nbsp;

```{r warning=FALSE, results = "hold"}
anova(m4, m7, test="Chisq") # Adding C (host plant) does not improve fit
anova(m4, m8, test="Chisq") # Adding A*B does not improve fit
```

### Best Fit

```{r warning=FALSE, results = "hold"}
M4 = multinom(flight_case ~ mass_per + sex_c, data = df)
model_table = calculate_P2(M4, "mass_per", "sex_c")
```
&nbsp;

Host plant was not a significant predictor, so we tested the wing-to-body ratio as a predictor next.

&nbsp;
&nbsp;

### Compare Models - predictors: % mass, sex, wing2body (w2b)

```{r warning=FALSE}
df$wing2body_c = df$wing2body - mean(df$wing2body) # re-center the w2b predictor
```

```{r warning=FALSE}
data = data.frame(R = df$flight_case, 
                  A = df$mass_per,
                  B = df$sex_c,
                  C = df$wing2body_c)
model_script = paste0(source_path,"generic multinomial models- multinom 1RF + 3 FF.R")
model_comparisonsAIC(model_script)
```

&nbsp;

```{r warning=FALSE, results = "hold"}
anova(m7, m12, test="Chisq") # adding A*C does not improve fit
anova(m7, m13, test="Chisq") # Adding B*C does not improve fit
```

### Best Fit

```{r warning=FALSE, results = "hold"}
M5 = multinom(flight_case ~ mass_per + sex_c + wing2body_c, data = df)
model_table = calculate_P3(M5)
```

&nbsp;

### Visualize Significant Multinomial Functions

```{r warning=FALSE, results = "hold", fig.width=10, fig.height=9}
# define a run_multinom_model function based on the best fit model
run_multinom_model = function(d) {
  m = multinom(flight_case ~ mass_per + sex_c + wing2body_c, trace=FALSE, data = d)
  model_table = calculate_P3(m, print_table=FALSE)
  return(model_table)
}

# determine which multinomial model equations are significant with a plot
par(mfrow=c(2,2))
mass_per_ML = get_significant_models(19) # % mass
  mtext("A. Mass", side=3, adj=0, line=0.5, cex=1.3, font=1)
sex_ML = get_significant_models(20) # sex
  mtext("B. Sex", side=3, adj=0, line=0.5, cex=1.3, font=1)
w2b_ML = get_significant_models(21) # wing2body
  mtext("C. Wing2Body", side=3, adj=0, line=0.5, cex=1.3, font=1)
```


### Plot Predicted Probabilities 

```{r warning=FALSE}
head(pp <- fitted(M4),3) # compute fitted values of the best fit model without wing-to-body ratio
```

```{r warning=FALSE, echo=FALSE}
df$index = 1:nrow(df)
females = df[df$sex=="F",]
males = df[df$sex=="M",]
Frows = females$index
Mrows = males$index
```

```{r warning=FALSE, echo=FALSE, fig.width=3.2*1.7*2, fig.height=2.8*2}
plot1 = function(df, pp) {
  plot(df$mass_per[Frows], pp[Frows,1], ylim=c(0,1), xlim=c(-40,104), col="red",
     type="l",
     ylab="Flight Case Probability", 
     xlab="Percent Mass Change From T1 to T2 (%)", lty=1, cex.axis=1.2, cex.lab=1.3)  
  points(df$mass_per[Mrows], pp[Mrows,1], col="red", type="l", cex=0.45, lty=2) 
  points(df$mass_per[Frows], pp[Frows,2], col="blue", type="l") 
  points(df$mass_per[Mrows], pp[Mrows,2], col="blue", type="l", cex=0.45, lty=2)
  points(df$mass_per[Frows], pp[Frows,3], col="darkgreen", cex=0.45, type="l", pch=16) 
  points(df$mass_per[Mrows], pp[Mrows,3], col="darkgreen", type="l", cex=0.45, lty=2) 
  points(df$mass_per[Frows], pp[Frows,4], col="darkred", type="l") 
  points(df$mass_per[Mrows], pp[Mrows,4], col="darkred", type="l", cex=0.45, lty=2) 
  mtext("A", side=3, adj=0.01, line=0.5, cex=1.5, font=2)
  text(68,0.7, labels="Flew in T1 only", col="blue")
  text(-27,0.9, labels="Did Not Fly", col="red")
  text(13,0.37, labels="Flew Twice", col="darkred")
  legend(83, 1.0,
         legend = c("female","male"),
         lty=1:2,
         col="black",
         cex=1.1)  
}
pp4 <- fitted(M4)
df4 <- df 
plot1(df4,pp4)
```

```{r warning=FALSE}
head(pp <- fitted(M5), 3) # compute fitted values of the best fit model with wing-to-body ratio
```

```{r warning=FALSE, echo=FALSE}
df$index = 1:nrow(df)
females = df[df$sex=="F",]
males = df[df$sex=="M",]
Frows = females$index
Mrows = males$index
```

```{r echo=FALSE, warning=FALSE}
plot2 = function(df, pp, PP, gradient=TRUE, circles=TRUE, stochasticity=TRUE, points=TRUE) {
  
  c1 = 0.65
  c2 = 1.2
  df$w2b_col = 0
  
  if (stochasticity) {
    plot(df$mass_per[Frows], pp[Frows,1], ylim=c(0,1), xlim=c(-40,104), col="red",type="l", 
       ylab="Flight Case Probability", xlab="Percent Change in Mass From T1 to T2 (%)", 
       lty=1, cex.axis=1.2, cex.lab=1.3) 
    points(df$mass_per[Frows], pp[Frows,2], col="blue", type="l")
    points(df$mass_per[Frows], pp[Frows,4], col="darkorange1", type="l") #darkred
  }
  if (points){
    plot(df$mass_per[Frows], PP[Frows,1], ylim=c(0,1), xlim=c(-40,104), col="red",type="l", 
       ylab="Flight Case Probability", xlab="Percent Change in Mass From T1 to T2 (%)", 
       lty=1, cex.axis=1.2, cex.lab=1.3) 
    points(df$mass_per[Frows], PP[Frows,2], col="blue", type="l")
    points(df$mass_per[Frows], PP[Frows,4], col="darkorange1", type="l") #darkred   
  }
  mtext(expression(italic("Females")), side=3, adj=0.05, line=-2, cex=1.5, font=2)
  mtext("B", side=3, adj=0.01, line=0.5, cex=1.5, font=2)
  text(64,0.7, labels="Flew in T1 only", col="blue", cex=c2)
  text(22,0.85, labels="Did Not Fly", col="red", cex=c2) 
  text(82,0.37, labels="Flew Twice", col="darkorange3", cex=c2) #maroon

  if (gradient) {
    rbPal <- colorRampPalette(c('black','red'))
    df$w2b_col <- rbPal(10)[as.numeric(cut(df$wing2body,breaks = 10))]
    points(df$mass_per[Frows], pp[Frows, 1], pch=20, col=df$w2b_col[Frows])
    
    rbPal <- colorRampPalette(c('black','royalblue1'))
    df$w2b_col <- rbPal(10)[as.numeric(cut(df$wing2body,breaks = 10))]
    points(df$mass_per[Frows], pp[Frows, 2], pch=20, col=df$w2b_col[Frows])
    
    rbPal <- colorRampPalette(c('black','orange')) # violetred1
    df$w2b_col <- rbPal(10)[as.numeric(cut(df$wing2body,breaks = 10))]
    points(df$mass_per[Frows], pp[Frows, 4], pch=20, col=df$w2b_col[Frows])
  }

 if (circles) {
  # Mark points in the graph with high wing2body ratio vs. points with low wing2body ratio.
  test <- df[with(df, order(wing2body)),] # ascending order

  small = test %>%
    filter(sex =="F", wing2body < 0.7184934)
  large = test %>%
    filter(sex == "F", wing2body > 0.7184934)
  srows = small$index
  lrows = large$index

  points(df$mass_per[lrows], pp[lrows,1], col="red", type="p", cex=c1, pch=16)
  points(df$mass_per[srows], pp[srows,1], col="red", type="p", cex=c1)
  # those with smaller wing2body ratio were more likely to NOT fly
  points(df$mass_per[lrows], pp[lrows,2], col="blue", type="p", cex=c1, pch=16)
  points(df$mass_per[srows], pp[srows,2], col="blue", type="p", cex=c1)
  points(df$mass_per[lrows], pp[lrows,4], col="darkred", type="p", cex=c1, pch=16)
  points(df$mass_per[srows], pp[srows,4], col="darkred", type="p", cex=c1)
 }
  
}
pp5 = pp
df5 = df
```

```{r echo=FALSE, warning=FALSE}
plot3 = function(df, pp, PP, gradient=TRUE, circles=TRUE, stochasticity=TRUE, points=TRUE) {
  
  c1 = 0.65
  c2 = 1.2
  df$w2b_col = 0

  if (stochasticity) {
    plot(df$mass_per[Mrows], pp[Mrows,1], ylim=c(-0.00,0.8+0.05), xlim=c(-25,58), col="red",type="l", 
       ylab=" ", xlab="Percent Mass Change from T1 to T2 (%)", 
       lty=1, cex.axis=1.2, cex.lab=1.3)
    points(df$mass_per[Mrows], pp[Mrows,2], col="blue", type="l", cex=0.45, lty=1)
    points(df$mass_per[Mrows], pp[Mrows,3], col="darkgreen", type="l", cex=0.45, lty=1) 
    points(df$mass_per[Mrows], pp[Mrows,4], col="darkorange1", type="l", cex=0.45, lty=1) # darkred
  }
  if (points) {
    plot(df$mass_per[Mrows], PP[Mrows,1], ylim=c(-0.00,0.8), xlim=c(-25,58), col="red",type="l", 
       ylab=" ", xlab="Percent Mass Change from T1 to T2 (%)", 
       lty=1, cex.axis=1.2, cex.lab=1.3)
    points(df$mass_per[Mrows], PP[Mrows,2], col="blue", type="l", cex=0.45, lty=1)
    points(df$mass_per[Mrows], PP[Mrows,3], col="darkgreen", type="l", cex=0.45, lty=1) 
    points(df$mass_per[Mrows], PP[Mrows,4], col="darkorange1", type="l", cex=0.45, lty=1) # darkred
  }
  mtext(expression(italic("Males")), side=3, adj=0.05, line=-2, cex=1.5, font=2)
  mtext("C", side=3, adj=0.01, line=0.5, cex=1.5, font=2)
  text(22,0.35, labels="Flew in T1 only", col="blue", cex=c2)
  text(52,0.17, labels="Did Not Fly", col="red", cex=c2)
  text(-16, 0, labels="Flew in T2 only", col="darkgreen", cex=c2)
  text(-17,0.69, labels="Flew Twice", col="darkorange3", cex=c2) # 13,0.77, maroon, 13,0.77

  if (circles) {
    legend(39, .81+0.05, 
           pch=c(16,1), 
           title="Wing-to-body", 
           legend=c("> mean", "< mean"), 
           cex=1.1)
  }
  if (gradient) {
    text(50,0.84, labels="Wing-to-body", cex=c2)
  }
  
  if (gradient) {
    rbPal <- colorRampPalette(c('black','red'))
    df$w2b_col <- rbPal(10)[as.numeric(cut(df$wing2body,breaks = 10))]
    points(df$mass_per[Mrows], pp[Mrows, 1], pch=20, col=df$w2b_col[Mrows])
    
    rbPal <- colorRampPalette(c('black','royalblue1'))
    df$w2b_col <- rbPal(10)[as.numeric(cut(df$wing2body,breaks = 10))]
    points(df$mass_per[Mrows], pp[Mrows, 2], pch=20, col=df$w2b_col[Mrows])
    
    rbPal <- colorRampPalette(c('black','palegreen2'))
    df$w2b_col <- rbPal(10)[as.numeric(cut(df$wing2body,breaks = 10))]
    points(df$mass_per[Mrows], pp[Mrows, 3], pch=20, col=df$w2b_col[Mrows])
    
    rbPal <- colorRampPalette(c('black','orange')) #violetred1
    df$w2b_col <- rbPal(10)[as.numeric(cut(df$wing2body,breaks = 10))]
    points(df$mass_per[Mrows], pp[Mrows, 4], pch=20, col=df$w2b_col[Mrows])
  }

 if (circles) {
  # Mark points in the graph with high wing2body ratio vs. points with low wing2body ratio. 
  test <- df[with(df, order(wing2body)),] # ascending order
  
  small = test %>%
    filter(sex =="M", wing2body < 0.7184934)
  large = test %>%
    filter(sex == "M", wing2body > 0.7184934)
  srows = small$index
  lrows = large$index
  
  points(df$mass_per[lrows], pp[lrows,1], col="red", type="p", cex=c1, pch=16)
  points(df$mass_per[srows], pp[srows,1], col="red", type="p", cex=c1)
  # those with smaller wing2body ratio were more likely to NOT fly
  points(df$mass_per[lrows], pp[lrows,2], col="blue", type="p", cex=c1, pch=16)
  points(df$mass_per[srows], pp[srows,2], col="blue", type="p", cex=c1)
  points(df$mass_per[lrows], pp[lrows,3], col="darkgreen", type="p", cex=c1, pch=16)
  points(df$mass_per[srows], pp[srows,3], col="darkgreen", type="p", cex=c1)
  points(df$mass_per[lrows], pp[lrows,4], col="darkred", type="p", cex=c1, pch=16)
  points(df$mass_per[srows], pp[srows,4], col="darkred", type="p", cex=c1)
 }
}
```

```{r echo=FALSE, fig.width=0.7, fig.height=1, warning=FALSE}
# Function to plot color bar
color.bar <- function(lut, min, max=77, nticks=3, ticks=seq(min, max, len=nticks), title='') {
    scale = (length(lut)-1)/(max-min)
    
    final_ticks=seq(min/100, max/100, len=nticks)
    #final_ticks = c("0.63", "0.66", "0.70", "0.74", "0.77")
    final_ticks = c("0.63", "0.70", "0.77")
    
    #dev.new(width=1.75, height=5)
    plot(c(0,10), c(min,max), type='n', bty='n', xaxt='n', xlab='', yaxt='n', ylab='', main=title)
    axis(2, ticks, las=1, labels=final_ticks)
    for (i in 1:(length(lut)-1)) {
     y = (i-1)/scale + min
     rect(0,y,10,y+1/scale, col=lut[i], border=NA)
    }
}
```

```{r warning=FALSE, echo=FALSE}
# generate small subset plots and scales in the top right hand corner
plot_histograms = function() {
  x = 0.48
  v <- c(x-0.13,x, 0.85, 0.90)
  par( fig=v, new=TRUE, mar=c(0,0,0,0) )
  hist(df$wing2body[Frows], col="white", main="", cex.axis=0.9) 
  text(0.64,45, labels="w2b", cex=0.9)

  x = 0.88
  v <- c(x-0.13, x, 0.85, 0.90)
  par( fig=v, new=TRUE, mar=c(0,0,0,0) )
  hist(df$wing2body[Mrows], col="white", main="", cex.axis=0.9, xlim=c(0.64,0.78)) 
  text(0.66,40, labels="w2b", cex=0.9)
}
plot_color_scale = function() {
  x = 0.98
  v <- c(x-0.025, x, 0.75, 0.86)
  par( fig=v, new=TRUE, mar=c(0,0,0,0) )
  color.bar(colorRampPalette(c("black", "grey"))(1000), 63)
}
```

```{r echo=FALSE, warning=FALSE, fig.width=3.2*1.7*2, fig.height=2.8*2}
par(mfrow=c(1,2), tcl=-0.5) # length of tick marks set at default

par(mai=c(1,0.85,0.4,0)) # bottom, right, top, left
plot2(df5,pp5, pp4, gradient=TRUE, circles=FALSE, stochasticity=TRUE, points=FALSE)

par(mai=c(1,0.6,0.4,0.05))
plot3(df5,pp5, pp4, gradient=TRUE, circles=FALSE, stochasticity=TRUE, points=FALSE)

plot_histograms()
plot_color_scale()
```

## Multinomial Modeling (Females Only)

### Egg Case

```{r echo=FALSE, warning=FALSE}
# key = response in T2 - response in T1
Event = c("laid eggs in both trials", "laid eggs in T2 only", "laid eggs in neither trials", "laid eggs in T1 only")
Encoding = c(2, 1,0,-1)
key = cbind(Event, Encoding)

kable(key) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))  %>%
  kable_classic(html_font = "Cambria") %>%
  add_header_above(c("Delta Egg Response Key" = 2 )) 
```

### Baseline

```{r warning=FALSE, results='hold'}
# filter for females &
# remove any missing values for flight case, mass percent change, and egg case between trials
df = d[with(d,!is.na(flight_case) & !is.na(mass_per) & !is.na(egg_case) & sex=="F"),]

# order the dataset by ascending mass percent change values
df = df[with(df, order(mass_per)),]

# relevel the flight case factors so as to set 0 as the first level.
df$flight_case = relevel(as.factor(df$flight_case), ref = "0")

unique(df$flight_case) # no female bug only flew in T2, so can drop factor "1"
df$flight_case = droplevels(df$flight_case)
```

### Null model

```{r warning=FALSE}
null <- multinom(flight_case ~ 1, data = df) 
```

### Comparing Models - predictors: % mass, egg diff, host

```{r warning=FALSE}
data <- data.frame(R = df$flight_case, 
         A = df$egg_case,
         B = df$mass_per,
         C = df$host_c)
model_script = paste0(source_path,"generic multinomial models- multinom 1RF + 3 FF.R")
model_comparisonsAIC(model_script)
```
&nbsp;

```{r warning=FALSE, results='hold'}
anova(m4, m7, test="Chisq") # Adding C does not improve fit
anova(m7, m13, test="Chisq") # Adding mass_per*host does not improve fit
```

&nbsp;

Host plant was not a significant predictor for females as well, so we tested the wing-to-body ratio as a predictor next.

&nbsp;
&nbsp;

### Comparing Models - predictors: % mass, egg diff, wing2body

```{r warning=FALSE, echo=FALSE}
data <- data.frame(R = df$flight_case, 
         A = df$egg_case,
         B = df$mass_per,
         C = df$wing2body)
model_script = paste0(source_path,"generic multinomial models- multinom 1RF + 3 FF.R")
model_comparisonsAIC(model_script)
```

```{r results='hold', warning=FALSE}
anova(m4, m7, test="Chisq") # adding wing2body does not include fit
anova(m7, m13, test="Chisq") # Adding A*C does not improve fit
anova(m7, m12, test="Chisq") # Adding B*C does not improve fit
```

### Best Fit

```{r results="hold"}
M6 = multinom(flight_case ~ mass_per + egg_case, data = df) # same top model 
model_table = calculate_P2(M6, "mass_per", "egg_case")
```

&nbsp;

```{r warning=FALSE, results = "hold", fig.width=11, fig.height=5}
# define a run_multinom_model function based on the best fit model
run_multinom_model = function(d) {
  m <- multinom(flight_case ~ mass_per + egg_case, trace=FALSE, data = d) 
  model_table = calculate_P2(m, "mass_per", "egg_case", print_table=FALSE)
  return(model_table)
}

# determine which multinomial model equations are significant with a plot
par(mfrow=c(1,2)) 
mass_per_ML = get_significant_modelsf(15) # mass_per 
  mtext("A. Mass", side=3, adj=0, line=0.5, cex=1.6, font=1)
egg_case_ML = get_significant_modelsf(16) # egg_case
  mtext("B. Egg \nResponse", side=3, adj=0, line=0.3, cex=1.6, font=1)
```

### Barplot

```{r}
data_fem = data_tested[data_tested$sex=="F",]
binary_counts <- table(data_fem$eggs_b, data_fem$trial_type)[,2:3]
```

```{r echo=FALSE, warning=FALSE}
par(mfrow=c(1,2))

barplot(binary_counts, ylab="Frequency",
        xlab="Trial", col=c("#b57783","#820331"),
        legend = c("no eggs", "yes eggs"), ylim=c(0,110), beside=TRUE)
```

&nbsp;
Notice that female bugs were laying more during the second trial (T2) than the first trial (T1).
&nbsp;

### Plot Predicted Probabilities 

```{r warning=FALSE}
head(pp <- fitted(M6),3)
```

```{r echo=FALSE, warning=FALSE}
df$index = 1:nrow(df)

eggT1 = df[df$egg_case == -1,]
egg0 = df[df$egg_case == 0,]
egg2 = df[df$egg_case == 2,]
eggT2 = df[df$egg_case == 1,]

eggT1_rows = eggT1$index
egg_0rows = egg0$index
egg_2rows = egg2$index
eggT2_rows = eggT2$index
```


```{r echo=FALSE, fig.width=3.2*1.7*2, fig.height=2.8*2}
plot4 = function(df, pp) {
  # only laid eggs in T1
  plot(df$mass_per[eggT1_rows], pp[eggT1_rows,1], ylim=c(0,1.05), xlim=c(-36,108), col="red", type="l", lty=1,main="Females Only", ylab="Flight Case Probability", xlab="Percent Change in Mass From T1 to T2 (g)")
  
  points(df$mass_per[eggT1_rows], pp[eggT1_rows,2], col="blue", type="l", lty=1, cex=0.45) 
  points(df$mass_per[eggT1_rows], pp[eggT1_rows,3], col="black", type="l", lty=1, cex=0.45) 
  # no egg change
  points(df$mass_per[egg_0rows], pp[egg_0rows,1], col="red", type="l", lty=2) # did not fly in either
  points(df$mass_per[egg_0rows], pp[egg_0rows,2], col="blue", type="l", lty=2) # flew in T1 only
  points(df$mass_per[egg_0rows], pp[egg_0rows,3], col="black", type="l", lty=2) # flew in both
  # eggs twice
  points(df$mass_per[egg_2rows], pp[egg_2rows,1], col="red", type="l", lty=4) # did not fly in either
  points(df$mass_per[egg_2rows], pp[egg_2rows,2], col="blue", type="l", lty=4) # flew in T1 only
  points(df$mass_per[egg_2rows], pp[egg_2rows,3], col="black", type="l", lty=4) # flew in both
  # only laid eggs in T2
  points(df$mass_per[eggT2_rows], pp[eggT2_rows,1], col="red", type="l", lty=3) # flew in neither
  points(df$mass_per[eggT2_rows], pp[eggT2_rows,2], col="blue", type="l", lty=3) # flew in T1 only
  points(df$mass_per[eggT2_rows], pp[eggT2_rows,3], col="black", type="l", lty=3) # flew in both
  
  text(70,0.68, labels="Flew in T1 only", col="blue") # 0.059,0.54
  text(-20,0.95, labels="Did Not Fly", col="red") # -0.036,0.95
  text(-20,0.55, labels="Flew Twice", col="black") # -0.02,0.35
  legend(76, 1.07,
         legend = c("laid eggs in T1","no eggs laid", "eggs laid twice", "laid eggs in T2"),
         lty=1:4,
         col="black",
         cex=0.8)
}
pp6 = fitted(M6)
df6 = df
plot4(df6,pp6) 
```

&nbsp;

# Flight Response Predictions

## Read Libraries 

```{r}
library(cvms) # cross-validating regressions
```

## Read Source Files 

```{r warning=FALSE, message=FALSE}
script_names = c("clean_flight_data-Fall2.R", # 1 function: clean_flight_data.Fall()
                 "unique_flight_data-Fall.R", # 1 function: create_delta_data.Fall()
                 "prediction_accuracy.R", # 1 function: calculate_accuracy()
                 "confusion_matrix.R") # 1 function: get_confusion_matrix()

for (script in script_names) { 
  path = paste0(source_path, script)
  source(path) 
}
```

## Read the Data

```{r warning=FALSE}
dataFall = clean_flight_data.Fall("data/full_data-Fall2019.csv")

# extract sets with an experimental design similar to the Winter tests
ongoing_data = dataFall[with(dataFall,!is.na(mass) & set_number > 71),]

# create delta data
d = create_delta_data.Fall(ongoing_data) 
```

## Compute predicted probabilities

```{r}
d <- d[with(d, order(mass_per)),]
```

```{r}
neither = c()
T1_rather_than_none = c()
T2_rather_than_none = c()
both_rather_than_none = c()

for (i in 1:nrow(d)) {
  m = d$mass_per[[i]]
  s = d$sex_c[[i]]
  #w = d$wing[[i]][1] / d$body[[i]][1]
  w = d$wing2body_c[i]
  top0 = exp(0) # just equals 1
  #top1 = exp(-17.862 + 0.041*m - 0.571*s + 23.558*w )
  #top2 = exp(-4.395 - 0.005*m - 9.580*s - 8.937*w)
  #top3 = exp(-19.931 + 0.018*m - 0.760*s + 28.019*w)
  top1 = exp(-0.935 + 0.041*m - 0.571*s + 23.739*w)
  top2 = exp(-8.177 - 0.005*m - 6.954*s - 6.595*w)
  top3 = exp(0.201 + 0.018*m - 0.760*s + 28.094*w)
  bottom = top0 + top1 + top2 + top3
  neither = c(neither, top0/bottom)
  T1_rather_than_none = c(T1_rather_than_none, top1/bottom)
  T2_rather_than_none = c(T2_rather_than_none, top2/bottom)
  both_rather_than_none = c(both_rather_than_none, top3/bottom)
}
```

## Plot predicted probabilities

```{r echo=FALSE}
d$index = 1:nrow(d)
females = df[df$sex=="F",]
males = df[df$sex=="M",]
Frows = females$index
Mrows = males$index
```

```{r echo=FALSE}
plot(d$mass_per[Frows], T1_rather_than_none[Frows], ylim=c(0,1), xlim=c(-40,104), col="blue", type="l",
     ylab="Flight Case Probability", xlab="Percent Change in Mass From T1 to T2 (%)", lty=1) #T1 only
points(d$mass_per[Mrows], T1_rather_than_none[Mrows], col="blue", type="l", lty=2)
points(d$mass_per[Frows], T2_rather_than_none[Frows], col="darkgreen", type="l", lty=1)
points(d$mass_per[Mrows], T2_rather_than_none[Mrows], col="darkgreen", type="l", lty=2)
points(d$mass_per[Frows], both_rather_than_none[Frows], col="maroon", type="l", lty=1)
points(d$mass_per[Mrows], both_rather_than_none[Mrows], col="maroon", type="l", lty=2)
points(d$mass_per[Frows], neither[Frows], col="red", type="l", lty=1)
points(d$mass_per[Mrows], neither[Mrows], col="red", type="l", lty=2)

text(-20,0.85, labels="Did Not Fly", col="red")
text(0,0.5, labels="Flew Twice", col="maroon")
text(40,0.25, labels="Flew in T1 only", col="blue")
text(30,0.1, labels="Flew in T2 only", col="darkgreen")

legend(84, 1.02,
       legend = c("female","male"),
       lty=1:2,
       col="black",
       cex=0.8)
```

## Overall and Grouped Accuracies

```{r}
probs = round(cbind(neither, T1_rather_than_none, T2_rather_than_none, both_rather_than_none),2)
summary_probs = cbind(as.character(d$flight_case), as.character(d$sex), probs)
colnames(summary_probs) = c("event", "sex", "none", "T1", "T2", "both")
dataframe = as.data.frame(summary_probs)
nrow(dataframe)
```

```{r}
# overall
acc = calculate_accuracy(dataframe,3,6)
paste("Overall prediction accuracy, ", round(acc,2))

# by sex
femdata = dataframe[dataframe$sex=="F",]
maledata = dataframe[dataframe$sex=="M",]

accF = calculate_accuracy(femdata,3,6)
paste("Female prediction accuracy, ", round(accF,2))
accM = calculate_accuracy(maledata,3,6)
paste("Male prediction accuracy, ", round(accM,2))
```

## Confusion Matrix

```{r}
acc_table = get_confusion_matrix(dataframe,3,6)
acc_table
confusion_matrix <- acc_table$'Confusion Matrix'[[1]]
confusion_matrix
plot_confusion_matrix(confusion_matrix, add_sums=FALSE)
```

# Females Only

```{r}
dfem = d[d$sex=="F",]
```
```{r}
dfem <- dfem[with(dfem, order(mass_diff)),]
```

```{r}
neither = c()
T1_rather_than_none = c()
both_rather_than_none = c()
for (i in 1:nrow(dfem)) {
  md = dfem$mass_diff[[i]]
  ed = dfem$egg_diff[[i]]
  top0 = exp(0) # just equals 1
  top1 = exp(-0.88 - 0.53*ed + 57.43*md )
  top2 = exp(-0.53 - 1.09*ed + 18.67*md)
  bottom = top0 + top1 + top2
  neither = c(neither, top0/bottom)
  T1_rather_than_none = c(T1_rather_than_none, top1/bottom)
  both_rather_than_none = c(both_rather_than_none, top2/bottom)
}
```

## Compute predicted probabilities

```{r}
probs = round(cbind(neither, T1_rather_than_none, both_rather_than_none),2)
summary_probs = cbind(as.character(dfem$flight_case), as.character(dfem$egg_diff), probs)
colnames(summary_probs) = c("event", "egg_diff", "none", "T1", "both")

egg2 = c(1,2,3,5,6,7,9,10,11,13)
noegg = c(4,8,12)

dataframe = as.data.frame(summary_probs)
dataframe$egg_cat = c(2,2,2,0,2,2,2,0,2,2,2,0,2)
```

## Plot predicted probabilities

```{r echo=FALSE}
plot(dfem$mass_diff[egg2], T1_rather_than_none[egg2], ylim=c(0,1), col="blue", type="l",
     ylab="Flight Case Probability", xlab="Change in Mass From T1 to T2 (g)", main="Females Only", lty=2) #T1 only
points(dfem$mass_diff[noegg], T1_rather_than_none[noegg], col="blue", type="l", lty=1)
points(dfem$mass_diff[egg2], both_rather_than_none[egg2], col="maroon", type="l", lty=2)
points(dfem$mass_diff[noegg], both_rather_than_none[noegg], col="maroon", type="l", lty=1)
points(dfem$mass_diff[egg2], neither[egg2], col="red", type="l", lty=2)
points(dfem$mass_diff[noegg], neither[noegg], col="red", type="l", lty=1)

text(0,0.75, labels="Did Not Fly", col="red")
text(0,0.35, labels="Flew Twice", col="maroon")
text(0,0.15, labels="Flew in T1 only", col="blue")
legend(0.01, 0.98,
       legend = c("no eggs","2 eggs"),
       lty=1:2,
       col="black",
       cex=0.8)
```

## Overall and Grouped Accuracies

```{r}
accF_eggs = calculate_accuracy(dataframe,3,5)
paste("Female prediction accuracy for mass diff and egg model, ", round(accF_eggs,2))
```

## Confusion Matrix

```{r}
acc_table = get_confusion_matrix(dataframe,3,5)
acc_table
confusion_matrix <- acc_table$'Confusion Matrix'[[1]]
confusion_matrix
plot_confusion_matrix(confusion_matrix, add_sums=FALSE)
```

```{r echo=FALSE}
#library(knitr) # table formatting
#library(dplyr) # data manipulation 
#library(ggformula) # ggplot plotting
#library(cowplot) # ggplot helper functions to arrange multi-panel figures
```




