---
title: "Predicting Fall"
author: "Anastasia Bernat"
date: "1/25/2021"
output: html_document
---

```{r setup, include=FALSE}
rm(list=ls())
dir = "~/Desktop/git_repositories/SBB-dispersal/avbernat_working_on/Dispersal/Winter_2020/stats/"
setwd(dir) 

library(dplyr)
library(cvms)


knitr::opts_chunk$set(echo = TRUE)
```

Using delta_yes_no.Rmd models to predict Fall delta flight response.

**Clean the data**

```{r}
source("src/clean_flight_data-Fall.R")
data = clean_flight_data.Fall("data/all_flight_data-Fall2019.csv")
```

**Keeping only the tests that were similar to Winter experiment design**

```{r}
data_mass = data %>%
  filter(!is.na(mass))

# 60 min trial
data60 = data_mass %>%
  filter(set_number < 53)

# 90 min trial
data90 = data_mass %>%
  filter(set_number < 72 & set_number > 52)

# ongoing trial
ongoing_data = data_mass %>%
  filter(set_number > 71)
```

**Create delta data**

```{r}
source("src/unique_flight_data-Fall.R")
d = create_delta_data.Fall(ongoing_data)
```

**Calculating individual predicted probabilities**

This is the format:

<div align="center">$\pi_j = \frac{e^{\alpha_j + \beta_j x}}{\sum_he^{\alpha_h + \beta_h x}}, j=1,...,J$</div> <br/>

Here are the equations:

- [1] "Where F = 1"
- [1] "log(pi_-1 / pi_1) = 5.81 + 0.05 Mass Percent Change + 4.93 Sex      Flew in T1, rather than T2"    
- [2] "log(pi_2 / pi_-1) = 1.14 + -0.02 Mass Percent Change + -0.21 Sex      Flew in both, rather than T1"
- [3] "log(pi_2 / pi_1) = 6.94 + 0.03 Mass Percent Change + 4.72 Sex      Flew in both, rather than T2"

- [4] "log(pi_-1 / pi_0) = -1.02 + 0.043 Mass Percent Change - 0.69 Sex      Flew in T1, rather than none"    
- [5] "log(pi_1 / pi_0) = -6.82 - 0.009 Mass Percent Change - 5.63 Sex      Flew in both, rather than none"
- [6] "log(pi_2 / pi_0) = 0.12 + 0.019 Mass Percent Change - 0.90 Sex      Flew in both, rather than none"

**Codifying individual predicted probability equations**

```{r}
d <- d[with(d, order(mass_per)),]
```

```{r}
neither = c()
T1_rather_than_none = c()
T2_rather_than_none = c()
both_rather_than_none = c()
for (i in 1:nrow(d)) {
  mp = d$mass_per[[i]]
  s = d$sex_c[[i]]
  top0 = exp(0) # just equals 1
  top1 = exp(-1.02 + 0.043*mp - 0.69*s)
  top2 = exp(-6.82 - 0.009*mp - 5.63*s)
  top3 = exp(0.12 + 0.019*mp- 0.90*s)
  bottom = top0 + top1 + top2 + top3
  neither = c(neither, top0/bottom)
  T1_rather_than_none = c(T1_rather_than_none, top1/bottom)
  T2_rather_than_none = c(T2_rather_than_none, top2/bottom)
  both_rather_than_none = c(both_rather_than_none, top3/bottom)
}
```

**Plotting predicted probabilities**

```{r}
d$index = 1:nrow(d)
females = d %>%
  filter(sex=="F")
males = d %>%
  filter(sex=="M")
Frows = females$index
Mrows = males$index
```

```{r}
plot(d$mass_per[Frows], T1_rather_than_none[Frows], ylim=c(0,1), xlim=c(-40,104), col="blue", type="l",
     ylab="Flight Case Probability", xlab="Percent Change in Mass From T1 to T2 (%)", lty=1) #T1 only
points(d$mass_per[Mrows], T1_rather_than_none[Mrows], col="blue", type="l", lty=2)
points(d$mass_per[Frows], T2_rather_than_none[Frows], col="darkgreen", type="l", lty=1)
points(d$mass_per[Mrows], T2_rather_than_none[Mrows], col="darkgreen", type="l", lty=2)
points(d$mass_per[Frows], both_rather_than_none[Frows], col="maroon", type="l", lty=1)
points(d$mass_per[Mrows], both_rather_than_none[Mrows], col="maroon", type="l", lty=2)
points(d$mass_per[Frows], neither[Frows], col="red", type="l", lty=1)
points(d$mass_per[Mrows], neither[Mrows], col="red", type="l", lty=2)

text(-20,0.9, labels="Did Not Fly", col="red")
text(0,0.5, labels="Flew Twice", col="maroon")
text(30,0.85, labels="Flew in T1 only", col="blue")
text(30,0.1, labels="Flew in T2 only", col="darkgreen")

legend(84, 1.02,
       legend = c("female","male"),
       lty=1:2,
       col="black",
       cex=0.8)
```


```{r echo=FALSE}
plot(d$mass_per, T1_rather_than_none, ylim=c(0,1), xlim=c(-25, 40), col="blue", ylab="Flight Case Probability", xlab="Percent Changes in Mass From T1 to T2 (g)", pch=2, cex=0.45) # T1 only
points(d$mass_per, T2_rather_than_none, col="darkgreen", pch=3, cex=0.45) # T2 only
points(d$mass_per, both_rather_than_none, col="maroon", cex=0.45) # flew in both
points(d$mass_per, neither, col="red", cex=0.45) # flew in neither

text(-20,0.9, labels="Did Not Fly", col="red")
text(0,0.5, labels="Flew Twice", col="maroon")
text(30,0.55, labels="Flew in T1 only", col="blue")
text(30,0.1, labels="Flew in T2 only", col="darkgreen")
```

Notice that the percent change masses are much lower than the scale we had in the Winter. This suggests that there are a lot of males and few females.

**Accuracy**

```{r}
probs = round(cbind(neither, T1_rather_than_none, T2_rather_than_none, both_rather_than_none),2)
summary_probs = cbind(as.character(d$flight_case), as.character(d$sex), probs)
colnames(summary_probs) = c("event", "sex", "none", "T1", "T2", "both")
dataframe = as.data.frame(summary_probs)
```

```{r}
calculate_accuracy = function(data, cfirst, clast) {
  
  probs = data[cfirst:clast]
  most_likely_events = colnames(probs)[apply(probs,1,which.max)]
  actual_events = c()
  
  for (i in 1:nrow(data)) {
    if (data[i,1] == "0") {
      actual_event = "none"
    }
    if (data[i,1] == "-1") {
      actual_event = "T1"
    }
    if (data[i,1] == "1") {
      actual_event = "T2"
    }
    if (data[i,1] == "2") {
      actual_event = "both"
    }
    actual_events = c(actual_events, actual_event)
  }
  
  events = actual_events == most_likely_events
  accurate_events = sum(events)
  total_events = length(events)
  accuracy = accurate_events / total_events
  
  return(accuracy)
}
```

**Overall and Grouped Accuracies**

```{r}
# overall
acc = calculate_accuracy(dataframe,3,6)
paste("Overall prediction accuracy, ", round(acc,2))

# by sex
femdata = dataframe %>%
  filter(sex=="F")
maledata = dataframe %>%
  filter(sex=="M")

accF = calculate_accuracy(femdata,3,6)
paste("Female prediction accuracy, ", round(accF,2))
accM = calculate_accuracy(maledata,3,6)
paste("Male prediction accuracy, ", round(accM,2))
```

It seems like for males, this model will lead you to overestimating its flight probability. Males will not only fly twice, but they will be most likely to fly twice. Additionally, predicting female flight is much less accurate because of egg-laying events. So, let's use the mass + eggs laid model to predict female flight case probabilities. 

**Confusion Matrix**

For details see: https://medium.com/analytics-vidhya/calculating-accuracy-of-an-ml-model-8ae7894802e

and  

https://stats.stackexchange.com/questions/179835/how-to-build-a-confusion-matrix-for-a-multiclass-classifier

4 X 4 Matrix 

```{r echo=FALSE}
get_confusion_matrix = function(data, cfirst, clast) {
  
  probs = data[cfirst:clast]
  most_likely_events = colnames(probs)[apply(probs,1,which.max)]
  actual_events = c()
  
  for (i in 1:nrow(data)) {
    if (data[i,1] == "0") {
      actual_event = "none"
    }
    if (data[i,1] == "-1") {
      actual_event = "T1"
    }
    if (data[i,1] == "1") {
      actual_event = "T2"
    }
    if (data[i,1] == "2") {
      actual_event = "both"
    }
    actual_events = c(actual_events, actual_event)
  }
  
  df = as.data.frame(cbind(actual_events, most_likely_events))
  #final_df = dplyr::count_(df, vars = c('actual_events','most_likely_events'))
  
  eval <- evaluate(df,
                 target_col = "actual_events",
                 prediction_cols = "most_likely_events",
                 type = "multinomial")
  
  return(eval)
}

```

```{r}
acc_table = get_confusion_matrix(dataframe,3,6)
acc_table
```

```{r}
confusion_matrix <- acc_table$'Confusion Matrix'[[1]]
confusion_matrix
plot_confusion_matrix(confusion_matrix, add_sums=TRUE)
```

For more on confusion maxtrices see: https://cran.r-project.org/web/packages/cvms/vignettes/Creating_a_confusion_matrix.html

**Females Only**

```{r}
d = d %>%
  filter(sex=="F")
```

*Problem with these models is that there is no case in which females only fly in T2 so that's a possible underestimation in our model.*

- [1] "log(pi_-1 / pi_1) = -1.15 + 0.66 Egg Change + 31.91 Mass Change       Flew in T1, rather than T2 trials"
- [2] "log(pi_-1 / pi_0) = -2.09 + 2.02 Egg Change + 44.75 Mass Change       Flew in T1, rather than none"
- [3] "log(pi_2 / pi_0) = -0.94 + 1.36 Egg Change + 12.85 Mass Change       Flew in T1, rather than none"

```{r}
d <- d[with(d, order(mass_diff)),]
```

```{r}
neither = c()
T1_rather_than_none = c()
both_rather_than_none = c()
for (i in 1:nrow(d)) {
  md = d$mass_diff[[i]]
  ed = d$egg_diff[[i]]
  top0 = exp(0) # just equals 1
  top1 = exp(-2.09 + 2.02*ed + 44.75*md )
  top2 = exp(-0.94 + 1.36*ed + 12.85*md)
  bottom = top0 + top1 + top2
  neither = c(neither, top0/bottom)
  T1_rather_than_none = c(T1_rather_than_none, top1/bottom)
  both_rather_than_none = c(both_rather_than_none, top2/bottom)
}
```

**Plotting predicted probabilities**

```{r}
plot(d$mass_diff, T1_rather_than_none, ylim=c(0,1), col="blue", type="l",
     ylab="Flight Case Probability", xlab="Change in Mass From T1 to T2 (g)", main="Females Only", lty=1) #T1 only
points(d$mass_diff, both_rather_than_none, col="maroon", type="l", lty=1)
points(d$mass_diff, neither, col="red", type="l", lty=1)

text(0,0.75, labels="Did Not Fly", col="red")
text(0,0.35, labels="Flew Twice", col="maroon")
text(0,0.15, labels="Flew in T1 only", col="blue")
```

**Accuracy**

```{r}
probs = round(cbind(neither, T1_rather_than_none, both_rather_than_none),2)
summary_probs = cbind(as.character(d$flight_case), as.character(d$egg_diff), probs)
colnames(summary_probs) = c("event", "egg_diff", "none", "T1", "both")

egg2 = c(1,2,3,5,6,7,9,10,11,13)
noegg = c(4,8,12)

dataframe = as.data.frame(summary_probs)
dataframe$egg_cat = c(2,2,2,0,2,2,2,0,2,2,2,0,2)
```

**Re-plot**

```{r echo=FALSE}
plot(d$mass_diff[egg2], T1_rather_than_none[egg2], ylim=c(0,1), col="blue", type="l",
     ylab="Flight Case Probability", xlab="Change in Mass From T1 to T2 (g)", main="Females Only", lty=2) #T1 only
points(d$mass_diff[noegg], T1_rather_than_none[noegg], col="blue", type="l", lty=1)
points(d$mass_diff[egg2], both_rather_than_none[egg2], col="maroon", type="l", lty=2)
points(d$mass_diff[noegg], both_rather_than_none[noegg], col="maroon", type="l", lty=1)
points(d$mass_diff[egg2], neither[egg2], col="red", type="l", lty=2)
points(d$mass_diff[noegg], neither[noegg], col="red", type="l", lty=1)

text(0,0.75, labels="Did Not Fly", col="red")
text(0,0.35, labels="Flew Twice", col="maroon")
text(0,0.15, labels="Flew in T1 only", col="blue")
legend(0.01, 0.98,
       legend = c("no eggs","2 eggs"),
       lty=1:2,
       col="black",
       cex=0.8)
```

**Calculate Accuracy**

```{r}
accF_eggs = calculate_accuracy(dataframe,3,5)
paste("Female prediction accuracy for mass diff and egg model, ", round(accF_eggs,2))
```
Female flight was underestimated, and, considering that the prediction for females is so low, this leads me to believe that *season* is an important factor in predicting flight case probability.

**Season**

```{r}
source("src/clean_flight_data.R") 
source("src/center_flight_data.R")
data <- read_flight_data("data/all_flight_data-Winter2020.csv")
data_all <- data[[1]]
data_tested <- data[[2]] 
source("src/unique_flight_data.R")
d_all <- create_delta_data(data_tested) # winter data
winter <- center_data(d_all, is_not_binded = FALSE)

source("src/clean_flight_data-Fall.R")
d = clean_flight_data.Fall("data/all_flight_data-Fall2019.csv")
source("src/unique_flight_data-Fall.R")
fall = create_delta_data.Fall(d)

fall$season <- "fall"
winter$season <- "winter"
```

```{r}
# just extract the first two trials 
fall$num_flew <- 0
fall$num_notflew <- 0
  
for(row in 1:length(fall$flew_b)){
    n_flew <- sum(fall$flew_b[[row]][1:2] == 1)
    fall$num_flew[[row]] <- n_flew 
    
    n_notflew <- sum(fall$flew_b[[row]][1:2] == 0)
    fall$num_notflew[[row]] <- n_notflew
    
}
```

```{r}
sd = rbind(fall, winter) # unique
```

```{r}
sd$f_prob<-sd$num_flew/(sd$num_flew+sd$num_notflew)
```

```{r}
library(binom)
####season vs. flight prob
data_temp <- aggregate(f_prob~season, data=sd, FUN=mean)
data_temp$trials <-c(sum(sd$num_flew[sd$season=="fall"]+sd$num_notflew[sd$season=="fall"]), sum(sd$num_flew[sd$season=="winter"]+sd$num_notflew[sd$season=="winter"]))

#calculate binomial confidence interval
data_temp$successes <- c(sum(sd$num_flew[sd$season=="fall"]), sum(sd$num_flew[sd$season=="winter"]))
data_temp$CI<-binom.confint(data_temp$successes, data_temp$trials, methods="exact")
```


```{r}
par(mai=c(1, 1, .2, .2), ps=22) # fig.height=2.7, fig.width=2.7

plot(data_temp$CI$mean~c(1,2), xaxt='n', ylab="Flight probability", xlab="Season", ylim=c(0,1), xlim=c(0.5,2.5), cex=3, pch=19, col=c("red", "blue"))

lines(x=xy.coords(x=c(1,1), y=c(data_temp$CI$lower[1], data_temp$CI$upper[1])), lwd=3, col="red")
lines(x=xy.coords(x=c(2,2), y=c(data_temp$CI$lower[2], data_temp$CI$upper[2])), lwd=3, col="blue")

axis(side=1, at=c(1,2), labels=c("fall", "winter"))

mtext(text=c("N=145", "N=281"), at=c(1,2), side=1, line=-1.2)
```

Seems like there is some difference by season for the first two trials of each bug (however, experimental setup differed). 

The low accuracies could also be due to age - a factor not controlled for. Possibly younger reproductively active females fly more than older reproductively active females? It's unclear. 

