---
title: "Modeling Flight Response Summary File"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
rm(list=ls())

# set working directory
dir = "~/Desktop/git_repositories/SBB-dispersal/avbernat_working_on/"
setwd(dir)

knitr::opts_chunk$set(echo = TRUE)
```

# Data Cleaning And Exploration

## Read Libraries

```{r message=FALSE, warning=FALSE}
library(lme4) # fit regressions 
library(rethinking) # Bayesian data analysis and plotting
library(popbio) # logistic regression plotting
library(binom) # binomial confidence intervals
#library(dplyr) # data manipulation 
#library(ggformula) # ggplot plotting
#library(cowplot) # ggplot helper functions to arrange multi-panel figures
```

## Read Source Files

```{r message=FALSE, warning=FALSE}
source_path = paste0(dir,"/Rsrc/")

script_names = c("center_flight_data.R", # 1 function: center_data()
                 "clean_flight_data.R", # 1 function: clean_flight_data()
                 "unique_flight_data.R", # 1 function: create_delta_data()
                 "compare_models.R", # 1 function: model_comparisonsAIC()
                 "AICprobabilities.R") # 1 function: AICprobs()

for (script in script_names) { 
  path = paste0(source_path, script)
  source(path) 
}
```

## Read the Data

```{r message=FALSE, warning=FALSE}
data <- read_flight_data("data/all_flight_data-Winter2020.csv")
data_all <- data[[1]]
data_tested <- data[[2]]
d <- create_delta_data(data_tested, tested_more_than_once=FALSE)
```

## Repeating Plot Parameters & Functions

```{r}
# scale/magnifications

c1 = 1.5 # size of points
c2 = 1.2 # size of text
c3 = 2 # size of title 

# compute confidence interval
get_CI = function(x,y,m) {
  x.seq = seq(min(x) - sd(x), max(x) + sd(x), length.out=100)
  prd <- data.frame(x=x.seq) # newdata
  err <- predict(m, newdata = prd, se.fit = TRUE)
  prd$lci <- err$fit - 1.96 * err$se.fit
  prd$fit <- err$fit
  prd$uci <- err$fit + 1.96 * err$se.fit
  mu_ci <- t(matrix(c(prd$lci,prd$uci), ncol=2))
  return(list(mu_ci, prd))
}

# tailoring variables for plotting
d$mass_block<-round(d$average_mass/0.005)*0.005
d$wing2body_block<-round(d$wing2body, digits=2)
d$days_block<-round(d$avg_days, digits=0)
```

## Across-Trial Flight Response

### Experimental Effects

```{r echo=FALSE, fig.height=2.7*2, fig.width=6*2}
# day tested: flight prob vs. days from start

par(mfrow=c(1,2))

# plot A (for linear regression)
dt = aggregate(flew_b ~ days_from_start, data=data_tested, FUN=mean)
x=dt$days_from_start
y=dt$flew_b
m <- glm(y ~ x, data=dt, family="gaussian")

shading = get_CI(x,y,m)
mu_ci=shading[[1]]
prd=shading[[2]]

plot(dt$flew_b~dt$days_from_start, 
     ylab="Flight Probability", xlab="Days From Start", 
     pch=19, 
     col=1, 
     cex=c1,
     cex.lab=c2)
abline(m, lty=2)
shade(mu_ci, lim = prd$x)
pval <- paste0("p = ", round(summary(m)$coefficients[8],3), "*")
text(11, 0.45, pval, cex=c2)
mtext("A", side=3, adj=0, line=0.5, cex=c3, font=1)


# plot B (for logistic regression)
m = glm(flew_b ~ days_from_start, data=data_tested, family = "binomial" )
pvalue <- paste0("p = ", round(summary(m)$coefficients[8],3), "*")

logi.hist.plot(data_tested$days_from_start,data_tested$flew_b,boxp=FALSE,type="hist",col="indianred",
               ylabel="Flight Probability", xlabel="Days From Start")
text(11, 0.45, pvalue, cex=c2)
mtext("B", side=3, adj=0, line=0.5, cex=c3, font=1)
```

```{r echo=FALSE, fig.height=2.7*2, fig.width=6*2}
# time tested: recording duration (HR) vs. trial start time (HR)

par(mfrow=c(1,2))

# plot C (all bugs)
x=data_tested$hr_start
y=data_tested$recording_duration/60/60
m <- glm(y ~ x, data=data_tested, family="gaussian") # marginal effect

shading = get_CI(x,y,m)
mu_ci=shading[[1]]
prd=shading[[2]]

plot(data_tested$hr_start, data_tested$recording_duration/60/60,
     xlab="Trial Start Time (HR)", 
     ylab="Recording Duration (HR)")
abline(m, lty=2)
shade(mu_ci, lim = prd$x)
pval <- paste0("p = ", round(summary(m)$coefficients[8],3))
text(15, 11, pval, cex=c2)
mtext("C", side=3, adj=0, line=0.5, cex=c3, font=1)


# plot D (filter out bugs that didn't fly)
dt=data_tested %>%
  filter(flew_b !=0)
x=dt$hr_start
y=dt$recording_duration/60/60
m <- glm(y ~ x, data=dt, family="gaussian") # marginal effect

shading = get_CI(x,y,m)
mu_ci=shading[[1]]
prd=shading[[2]]

plot(dt$hr_start, dt$recording_duration/60/60,
     xlab="Trial Start Time (HR)", 
     ylab="Recording Duration (HR)")
abline(m, lty=2)
shade(mu_ci, lim = prd$x)
pval <- paste0("p = ", round(summary(m)$coefficients[8],3), "*")
text(15, 11, pval, cex=c2)
mtext("D", side=3, adj=0, line=0.5, cex=c3, font=1)
```

There was a negative of day a bug was tested but only when the full dataset is considered (not the unique dataset).
There was a negative effect of the trial start time but only after removing bugs that didn't fly.

### Single-Variate Effects

```{r echo=FALSE, fig.height=2.7*2*2, fig.width=6*2}

par(mfrow=c(2,2))

# plot A (flight probability grouped by sex)

data_temp <- aggregate(flew_prob~sex, data=d, FUN=mean)
data_temp$trials <-c(sum(d$num_flew[d$sex=="F"]+d$num_notflew[d$sex=="F"]), sum(d$num_flew[d$sex=="M"]+d$num_notflew[d$sex=="M"]))

# calculate binomial confidence interval
data_temp$successes <- c(sum(d$num_flew[d$sex=="F"]), sum(d$num_flew[d$sex=="M"]))
data_temp$CI<-binom.confint(data_temp$successes, data_temp$trials, methods="exact")


plot(data_temp$CI$mean~c(1,2), xaxt='n', 
     ylab="Flight probability", 
     xlab="Sex", ylim=c(0,1), 
     xlim=c(0.5,2.5), 
     cex=3, pch=19, col=c("red", "blue"))
lines(x=xy.coords(x=c(1,1), y=c(data_temp$CI$lower[1], data_temp$CI$upper[1])), lwd=3, col="red")
lines(x=xy.coords(x=c(2,2), y=c(data_temp$CI$lower[2], data_temp$CI$upper[2])), lwd=3, col="blue")
axis(side=1, at=c(1,2), labels=c("female", "male"))
mtext(text=c("N=120", "N=213"), at=c(1,2), side=1, line=-1.2)
mtext("A", side=3, adj=0, line=0.5, cex=c3, font=1)

# plot B (wing2body ratio vs. flew_prob grouped by sex)

data_temp<-aggregate(flew_prob~sex*wing2body_block, data=d, FUN=mean)
data_temp$n<-aggregate(flew_prob~sex*wing2body_block, data=d, FUN=length)$flew_prob

d1 = data_temp[data_temp$sex=="F",]
d2 = data_temp[data_temp$sex=="M",]

# females
x=d1[,"wing2body_block"]
y=d1[,"flew_prob"] 
m1 <- lm(y ~ x, data=d1)
shading=get_CI(x,y,m1)
mu_ci1=shading[[1]]
prd1=shading[[2]]

# males
x=d2[,"wing2body_block"]
y=d2[,"flew_prob"] 
m2 <- lm(y ~ x, data=d2)
shading=get_CI(x,y,m2)
mu_ci2=shading[[1]]
prd2=shading[[2]]

plot(data_temp$flew_prob~data_temp$wing2body_block, 
     ylab="Flight probability", xlab="wing-to-body ratio", 
     pch=19, col=c("red","blue")[as.factor(data_temp$sex)], 
     cex=2, ylim=c(0,1))
abline(m1, lty=2, col="indianred1")
abline(m2, lty=2, col="darkblue")
shade(mu_ci1, lim = prd1$x, col=col.alpha("indianred1"))
shade(mu_ci2, lim = prd2$x) #, col=col.alpha("darkblue"))
pval1 <- paste0("p = ", round(summary(m1)$coefficients[8],2))
pval2 <- paste0("p = ", round(summary(m2)$coefficients[8],5), "*")
text(0.72, 0.05, pval1, cex=1, col="darkred")
text(0.67, 0.6, pval2) #, cex=c4, col="darkblue")
legend(0.64, 1, legend=c("male", "female"), pch=19, col=c("blue", "red"), y.intersp=1.5, pt.cex=2, cex=.8)


# plot C (mass vs. flew_prob grouped by sex)
data_temp<-aggregate(flew_prob~sex*mass_block, data=d, FUN=mean)
data_temp$n<-aggregate(flew_prob~sex*mass_block, data=d, FUN=length)$flew_prob

d1 = data_temp[data_temp$sex=="F",]
d2 = data_temp[data_temp$sex=="M",]

# females
x=d1[,"mass_block"]
y=d1[,"flew_prob"] 
m1 <- lm(y ~ x, data=d1)
shading=get_CI(x,y,m1)
mu_ci1=shading[[1]]
prd1=shading[[2]]

# males
x=d2[,"mass_block"]
y=d2[,"flew_prob"] 
m2 <- lm(y ~ x, data=d2)
shading=get_CI(x,y,m2)
mu_ci2=shading[[1]]
prd2=shading[[2]]

plot(data_temp$flew_prob~data_temp$mass_block, 
     ylab="Flight probability", 
     xlab="Mass (g)", 
     pch=19, 
     col=c("red","blue")[as.factor(data_temp$sex)], 
     cex=2, 
     ylim=c(0,1), xlim=c(0.02, 0.18))
legend(0.14, 1, legend=c("male", "female"), pch=19, col=c("blue", "red"), y.intersp=1.5, pt.cex=2, cex=.8)
abline(m1, lty=2) #, col="indianred1")
abline(m2, lty=2, col="darkblue")
shade(mu_ci1, lim = prd1$x) #, col=col.alpha("indianred1"))
shade(mu_ci2, lim = prd2$x, col=col.alpha("darkblue"))
pval1 <- paste0("p = ", round(summary(m1)$coefficients[8],3), "*")
pval2 <- paste0("p = ", round(summary(m2)$coefficients[8],2))
text(0.11, 0.5, pval1) #, cex=c4, col="indianred1")
text(0.1, 0.9, pval2, cex=1, col="darkblue")
mtext("C", side=3, adj=0, line=0.5, cex=c3, font=1)
```




