---
title: "Principal component analysis (PCA)"
author: "Anastasia Bernat"
date: "8/12/2020 - 12/16/2020"
output: html_document
---

```{r setup, include=FALSE}
rm(list=ls())
dir = "~/Desktop/git_repositories/SBB-dispersal/avbernat_working_on/Dispersal/Winter_2020/stats/"
setwd(dir) 

library(FactoMineR)
library(factoextra)
library(corrplot)
library(readr)
library(dplyr)

knitr::opts_chunk$set(echo = TRUE)
```

## Winter 2020 Flight Trials {.tabset}

Flight Trials Winter 2020 were conducted from 2/17/2020 - 3/10/2020. Soapberry bugs were flight tested twice for multiple hours in the flight mill and observed from 8 AM to (5-8 PM) each day. 

### Size-adjusted CPCA (notes)

PCAs are good to...

- **Isolate changes in allometry**: "Morphological relationships change with overall body size and body size often varies among populations. Therefore, quantitative analyses of individual traits from organisms in different populations or environments (e.g. in studies of phenotypic plasticity) often adjust for differences in body size to isolate changes in allometry." (McCoy et al, 2006)

- **Help define scaling of traits with body size with PC1**: "For most studies of morphological plasticity and size correction, it is only the first CPC (CPC1) that is of interest because it describes the general scaling of traits with body size." (McCoy et al, 2006)

- **Define size patterns and size corrections**: "If the first principal components are not shared (Fig. 1c), then the patterns of morphological variation are so fundamentally different that there can be no globally applied size correction because 'size' does not have a common meaning across groups. On the other hand, if all principal components are common to all groups (Fig. 1a), then the groups show identical patterns of within-group covariation, indicating the same allometries of traits with size in all groups." (McCoy et al, 2006) AB: Looking at the figures....I would say differences in size but not shape.

- **Define common body size dimensions to evaluate group differences**: "As long as the groups to be compared share CPC1 (Fig. 1a, b) and all the traits have strong loadings on CPC1, then it can be interpreted as a common body size dimension: i.e., a 'size-axis' that can be used to evaluate between-group differences in morphology that go beyond simple (isometric) changes in size" (McCoy et al, 2006)

- **Correct for the effects of a growth axis**: "...where the main goal is to correct for the effects of a growth axis that explains a large fraction of the total variation (e.g., >90%)" (McCoy et al, 2006)

### PCA Functions 

```{r}
Round <- function(number){
  # for plotting
  x <- round(number, 1)
  if(x%%1 == 0){
    return(paste(as.character(x), ".0", sep = ""))
  }
  else{
    return(x)
  }
}
```

```{r}
PCA_graphs <- function(dataset, PCA_title){
    # cos2 and the alpha.var: alpha.var colours variables by cos2 
    # (importance of most important PC to variable), 
    # see https://personal.utdallas.edu/~herve/abdi-awPCA2010.pdf
  
    GFpca <- PCA(dataset, scale.unit = TRUE, graph = TRUE, ncp = 10)
    
    eig.val <- get_eigenvalue(GFpca)
    var.val <- GFpca$var
    print(eig.val) #will only show in in console
    print(var.val)
    
    scree <- fviz_eig(GFpca, addlabels = TRUE, ylim = c(0, 100))
    print(scree)
    
    labX <- paste("PC1 (", Round(eig.val[1, 2]), "%)", sep = "")
    labY <- paste("PC1 (", Round(eig.val[2, 2]), "%)", sep = "")
    leplot <- fviz_pca_biplot(GFpca, geom.id = c("point"), 
                              geom.var = c("arrow", "text"), 
                              alpha.var = "cos2",
                              label = "var", repel = T, 
                              col.ind = "gray", col.var = "black")
    print(leplot)
    
    ggpubr::ggpar(leplot, title = PCA_title, xlab = labX, ylab = labY, 
                  ggtheme = theme_classic(), font.main = c(20, "bold"), 
                  font.x = 14, font.y = 14, font.tickslab = 12
                  )
    
    D = cor(dataset)
    test <- cor.mtest(dataset)$p
    par(mfrow=c(1,2))
    corrplot.mixed(D,lower.col = "black", number.cex = .7, p.mat=test, sig.level=0.05)
    corrplot.mixed(D,lower.col = "black", number.cex = .7)
    
    return(GFpca)
}
```

### Reading and Standardizaing Data

```{r message=FALSE}
source("src/clean_flight_data.R") # Loads and cleans data
source("src/center_flight_data.R") # Centers data

data <- read_flight_data("data/all_flight_data-Winter2020.csv")
data_all <- data[[1]]
data_tested <- data[[2]]

# Filter out duplicates (PCA applies to data tables where rows are considered as individuals)
data_tested <- data_tested[data_tested$trial_type == "T1",]

# Standardize variables
data_tested$thorax_s <- (data_tested$thorax-mean(data_tested$thorax, na.rm=TRUE)) / sd(data_tested$thorax, na.rm=TRUE)
data_tested$body_s <- (data_tested$body-mean(data_tested$body, na.rm=TRUE)) / sd(data_tested$body, na.rm=TRUE)
data_tested$wing_s <- (data_tested$wing-mean(data_tested$wing, na.rm=TRUE)) / sd(data_tested$wing, na.rm=TRUE)
data_tested$beak_s <- (data_tested$beak-mean(data_tested$beak, na.rm=TRUE)) / sd(data_tested$beak, na.rm=TRUE)
data_tested$mass_s <- (data_tested$mass-mean(data_tested$mass, na.rm=TRUE)) / sd(data_tested$mass, na.rm=TRUE)
```

### Morphology

Thorax, body, wing, and beak lengths only considered. 

```{r}
d <- data_tested %>%
  select(thorax_s, body_s, wing_s, beak_s) %>%
  filter(!is.na(body_s))
colnames(d) <- c("thorax", "body", "wing", "beak")

MorphPCA = PCA_graphs(d, "(a) ")
```

```{r}
vars <- MorphPCA$ind
par(mfrow=c(2,3))

#Correlations
m <- lm(d$thorax ~ vars$coord[,1])
plot(vars$coord[,1], d$thorax, xlab="PC1", ylab="thorax (mm)")
cor_val <- paste("R = ", round(cor(vars$coord[,1], d$thorax), 2))
text(-2,2, cor_val)
abline(m, lty=2)

m <- lm(d$body ~ vars$coord[,1])
plot(vars$coord[,1], d$body, xlab="PC1", ylab="body (mm)")
cor_val <- paste("R = ", round(cor(vars$coord[,1], d$body), 2))
text(-2,2, cor_val)
abline(m, lty=2)

m <- lm(d$wing ~ vars$coord[,1])
plot(vars$coord[,1], d$wing, xlab="PC1", ylab="wing (mm)")
cor_val <- paste("R = ", round(cor(vars$coord[,1], d$wing), 2))
text(-2,2, cor_val)
abline(m, lty=2)

m <- lm(d$beak ~ vars$coord[,1])
plot(vars$coord[,1], d$beak, xlab="PC1", ylab="beak (mm)")
cor_val <- paste("R = ", round(cor(vars$coord[,1], d$beak), 2))
text(-2,2, cor_val)
abline(m, lty=2)

m <- lm(d$beak ~ vars$coord[,2])
plot(vars$coord[,2], d$beak, xlab="PC2", ylab="beak (mm)")
cor_val <- paste("R = ", round(cor(vars$coord[,2], d$beak), 2))
text(-0.5,3, cor_val)
abline(m, lty=2)
```


```{r}
#Plot by host
host <- data_tested %>%
  select(host_plant)

d$host_plant <- host$host_plant
fviz_pca_ind(MorphPCA,
             geom.ind = "point",
             col.ind = d$host_plant,
             legend.title = "Host Plant",
             addEllipses=TRUE)

# Plot by sex
sex <- data_tested %>%
  select(sex) 

d$sex <- sex$sex
fviz_pca_ind(MorphPCA,
             geom.ind = "point",
             col.ind = d$sex,
             legend.title = "Sex",
             addEllipses=TRUE)

# Plot by distance from the sympatric zone
sym_dist <- data_tested %>%
  select(sym_dist)

d$sym_dist <- sym_dist$sym_dist
fviz_pca_ind(MorphPCA,
             geom.ind = "point",
             col.ind = d$sym_dist,
             legend.title = "Sympatric Zone Distance")

# Plot by population
pop <- data_tested %>%
  select(population)

d$population <- pop$pop
fviz_pca_ind(MorphPCA,
             geom.ind = "point",
             col.ind = d$population,
             legend.title = "Population",
             addEllipses=TRUE)
```

### Morphology w/ Mass

```{r}
d <- data_tested %>%
  select(thorax_s, body_s, wing_s, beak_s, mass_s) %>%
  filter(!is.na(mass_s))
colnames(d) <- c("thorax", "body", "wing", "beak", "mass")

MorphPCA = PCA_graphs(d, "(b) ")
```

### Morphology w/ Wing2Body

```{r}
# No mass
d <- data_tested %>%
  select(thorax_s, wing2body_s, beak_s) %>%
  filter(!is.na(beak_s))
colnames(d) <- c("thorax", "wing2body", "beak")

MorphPCA = PCA_graphs(d, "(c) ")

# Mass
d <- data_tested %>%
  select(thorax_s, wing2body_s, beak_s, mass_s) %>%
  filter(!is.na(mass_s))
colnames(d) <- c("thorax", "wing2body", "beak", "mass")

MorphPCA = PCA_graphs(d, "(d) ")
```
