---
title: "Soapberry Bug Flight Dispersal Spatial Data Visualization - Winter2020"
author: "Anastasia Bernat"
fontsize: 11pt
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document:
    df_print: kable
    fig_caption: no
    fig_height: 3.5
    fig_width: 3.5
    highlight: monochrome
geometry: margin=0.75in
fig_crop: no
---

```{r setup, include=FALSE}
rm(list=ls())
dir = "~/Desktop/git_repositories/SBB-dispersal/avbernat_working_on/Dispersal/Winter_2020/stats/"
setwd(dir) 

# Load some commonly used packages.
library(MASS)
library(gridExtra)
library(tidyverse)
library(mosaic)
library(broom)
library(dplyr)
library(knitr)
library(readr)
library(chron)
library(stringr)

# Set numerical output display parameters
options(width=70, digits=4, scipen=8)
# Set R output size a bit smaller than default
knitr::opts_chunk$set(size='small', prompt=FALSE, comment="")
# set plot theme to black and white
ggplot2::theme_set(ggplot2::theme_bw())
ggplot2::theme_update(panel.grid.major = ggplot2::element_line(colour = "grey75"))

# Spatial Libraries
library(sf)
library(sp)
library(lubridate)
library(tmap)
library(geojsonR)
library(geojsonio)
library(lwgeom) # lwgeom is known as the light-weight geometry library used by 'PostGIS' 
library(cleangeo)
library(maps)
library(shinyjs)
library(tmaptools)
```

## Read the data.

```{r}
source("src/clean_flight_data.R")
data <- read_flight_data("data/all_flight_data-Winter2020.csv")
all_data <- data[[1]]
data_all <- data[[2]] # bugs tested
```

```{r}
# data_all$flew_b<-0
# data_all$flew_b[data_all$flew=="Y"]<-1
# 
# data_all$eggs_b<-0
# data_all$eggs_b[data_all$EWM=="Y"]<-1
# 
# data_all$sym_dist<-abs(data_all$latitude-25.49197)
# 
# data_all
```

## Filter for only the columns you need.

```{r}
# filtered_data <- full_data[,c("ID", "filename", "chamber", "set_number", "average_speed",
#                         "total_flight_time", "distance", "shortest_flying_bout", 
#                         "longest_flying_bout", "portion_flying", "max_speed",
#                         "sex", "population", "site", "host_plant", "flew", "flew_b", flight_type", 
#                         "mass", "eggs", "latitude", "longitude")]
filtered_data <- select(data_all, ID, filename, chamber, set_number, average_speed, # recently added flew_b
                        total_flight_time, distance, shortest_flying_bout, 
                        longest_flying_bout, max_speed,
                        sex, population, site, host_plant, host_temp, flew, flew_b, flight_type, mass, eggs_b, 
                        total_eggs, beak, thorax, wing, body, sym_dist, tested, w_morph,
                        lat = latitude, 
                        lon = longitude)

# Get rid of NA's in long and lat columns
filtered_data <- filter(filtered_data, !is.na(lat), !is.na(lon))
glimpse(filtered_data)
```

## Convert to a spatial point object and plot.

```{r}
dispersal_data <- st_as_sf(filtered_data,
                           coords = c("lon", "lat"),
                           crs = 4326,
                           agr = "constant")

dispersal_data <- st_transform(dispersal_data, 2236)
plot(dispersal_data)
```

This shows us the data, but visually it's hard to read. Let's try mapping multipolygon and polygon shapes.

## Create mappable multipolygon shapes for designated areas. Sites to use:

(i) Go to OpenStreetMaps: https://nominatim.openstreetmap.org

(ii) Type in the area you want your multipolygon for.

(iii) Find the OSM ID and copy it.

(iv) Go to Polygon Creation: http://polygons.openstreetmap.fr/index.py

(v) Past the OSM ID into the 'Id of relation' box.

(vi) Click on 'image' to confirm that is the shape you want on OpenStreetMaps or Google Maps.

(vii) Click on GeoJSON and copy the url.

(viii) Find a spatial reference number or (espg) for Florida. I went with 2236, East Florida as pictured here on OpenStreetMaps: https://maps.omniscale.com/en/openstreetmap/epsg-2236. 

```{r}
florida <- "http://polygons.openstreetmap.fr/get_geojson.py?id=162050&params=0"

# Just For Florida

# URL Florida
florida_read <- st_read(florida)
FL_url <- mutate(florida_read, population = "Florida")
FL_url_transformed <- st_transform(FL_url, 2236)

# State Maps Data
states <- st_as_sf(map("state", plot = FALSE, fill = TRUE))
florida_state <- states %>%
  filter(ID == "florida")
FL <- mutate(florida_state, State = "Florida")
FL_transformed <- st_transform(FL, 2236)
```


(ix) Convert the GeoJSON dat into a nested list using geojson_read() or lapply() and st_read as the second argument. lapply() and st_read() are much more efficient. The map() function is also similar to lapply(), but for some reason it does not work here. Similar to pandas in Python, the map() function transforms its input by applying a function to each element and returning a vector (or whichever class you want) with the same length as the input.

(x) If there is a gap in one or more of the polygons created, use st_make_valid() to bypass the error and make the polygon valid.

(xi) Transform the GEOMETRY COLLECTIONS to MULTIPOLYGONS using st_cast().

```{r}
# For Other Populations and Their Respective Counties

lake_wales <- "http://polygons.openstreetmap.fr/get_geojson.py?id=117732&params=0"
polk <- "http://polygons.openstreetmap.fr/get_geojson.py?id=1214525&params=0"
  
lake_placid <- "http://polygons.openstreetmap.fr/get_geojson.py?id=117660&params=0"
highlands <- "http://polygons.openstreetmap.fr/get_geojson.py?id=1210702&params=0"
  
homestead <-"http://polygons.openstreetmap.fr/get_geojson.py?id=1216731&params=0" 
miami_dade <- "http://polygons.openstreetmap.fr/get_geojson.py?id=1210692&params=0"

gainesville <- "http://polygons.openstreetmap.fr/get_geojson.py?id=118870&params=0"
alachua <- "http://polygons.openstreetmap.fr/get_geojson.py?id=1210739&params=0"

leesburg <- "http://polygons.openstreetmap.fr/get_geojson.py?id=117476&params=0"
lake <- "http://polygons.openstreetmap.fr/get_geojson.py?id=389026&params=0"
  
# Have the same counties (Monroe), so kept on an island level. 
north_key_largo <- "http://polygons.openstreetmap.fr/get_geojson.py?id=119071&params=0" 
key_largo <- "http://polygons.openstreetmap.fr/get_geojson.py?id=119070&params=0"
plantation_key <-  "http://polygons.openstreetmap.fr/get_geojson.py?id=117963&params=0"

populations <- c("Lake Wales", "Lake Placid", "Homestead", "North Key Largo",
           "Gainesville", "Leesburg", "Key Largo", "Plantation Key")

counties <- c("Polk", "Highlands", "Miami-Dade", "Alachua", "Lake", "North Key Largo", "Key Largo", "Plantation Key")

pop_urls <- c(lake_wales, lake_placid, homestead, north_key_largo, gainesville,
          leesburg, key_largo, plantation_key)
county_urls <- c(polk, highlands, miami_dade, alachua, lake, north_key_largo, key_largo, plantation_key)
```

```{r}
# For Populations
polylist <- lapply(pop_urls, st_read)
sf_df <- do.call(rbind, polylist)
pops <- cbind(sf_df, populations)
df_transformed <- st_transform(pops, 2236)

# For Counties
countylist <- lapply(county_urls, st_read)
sf_df_county <- do.call(rbind, countylist)
counties_df <- cbind(sf_df_county, counties)
counties_transformed <- st_transform(counties_df, 2236)

# Geometry reports

poly_valid_report <- st_is_valid(df_transformed)
poly_valid_report # see that LW, LP, GV, LB, and PK are invalid

county_valid_report <- st_is_valid(counties_transformed)
county_valid_report

# Make any invalid polygons valid
df_final_counties <- st_make_valid(counties_transformed) %>%
  st_cast("MULTIPOLYGON") 

# Check the final report
county_valid_report2 <- st_is_valid(df_final_counties)
county_valid_report2
```

```{r}
# this stopped working so don't run:

df_final_pop <- st_make_valid(df_transformed) %>% # stopped working but not necessary for map making
  st_cast("MULTIPOLYGON") 

poly_valid_report2 <- st_is_valid(df_final_pop)
poly_valid_report2

tm_shape(df_final_pop) +
  tm_polygons("populations") +
  tm_basemap("Hydda.Base")
```

```{r}
tmap_mode("view")
tmap_mode("plot")

# test
tm_shape(df_final_counties) +
  tm_polygons("counties") +
  tm_basemap("Hydda.Base")
```


(xii) Combine the comprehensive sf dataframe with the sf dataframe mapping the populations to the polygon shapes.

(xiii) Write your data to a csv or gpkg file using write_sf(). (Still unsure what application opens gpkg files)

```{r}
# Add Counties to the Populations
combined <- st_join(df_final_counties, dispersal_data) 

#write_sf(combined, "flight_spatial_data.csv", layer_options="GEOMETRY=AS_WKT")
#write_sf(combined_df, "combined2.gpkg")
```

(xiv) Filter for the datasets you want to work with. This is a good point to customize you datasets as well. Then, calculate proportion flew and site names.

```{r}
# Only those that flew (this also removes distance NA and speed NA values):
combined_filtered <- combined %>%
  filter(average_speed !=0)
dispersal_filtered <- dispersal_data %>%
  filter(average_speed !=0)

# Remove bugs not tested and L or LS

experiment_bugs <- combined %>%
  filter(tested == "yes" | tested == "no" & w_morph == "S")

# Remove NA mass vales
mass_data <- combined_filtered %>%
  filter(!is.na(mass))
```

```{r}
# Plots for ASN 2021:

## mainland
big_bb <- st_bbox(FL_transformed)
big_bb[2] = big_bb[2]-100000

dispersal_main = dispersal_filtered %>%
  filter(host_temp=="K. elegans")

colors = c("lightblue", "darkolivegreen3") # this one takes a few minutes (~4 min) to load
tm_shape(FL_transformed, bbox=big_bb) +
  tm_graticules(n.x = 3, n.y = 7, labels.format = list(digits=1), col= "grey") +
  tm_fill(col="white", alpha= 0.3) +
  tm_polygons("State") +
  tm_scale_bar(position = c("left", "bottom")) +
tm_shape(combined_filtered) +
  tm_polygons("host_temp", tm_legend="Host Plant", palette=colors) +
  tm_basemap("Hydda.Base") +
tm_shape(dispersal_main) +
  tm_dots(alpha = 0.8, size=0.2, shape=2, title = 'Collection Sites', labels = "geometry", legend.show=TRUE) 

## keys
keys_bb <- st_bbox(combined_filtered %>% filter(counties %in% c("North Key Largo", "Key Largo", "Plantation Key")) )

dispersal_keys = dispersal_filtered %>%
  filter(host_temp=="C. corindum")

tm_shape(combined_filtered, bbox=keys_bb) + 
    tm_graticules(n.x = 3, n.y = 7, labels.format = list(digits=1), col= "grey") +
    tm_polygons("host_temp", labels = NULL, legend.show = FALSE, palette=colors) +
    tm_scale_bar(position = c("right", "bottom")) +
tm_shape(dispersal_keys) +
  tm_dots(alpha = 0.9, size= 0.2, shape=2, title = 'Collection Sites', labels = "geometry", legend.show=TRUE) 
#+ tm_text("zmap_site", size = 0.4, xmod = -1.7, ymod = 0.15)

```

```{r}
# Prop flew --> rewrite to make it cleaner
cal_prop_flew <- function(data, site_name) {
  d <- filter(data, population == site_name)
  data$flew_prop[data$population == site_name] <- sum(d$flew_b) / nrow(d)
}

# For tested bugs:

NKL <- filter(combined_filtered, population == "North Key Largo")
KL <- filter(combined_filtered, population == "Key Largo")
PK <- filter(combined_filtered, population == "Plantation Key")
HS <- filter(combined_filtered, population == "Homestead")
LP <- filter(combined_filtered, population == "Lake Placid")
LW <- filter(combined_filtered, population == "Lake Wales")
LB <- filter(combined_filtered, population == "Leesburg")
GV <- filter(combined_filtered, population == "Gainesville")

combined_filtered$flew_prop[combined_filtered$population == "North Key Largo"] <- sum(NKL$flew_b) / nrow(NKL)
combined_filtered$flew_prop[combined_filtered$population == "Key Largo"] <- sum(KL$flew_b) / nrow(KL)
combined_filtered$flew_prop[combined_filtered$population == "Plantation Key"] <- sum(PK$flew_b) / nrow(PK)
combined_filtered$flew_prop[combined_filtered$population == "Homestead"] <- sum(HS$flew_b) / nrow(HS)
combined_filtered$flew_prop[combined_filtered$population == "Lake Placid"] <- sum(LP$flew_b) / nrow(LP)
combined_filtered$flew_prop[combined_filtered$population == "Lake Wales"] <- sum(LW$flew_b) / nrow(LW)
combined_filtered$flew_prop[combined_filtered$population == "Leesburg"] <- sum(LB$flew_b) / nrow(LB)
combined_filtered$flew_prop[combined_filtered$population == "Gainesville"] <- sum(GV$flew_b) / nrow(GV)

# For tested and non tested bugs:

NKL <- filter(experiment_bugs, population == "North Key Largo")
KL <- filter(experiment_bugs, population == "Key Largo")
PK <- filter(experiment_bugs, population == "Plantation Key")
HS <- filter(experiment_bugs, population == "Homestead")
LP <- filter(experiment_bugs, population == "Lake Placid")
LW <- filter(experiment_bugs, population == "Lake Wales")
LB <- filter(experiment_bugs, population == "Leesburg")
GV <- filter(experiment_bugs, population == "Gainesville")

combined_filtered$flew_prop_all[combined_filtered$population == "North Key Largo"] <- sum(NKL$flew_b) / nrow(NKL)
combined_filtered$flew_prop_all[combined_filtered$population == "Key Largo"] <- sum(KL$flew_b) / nrow(KL)
combined_filtered$flew_prop_all[combined_filtered$population == "Plantation Key"] <- sum(PK$flew_b) / nrow(PK)
combined_filtered$flew_prop_all[combined_filtered$population == "Homestead"] <- sum(HS$flew_b) / nrow(HS)
combined_filtered$flew_prop_all[combined_filtered$population == "Lake Placid"] <- sum(LP$flew_b) / nrow(LP)
combined_filtered$flew_prop_all[combined_filtered$population == "Lake Wales"] <- sum(LW$flew_b) / nrow(LW)
combined_filtered$flew_prop_all[combined_filtered$population == "Leesburg"] <- sum(LB$flew_b) / nrow(LB)
combined_filtered$flew_prop_all[combined_filtered$population == "Gainesville"] <- sum(GV$flew_b) / nrow(GV)
```

```{r}
# Creating site names for the map: 

# Large Map (Non-Keys):
pop_abbrev <- gsub("[^A-Z]","",dispersal_filtered$population) # Extract only capital letters 

dispersal_filtered$lmap_site <- paste0(pop_abbrev, ", ", dispersal_filtered$site)

for (row in 1:nrow(dispersal_filtered)) {
  if (dispersal_filtered$population[row] == "North Key Largo" |
      dispersal_filtered$population[row] == "Key Largo" |
      dispersal_filtered$population[row] == "Plantation Key") {
    dispersal_filtered$lmap_site[row] <- " "
      }
}

dispersal_filtered$lmap_site <- gsub("\\s", " ", format(dispersal_filtered$lmap_site,
                                                        width=max(nchar(dispersal_filtered$lmap_site)))) 

# Small Zoon Map (Keys):
pop_abbrev <- gsub("[^A-Z]","",dispersal_filtered$population) 
dispersal_filtered$zmap_site <- paste0(pop_abbrev, ", ", dispersal_filtered$site)
dispersal_filtered$zmap_site <- gsub("\\s", " ", format(dispersal_filtered$zmap_site,
                                                        width=max(nchar(dispersal_filtered$zmap_site)))) 
```

(xiv) Distance variable only: Create distance buffers using st_buffer(). Buffers are polygons representing the area within a given distance of a geometric feature. st_buffer() requires at least two arguments: an input geometry and a distance, provided in the units of the CRS (in this case meters):

## Build the distance buffer map

```{r}
################################################################################
# FUNCTION: get_distance_buffer | extracts the distance buffer of a site
#
# INPUTS:   data as a data.frame and site_name as a character
#
# OUTPUTS:  the preserved spatial data.frame and its distance buffer to be used
#           for mapping
################################################################################

get_distance_buffer <- function(data, site_name) { 
  d <- filter(data, population == site_name | site == site_name)
  d_buff <- st_buffer(d, dist=max(d$distance))
  d_buff <- d_buff %>%
    select(distance, geometry) # can probably include population and site for easier accesss
}

# Population Buffers
pops <- unique(dispersal_filtered$population)
pops

NKL_buffer <- get_distance_buffer(dispersal_filtered, "North Key Largo")
KL_buffer <- get_distance_buffer(dispersal_filtered, "Key Largo")
PK_buffer <- get_distance_buffer(dispersal_filtered, "Plantation Key")
HS_buffer <- get_distance_buffer(dispersal_filtered, "Homestead")
LB_buffer <- get_distance_buffer(dispersal_filtered, "Leesburg")
LW_buffer <- get_distance_buffer(dispersal_filtered, "Lake Wales")
LP_buffer <- get_distance_buffer(dispersal_filtered, "Lake Placid")
GV_buffer <- get_distance_buffer(dispersal_filtered, "Gainesville")

pop_buffer_data <- rbind(NKL_buffer, KL_buffer, PK_buffer, HS_buffer, LB_buffer, LW_buffer, LP_buffer, GV_buffer)

# Site Buffers
sites <- unique(dispersal_filtered$site)
sites

KLMRL_buffer <- get_distance_buffer(dispersal_filtered, "KLMRL")
JP_buffer <- get_distance_buffer(dispersal_filtered, "JP")
ND_buffer <- get_distance_buffer(dispersal_filtered, "N. Dagny")
CH_buffer <- get_distance_buffer(dispersal_filtered, "Charlemagne")
DDF_buffer <- get_distance_buffer(dispersal_filtered, "DD front")
MM_buffer <- get_distance_buffer(dispersal_filtered, "MM165")
PA_buffer <- get_distance_buffer(dispersal_filtered, "Polk Ave")
St_buffer <- get_distance_buffer(dispersal_filtered, "23rd & 8th")
F_buffer <- get_distance_buffer(dispersal_filtered, "Founder's")
CF_buffer <- get_distance_buffer(dispersal_filtered, "Carysfort")
DT_buffer <- get_distance_buffer(dispersal_filtered, "Dagny Trellis")
MSt_buffer <- get_distance_buffer(dispersal_filtered, "Mount & 8th")
DD_buffer <- get_distance_buffer(dispersal_filtered, "Dynamite Docks")
StM_buffer <- get_distance_buffer(dispersal_filtered, "110N Main")
AL_buffer <- get_distance_buffer(dispersal_filtered, "Aregood Ln")
SWSt_buffer <- get_distance_buffer(dispersal_filtered, "SW 296th St")

site_buffer_data <- rbind(KLMRL_buffer, JP_buffer, ND_buffer, CH_buffer,
                          DDF_buffer, MM_buffer, PA_buffer, St_buffer, 
                          F_buffer, CF_buffer, DT_buffer, MSt_buffer, 
                          DD_buffer, StM_buffer, AL_buffer, SWSt_buffer)
```

```{r}
# Island vs. Mainland & Island vs. Island Crossings

zmap.filename = "keys_dist_map1.png"

keys_bb <- st_bbox(combined_filtered %>% filter(counties %in% c("North Key Largo", "Key Largo", "Plantation Key")) )

m21 <- tm_shape(combined_filtered, bbox=keys_bb) + # only place where it would matter is in the keys.
    tm_graticules(n.x = 3, n.y = 7, labels.format = list(digits=1), col= "azure2") +
    tm_polygons("host_temp", labels = NULL, legend.show = FALSE, palette=colors) +
    tm_scale_bar(position = c("right", "bottom")) +
tm_shape(site_buffer_data) + # NKL_buffer
  tm_fill("distance", alpha=0.02, legend.show = FALSE, col = "lightblue") + # can change the color, col = "blue"
tm_shape(dispersal_filtered) +
  tm_dots(alpha = 0.9, title = 'Collection Sites', labels = "geometry") +
  tm_text("zmap_site", size = 0.4, xmod = -1.7, ymod = 0.15)

m21

tmap_save(m21, zmap.filename )
```

```{r}
h.keys_bb <- st_bbox(combined_filtered %>% filter(counties %in% c("Miami-Dade"," North Key Largo", "Key Largo", "Plantation Key")) )
h.keys_bb[4] <- h.keys_bb[4] - 169000
h.keys_bb[1] <- h.keys_bb[1] + 40000
h.keys_bb[3] <- h.keys_bb[3] - 60000
m22 <- tm_shape(combined_filtered, bbox=h.keys_bb) + # only place where it would matter is in the keys.
    tm_graticules(n.x = 3, n.y = 7, labels.format = list(digits=1), col= "azure2") +
    tm_fill("distance", labels = NULL, legend.show = FALSE) +
    tm_scale_bar(position = c("right", "bottom")) +
tm_shape(site_buffer_data) + # NKL_buffer
  tm_fill("distance", alpha=0.02, legend.show = FALSE, col= "lightskyblue1") +
tm_shape(dispersal_filtered) +
  tm_dots(alpha = 0.9, title = 'Collection Sites', labels = "geometry") +
  tm_text("zmap_site", size = 0.4, xmod = -1.7, ymod = 0.15)

m22
```

```{r}
mainland_bb <- st_bbox(combined_filtered %>% filter(counties %in% c("Polk", "Highlands", "Miami-Dade", "Lake", "Alachua")) ) # same thing with or without this

tm_shape(FL_transformed, bbox=mainland_bb) +
  tm_fill(col="darkolivegreen4", alpha= 0.3) +
  tm_polygons("State") +
tm_shape(combined_filtered, bbox=mainland_bb) +
    tm_graticules(n.x = 3, n.y = 7, labels.format = list(digits=1)) + #, col= "azure2") +
    tm_fill("distance", labels = NULL, legend.show = FALSE) +
    tm_scale_bar(position = c("left", "bottom")) +
tm_shape(site_buffer_data) + 
  tm_fill("distance", alpha=0.02, legend.show = FALSE, col="blue") +
tm_shape(dispersal_filtered) + # can also ommit the keys before map this 
  tm_dots(alpha = 0.7, title = 'Collection Sites', labels = "geometry") +
  tm_text("lmap_site", size = 0.4, xmod = -0.15, ymod = -0.35)
```

## Build the rest of the maps!

```{r}
# ORIGINAL VERSION (WIDE)
map_data = combined_filtered
var = "mass"
main.title = "Mass Across Latitude"
label.title = "Mass (g)"
map.filename = "winter_test6.png"


m <- tm_shape(FL_transformed) +
  tm_fill(col="darkolivegreen4", alpha= 0.3) +
  tm_polygons("State") +
#    tm_text("State", col= "black", xmod = 1, ymod = 3, fontface="bold") +
tm_shape(map_data) +
  tm_fill(var, title = label.title, legend.hist = TRUE) + #, palette="Oranges"
  tm_basemap("Hydda.Base") +
#  tm_style("natural") +
  tm_layout(main.title = main.title,
            legend.hist.size = 2.5,
            legend.outside = TRUE,
            compass.type = "4star",
            asp = 1, # legend.inside vs. outside
            inner.margins=c(0.06, 0.01, 0, 0.1)) + # c(bottom, left, top, and right
  #bg.color= "aliceblue") + # or aliceblue as bg.color
    tm_compass(show.labels = 1, 
             cardinal.directions = c("N", "E", "S", "W"),
             position = c(0.8, 0.77), 
             size = 4, 
             color.light= "grey90",
             color.dark= "grey30") +
  tm_scale_bar(position = c("left", "bottom")) +
  tm_graticules() +
tm_shape(dispersal_filtered) + # can also ommit the keys before map this 
  tm_dots(alpha = 0.7, title = 'Collection Sites', labels = "geometry") +
  tm_text("lmap_site", size = 0.4, xmod = 1.2, ymod = -0.3)

tmap_save(m, map.filename) 
```

```{r}
# CLOSER VERSION
map_data = combined_filtered #mass_data #
var = "flew_prop"
main.title = "Prop Flew Across Latitude"
label.title = "Prop Flew"
map.filename = "propflew_map2.png"


m1 <- tm_shape(FL_transformed) +
  tm_fill(col="darkolivegreen4", alpha= 0.3) +
  tm_polygons("State") +
#    tm_text("State", col= "black", xmod = 1, ymod = 3, fontface="bold")
  tm_view(set.bounds = c(1, 29, 1, 85)) +
tm_shape(map_data) +
  tm_graticules(n.x = 4, n.y = 7, labels.format = list(digits=1)) +
  tm_fill(var, title = label.title) + #, palette="Oranges" | legend.hist = TRUE
  tm_layout(#main.title = main.title,
            #main.title.position = c("center", "top"),
            legend.hist.size = 2,
            legend.position = c("left", "center"),
            legend.height = 0.8,
            compass.type = "4star",
            asp = 1, # legend.inside vs. outside
            inner.margins=c(0.06, 0.01, 0, 0.13)) +
      tm_compass(show.labels = 1, 
             cardinal.directions = c("N", "E", "S", "W"),
             position = c(0.75, 0.76), 
             size = 4, 
             color.light= "grey90",
             color.dark= "grey30") +
  tm_scale_bar(position = c("left", "bottom")) +
tm_shape(dispersal_filtered) +
  tm_dots(alpha = 0.7, title = 'Collection Sites', labels = "geometry") +
  tm_text("lmap_site", size = 0.4, xmod = 1.27, ymod = -0.1)

tmap_save(m1, map.filename, width=1260, height=1080, asp=0) # 1920, 1080
```

```{r}
# KEYS MAP
zmap.filename = "keys_propflewall_map.png"
keys_bb <- st_bbox(combined_filtered %>% filter(counties %in% c("North Key Largo", "Key Largo", "Plantation Key")) )

m20 <- tm_shape(combined_filtered, bbox=keys_bb) + # only place where it would matter is in the keys.
    tm_graticules(n.x = 3, n.y = 7, labels.format = list(digits=1), col= "azure2") +
    tm_fill(var, labels = NULL, legend.show = FALSE) +
    tm_scale_bar(position = c("right", "bottom")) +
tm_shape(dispersal_filtered) +
  tm_dots(alpha = 0.9, title = 'Collection Sites', labels = "geometry") +
  tm_text("zmap_site", size = 0.4, xmod = -1.7, ymod = 0.15)

m20
tmap_save(m20, zmap.filename )
```


```{r}
# COMBINATION OF MAPS 
m_full <- tmap_arrange(m1, m20, ncol = 2, widths= c(0.65, 0.35))
tmap_save(m_full, map.filename, width=1920, height=1080, asp=0)
```


Extra sources and notes:

?tmap_icons()
Specifies icons from a png images, which can be used as markers in thematic maps. The function marker_icon is the specification of the default marker.


grids: https://geocompr.github.io/post/2019/tmap-grid/

tm_fill
tm_layout
tm_legend
tmap_arrange: https://rdrr.io/cran/tmap/man/tmap_arrange.html
tm_shape and bbox: https://www.jla-data.net/eng/adjusting-bounding-box-of-a-tmap-map/ | https://rdrr.io/cran/tmaptools/man/bb.html


more mapping layouts: https://mran.revolutionanalytics.com/snapshot/2015-07-13/web/packages/tmap/vignettes/tmap-nutshell.html

creating beautiful maps: http://zevross.com/blog/2018/10/02/creating-beautiful-demographic-maps-in-r-with-the-tidycensus-and-tmap-packages/

tmap_save

title? Florida Keys National Marine Sanctuary
