---
title: 'Modeling Flight Distance'
author: "Anastasia Bernat"
date: "5/4/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
rm(list=ls())
setwd("~/Desktop/Rstats-winter2020/")

library(lme4)

library(dplyr)
library(tidyselect)

library(ggplot2)
library(glmnet)
library(ggplotify)
library(gridExtra)
library(ggformula)
library(randomcoloR)

knitr::opts_chunk$set(echo = TRUE)
```

## Winter 2020 Flight Trials: Distance Flight Modeling {.tabset}

Flight Trials Winter 2020 Dataset was conducted from 2/17/2020 - 3/10/2020. Soapberry bugs were flight tested twice for multiple hours in the flight mill and observed from 8 AM to (5-8 PM) each day. Used multivariate (glm) and mixed effect modeling (glmer) to analyze the flight results.

### All Data

**Cleaning the Data**

```{r}
rm(list=ls())
output_col = FALSE # Recommend changing this to TRUE if working in Base R or RStudio, and FALSE if generating an html
source("src/clean_flight_data.R") # Script that loads and cleans up the data
source("src/regression_output.R") # A script that cleans up regression outputs and prints in color or black and white
source("src/center_flight_data.R")
source("get_warnings.R")

data <- read_flight_data("data/all_flight_data-Winter2020.csv")
data_all <- data[[1]]
data_tested <- data[[2]]

### Remove everyone who didn't fly (then remove distances = 0, if that didn't work fully)
data_flew <- data_tested[data_tested$flew_b == 1, ]
data_flew <- data_flew[data_flew$distance > 0, ]
data_flew <- center_data(data_flew)

###Break up by trial type
d_T1 <-data_flew[data_flew$trial_type=="T1",] 
d_T1 <- center_data(d_T1)
d_T2 <-data_flew[data_flew$trial_type=="T2",]
d_T2 <- center_data(d_T2)

###Break up by sex 
d_fem <-data_flew[data_flew$sex=="F",] 
d_fem <- center_data(d_fem)
d_male <-data_flew[data_flew$sex=="M",]
d_male <- center_data(d_male)
```

### All Data Plots

```{r}
h1 <-  as.grob(expression(
  hist(data_flew$distance)))
p1 <- as.grob(expression(
  plot(distance ~ mass_c, data=data_flew)))
p2 <- as.grob(expression(
  plot(distance ~ days_from_start_c, data=data_flew)))
p3 <- as.grob(expression(
  plot(distance ~ min_from_IncStart, data=data_flew)))
p4 <- as.grob(expression(
  plot(distance ~ wing_c, data=data_flew)))
p5 <- as.grob(expression(
  plot(distance ~ beak_c, data=data_flew)))
p6 <- as.grob(expression(
  plot(distance ~ thorax_c, data=data_flew)))
p7 <- as.grob(expression(
  plot(distance ~ body_c, data=data_flew)))

grid.arrange(h1,p1,p2,p3,p4,p5,p6,p7, ncol=3)
```

```{r}
gf_point(distance~sym_dist, col=~host_plant, alpha=~sex_c, data=data_flew) 
summary <- aggregate(distance~sym_dist*sex*host_plant, data=data_flew, FUN=mean)
plot(summary$distance~summary$sym_dist, 
     col=c(1,2)[as.factor(summary$sex)], # Female = Black, Male = Red
     pch=c(19,22)[as.factor(summary$host_plant)], # Filled circle is GRT, Open square is BV
     main="Observed Data: distance ~ sex*host_plant*sym_dist",
     xlab = "Distance from Sympatric Zone (°)",
     ylab= "Distance Flew", 
     #sub=eq_glmer
     ) 
legend("topright",
       legend = c("F and K.elegans", "M and C.corindum", "F and C.corindum","M and K.elegans"),
       #inset=c(-0.27,0.2),
       col= 1:2,
       pch = c(0,16,19),
       title="Groups")
p <- ggplot(data_flew, aes(x=sym_dist, y=distance, color=host_plant)) + 
  geom_violin()
p + stat_summary(fun=mean, geom="point", shape=23, size=2)
```

### Trial 1

[Testing link and covariates](#test)

[Without Mass](#wo.mass)

[With Mass](#w.mass)

[Min From Inc Start ](#timestart)

[Morphology](#morph)

<a id="test"></a>

**Testing link functions**

```{r}
model_test<-glm(distance~chamber, data=d_T1, family=Gamma(link="log")) #equivalent to using log link function in Gamma is to log transform distance, except it won't fuss about non-0 values
summary(model_test)
plot(model_test)
```

**Testing experimental covariates**
```{r}
####### Effect of chamber A-4
d_T1$chamber <- relevel(d_T1$chamber, ref="A-2")
tidy_regression(glm(distance~chamber, data=d_T1, family=Gamma(link="log")), is_color=output_col)

####### No effect of test date
tidy_regression(glm(distance~days_from_start_c, data=d_T1, family=Gamma(link="log")), is_color=output_col)

####### Effect of Minutes from IncStart time
tidy_regression(glm(distance~min_from_IncStart_c, data=d_T1, family=Gamma(link="log")), is_color=output_col)
```

<a id="wo.mass"></a>

**Without Mass**

```{r message=FALSE}
data<-data.frame(R=d_T1$distance, A=d_T1$host_c, B=d_T1$sex_c, C=d_T1$sym_dist, X=d_T1$chamber)

source("src/compare_models.R")
errors <- withWarnings(model_comparisonsAIC("src/generic models-Gamma glmer 3-FF log link.R"))
cat("Number of models that failed to converge:", length(errors$warnings))
```

```{r}
anova(m2, m6, test="Chisq") # Adding C does not improve fit
anova(m2, m4, test="Chisq") # Adding A does not improve fit
```

```{r}
best.fit <- glm(distance ~ sex_c, family = Gamma(link = "log"), data = d_T1)
tidy_regression(best.fit, is_color=output_col)
```

* no effect of sex

```{r}
s.test <- paste("pval: ", shapiro.test(residuals(best.fit))$p.value)
qqnorm(resid(best.fit))
qqline(resid(best.fit))
text(-1, 0.1, s.test)
```

<a id="w.mass"></a>

**With Mass**

```{r}
d_T1 <- d_T1 %>%
  filter(!is.na(mass))
d_T1 <- center_data(d_T1)

#data<-data.frame(R=d_T1$distance, A=d_T1$host_c, B=d_T1$sex_c, C=d_T1$sym_dist_s, D=d_T1$mass_s, X=d_T1$chamber)
data<-data.frame(R=d_T1$distance, A=d_T1$host_c, B=d_T1$sex_c, C=d_T1$mass_c, X=d_T1$chamber)

source("src/compare_models.R")
errors <- withWarnings(model_comparisonsAIC("src/generic models-Gamma glmer 3-FF log link.R")) 
cat("Number of models that failed to converge:", length(errors$warnings)) # 50 models failed to converge if keep sym_dist_s, but it doesn't show up in top models so can remove it | After removing it get 3 models that fail to converge
```

```{r}
# anova(m20, m30, test="Chisq") # Adding A does not improve fit
# anova(m20, m31, test="Chisq") # Adding C does not improve fit
# anova(m30, m50, test="Chisq") # Adding A*D does not improve fit

anova(m10, m13, test="Chisq") # Adding A does not improve fit
```

```{r}
best.fit <- glm(distance ~ sex_c * mass_c, family = Gamma(link = "log"), data = d_T1) # model failed to converge using glmer and (1|chamber)
#best.fit <- glmer(distance ~ sex_c * mass_s + (1|chamber), family = Gamma(link = "log"), data = d_T1)# model still fails to converge even when you standardize mass.
tidy_regression(best.fit, is_color=output_col)
```
* positive effect of sex where if F then more likely to disperse farther than M
* no effect of mass
* negative effect of sex*mass where if female and heavy then less likely to disperse far

```{r}
s.test <- paste("pval: ", shapiro.test(residuals(best.fit))$p.value)
qqnorm(resid(best.fit))
qqline(resid(best.fit))
text(-1, 0.1, s.test)
```

<a id="timestart"></a>

**Min From Inc Start**

```{r}
# sym dist and host isn't really showing up before so left them out
# less models fail if use standardized variables and it fixes the scaling issues.
data<-data.frame(R=d_T1$distance, A=d_T1$min_from_IncStart_s, B=d_T1$sex_c, C=d_T1$mass_s, X=d_T1$chamber)

source("src/compare_models.R")
errors <- withWarnings(model_comparisonsAIC("src/generic models-Gamma glmer 3-FF log link.R"))
cat("Number of models that failed to converge:", length(errors$warnings)) 
```

```{r}
best.fit <- glmer(distance ~ min_from_IncStart_s * sex_c + sex_c * mass_s  + (1|chamber), family = Gamma(link = "log"), data = d_T1) # converges! 
tidy_regression(best.fit, is_color=output_col)
```

* negative effect of min from start, where the later in the day the bug is tested, the less distance the bug will fly
* positive effect from sex where if F then more likely to disperse far
* no effect of mass

* negative effect of min from start*sex, where if F and tested late in the day then less likely to disperse far
* negative effect of sex*mass where the more mass and F then less likely to disperse far

```{r}
s.test <- paste("pval: ", shapiro.test(residuals(best.fit))$p.value)
qqnorm(resid(best.fit))
qqline(resid(best.fit))
text(-1, 0.1, s.test)
```

<a id="morph"></a>

**Morphology**

```{r}
d_T1 <- d_T1 %>%
  filter(!is.na(body))
d_T1 <- center_data(d_T1)

data<-data.frame(R=d_T1$distance, A=d_T1$body_c, B=d_T1$wing_c, C=d_T1$thorax_c, X=d_T1$chamber) # singular fit if run with min_from_IncStart meaning the variance is near 0

source("src/compare_models.R")
errors <- withWarnings(model_comparisonsAIC("src/generic models-Gamma glmer 3-FF log link.R"))
cat("Number of models that failed to converge:", length(errors$warnings)) 
```

```{r}
anova(m1, m4, test="Chisq") # Adding B does not improve fit
anova(m1, m6, test="Chisq") # Adding C does not improve fit
```

```{r}
best.fit <- glmer(distance ~ body_c + (1|chamber), family = Gamma(link = "log"), data = d_T1) # model failed to converge if body is centered or standardized
tidy_regression(best.fit, is_color=output_col)
```
* positive effect of body length where the longer the body the more likely the bug disperse farther

```{r}
d_T1 <- d_T1 %>%
  filter(!is.na(mass))
d_T1 <- center_data(d_T1)

data<-data.frame(R=d_T1$distance, A=d_T1$wing2body_c, B=d_T1$thorax_c, X=d_T1$chamber) 

source("src/compare_models.R")
errors <- withWarnings(model_comparisonsAIC("src/generic models-Gamma glmer 2-FF log link.R"))
cat("Number of models that failed to converge:", length(errors$warnings)) 
```

```{r}
anova(m2, m4, test="Chisq") # Adding A does not improve fit
anova(m0, m2, test="Chisq") # Adding B does improve fit
anova(m3, m4, test="Chisq") # Adding A*B does not improve fit
```

```{r}
best.fit <- glmer(distance ~ thorax_c +(1|chamber), family = Gamma(link = "log"), data = d_T1) 
tidy_regression(best.fit, is_color=output_col)
```

* positive effect of thorax length, where the longer the length the farther the dispersal

```{r}
s.test <- paste("pval: ", shapiro.test(residuals(best.fit))$p.value)
qqnorm(resid(best.fit))
qqline(resid(best.fit))
text(-1, 0.1, s.test)
```

### Trial 1 Plots 

```{r}
h1 <-  as.grob(expression(
  hist(d_T1$distance)))
p1 <- as.grob(expression(
  plot(distance ~ mass_c, data=d_T1)))
p2 <- as.grob(expression(
  plot(distance ~ days_from_start_c, data=d_T1)))
p3 <- as.grob(expression(
  plot(distance ~ min_from_IncStart, data=d_T1)))
p4 <- as.grob(expression(
  plot(distance ~ wing_c, data=d_T1)))
p5 <- as.grob(expression(
  plot(distance ~ beak_c, data=d_T1)))
p6 <- as.grob(expression(
  plot(distance ~ thorax_c, data=d_T1)))
p7 <- as.grob(expression(
  plot(distance ~ body_c, data=d_T1)))

grid.arrange(h1,p1,p2,p3,p4,p5,p6,p7, ncol=3)
```

```{r}
gf_point(distance~sym_dist, col=~host_plant, alpha=~sex_c, data=d_T1) 
summary <- aggregate(distance~sym_dist*sex*host_plant, data=d_T1, FUN=mean)
plot(summary$distance~summary$sym_dist, 
     col=c(1,2)[as.factor(summary$sex)], # Female = Black, Male = Red
     pch=c(19,22)[as.factor(summary$host_plant)], # Filled circle is GRT, Open square is BV
     main="T1 Observed Data: distance ~ sex*host_plant*sym_dist",
     xlab = "Distance from Sympatric Zone (°)",
     ylab= "Distance Flew", 
     #sub=eq_glmer
     ) 
legend("topright",
       legend = c("F and K.elegans", "M and C.corindum", "F and C.corindum","M and K.elegans"),
       #inset=c(-0.27,0.2),
       col= 1:2,
       pch = c(0,16,19),
       title="Groups")
```

### Trial 2

[Testing link and covariates](#test2)

[Without Mass](#wo.mass2)

[With Mass](#w.mass2)

[Morphology](#morph2)

<a id="test2"></a>

**Testing link functions**

```{r}
model_test<-glm(distance~chamber, data=d_T2, family=Gamma(link="log")) 
summary(model_test)
plot(model_test)
```

**Testing experimental Covariates**

```{r}
####### No effect of chamber
tidy_regression(glm(distance~chamber, data=d_T2, family=Gamma(link="log")), is_color=output_col)

####### No effect of test date
tidy_regression(glm(distance~days_from_start, data=d_T2, family=Gamma(link="log")), is_color=output_col)

####### No effect of Minutes from IncStart time
tidy_regression(glm(distance~min_from_IncStart, data=d_T2, family=Gamma(link="log")), is_color=output_col)
```

<a id="wo.mass2"></a>

**Without Mass**

```{r, message=FALSE}
data<-data.frame(R=d_T2$distance, A=d_T2$host_c, B=d_T2$sex_c, C=d_T2$sym_dist, X=d_T2$chamber)

source("src/compare_models.R")
errors <- withWarnings(model_comparisonsAIC("src/generic models-Gamma glmer 3-FF log link.R"))
cat("Number of models that failed to converge:", length(errors$warnings)) 
```

```{r}
anova(m0, m1, test="Chisq") # Adding A does not improve fit
anova(m1, m4, test="Chisq") # Adding B does not improve fit
anova(m1, m5, test="Chisq") # Adding C does not improve fit
```

```{r}
best.fit <- glm(distance ~ 1, family = Gamma(link = "log"), data = d_T2) 
tidy_regression(best.fit, is_color=output_col)
```

```{r}
s.test <- paste("pval: ", shapiro.test(residuals(best.fit))$p.value)
qqnorm(resid(best.fit))
qqline(resid(best.fit))
text(-1, 0.1, s.test)
```

<a id="w.mass2"></a>

**With Mass**

```{r, message=FALSE}
d_T2 <- d_T2 %>%
  filter(!is.na(mass))
d_T2 <- center_data(d_T2)

data<-data.frame(R=d_T2$distance, A=d_T2$host_c, B=d_T2$sex_c, C=d_T2$sym_dist, D=d_T2$mass_c, X=d_T2$chamber)

source("src/compare_models.R")
errors <- withWarnings(model_comparisonsAIC("src/generic models-Gamma glmer 4-FF log link.R"))
cat("Number of models that failed to converge:", length(errors$warnings)) 
```

```{r}
anova(m0, m4, test="Chisq") # Adding D does not improve fit
anova(m0, m1, test="Chisq") # Adding A does not improve fit
anova(m0, m3, test="Chisq") # Ading  C does not improve fit
anova(m0, m2, test="Chisq") # Adding B does not improve fit; null model the best model
```

```{r}
best.fit <- glm(distance ~ 1, family = Gamma(link = "log"), data = d_T2) 
tidy_regression(best.fit, is_color=output_col)
```

```{r}
s.test <- paste("pval: ", shapiro.test(residuals(best.fit))$p.value)
qqnorm(resid(best.fit))
qqline(resid(best.fit))
text(-1, 0.1, s.test)
```

<a id="morph2"></a>

**Morphology**

```{r, message=FALSE}
d_T2 <- d_T2 %>%
  filter(!is.na(body))
d_T2 <- center_data(d_T2)

data<-data.frame(R=d_T2$distance, A=d_T2$body_c, B=d_T2$wing_c, C=d_T2$thorax_c, X=d_T2$chamber)

source("src/compare_models.R")
errors <- withWarnings(model_comparisonsAIC("src/generic models-Gamma glmer 3-FF log link.R"))
cat("Number of models that failed to converge:", length(errors$warnings)) 
```

```{r}
anova(m0, m2, test="Chisq") # Adding B does not improve fit
anova(m0, m3, test="Chisq") # Adding C does not improve fit
```

```{r}
best.fit <- glm(distance ~ 1, family = Gamma(link = "log"), data = d_T2) 
tidy_regression(best.fit, is_color=output_col)
```

```{r}
s.test <- paste("pval: ", shapiro.test(residuals(best.fit))$p.value)
qqnorm(resid(best.fit))
qqline(resid(best.fit))
text(-1, 0.1, s.test)
```

```{r warning=FALSE}
data<-data.frame(R=d_T2$distance, A=d_T2$wing2body, B=d_T2$thorax_c, X=d_T2$chamber)

source("src/compare_models.R")
errors <- withWarnings(model_comparisonsAIC("src/generic models-Gamma glmer 2-FF log link.R"))
cat("Number of models that failed to converge:", length(errors$warnings)) 
```

```{r}
anova(m0, m2, test="Chisq") # Adding B does not improve fit
anova(m0, m3, test="Chisq") # Adding C does not improve fit
```

```{r}
best.fit <- glm(distance ~ 1, family = Gamma(link = "log"), data = d_T2) 
tidy_regression(best.fit, is_color=output_col)
```

```{r}
s.test <- paste("pval: ", shapiro.test(residuals(best.fit))$p.value)
qqnorm(resid(best.fit))
qqline(resid(best.fit))
text(-1, 0.1, s.test)
```

### Trial 2 Plots

```{r}
h1 <-  as.grob(expression(
  hist(d_T2$distance)))
p1 <- as.grob(expression(
  plot(distance ~ mass_c, data=d_T2)))
p2 <- as.grob(expression(
  plot(distance ~ days_from_start_c, data=d_T2)))
p3 <- as.grob(expression(
  plot(distance ~ min_from_IncStart, data=d_T2)))
p4 <- as.grob(expression(
  plot(distance ~ wing_c, data=d_T2)))
p5 <- as.grob(expression(
  plot(distance ~ beak_c, data=d_T2)))
p6 <- as.grob(expression(
  plot(distance ~ thorax_c, data=d_T2)))
p7 <- as.grob(expression(
  plot(distance ~ body_c, data=d_T2)))

grid.arrange(h1,p1,p2,p3,p4,p5,p6,p7, ncol=3)
```


```{r}
gf_point(distance~sym_dist, col=~host_plant, alpha=~sex_c, data=d_T2) 
summary <- aggregate(distance~sym_dist*sex*host_plant, data=d_T2, FUN=mean)
plot(summary$distance~summary$sym_dist, 
     col=c(1,2)[as.factor(summary$sex)], # Female = Black, Male = Red
     pch=c(19,22)[as.factor(summary$host_plant)], # Filled circle is GRT, Open square is BV
     main="T2 Observed Data: distance ~ sex*host_plant*sym_dist",
     xlab = "Distance from Sympatric Zone (°)",
     ylab= "Distance Flew", 
     #sub=eq_glmer
     ) 
legend("topright",
       legend = c("F and K.elegans", "M and C.corindum", "F and C.corindum","M and K.elegans"),
       #inset=c(-0.27,0.2),
       col= 1:2,
       pch = c(0,16,19),
       title="Groups")
```

### Females

[Testing Link Functions](#ftest)

[Without Mass](#fwomass)

[With Mass](#fmass)

[Minutes from IncStart](#fmin)

[Morphology](#fmorph)

[Eggs](#feggs)

<a id="ftest"></a>

**Testing link functions**

```{r}
model_test<-glm(distance~chamber, data=d_fem, family=Gamma(link="log")) # glm.fit algorithm did not converge
summary(model_test)
plot(model_test)
```

**Testing covariates**

```{r}
####### Effect of chamber A-1, B-2, and marginal effect of B-4 (also algorithm did not converge)
d_fem$chamber <- relevel(d_fem$chamber, ref="A-2")
tidy_regression(glm(distance~chamber, data=d_fem, family=Gamma(link="log")), is_color=output_col)

####### No effect of test date
tidy_regression(glm(distance~days_from_start, data=d_fem, family=Gamma(link="log")), is_color=output_col)

####### Marginal effect of Minutes from IncStart time
tidy_regression(glm(distance~min_from_IncStart, data=d_fem, family=Gamma(link="log")), is_color=output_col)

#### No effect of eggs laid
tidy_regression(glm(distance~eggs_b, data=d_fem, family=Gamma(link="log")), is_color=output_col)

### No effect of total eggs laid
tidy_regression(glm(distance~total_eggs, data=d_fem, family=Gamma(link="log")), is_color=output_col)

```

<a id="fwomass"></a>

**Without Mass**

```{r, message=FALSE}
data<-data.frame(R=d_fem$distance, A=d_fem$host_c, B=d_fem$sym_dist, X=d_fem$chamber)

source("src/compare_models.R")
errors <- withWarnings(model_comparisonsAIC("src/generic models-Gamma glmer 2-FF log link.R"))
cat("Number of models that failed to converge:", length(errors$warnings)) 
```

```{r}
anova(m0, m2, test="Chisq") # Adding B does not improve fit
anova(m0, m1, test="Chisq") # Adding A does not improve fit
anova(m0, m3, test="Chisq") # Adding C does not improve fit
```

```{r}
best.fit <- glm(distance ~ 1, family = Gamma(link = "log"), data = d_fem) 
tidy_regression(best.fit, is_color=output_col)
```

```{r}
s.test <- paste("pval: ", shapiro.test(residuals(best.fit))$p.value)
qqnorm(resid(best.fit))
qqline(resid(best.fit))
text(-1, 0.1, s.test)
```


<a id="fmass"></a>

**With Mass**

```{r, message=FALSE}
d_fem <- d_fem %>%
  filter(!is.na(mass))
d_fem <- center_data(d_fem)

data<-data.frame(R=d_fem$distance, A=d_fem$host_c, B=d_fem$sym_dist, C=d_fem$mass_c, X=d_fem$chamber)

source("src/compare_models.R")
errors <- withWarnings(model_comparisonsAIC("src/generic models-Gamma glmer 3-FF log link.R")) 
cat("Number of models that failed to converge:", length(errors$warnings))
```

```{r}
anova(m3, m6, test="Chisq") # Adding B does not improve fit
anova(m3, m5, test="Chisq") # Adding A does not improve fit
```

```{r}
best.fit <- glm(distance ~ mass_c, family = Gamma(link = "log"), data = d_fem) 
tidy_regression(best.fit, is_color=output_col)
```

* no effect of mass

```{r}
s.test <- paste("pval: ", shapiro.test(residuals(best.fit))$p.value)
qqnorm(resid(best.fit))
qqline(resid(best.fit))
text(-1, 0.1, s.test)
```


<a id="fmin"></a>

**Min From Inc Start**

```{r message=FALSE}
data<-data.frame(R=d_fem$distance, A=d_fem$min_from_IncStart_s, B=d_fem$sym_dist_s, C=d_fem$mass_s, X=d_fem$chamber)

source("src/compare_models.R")
errors <- withWarnings(model_comparisonsAIC("src/generic models-Gamma glmer 3-FF log link.R"))
cat("Number of models that failed to converge:", length(errors$warnings))
```

```{r}
anova(m0, m1, test="Chisq") # Adding A does not improve fit
anova(m0, m3, test="Chisq") # Adding C does not impove fit
anova(m0, m2, test="Chisq") # Adding B does not improve fit
```

```{r}
best.fit <- glm(distance ~ 1, family = Gamma(link = "log"), data = d_fem) 
tidy_regression(best.fit, is_color=output_col)
```

```{r}
s.test <- paste("pval: ", shapiro.test(residuals(best.fit))$p.value)
qqnorm(resid(best.fit))
qqline(resid(best.fit))
text(-1, 0.1, s.test)
```

<a id="fmorph"></a>

**Morphology**

```{r message=FALSE}
d_fem <- d_fem %>%
  filter(!is.na(mass))
d_fem <- center_data(d_fem)

data<-data.frame(R=d_fem$distance, A=d_fem$body_c, B=d_fem$wing_c, C=d_fem$thorax_c, X=d_fem$chamber) 

source("src/compare_models.R")
errors <- withWarnings(model_comparisonsAIC("src/generic models-Gamma glmer 3-FF log link.R"))
cat("Number of models that failed to converge:", length(errors$warnings)) 
```

```{r}
anova(m13, m16, test="Chisq") # Adding A*C does not improve fit
anova(m15, m16, test="Chisq") # Adding B*C does improve fit
```

```{r}
best.fit <- glmer(distance ~ wing_c*thorax_c + body_c + (1|chamber), family = Gamma(link = "log"), data = d_fem) 
tidy_regression(best.fit, is_color=output_col)
```
* no effect of wing length
* positive effect of thorax where the wider the thorax, the more likely the bug will disperse farther
* no effect of body
* negative effect of wing*thorax where the wider the thorax and longer the wings, the more likely the bug will disperse shorter distances

```{r}
s.test <- paste("pval: ", shapiro.test(residuals(best.fit))$p.value)
qqnorm(resid(best.fit))
qqline(resid(best.fit))
text(-1, 0.1, s.test)
```

```{r message=FALSE}
d_fem <- d_fem %>%
  filter(!is.na(mass))
d_fem <- center_data(d_fem)

data<-data.frame(R=d_fem$distance, A=d_fem$wing2body, B=d_fem$thorax_c, X=d_fem$chamber)

source("src/compare_models.R")
errors <- withWarnings(model_comparisonsAIC("src/generic models-Gamma glmer 2-FF log link.R"))
cat("Number of models that failed to converge:", length(errors$warnings)) 
```

```{r}
best.fit <- glmer(distance ~ wing2body*thorax_c + (1|chamber), family = Gamma(link = "log"), data = d_fem) 
tidy_regression(best.fit, is_color=output_col) # the coeffients are crazy large
```

* no effect of wing2body
* positive effect of thorax where the wider the thorax the more likely the bug flies far
* negative effect of wing2body*thorax where the wider the thorax and longer the wings in comparison to the body, the less likely the bugs fly far (contradicting)

```{r}
s.test <- paste("pval: ", shapiro.test(residuals(best.fit))$p.value)
qqnorm(resid(best.fit))
qqline(resid(best.fit))
text(-1, 0.1, s.test)
```

<a id="feggs"></a>

**Eggs**

```{r message=FALSE}
d_fem <- d_fem %>%
  filter(!is.na(mass))
d_fem <- center_data(d_fem)

data<-data.frame(R=d_fem$distance, A=d_fem$eggs_b, B=d_fem$mass_c, C=d_fem$thorax_c, X=d_fem$chamber) # adding D=d_fem$wing2body_c leads to a massive converging error

source("src/compare_models.R")
errors <- withWarnings(model_comparisonsAIC("src/generic models-Gamma glmer 3-FF log link.R"))
cat("Number of models that failed to converge:", length(errors$warnings)) 
```

```{r}
anova(m13, m16, test="Chisq") # Adding A*C does not improve fit
```

```{r}
best.fit <- glmer(distance ~  mass_c * thorax_c + eggs_b +  (1|chamber), family = Gamma(link = "log"), data = d_fem) 
tidy_regression(best.fit, is_color=output_col) # the coeffients are crazy large
```
* negative effect of mass where the larger the mass the less likely the bug will disperse far
* positive effect of thorax whre the wider the thorax the more likely the bug will disperse far
* positive effect of eggs laid on test day where if female laid eggs then more likely to disperse far
* negative effect of mass*thorax where if heavy and have a wide thorax then less likley to disperse far (conflicting)

```{r}
s.test <- paste("pval: ", shapiro.test(residuals(best.fit))$p.value)
qqnorm(resid(best.fit))
qqline(resid(best.fit))
text(-1, 0.1, s.test)
```

### Female Plots 

```{r}
h1 <-  as.grob(expression(
  hist(d_fem$distance)))
p1 <- as.grob(expression(
  plot(distance ~ mass_c, data=d_fem)))
p2 <- as.grob(expression(
  plot(distance ~ days_from_start_c, data=d_fem)))
p3 <- as.grob(expression(
  plot(distance ~ min_from_IncStart, data=d_fem)))
p4 <- as.grob(expression(
  plot(distance ~ wing_c, data=d_fem)))
p5 <- as.grob(expression(
  plot(distance ~ beak_c, data=d_fem)))
p6 <- as.grob(expression(
  plot(distance ~ thorax_c, data=d_fem)))
p7 <- as.grob(expression(
  plot(distance ~ body_c, data=d_fem)))
p8 <- as.grob(expression(
  plot(distance ~ eggs_b, data=d_fem)))
grid.arrange(h1,p1,p2,p3,p4,p5,p6,p7,p8, ncol=3)
```

```{r}
gf_point(distance~sym_dist, col=~host_plant, alpha=~sex_c, data=d_fem) 
summary <- aggregate(distance~sym_dist*sex*host_plant, data=d_fem, FUN=mean)
plot(summary$distance~summary$sym_dist, 
     #col=c(1,2)[as.factor(summary$sex)], # Female = Black, Male = Red
     pch=c(19,22)[as.factor(summary$host_plant)], # Filled circle is GRT, Open square is BV
     main="Female Observed Data: distance ~ sex*host_plant*sym_dist",
     xlab = "Distance from Sympatric Zone (°)",
     ylab= "Distance Flew", 
     #sub=eq_glmer
     ) 
legend("topright",
       legend = c("F and K.elegans","F and C.corindum"),
       #inset=c(-0.27,0.2),
       col= 1,
       pch = c(0,16,19),
       title="Groups")
```

### Males 

[Testing Link Functions](#mtest)

[Without Mass](#mwomass)

[With Mass](#mmass)

[Minutes from IncStart](#mmin)

[Morphology](#mmorph)

<a id="mtest"></a>

**Testing link functions**

```{r}
model_test<-glm(distance~chamber, data=d_male, family=Gamma(link="log"))
summary(model_test)
plot(model_test)
```

**Testing covariates**

```{r}
####### Effect of chamber A-4
d_male$chamber <- relevel(d_male$chamber, ref="A-2")
tidy_regression(glm(distance~chamber, data=d_male, family=Gamma(link="log")), is_color=output_col)

####### No effect of test date
tidy_regression(glm(distance~days_from_start, data=d_male, family=Gamma(link="log")), is_color=output_col)

####### Effect of Minutes from IncStart time
tidy_regression(glm(distance~min_from_IncStart, data=d_male, family=Gamma(link="log")), is_color=output_col)

#### No effect of eggs laid
tidy_regression(glm(distance~eggs_b, data=d_male, family=Gamma(link="log")), is_color=output_col)

### No effect of total eggs laid
tidy_regression(glm(distance~total_eggs, data=d_male, family=Gamma(link="log")), is_color=output_col)

```

<a id="mwomass"></a>

**Without Mass**

```{r, message=FALSE}
data<-data.frame(R=d_male$distance, A=d_male$host_c, B=d_male$sym_dist, X=d_male$chamber)

source("src/compare_models.R")
errors <- withWarnings(model_comparisonsAIC("src/generic models-Gamma glmer 2-FF log link.R"))
cat("Number of models that failed to converge:", length(errors$warnings)) 
```

```{r}
anova(m0, m2, test="Chisq") # Adding B does not improve fit
anova(m0, m1, test="Chisq") # Adding A does not improve fit
anova(m0, m3, test="Chisq") # Adding C does not improve fit
```

```{r}
best.fit <- glm(distance ~ 1, family = Gamma(link = "log"), data = d_male) 
tidy_regression(best.fit, is_color=output_col)
```

```{r}
s.test <- paste("pval: ", shapiro.test(residuals(best.fit))$p.value)
qqnorm(resid(best.fit))
qqline(resid(best.fit))
text(-1, 0.1, s.test)
```

<a id="mmass"></a>

**With Mass**

```{r, message=FALSE}
d_male <- d_male %>%
  filter(!is.na(mass))
d_male <- center_data(d_male)

data<-data.frame(R=d_male$distance, A=d_male$host_c, B=d_male$sym_dist, C=d_male$mass_c, X=d_male$chamber)

source("src/compare_models.R")
errors <- withWarnings(model_comparisonsAIC("src/generic models-Gamma glmer 3-FF log link.R")) 
cat("Number of models that failed to converge:", length(errors$warnings)) 
```

```{r}
anova(m0, m3, test="Chisq") # Adding C marginally improves fit
anova(m3, m5, test="Chisq") # Adding A does not improve fit
```

```{r}
best.fit <- glm(distance ~ mass_c, family = Gamma(link = "log"), data = d_male) 
tidy_regression(best.fit, is_color=output_col)
```

* marginal positive effect of mass, where the heavier the male bug, the more likely the male bug flies far (this could be because of thorax wider so has more muscle power?)

```{r}
s.test <- paste("pval: ", shapiro.test(residuals(best.fit))$p.value)
qqnorm(resid(best.fit))
qqline(resid(best.fit))
text(-1, 0.1, s.test)
```

<a id="mmin"></a>

**Min From Inc Start**

```{r message=FALSE}
data<-data.frame(R=d_male$distance, A=d_male$min_from_IncStart_s, B=d_male$sym_dist_s, C=d_male$mass_s, X=d_male$chamber)

source("src/compare_models.R")
errors <- withWarnings(model_comparisonsAIC("src/generic models-Gamma glmer 3-FF log link.R"))
cat("Number of models that failed to converge:", length(errors$warnings))
```

```{r}
anova(m5, m9, test="Chisq") # adding A*C does not improve fit
anova(m3, m5, test="Chisq") # Adding A does not improve fit
anova(m0, m1, test="Chisq") # Adding A does not improve fit
anova(m0, m3, test="Chisq") # Adding C marginally improves fit

# Do we go with anova or AIC? I went with anova step-wise comparisons
```

```{r}
best.fit <- glmer(distance ~ mass_s + (1|chamber), family = Gamma(link = "log"), data = d_fem) 
tidy_regression(best.fit, is_color=output_col)
```

* no effect of mass

```{r}
s.test <- paste("pval: ", shapiro.test(residuals(best.fit))$p.value)
qqnorm(resid(best.fit))
qqline(resid(best.fit))
text(-1, 0.1, s.test)
```

<a id="mmorph"></a>

**Morphology**

```{r message=FALSE}
d_male <- d_male %>%
  filter(!is.na(body))
d_male <- center_data(d_male)

data<-data.frame(R=d_male$distance, A=d_male$body_c, B=d_male$wing_c, C=d_male$thorax_c, X=d_male$chamber) 

source("src/compare_models.R")
errors <- withWarnings(model_comparisonsAIC("src/generic models-Gamma glmer 3-FF log link.R"))
cat("Number of models that failed to converge:", length(errors$warnings)) 
```

```{r}
anova(m9, m12, test="Chisq") # Adding B does not improve fit
```

```{r}
best.fit <- glmer(distance ~ body_c*thorax_c + (1|chamber), family = Gamma(link = "log"), data = d_male)
tidy_regression(best.fit, is_color=output_col)
```
* positive effect of body, where the longer the body the more likely the bug will disperse farther
* negative effect of thorax where the wider the thorax, the less likely the bug will disperse farther

* negative effect of body*thorax where the wider the thorax and longer the body, the less likely the bug will disperse farther distances

```{r}
s.test <- paste("pval: ", shapiro.test(residuals(best.fit))$p.value)
qqnorm(resid(best.fit))
qqline(resid(best.fit))
text(-1, 0.1, s.test)
```

```{r message=FALSE}
data<-data.frame(R=d_male$distance, A=d_male$wing2body, B=d_male$thorax_c, X=d_male$chamber)

source("src/compare_models.R")
errors <- withWarnings(model_comparisonsAIC("src/generic models-Gamma glmer 2-FF log link.R"))
cat("Number of models that failed to converge:", length(errors$warnings)) 
```

```{r}
anova(m0, m2, test="Chisq") # Adding B marginally improves fit
anova(m0, m1, test="Chisq") # Adding A does not improve fit
```

```{r}
best.fit <- glmer(distance ~ thorax_c + (1|chamber), family = Gamma(link = "log"), data = d_male) 
tidy_regression(best.fit, is_color=output_col)
```

*marginal positive effect of thorax where the wider the thorax the more likely the bug will disperse far.

```{r}
s.test <- paste("pval: ", shapiro.test(residuals(best.fit))$p.value)
qqnorm(resid(best.fit))
qqline(resid(best.fit))
text(-1, 0.1, s.test)
```


```{r message=FALSE}
data<-data.frame(R=d_male$distance, A=d_male$wing2body_s, B=d_male$thorax_s, C=d_male$mass_s, X=d_male$chamber)

source("src/compare_models.R")
errors <- withWarnings(model_comparisonsAIC("src/generic models-Gamma glmer 3-FF log link.R"))
cat("Number of models that failed to converge:", length(errors$warnings)) 
```

```{r}
anova(m6, m10, test="Chisq") # Adding B*C improves fit
```

```{r}
best.fit <- glmer(distance ~ mass_s * thorax_s + (1|chamber), family = Gamma(link = "log"), data = d_male)
tidy_regression(best.fit, is_color=output_col)
```

* no effect of mass
* no effect of thorax
* negative effect of thorax and mass where if weighs more and wider thorax, then less likely to disperse far.

```{r}
s.test <- paste("pval: ", shapiro.test(residuals(best.fit))$p.value)
qqnorm(resid(best.fit))
qqline(resid(best.fit))
text(-1, 0.1, s.test)
```


### Male Plots 

```{r}
h1 <-  as.grob(expression(
  hist(d_male$distance)))
p1 <- as.grob(expression(
  plot(distance ~ mass_c, data=d_male)))
p2 <- as.grob(expression(
  plot(distance ~ days_from_start_c, data=d_male)))
p3 <- as.grob(expression(
  plot(distance ~ min_from_IncStart, data=d_male)))
p4 <- as.grob(expression(
  plot(distance ~ wing_c, data=d_male)))
p5 <- as.grob(expression(
  plot(distance ~ beak_c, data=d_male)))
p6 <- as.grob(expression(
  plot(distance ~ thorax_c, data=d_male)))
p7 <- as.grob(expression(
  plot(distance ~ body_c, data=d_male)))

grid.arrange(h1,p1,p2,p3,p4,p5,p6,p7, ncol=3)
```

```{r}
gf_point(distance~sym_dist, col=~host_plant, alpha=~sex_c, data=d_male) 
summary <- aggregate(distance~sym_dist*sex*host_plant, data=d_male, FUN=mean)
plot(summary$distance~summary$sym_dist, 
     #col=c(1,2)[as.factor(summary$sex)], # Female = Black, Male = Red
     pch=c(19,22)[as.factor(summary$host_plant)], # Filled circle is GRT, Open square is BV
     main="Male Observed Data: distance ~ sex*host_plant*sym_dist",
     xlab = "Distance from Sympatric Zone (°)",
     ylab= "Distance Flew", 
     #sub=eq_glmer
     ) 
legend("topright",
       legend = c("M and K.elegans","M and C.corindum"),
       #inset=c(-0.27,0.2),
       col= 1,
       pch = c(0,16,19),
       title="Groups")
```