---
title: "Binomial Modeling: flight_yes_no"
author: "Anastasia Bernat"
date: "3/30/2020"
output: html_document
---

# Winter 2020 Flight Trials 

```{r setup, include=FALSE}
rm(list=ls())
setwd("~/Desktop/Rstats-winter2020/")

library(lme4)
library(lubridate)
library(dplyr)
library(glmnet)
library(chron)
library(ggplotify)
library(gridExtra)
library(ggformula)
library(tidyselect)

knitr::opts_chunk$set(echo = TRUE)
```

# Reading the data

Flight Trials Winter 2020 Dataset was conducted from 2/17/2020 - 3/10/2020. Soapberry bugs were flight tested twice in the flight mill and observed from 8 AM to (5-8 PM) each day.

# Loading, cleaning, and recoding the data

```{r}
# Script that loads and cleans up the data
source("clean_flight_data.R")
data_all <- read_flight_data("complete_flight_data-Winter2020.csv")
```

# GLM Modeling: Multi-hour trials

```{r}
glmer(flew_b~sex_c*host_c + (1|ID) + (1|trial_type), data=data_all, family=binomial)
model<-glmer(flew_b~host_c*sex_c + (1|ID) + (1|trial_type), data=data_all, family=binomial)
summary(model)
getME(model, "lower")
```

## All trials

```{r}
# Experimental Set-Up Covariates:

####### No effect of chamber
summary(glm(flew_b~chamber, data=data_all, family=binomial))

####### Effect of test date (but no neffect of test date when you split between T1 and T2)
summary(glm(flew_b~days_from_start_c, data=data_all, family=binomial))

####### No effect of test time
summary(glm(flew_b~min_from_IncStart_c, data=data_all, family=binomial))

```

```{r}
# Biological Effects:

####### (Strong) Effect of mass
summary(glm(flew_b~mass_c, data=data_all, family=binomial))

####### Effect of number of eggs laid
summary(glm(flew_b~total_eggs, data=data_all, family=binomial))

####### Effect of whether eggs were laid or not
summary(glm(flew_b~eggs_b, data=data_all, family=binomial))
```

```{r}
# Morphology Effects:

####### Effect of beak length
summary(glm(flew_b~beak_c, data=data_all, family=binomial))

####### Effect of thorax length
summary(glm(flew_b~thorax_c, data=data_all, family=binomial))

####### Effect of body length
summary(glm(flew_b~body_c, data=data_all, family=binomial))

####### No effect of wing length
summary(glm(flew_b~wing_c, data=data_all, family=binomial))

####### No effect of wing morph (check how annotated the wing morph) - don't include it
#summary(glm(flew_b~w_morph_c, data=data_all, family=binomial)) # but close p val = 0.0512

```

## TRIAL 1: GLM Binomial Modeling 

```{r}
data_T1 <-data_all[data_all$trial_type=="T1",]

####### No effect of chamber
summary(glm(flew_b~chamber, data=data_T1, family=binomial))

####### No effect of test date
summary(glm(flew_b~days_from_start_c, data=data_T1, family=binomial))

####### No effect of test time
summary(glm(flew_b~min_from_IncStart, data=data_T1, family=binomial))
```

```{r}
# Remove any missing masses
missing_mass <- subset(data_T1, is.na(data_T1$mass))
data_T1 <- data_T1 %>%
  filter(!is.na(mass))

# Recentering
data_T1$lat_c<-data_T1$latitude-mean(data_T1$latitude)
data_T1$sym_dist<-abs(data_T1$latitude-25.49197)
data_T1$mass_c <- data_T1$mass-mean(data_T1$mass, na.rm=TRUE) 

R = data_T1$flew_b
A = data_T1$host_c
B = data_T1$sex_c
C = data_T1$sym_dist
D = data_T1$mass_c # as a covariate 
##I would not compete these against each other - because mass is determined in part by A, B, and C, I would not add it until after we've tested the other three.

data<-data.frame(R, A, B, C, D)

# Automatic stepwise functions using AIC to decide between models
model.null <- glm(R~1, family=binomial, data=data)
model.large <-glm(R~A*B*C*D, family=binomial, data=data)

# Backwards
#m_backwards <- step(model.large, direction = "backward")
```
For backwards selection:

- Minimum AIC value from the backwards selection function:  m_backwards <- glm(formula = R ~ A + B + C + D + A:D + C:D, family = binomial, data = data), AIC: 400.48

```{r}
# Forwards
#m_forwards <- step(model.null, scope = ~ A*B*C*D, direction = "forward")
```

For forwards selection:

- Minimum AIC value from the forwards selection function:  m_forwards <- glm(formula = R ~ D + B, family = binomial, data = data_T1), AIC: 404.83

```{r}
# Glmnet is a package that fits a generalized linear model via penalized maximum likelihood.
# source("glmnet_analyses-3FF.R")
# See that 4 = D = mass has a much greater coefficient than the rest
# The cross validation curve has large error bars.
```

```{r}
# TRIAL 1 WITHOUT MASS AS A COVARIATE

#run AICprobs script
source("AICprobabilities.R")
source("generic models-binomial glm 3-FF.R")
AICs <- sort(summary$AIC)
models_init <- sort(P, decreasing=TRUE, index.return=TRUE)
top <- length(models_init$x[which(models_init$x>0.05)])

AICs <- AICs[1:top]
models <- models_init$ix[1:top]
probs <- models_init$x[1:top]
rbind(AICs, models, probs)
```

**All models with  >0.05 probability.**

m8 <- glm(formula = R ~ A * B, family = binomial, data = data)
m11 <- glm(formula = R ~ A * B + C, family = binomial, data = data)
m15 <- glm(formula = R ~ A * B + B * C, family = binomial, data = data)
m2 <- glm(formula = R ~ B, family = binomial, data = data)
m14 <- glm(formula = R ~ A * B + A * C, family = binomial, data = data)

```{r}
anova(m8, m11, test="Chisq") ## Adding C does not improve fit
anova(m8, m15, test="Chisq") ## Adding B*C interaction does not improve fit
anova(m8, m2, test="Chisq") ## Removing A does improve fit
```

```{r}
model_T1<-glm(flew_b~sex_c*host_c, family=binomial, data=data_T1)
summary(model_T1)
```
* strong negative effect of sex, where if you are female you are less likely to fly.
* no longer an effect of host or distance from sympatric zone. (last season there was)
* strong interaction between sex and host where if female and from GRT, then more likely to fly.

```{r}
model_T1_final <-glmer(flew_b~sex_c*host_c + (1|population), family=binomial, data=data_T1) #  boundary (singular) fit: see ?isSingular error
summary(model_T1_final) 
## Did not change effect estimates or improve model fit
```

```{r}
# TRIAL 1 WITH MASS AS A COVARIATE

#run AICprobs script
source("AICprobabilities.R")
source("generic models-binomial glm 4-FF.R")
AICs <- sort(summary$AIC)
models_init <- sort(P, decreasing=TRUE, index.return=TRUE)
top <- length(models_init$x[which(models_init$x>0.05)])

AICs <- AICs[1:top]
models <- models_init$ix[1:top]
probs <- models_init$x[1:top]
rbind(AICs, models, probs)
```

m26 <- glm(formula = R ~ A * D + B, family = binomial, data = data)
m63 <- glm(formula = R ~ A * D + C * D + B, family = binomial, data = data)

```{r}
anova(m26, m63, test="Chisq") ## Adding the C*D interaction does not improve fit
```

Comparing models: Because most changes in deviance are positive, meaning that Model 2 had a higher deviance value, this suggests that the Model 1 fits better Model 2.

```{r}
model_T1_m<-glm(flew_b~host_c*mass_c + sex_c, family=binomial, data=data_T1)
summary(model_T1_m)
```

* no longer a strong negative effect of being far from the sympatric zone; 
* no effect of host plant
* strong negative effect of mass and has a large coefficient, where the heavier the bug is the much less likely the bug will fly.
* host*sex interaction shows a strong positive effect, such that if from GRT and are female then more likely to be dispersive
* marginal effect of sex, where if you are female you are less likely to fly

```{r}
model_T1_mfinal <-glmer(flew_b~host_c*sex_c + mass_c + (1|population), 
                    family=binomial, data=data_T1) #  boundary (singular) fit: see ?isSingular error
summary(model_T1_mfinal) 
## Did change effect estimates, sex no longer has an effect, the host*sex interaction has a much smaller effect but still significant, and the mass effect has increased.
```

## Glmnet Plotting for Trial 1

```{r}
# Glmnet is a package that fits a generalized linear model via penalized maximum likelihood.
# source("glmnet_analyses-4FF.R")
# See that 4 = D = mass has a much greater coefficient than the rest
# The cross validation curve has large error bars.
```

## T2: GLM Binomial Modeling | Mass as a covariate 

```{r}
source("clean_flight_data.R")
data_all <- read_flight_data("complete_flight_data-Winter2020.csv")
data_T2 <-data_all[data_all$trial_type=="T2",]

####### No effect of chamber
summary(glm(flew_b~chamber, data=data_T2, family=binomial))

####### No effect of test date
summary(glm(flew_b~days_from_start_c, data=data_T2, family=binomial))

####### No effect of test time
summary(glm(flew_b~min_from_IncStart, data=data_T2, family=binomial))
```

```{r}
# Recentering
data_T2$lat_c<-data_T2$latitude-mean(data_T2$latitude)
data_T2$sym_dist<-abs(data_T2$latitude-25.49197)
data_T2$mass_c <- data_T2$mass-mean(data_T2$mass, na.rm=TRUE) 

R = data_T2$flew_b
A = data_T2$host_c
B = data_T2$sex_c
C = data_T2$sym_dist
#D = data_T2$mass_c # as a covariate 

#data<-data.frame(R, A, B, C, D)
data<-data.frame(R, A, B, C)
head(data) #kable()

# Automatic stepwise functions using AIC to decide between models
model.null <- glm(R~1, family=binomial, data = data_T2)
model.large <-glm(R~A*B*C, family=binomial, data=data_T2)

# Backwards
#step(model.large, direction = "backward")
```

For Backwards:

best_fit <- glm(formula = R ~ A + B + C + D + A:B + A:C + B:C + A:B:C
AIC: 364.7

```{r}
# Forwards
#step(model.null, scope = ~ (A*B*C*D), direction = "forward")
```
For forwards:

best_fit <- glm(formula = R ~ D + A
AIC:356.7

```{r}
# TRIAL 2 WITHOUT MASS AS COVARIATE
#run AICprobs script
source("AICprobabilities.R")
source("generic models-binomial glm 3-FF.R")
AICs <- sort(summary$AIC)
models_init <- sort(P, decreasing=TRUE, index.return=TRUE)
top <- length(models_init$x[which(models_init$x>0.05)])

AICs <- AICs[1:top]
models <- models_init$ix[1:top]
probs <- models_init$x[1:top]
rbind(AICs, models, probs)
```

m2 <- glm(formula = R ~ B, family = binomial, data = data)
m4 <- glm(formula = R ~ A + B, family = binomial, data = data)
m6 <- glm(formula = R ~ B + C, family = binomial, data = data)
m7 <- glm(formula = R ~ A + B + C, family = binomial, data = data)
m8 <- glm(formula = R ~ A * B, family = binomial, data = data)
m10 <- glm(formula = R ~ B * C, family = binomial, data = data)

```{r}
anova(m2, m4, test="Chisq") # Adding A does not improve fit
anova(m2, m6, test="Chisq") # Adding C does not improve fit
anova(m2, m7, test="Chisq") # Adding A and B does not improve fit
anova(m2, m8, test="Chisq") # Adding an A*B interaction term does not improve fit
anova(m2, m10, test="Chisq") # Adding a B*C interaction term does not improve fit
```

```{r}
model_T2 <-glm(flew_b~sex_c, family=binomial, data=data_T2)
summary(model_T2)
```

* sex is the only significant and strong effect, so that if you are female you are much less likely to disperse/fly

```{r}
## Consider covariates
model_T2_final <-glmer(flew_b~sex_c + (1|population), family=binomial, data=data_T2)
summary(model_T2_final)
## Did not change effect estimates or improve fit
```

```{r}
# TRIAL 2 MASS AS COVARIATE

source("clean_flight_data.R")
data_all <- read_flight_data("complete_flight_data-Winter2020.csv")
data_T2 <-data_all[data_all$trial_type=="T2",]

# Recentering
data_T2$lat_c<-data_T2$latitude-mean(data_T2$latitude)
data_T2$sym_dist<-abs(data_T2$latitude-25.49197)
data_T2$mass_c <- data_T2$mass-mean(data_T2$mass, na.rm=TRUE) 

R = data_T2$flew_b
A = data_T2$host_c
B = data_T2$sex_c
C = data_T2$sym_dist
D = data_T2$mass_c # as a covariate 

data<-data.frame(R, A, B, C, D)
head(data) #kable()

# Automatic stepwise functions using AIC to decide between models
model.null <- glm(R~1, family=binomial, data = data_T2)
model.large <-glm(R~A*B*C*D, family=binomial, data=data_T2)

# Backwards
#step(model.large, direction = "backward")
```

```{r}
# TRIAL 2 WITHOUT MASS AS COVARIATE
#run AICprobs script
source("AICprobabilities.R")
source("generic models-binomial glm 4-FF.R")
AICs <- sort(summary$AIC)
models_init <- sort(P, decreasing=TRUE, index.return=TRUE)
top <- length(models_init$x[which(models_init$x>0.05)])

AICs <- AICs[1:top]
models <- models_init$ix[1:top]
probs <- models_init$x[1:top]
rbind(AICs, models, probs)
```
m7 <- glm(formula = R ~ A + D, family = binomial, data = data)
m4 <- glm(formula = R ~ D, family = binomial, data = data)
m9 <- glm(formula = R ~ B + D, family = binomial, data = data)
m12 <- glm(formula = R ~ A + B + D, family = binomial, data = data)
```{r}
anova(m7, m4, test="Chisq") # Removing A does improve fit (neg Dev)
anova(m7, m9, test="Chisq") # REplacing A with B does improve fit (neg Dev)
anova(m7, m12, test="Chisq") # Adding B does not improve fit
anova(m4, m9, test="Chisq") # Adding B does not improve fit
```

```{r}
model_T2_mass <-glm(flew_b~mass_c, family=binomial, data=data_T2)
summary(model_T2_mass)
```

* mass is the only significant and extremely strong effect, so that if you are heavy you are much less likely to disperse/fly

```{r}
## Consider covariates
model_T2_mass_final <-glmer(flew_b~mass_c + (1|population), family=binomial, data=data_T2)
summary(model_T2_mass_final)
## Did not change effect estimates or improve fit
```


Mass is really dominating everything, so let's split by sex after looking at Trial 2.

# ------------------------------------------------------------------------------------

# SPLITTING BY SEX (due to strong effect of mass)

```{r}
source("clean_flight_data.R")
data_all <- read_flight_data("complete_flight_data-Winter2020.csv")
data_fem <- data_all[data_all$sex=="F",]
```

## FEMALE: GLM Binomial Modeling 

```{r}
####### Marginal effect of chamber B-4 (ah this again!)
summary(glm(flew_b~chamber, data=data_fem, family=binomial))

####### No effect of test date
summary(glm(flew_b~days_from_start, data=data_fem, family=binomial))

####### No effect of test time
summary(glm(flew_b~min_from_IncStart, data=data_fem, family=binomial))
```

```{r}
# Recentering
data_fem$lat_c<-data_fem$latitude-mean(data_fem$latitude)
data_fem$sym_dist<-abs(data_fem$latitude-25.49197)
data_fem$mass_c <- data_fem$mass-mean(data_fem$mass, na.rm=TRUE) 

R = data_fem$flew_b
A = data_fem$host_c
B = data_fem$sym_dist

data<-data.frame(R, A, B)
head(data) #kable()

# Automatic stepwise functions using AIC to decide between models
model.null <- glm(R~1, family=binomial, data = data_fem)
model.large <-glm(R~A*B, family=binomial, data = data_fem)

# Backwards
#step(model.large, direction = "backward")
```

```{r}
## Forwards
#step(model.null, scope = ~ (A*B*C), direction = "forward")
```
For forwards and backwards:

best_fit <- glm(formula = R ~ 1, family = binomial, data = data_fem)
AIC:266.3

```{r}
# FEM DATA WITHOUT MASS AS COVARIATE
#run AICprobs script
source("AICprobabilities.R")
source("generic models-binomial glm 2-FF.R")
AICs <- sort(summary$AIC)
models_init <- sort(P, decreasing=TRUE, index.return=TRUE)
top <- length(models_init$x[which(models_init$x>0.05)])

AICs <- AICs[1:top]
models <- models_init$ix[1:top]
probs <- models_init$x[1:top]
rbind(AICs, models, probs) # why does m5 show up if it doesn't exist?
```

m1 <- glm(formula = R ~ A, family = binomial, data = data) *host*
m2 <- glm(formula = R ~ B, family = binomial, data = data) *sym_dist*
m3 <- glm(formula = R ~ A + B, family = binomial, data = data) *host and sym_dist*
m4 <- glm(formula = R ~ A * B, family = binomial, data = data) *host x sym_dist*

```{r}
anova(m1, m2, test="Chisq") # Replacing A with B improves fit 
anova(m1, m3, test="Chisq") # Adding B does not improve fit
anova(m2, m3, test="Chisq") # Adding A does not improve fit
```

```{r}
model_fem <-glm(flew_b~sym_dist, family=binomial, data=data_fem)
summary(model_fem) 
```

* but effect of sym_dist not significant...

```{r}
## Consider covariates
model_fem_final <-glmer(flew_b~sym_dist + (1|population) + (1|trial_type), family=binomial, data=data_fem) # boundary (singular) fit: see ?isSingular
summary(model_fem_final)
## Barely changed effect estimates and does not improve fit
```

```{r}
# FEM DATA: MASS AS COVARIATE 

source("clean_flight_data.R")
data_all <- read_flight_data("complete_flight_data-Winter2020.csv")
data_fem <- data_all[data_all$sex=="F",]

# Recentering
data_fem$lat_c<-data_fem$latitude-mean(data_fem$latitude)
data_fem$sym_dist<-abs(data_fem$latitude-25.49197)
data_fem$mass_c <- data_fem$mass-mean(data_fem$mass, na.rm=TRUE) 

R = data_fem$flew_b
A = data_fem$host_c
B = data_fem$sym_dist
C = data_fem$mass_c


data<-data.frame(R, A, B, C)
head(data) #kable()

# Automatic stepwise functions using AIC to decide between models
model.null <- glm(R~1, family=binomial, data = data_fem)
model.large <-glm(R~A*B*C, family=binomial, data = data_fem)

# Backwards
#step(model.large, direction = "backward")
```


```{r}
#run AICprobs script
source("AICprobabilities.R")
source("generic models-binomial glm 3-FF.R")
AICs <- sort(summary$AIC)
models_init <- sort(P, decreasing=TRUE, index.return=TRUE)
top <- length(models_init$x[which(models_init$x>0.05)])

AICs <- AICs[1:top]
models <- models_init$ix[1:top]
probs <- models_init$x[1:top]
rbind(AICs, models, probs)
```

m3 <- glm(formula = R ~ C, family = binomial, data = data)
m6 <- glm(formula = R ~ B + C, family = binomial, data = data)
m5 <- glm(formula = R ~ A + C, family = binomial, data = data)
m7 <- glm(formula = R ~ A + B + C, family = binomial, data = data)
m9 <- glm(formula = R ~ A * C, family = binomial, data = data)
m10 <- glm(formula = R ~ B * C, family = binomial, data = data)
m12 <- glm(formula = R ~ A * C + B, family = binomial, data = data)

```{r}
anova(m3, m6, test="Chisq") # Adding B does not improve fit
anova(m3, m5, test="Chisq") # Adding A does not improve fit
anova(m3, m7, test="Chisq") # Adding A and B does not improve fit
anova(m3, m9, test="Chisq") # Adding A*C interaction does not improve fit
anova(m3, m10, test="Chisq") # Adding B*C interaction does not improve fit
anova(m3, m12, test="Chisq") # Adding A*C and B does not improve fit
```

```{r}
model_fem_mass <-glm(flew_b~mass_c, family=binomial, data=data_fem)
summary(model_fem_mass) 
```

* large effect of mass, where the heavier the bug is the less likely the bus will fly/disperse.

```{r}
## Consider covariates
model_fem_mass_final <-glmer(flew_b~mass_c + (1|population) + (1|trial_type), family=binomial, data=data_fem) # boundary (singular) fit: see ?isSingular
summary(model_fem_mass_final)
## Did not change effect estimates and does not improve fit
```

## MALE: GLM Binomial Modeling 

```{r}
source("clean_flight_data.R")
data_all <- read_flight_data("complete_flight_data-Winter2020.csv")
data_male <- data_all[data_all$sex=="M",]
```

```{r}
####### No effect of chamber B-4 
summary(glm(flew_b~chamber, data=data_male, family=binomial))

####### Effect of test date (as seen in the data_all)
summary(glm(flew_b~days_from_start_c, data=data_male, family=binomial))

####### No effect of test time (but close p = 0.09443)
summary(glm(flew_b~min_from_IncStart, data=data_male, family=binomial))

# seems like males are very time sensitive
```

```{r}
# MALE DATA WITHOUT MASS AS COVARIATE

# Remove any missing masses
data_male <- data_male %>%
  filter(!is.na(mass))

# Recentering
data_male$lat_c<-data_male$latitude-mean(data_male$latitude)
data_male$sym_dist<-abs(data_male$latitude-25.49197)
data_male$mass_c <- data_male$mass-mean(data_male$mass, na.rm=TRUE) 

R = data_male$flew_b
A = data_male$host_c
B = data_male$sym_dist
#C = data_male$mass_c # as a covariate 
##I would not compete these against each other - because mass is determined in part by A, B, and C, I would not add it until after we've tested the other three.

data<-data.frame(R, A, B)

# Automatic stepwise functions using AIC to decide between models
model.null <- glm(R~1, family=binomial, data=data)
model.large <-glm(R~A*B, family=binomial, data=data)

# Backwards
#m_backwards <- step(model.large, direction = "backward")
```

```{r}
## Forwards
#step(model.null, scope = ~ (A*B), direction = "forward")
```

```{r}
#run AICprobs script
source("AICprobabilities.R")
source("generic models-binomial glm 2-FF.R")
AICs <- sort(summary$AIC)
models_init <- sort(P, decreasing=TRUE, index.return=TRUE)
top <- length(models_init$x[which(models_init$x>0.05)])

AICs <- AICs[1:top]
models <- models_init$ix[1:top]
probs <- models_init$x[1:top]
rbind(AICs, models, probs) # why does m5 show up if it doesn't exist?
```

m1 <- glm(formula = R ~ A, family = binomial, data = data) *host_c*
m3 <- glm(formula = R ~ A + B, family = binomial, data = data)
m4 <- glm(formula = R ~ A * B, family = binomial, data = data)
m2 <- glm(formula = R ~ B, family = binomial, data = data) *sym_dist*

```{r}
anova(m1, m3, test="Chisq") # Adding B does not improve fit 
anova(m1, m4, test="Chisq") # Adding A*B does not improve fit
anova(m3, m4, test="Chisq") # Adding A*B does not improve fit
anova(m1, m2, test="Chisq") # Replacing A with B does improve fit
anova(m2, m3, test="Chisq") 
anova(m2, m4, test="Chisq") 
```

```{r}
model_male<-glm(flew_b~host_c, family=binomial, data=data_male)
summary(model_male)
```

* Strong negative effect if from GRT

```{r}
## Consider covariates
model_male_final <-glmer(flew_b~host_c + (1|population) + (1|trial_type), family=binomial, data=data_male) # no error
summary(model_male_final)
## Q: Changed the effect slightly and improved the model and in turn made host notsignificant...
```


```{r}
# MALE DATA WITH MASS AS COVARIATE

source("clean_flight_data.R")
data_all <- read_flight_data("complete_flight_data-Winter2020.csv")
data_male <- data_all[data_all$sex=="M",]

# Remove any missing masses
data_male <- data_male %>%
  filter(!is.na(mass))

# Recentering
data_male$lat_c<-data_male$latitude-mean(data_male$latitude)
data_male$sym_dist<-abs(data_male$latitude-25.49197)
data_male$mass_c <- data_male$mass-mean(data_male$mass, na.rm=TRUE) 

R = data_male$flew_b
A = data_male$host_c
B = data_male$sym_dist
C = data_male$mass_c # as a covariate 
##I would not compete these against each other - because mass is determined in part by A, B, and C, I would not add it until after we've tested the other three.

data<-data.frame(R, A, B, C)

# Automatic stepwise functions using AIC to decide between models
model.null <- glm(R~1, family=binomial, data=data)
model.large <-glm(R~A*B*C, family=binomial, data=data)

# Backwards
#m_backwards <- step(model.large, direction = "backward")
```

```{r}
#run AICprobs script
source("AICprobabilities.R")
source("generic models-binomial glm 3-FF.R")
AICs <- sort(summary$AIC)
models_init <- sort(P, decreasing=TRUE, index.return=TRUE)
top <- length(models_init$x[which(models_init$x>0.05)])

AICs <- AICs[1:top]
models <- models_init$ix[1:top]
probs <- models_init$x[1:top]
rbind(AICs, models, probs)
```

m5 <- glm(formula = R ~ A + C, family = binomial, data = data) *host + mass*
m7 <- glm(formula = R ~ A + B + C, family = binomial, data = data)
m13 <- glm(formula = R ~ B * C + A, family = binomial, data = data)
m9 <- glm(formula = R ~ A * C, family = binomial, data = data)
m16 <- glm(formula = R ~ A * C + B * C, family = binomial, data = data)
m11 <- glm(formula = R ~ A * B + C, family = binomial, data = data)
m12 <- glm(formula = R ~ A * C + B, family = binomial, data = data)
m15 <- glm(formula = R ~ A * B + B * C, family = binomial, data = data)

```{r}
anova(m5, m7, test="Chisq") # Adding B does not improve fit 
anova(m5, m13, test="Chisq") # Adding B*C does not improve fit
anova(m5, m9, test="Chisq") # Adding A*C does not improve fit
anova(m5, m16, test="Chisq") # Adding A*c and B*C does not improve fit
anova(m5, m11, test="Chisq") # Adding A*B does not improve fit
anova(m5, m12, test="Chisq") # Adding A*C and B does not improve fit
anova(m5, m15, test="Chisq") # Adding A*B and B*C does not irmpve fit
```

```{r}
model_male_mass <-glm(flew_b~host_c + mass_c, family=binomial, data=data_male)
summary(model_male_mass)
```

* Strong negative effect if from GRT
* Strong negative effect of mass, that if weigh more then less likely to disperse

```{r}
## Consider covariates
model_male_mass_final <-glmer(flew_b~host_c + mass_c+ (1|population) + (1|trial_type), family=binomial, data=data_male) # no error
summary(model_male_mass_final)
## Changed the effects slightly and in turn made host notsignificant...
```


**SUMMARY: Other factors I have not addressed - egg laying which is specific to females, males have an effect of test date, and mass effects for both males and females.**


# ----------------------------------------------------------------------------------

# ADDING NEW COVARIATES TO SEX SPLIT

Let's add these other measurements as covariates...

```{r}
source("clean_flight_data.R")
data_all <- read_flight_data("complete_flight_data-Winter2020.csv")
data_fem <- data_all[data_all$sex=="F",]
data_male <- data_all[data_all$sex=="M",]

# Recentering --> can create a script for this --> recentering on groups

data_male$lat_c<-data_male$latitude - mean(data_male$latitude)
data_male$sym_dist<-abs(data_male$latitude-25.49197)
data_male$mass_c <- data_male$mass - mean(data_male$mass, na.rm=TRUE) 
data_male$beak_c <- data_male$beak-mean(data_male$beak, na.rm=TRUE)
data_male$thorax_c <- data_male$thorax-mean(data_male$thorax, na.rm=TRUE)
data_male$body_c <- data_male$body-mean(data_male$body, na.rm=TRUE) 
data_male$wing_c <- data_male$wing-mean(data_male$wing, na.rm=TRUE) 

data_fem$lat_c<-data_fem$latitude - mean(data_fem$latitude)
data_fem$sym_dist<-abs(data_fem$latitude-25.49197)
data_fem$mass_c <- data_fem$mass - mean(data_fem$mass, na.rm=TRUE) 
data_fem$beak_c <- data_fem$beak-mean(data_fem$beak, na.rm=TRUE)
data_fem$thorax_c <- data_fem$thorax-mean(data_fem$thorax, na.rm=TRUE)
data_fem$body_c <- data_fem$body-mean(data_fem$body, na.rm=TRUE) 
data_fem$wing_c <- data_fem$wing-mean(data_fem$wing, na.rm=TRUE) 

```

```{r}
# Biological Effects:

####### MASS 
summary(glm(flew_b~mass_c, data=data_fem, family=binomial)) # F | (Strong) Effect of mass 
summary(glm(flew_b~mass_c, data=data_male, family=binomial)) # M | (Strong) Effect of mass 

####### EGGS
summary(glm(flew_b~total_eggs_c, data=data_fem, family=binomial)) # F | Effect of number of eggs laid
summary(glm(flew_b~eggs_b, data=data_fem, family=binomial))  #F | Effect of whether eggs were laid or not
```

```{r}
# Morphology Effects:

####### BEAK LENGTH
summary(glm(flew_b~beak_c, data=data_fem, family=binomial)) # F | Effect of beak length
summary(glm(flew_b~beak_c, data=data_male, family=binomial)) # M | No effect of beak length

####### THORAX LENGTH
summary(glm(flew_b~thorax_c, data=data_fem, family=binomial)) # F | No effect of thorax length
summary(glm(flew_b~thorax_c, data=data_male, family=binomial)) #  M | No effect of thorax length

####### BODY LENGTH
summary(glm(flew_b~body_c, data=data_fem, family=binomial)) # F |Effect of body length
summary(glm(flew_b~body_c, data=data_male, family=binomial)) # N |No effect of body length

####### WING LENGTH
summary(glm(flew_b~wing_c, data=data_fem, family=binomial)) # F | Effect of wing length
summary(glm(flew_b~wing_c, data=data_male, family=binomial)) # M | Effect of wing length

####### No effect of wing morph (check how annotated the wing morph) - don't include it
#summary(glm(flew_b~w_morph_c, data=data_all, family=binomial)) # but close p val = 0.0512

```

# Female Data | Remodeling with new covariates 

```{r}
# Remove any missing masses
data_fem <- data_fem %>%
  filter(!is.na(mass))

R = data_fem$flew_b
A = data_fem$host_c
B = data_fem$sym_dist
C = data_fem$mass_c
D = data_fem$eggs_b
  
data<-data.frame(R, A, B, C, D)
head(data) #kable()

# Automatic stepwise functions using AIC to decide between models
model.null <- glm(R~1, family=binomial, data = data)
model.large <-glm(R~A*B*C*D, family=binomial, data = data)

# Backwards
#step(model.large, direction = "backward")
```

For backwards:
best_fit <- glm(formula = R ~ A + B + C + D + A:C + B:C + B:D, family = binomial, 
    data = data_male)
AIC: 222.7

```{r}
## Forwards
#step(model.null, scope = ~ (A*B*C*D), direction = "forward")
```

For forwards:

best_fit <- glm(formula = R ~ D, family = binomial, data = data)
AIC: 222.4

```{r}
#run AICprobs script
source("AICprobabilities.R")
source("generic models-binomial glm 4-FF.R")
AICs <- sort(summary$AIC)
models_init <- sort(P, decreasing=TRUE, index.return=TRUE)
top <- length(models_init$x[which(models_init$x>0.05)])

AICs <- AICs[1:top]
models <- models_init$ix[1:top]
probs <- models_init$x[1:top]
rbind(AICs, models, probs)
```

m4 <-  glm(formula = R ~ D, family = binomial, data = data) *eggs_b*
m20 <- glm(formula = R ~ B * D, family = binomial, data = data) *sym_dist x eggs_b*
m9 <- glm(formula = R ~ B + D, family = binomial, data = data) *sym_dist and eggs_b*

```{r}
anova(m4, m20, test="Chisq") # Replacing B with B*D does not improve fit
anova(m4, m9, test="Chisq") # Adding B does not improve fit
anova(m4, m10, test="Chisq") # Adding C does not improve fit
```

```{r}
fem_model<-glm(flew_b~eggs_b, family=binomial, data=data_fem)
summary(fem_model)
```

* Strong negative effect if laid eggs that day

Looks like eggs are overpowering mass in this modeling example

```{r}
ffem_model<-glmer(flew_b~eggs_b + (1|population) + (1|trial_type), family=binomial, data=data_male) # no error
summary(ffem_model) # still a strong negative effect if laid eggs that day
```

Host seems to not be effecing female flight...let's remove it and replace it with total_eggs laid

```{r}
source("clean_flight_data.R")
data_all <- read_flight_data("complete_flight_data-Winter2020.csv")
data_fem <- data_all[data_all$sex=="F",]

# Remove any missing masses
data_fem <- data_fem %>%
  filter(!is.na(mass))

R = data_fem$flew_b
A = data_fem$total_eggs_c
B = data_fem$sym_dist
C = data_fem$mass_c
D = data_fem$eggs_b
  
data<-data.frame(R, A, B, C, D)
head(data) #kable()

# Automatic stepwise functions using AIC to decide between models
model.null <- glm(R~1, family=binomial, data = data)
model.large <-glm(R~A*B*C*D, family=binomial, data = data)

# Backwards
#step(model.large, direction = "backward")
```

```{r}
## Forwards
#step(model.null, scope = ~ (A*B*C*D), direction = "forward")
```

```{r}
#run AICprobs script
source("AICprobabilities.R")
source("generic models-binomial glm 4-FF.R")
AICs <- sort(summary$AIC)
models_init <- sort(P, decreasing=TRUE, index.return=TRUE)
top <- length(models_init$x[which(models_init$x>0.05)])

AICs <- AICs[1:top]
models <- models_init$ix[1:top]
probs <- models_init$x[1:top]
rbind(AICs, models, probs)

# error? "models are not all fitted to the same number of observations"
```

m43 <-  glm(formula = R ~ A * B + B * D, family = binomial, data = data) *total_eggs_c, sym_dist, eggs_b*
m72 <- glm(formula = R ~ A * B + A * D + B * D, family = binomial, data = data) *total_eggs_c, sym_dist, eggs_b*
m50 <- glm(formula = R ~ A * D + B * D, family = binomial, data = data) *total_eggs_c, sym_dist, eggs_b*

```{r}
anova(m43, m72, test="Chisq") # Replacing B*D with A*D does not improve fit
anova(m43, m50, test="Chisq") # Replacing A*B with A*D does improve fit
anova(m50, m72, test="Chisq") # Adding A*B does not improve fit
m50
```

```{r}
fem_model<-glm(flew_b~total_eggs_c*eggs_b + sym_dist*eggs_b, family=binomial, data=data_fem) 
summary(fem_model)
```

* strong negative effect from eggs_b alone, no other significant effects 

```{r}
ffem_model<-glmer(flew_b~total_eggs_c*eggs_b + sym_dist*eggs_b + (1|population) + (1|trial_type), family=binomial, data=data_fem) 
summary(ffem_model) 
# did not change effects or improve the fit
```


# Male Data | Remodeling with new covariates 

```{r}
source("clean_flight_data.R")
data_all <- read_flight_data("complete_flight_data-Winter2020.csv")
data_male <- data_all[data_all$sex=="M",]

# Remove any missing masses
data_male <- data_male %>%
  filter(!is.na(mass))

# Recentering --> can create a script for this --> recentering on groups

data_male$lat_c<-data_male$latitude - mean(data_male$latitude)
data_male$sym_dist<-abs(data_male$latitude-25.49197)
data_male$mass_c <- data_male$mass - mean(data_male$mass, na.rm=TRUE) 
data_male$beak_c <- data_male$beak-mean(data_male$beak, na.rm=TRUE)
data_male$thorax_c <- data_male$thorax-mean(data_male$thorax, na.rm=TRUE)
data_male$body_c <- data_male$body-mean(data_male$body, na.rm=TRUE) 
data_male$wing_c <- data_male$wing-mean(data_male$wing, na.rm=TRUE) 

# recap of covariates with significant effects
summary(glm(flew_b~days_from_start_c, data=data_male, family=binomial)) # days_from_start_c
summary(glm(flew_b~mass_c, data=data_male, family=binomial)) # mass_c
```

```{r}
R = data_male$flew_b
A = data_male$host_c
B = data_male$sym_dist
C = data_male$mass_c
D = data_male$days_from_start_c 

data<-data.frame(R, A, B, C, D)

# Automatic stepwise functions using AIC to decide between models
model.null <- glm(R~1, family=binomial, data=data)
model.large <-glm(R~A*B*C*D, family=binomial, data=data)

# Backwards
#m_backwards <- step(model.large, direction = "backward")
```

```{r}
## Forwards
#step(model.null, scope = ~ (A*B), direction = "forward")
```

```{r}
#run AICprobs script
source("AICprobabilities.R")
source("generic models-binomial glm 4-FF.R")
AICs <- sort(summary$AIC)
models_init <- sort(P, decreasing=TRUE, index.return=TRUE)
top <- length(models_init$x[which(models_init$x>0.05)])

AICs <- AICs[1:top]
models <- models_init$ix[1:top]
probs <- models_init$x[1:top]
rbind(AICs, models, probs)
```

m13 <-  glm(formula = R ~ A + C + D, family = binomial, data = data) *host, mass, days_from_start*
m27 <- glm(formula = R ~ A * D + C, family = binomial, data = data)*host, sym_dist, days_from_start, mass*
m15 <- glm(formula = R ~ A + B + C + D, family = binomial, data = data) *host, sym_dist, days_from_start, mass*

```{r}
anova(m13, m27, test="Chisq") # Adding A*D does not improve fit
anova(m13, m15, test="Chisq") # Adding B effect does not improve fit
m13
```
R = data_male$flew_b
A = data_male$host_c
B = data_male$sym_dist
C = data_male$mass_c
D = data_male$days_from_start_c 

```{r}
male_model1<-glm(flew_b~host_c + mass_c + days_from_start_c, family=binomial, data=data_male) 
summary(male_model1) 
```

* strong negative effect from mass;
* strong negative effect of host_c;
* strong negative effect of days_from_start_c;

```{r}
male_model2<-glmer(flew_b~host_c + mass_c + days_from_start_c + (1|population) + (1|trial_type), family=binomial,
                   data=data_male) 
summary(male_model2) 
# did not improve fit but did change the effect significance
```
* strong negative effect from mass;
* strong negative effect for days_from_start;
* but now marginal effect from host

* less likely to fly if you are from GRT, have more mass, and fly later (older bug..?)


