---
title: 'Modeling Flight Speed'
author: "Anastasia Bernat"
date: "5/4/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
rm(list=ls())
setwd("~/Desktop/Rstats-winter2020/")

library(lme4)
library(rethinking)

library(dplyr)
library(tidyselect)
library(stringr)

library(glmnet)
library(ggplotify)
library(gridExtra)
library(ggformula)
library(randomcoloR)

knitr::opts_chunk$set(echo = TRUE)
```

## Winter 2020 Flight Trials: Speed Flight Modeling {.tabset}

Flight Trials Winter 2020 Dataset was conducted from 2/17/2020 - 3/10/2020. Soapberry bugs were flight tested twice for multiple hours in the flight mill and observed from 8 AM to (5-8 PM) each day. Used multivariate (glm) and mixed effect modeling (glmer) to analyze the flight results.

### All Flyers

[Cleaning Data](#clean)

[Testing Covariates](#test)

[Without Mass](#wo.mass)

[With Mass](#w.mass)

[Morphology](#morph)

<a id="clean"></a>

**Cleaning the Data**

```{r}
rm(list=ls())
output_col = FALSE # Recommend changing this to TRUE if working in Base R or RStudio, and FALSE if generating an html
source("src/clean_flight_data.R") # Script that loads and cleans up the data
source("src/regression_output.R") # A script that cleans up regression outputs and prints in color or black and white
source("src/center_flight_data.R")
source("get_warnings.R")

data <- read_flight_data("data/all_flight_data-Winter2020.csv")
data_all <- data[[1]]
data_tested <- data[[2]]

### Remove everyone who didn't fly (then remove distances = 0, if that didn't work fully)
data_flew_all <- data_tested[data_tested$flew_b == 1, ] 
```

```{r}
### Check for low speeds
low_speeds <- data_flew_all %>%
  filter(average_speed <0.05)

### Check for high speeds
high_speeds <- data_flew_all %>%
  filter(average_speed >0.65)

low_speeds$flight_type# have 7 bugs with average_speed = 0 but were marked as bursters (this could be just something very short (second burst) - not enough to grant a calculation) - I decided to remove them. But one bug was continuous and had 0 distance and 0 speeds - that was bug 196 T2 set011-3-03-2020-A3_196.txt
high_speeds$flight_type # 3 bugs - also bursters. Could also be short explosive bursts but not true to the biology of these bugs (more like us blowing on them).
```

```{r}
### Remove outliers
data_flew <- data_flew_all %>%
  filter(average_speed > 0.05) %>%
  filter(average_speed < 0.65)
  
data_flew <- center_data(data_flew)
```

<a id="test"></a>

**Testing Models and Covariates**

```{r}
test_model<-lmer(average_speed~host_c*sex_c*sym_dist + (1|ID), data=data_flew) #it converges - Run one set of models with ID as a random factor.
getME(test_model, "lower")
```

```{r echo=FALSE}
#######testing some covariates:
data_flew$chamber <- relevel(data_flew$chamber, ref="A-2")

####### Effect of chamber B-2 and B-4
tidy_regression(lmer(average_speed~chamber + (1|ID), data=data_flew), is_color=output_col) ###Possibly reductions in speed 

####### No effect of test date
tidy_regression(lmer(average_speed~days_from_start + (1|chamber), data=data_flew), is_color=output_col)

####### No effect of test time
tidy_regression(lmer(average_speed~min_from_IncStart + (1|trial_type), data=data_flew), is_color=output_col)

```

<a id="wo.mass"></a>

**Without Mass**

```{r message=FALSE}
data<-data.frame(R=data_flew$average_speed,
                 A=data_flew$host_c, 
                 B=data_flew$sex_c, 
                 C=data_flew$sym_dist, 
                 X=data_flew$ID)

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian glmer 1-RF + 3-FF REMLF.R")
```

```{r}
anova(m0, m2, test="Chisq") # Adding B improves fit
anova(m2, m4, test="Chisq") # Adding Adoes not improve fit
anova(m2, m6, test="Chisq") # Adding C does not improve fit
```

```{r}
speed_model <- lmer(average_speed ~ sex_c + (1 | ID), data=data_flew, REML=FALSE)
tidy_regression(speed_model, is_color=output_col)
```

* marginal effect of sex

```{r}
s.test <- paste("pval: ", shapiro.test(residuals(speed_model))$p.value)
qqnorm(resid(speed_model))
qqline(resid(speed_model))
text(-2, 0.1, s.test)
```

<a id="w.mass"></a>

**With Mass**

```{r}
data_flew <- data_flew %>%
  filter(!is.na(mass))
data_flew <- center_data(data_flew)

data<-data.frame(R=data_flew$average_speed,
                 A=data_flew$host_c, 
                 B=data_flew$sex_c, 
                 C=data_flew$sym_dist, 
                 D=data_flew$mass_c,
                 X=data_flew$chamber) 

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian glmer 1-RF + 4-FF REMLF.R")
```


```{r}
anova(m0, m4, test="Chisq") # Adding D does improve fit
anova(m0, m2, test="Chisq") # Adding B does improve fit
```

```{r}
speed_model <- lmer(average_speed ~ sex_c + (1 | chamber), data=data_flew, REML=FALSE)
tidy_regression(speed_model, is_color=output_col)
```

* positive effect of sex, so that if F then faster 

```{r}
s.test <- paste("pval: ", shapiro.test(residuals(speed_model))$p.value)
qqnorm(resid(speed_model))
qqline(resid(speed_model))
text(-2, 0.1, s.test)
```

<a id="morph"></a>

**Morphology**

```{r message=FALSE}
data_flew <- data_flew %>%
  filter(!is.na(mass)) %>%
  filter(!is.na(body))
data_flew <- center_data(data_flew)

data<-data.frame(R=data_flew$average_speed,
                 A=data_flew$thorax_c, 
                 B=data_flew$body_c, 
                 C=data_flew$wing_c,
                 X=data_flew$ID) 

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian glmer 1-RF + 3-FF REMLF.R")
```

```{r}
anova(m0, m1, test="Chisq") # Adding A does improve fit
anova(m0, m2, test="Chisq") # Adding B does improve fit
anova(m0, m3, test="Chisq") # Adding C does improve fit
anova(m1, m5, test="Chisq") # Addihng C does not improve fit
anova(m1, m4, test="Chisq") # Adding B does not improve fit
```

```{r}
speed_morph_model <- lmer(average_speed ~ thorax_c + (1|ID), data=data_flew, REML=FALSE)
tidy_regression(speed_morph_model, is_color=output_col)
```

* positive effect of thorax, where the longer the thorax, the faster the speed

```{r}
s.test <- paste("pval: ", shapiro.test(residuals(speed_morph_model))$p.value)
qqnorm(resid(speed_morph_model))
qqline(resid(speed_morph_model))
text(-2, 0.1, s.test)
```

### All Flyers Plots

```{r}
gf_histogram(~average_speed,data=data_flew_all, col=~flight_type) # before filtering outliers
gf_histogram(~average_speed,data=data_flew, col=~flight_type) # after filtering outliers
```

```{r}
source("src/plotting-lm.R")
source("src/plotting-lm2.R")
```

```{r echo=FALSE, warning=FALSE}
#, fig.width=2.8, fig.height=2.3}

lm_plot(data_flew, "sym_dist", "average_speed", "Distance From Sympatric Zone", "Average Speed (m/s)", "Average Speed by Distance From Sympatric Zone (°)")

lm_plot2(data_flew, "sex_c", "average_speed", "Sex", "Average Speed (m/s)", "Average Speed by Sex")
axis(1, at = seq(-1,1,.5),labels=c("M", " ", " ", " ", "F"))

lm_plot2(data_flew, "host_c", "average_speed", "Host Plant", "Average Speed (m/s)", "Average Speed by Host Plant")
axis(1, at = seq(-1,1,.5),labels=c("BV", " ", " ", " ", "GRT"))

data_flew <- data_flew %>%
  filter(!is.na(mass_c))
lm_plot(data_flew, "mass_c", "average_speed", "Mass",  "Average Speed (m/s)", "Average Speed by Mass")
```

```{r echo=FALSE, warning=FALSE}

#, fig.width=2.8, fig.height=2.3}

lm_plot(data_flew, "thorax_c", "average_speed", "Thorax Length (mm)",  "Average Speed (m/s)", "Average Speed by Thorax Length") # best model
lm_plot(data_flew, "body_c", "average_speed", "Body Length (mm)", "Average Speed (m/s)", "Average Speed by Body Length")
lm_plot(data_flew, "wing_c", "average_speed", "Wing Length (mm)", "Average Speed (m/s)", "Average Speed by Wing Length")
lm_plot(data_flew, "beak_c", "average_speed", "Beak Length (mm)", "Average Speed (m/s)", "Average Speed by Beak Length")
```

### Delta Speed

```{r}
rm(list=ls())
output_col = FALSE # Recommend changing this to TRUE if working in Base R or RStudio, and FALSE if generating an html
source("src/clean_flight_data.R") # Script that loads and cleans up the data
source("src/regression_output.R") # A script that cleans up regression outputs and prints in color or black and white
source("src/center_flight_data.R")
source("get_warnings.R")

data <- read_flight_data("data/all_flight_data-Winter2020.csv")
data_all <- data[[1]]
data_tested <- data[[2]]

### Remove everyone who didn't fly (then remove distances = 0, if that didn't work fully)
data_flew_all <- data_tested[data_tested$flew_b == 1, ] 

### Remove outliers
data_flew <- data_flew_all %>%
  filter(average_speed > 0.05) %>%
  filter(average_speed < 0.65)
  
data_flew <- center_data(data_flew)
```

```{r}
d <- data_flew %>%
   group_by(ID, sex,population, site, host_plant, latitude, longitude, total_eggs, 
            beak, thorax, wing, body, w_morph, morph_notes, tested,
            host_c, sex_c, w_morph_c, lat_c, sym_dist, sym_dist_s, total_eggs_c, 
            beak_c, thorax_c, thorax_s, body_c, wing_c, wing2body, wing2body_c, wing2body_s) %>%
   summarise_all(funs(list(na.omit(.))))

d$num_flew <- 0
d$num_notflew <- 0
d$average_mass <- 0

d$mass_diff <- 0
d$flew_diff <- 0
d$dist_diff <- 0
d$speed_diff <- 0

for(row in 1:length(d$flew_b)){
  
  n_flew <- sum(d$flew_b[[row]] == 1) # total number of times did fly among trails
  d$num_flew[[row]] <- n_flew 
  
  n_notflew <- sum(d$flew_b[[row]] == 0) # total number of times did not fly among trails
  d$num_notflew[[row]] <- n_notflew
  
  avg_mass <- mean(d$mass[[row]])
  d$average_mass[[row]] <- avg_mass
  
  # mass, flight, distance, and speed changes between T1 and T2
  
  d$mass_diff[[row]] <- d$mass[[row]][2] - d$mass[[row]][1]  # T2 - T1
  d$flew_diff[[row]] <- d$flew_b[[row]][2] - d$flew_b[[row]][1]  # T2 - T1
  d$dist_diff[[row]] <- d$distance[[row]][2] - d$distance[[row]][1]  # T2 - T1
  d$speed_diff[[row]] <- d$average_speed[[row]][2] - d$average_speed[[row]][1]  # T2 - T1
  d$egg_diff[[row]] <- d$eggs_b[[row]][2] - d$eggs_b[[row]][1]  # T2 - T1

}

d <- select(d, -filename, -channel_letter, -set_number)

d # NA's generated are good because that means it's accounted only for bugs that flew in both trials

# Filter out bugs that ONLY flew in T1:
rows_remove <- c()
for (row in 1:nrow(d)){
  if (length(d$trial_type[[row]]) < 2) {
    rows_remove <- c(rows_remove, row)
  }
}
d <- d[-rows_remove, ]
d
```

#### Speed Diff ~ Mass Diff

```{r}
## Test
tidy_regression(lm(speed_diff ~ mass_diff, data=d), is_color = output_col) # no effect of mass_diff AFTER you filter out the outliers. If don't filter out the outliers then this relationship is significant and that data is residuals are not normal
```

```{r}
hist(d$speed_diff) # much more normally distributed data and later will see the residuals pass the Shapiro-Wilk test
```

**No mass_diff**

```{r message=FALSE}
R = d$speed_diff
A = d$host_c
B = d$sex_c
C = d$sym_dist
D = d$mass_diff 
X = d$population 

data<-data.frame(R, A, B, C, D, X)
source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian glmer 1-RF + 3-FF REMLF.R")
```

```{r}
anova(m7, m13, test="Chisq") # Adding B*C does not improve fit
```

```{r}
speed_model <- lmer(speed_diff ~ host_c + sex_c + sym_dist  + (1 | population), data=d, REML=FALSE)
tidy_regression(speed_model, is_color=output_col)
```

* (marginal) negative effect of sex where if F then more likely to fly slower in T2 than T1 
* (marginal) negative effect of host where if from GRT then more likely to fly slower in T2 than T1
* negative effect of sym_dist where if farther from Homestead then more likely to fly slower in T2 than T1

```{r}
s.test <- paste("pval: ", shapiro.test(residuals(speed_model))$p.value)
qqnorm(resid(speed_model))
qqline(resid(speed_model))
text(-1, 0.1, s.test)
```

**Mass_diff**

```{r message=FALSE}
d <- d %>%
  filter(!is.na(mass_diff))

R = d$speed_diff
A = d$host_c
B = d$sex_c
C = d$sym_dist
D = d$mass_diff 
X = d$population 

data<-data.frame(R, A, B, C, D, X)

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian glmer 1-RF + 4-FF REMLF.R")
```

```{r}
anova(m11, m28, test="Chisq") # Adding B*C does not improve fit
```

```{r}
speed_model <- lmer(speed_diff ~ host_c + sex_c + sym_dist + (1 | population), data=d, REML=FALSE)
tidy_regression(speed_model, is_color=output_col)
```

Same top model as before. All of these variables have weak effects.

So what if it was only females, since they had the greatest mass changes?

### Delta Speed - Females Only

```{r}
d_fem <- d[d$sex == "F",] # very low sample size 
```

**No mass_diff**

```{r message=FALSE}
R = d_fem$speed_diff
A = d_fem$host_c
B = d_fem$sym_dist
C = d_fem$mass_diff 
X = d_fem$population 

data<-data.frame(R, A, B, C, X)
source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian glmer 1-RF + 2-FF REMLF.R")
```

```{r}
anova(m0, m1, test="Chisq") # Adding A does not improve fit
anova(m0, m2, test="Chisq") # Adding B does not improve fit
```

```{r}
speed_model <- lmer(speed_diff ~ 1 + (1 | population), data=d_fem, REML=FALSE)
tidy_regression(speed_model, is_color=output_col)
```

**mass_diff**

```{r message=FALSE}
source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian glmer 1-RF + 3-FF REMLF.R")
```

```{r}
anova(m0, m1, test="Chisq") # strange. get error "refitting model(s) with ML (instead of REML)" even though these have been refit to REML=FALSE...
anova(m0, m3, test="Chisq") # Adding C does not improve fit
anova(m0, m2, test="Chisq") # Adding B does not improve fit
```

```{r}
speed_model <- lmer(speed_diff ~ 1 + (1 | population), data=d_fem, REML=FALSE)
tidy_regression(speed_model, is_color=output_col)
```

```{r}
s.test <- paste("pval: ", shapiro.test(residuals(speed_model))$p.value)
qqnorm(resid(speed_model))
qqline(resid(speed_model))
text(-1, 0.1, s.test)
```


### Continuous Flyers

[Cleaning Data](#cleandata)

[Testing Covariates](#testing)

[Without Mass](#wo.massc)

[With Mass](#w.massc)

[Morphology](#morphc)

<a id="cleandata"></a>

**Cleaning the data**

```{r}
rm(list=ls())
output_col = FALSE 
source("src/clean_flight_data.R") 
source("src/regression_output.R") 
source("src/center_flight_data.R")
source("get_warnings.R")

data <- read_flight_data("data/all_flight_data-Winter2020.csv")
data_all <- data[[1]]
data_tested <- data[[2]]

### Remove everyone who didn't fly (then remove distances = 0, if that didn't work fully)
data_flew <- data_tested[data_tested$flew_b == 1, ]
data_flew <- data_flew[data_flew$average_speed > 0, ]
data_flew <- center_data(data_flew)

### Break up by flight type
dC<-data_flew[data_flew$flight_type=="C" | data_flew$flight_type=="BC" | data_flew$flight_type=="CB",] # this includes BC or CB flyers 
dC <- center_data(dC)
```

<a id="testing"></a>

**Testing Models and Covariates**

```{r}
test_model<-lmer(average_speed~host_c*sex_c*sym_dist + (1|ID), data=dC) #it converges - Run one set of models with ID as a random factor.
getME(test_model, "lower")
```

```{r echo=FALSE}
#######testing some covariates:
dC$chamber <- relevel(dC$chamber, ref="A-2")

####### Effect of chamber B-4, B-3, B-2
tidy_regression(lmer(average_speed~chamber + (1|ID), data=dC), is_color=output_col) ###Possibly reductions in speed 

####### No effect of test date
tidy_regression(lmer(average_speed~days_from_start + (1|ID), data=dC), is_color=output_col)

####### No effect of test time
tidy_regression(lmer(average_speed~min_from_IncStart + (1|ID), data=dC), is_color=output_col)
```

<a id="wo.massc"></a>

**Without Mass**

```{r}
data<-data.frame(R=dC$average_speed,
                 A=dC$host_c, 
                 B=dC$sex_c, 
                 C=dC$sym_dist, 
                 X=dC$ID)

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian glmer 1-RF + 3-FF REMLF.R")
```

```{r}
anova(m2, m4, test="Chisq") # Adding A does not improve fit
anova(m2, m6, test="Chisq") # Adding B does not improve fit
```

```{r}
continuous_model <- lmer(average_speed ~ sex_c + (1|ID), data=dC, REML=FALSE)
tidy_regression(continuous_model, is_color=output_col)

continuous_model <- lmer(average_speed ~ sex_c + (1|chamber) + (1|ID), data=dC, REML=FALSE)
tidy_regression(continuous_model, is_color=output_col) # Adding chamber improves fit
```

* positive effect of sex_c, where if a female continuous-flyer bug then more likely to fly faster on average 

```{r}
s.test <- paste("pval: ", shapiro.test(residuals(continuous_model))$p.value)
qqnorm(resid(continuous_model))
qqline(resid(continuous_model))
text(-1, 0.1, s.test)
```

<a id="w.massc"></a>

**With Mass**

```{r}
data<-data.frame(R=dC$average_speed,
                 A=dC$host_c, 
                 B=dC$sex_c, 
                 C=dC$sym_dist,
                 D=dC$mass_c,
                 X=dC$ID)

source("src/compare_models.R")
errors <- withWarnings(model_comparisonsAIC("src/generic models-gaussian glmer 1-RF + 4-FF REMLF.R"))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m20, m30, test="Chisq") # Adding A does not improve fit
anova(m20, m31, test="Chisq") # Adding C does not improve fit
```

```{r}
continuous_model <- lmer(average_speed ~ sex_c*mass_c + (1|ID), data=dC)
tidy_regression(continuous_model, is_color=output_col)

continuous_model <- lmer(average_speed ~ sex_c*mass_c + (1|chamber) + (1|ID), data=dC)
tidy_regression(continuous_model, is_color=output_col) # Adding chamber improves model
```

* no effect of sex_c
* no effect of mass_c
* negative effect of sex_c*mass_c, where if female and heavier, then slower avg speed (make sense because probably due to egg-laying)

```{r}
s.test <- paste("pval: ", shapiro.test(residuals(continuous_model))$p.value)
qqnorm(resid(continuous_model))
qqline(resid(continuous_model))
text(-1, 0.1, s.test)
```


<a id="morphc"></a>

**Morphology**

```{r}
dC<-data_flew[data_flew$flight_type=="C",] 
dC <- dC %>%
  filter(!is.na(body))
dC <- center_data(dC)

data<-data.frame(R=dC$average_speed,
                 A=dC$thorax_c, 
                 B=dC$body_c, 
                 C=dC$wing_c,
                 X=dC$ID)

source("src/compare_models.R")
errors <- withWarnings(model_comparisonsAIC("src/generic models-gaussian glmer 1-RF + 3-FF REMLF.R"))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m10, m13, test="Chisq") # Adding A does not improve fit
```

```{r}
continuous_morph_model <- lmer(average_speed ~ body_c*wing_c + (1|ID), data=dC)
tidy_regression(continuous_morph_model, is_color=output_col)

continuous_morph_model <- lmer(average_speed ~ body_c*wing_c + (1|ID) + (1|chamber), data=dC)
tidy_regression(continuous_morph_model, is_color=output_col) # improves model
```

* positive effect of body length, where the longer the body the faster the avg speed
* negative effect of wing length, where the longer the wing the slower the avg speed
* neg eggect of body* wing, where the longer the wing *body  the slower the avg speed

```{r}
s.test <- paste("pval: ", shapiro.test(residuals(continuous_morph_model))$p.value)
qqnorm(resid(continuous_morph_model))
qqline(resid(continuous_morph_model))
text(-1, 0.1, s.test)
```

### Continuous Flyers Plots

```{r}
gf_histogram(~average_speed,data=dC)
```

```{r}
source("src/plotting-lm.R")
source("src/plotting-lm2.R")
```

```{r echo=FALSE, warning=FALSE}

#, fig.width=2.8, fig.height=2.3}

lm_plot(dC, "sym_dist", "average_speed", "Distance From Sympatric Zone", "Average Speed (m/s)", "Average Speed by Distance From Sympatric Zone")

lm_plot2(dC, "sex_c", "average_speed", "Sex", "Average Speed (m/s)", "Average Speed by Sex")
axis(1, at = seq(-1,1,.5),labels=c("M", " ", " ", " ", "F"))

lm_plot2(dC, "host_c", "average_speed", "Host Plant", "Average Speed (m/s)", "Average Speed by Host Plant")
axis(1, at = seq(-1,1,.5),labels=c("BV", " ", " ", " ", "GRT"))

dC <- dC %>%
  filter(!is.na(mass_c))
lm_plot(dC, "mass_c", "average_speed", "Mass", "Average Speed (m/s)", "Average Speed by Mass")
```

```{r echo=FALSE, warning=FALSE}

#, fig.width=2.8, fig.height=2.3}

lm_plot(dC, "thorax_c", "average_speed", "Thorax Length (mm)",  "Average Speed (m/s)", "Average Speed by Thorax Length") # best model
lm_plot(dC, "body_c", "average_speed", "Body Length (mm)", "Average Speed (m/s)", "Average Speed by Body Length")
lm_plot(dC, "wing_c", "average_speed", "Wing Length (mm)", "Average Speed (m/s)", "Average Speed by Wing Length")
lm_plot(dC, "beak_c", "average_speed", "Beak Length (mm)", "Average Speed (m/s)", "Average Speed by Beak Length")
```

### Bursters

[Cleaning Data](#cleanthedata)

[Testing Covariates](#testingcov)

[Without Mass](#wo.massb)

[With Mass](#w.massb)

[Morphology](#morphb)

<a id="cleanthedata"></a>

**Cleaning the data**

```{r}
rm(list=ls())
output_col = FALSE 
source("src/clean_flight_data.R") 
source("src/regression_output.R") 
source("src/center_flight_data.R")
source("get_warnings.R")

data <- read_flight_data("data/all_flight_data-Winter2020.csv")
data_all <- data[[1]]
data_tested <- data[[2]]

### Remove everyone who didn't fly (then remove distances = 0, if that didn't work fully)
data_flew <- data_tested[data_tested$flew_b == 1, ]
data_flew <- data_flew %>%
  filter(average_speed > 0) %>%
  filter(average_speed < 0.65)
data_flew <- center_data(data_flew)

### Break up by flight type
dB<-data_flew[data_flew$flight_type=="B",] # this exclusdes BC or CB flyers 
dB <- center_data(dB)
```

<a id="testingcov"></a>

**Testing Models and Covariates**

```{r}
test_model<-lmer(average_speed~host_c*sex_c*sym_dist + (1|ID), data=dB) #it converges - Run one set of models with ID as a random factor.
getME(test_model, "lower")
```

```{r echo=FALSE}
#######testing some covariates:
dB$chamber <- relevel(dB$chamber, ref="A-2")

####### Maringal effect of chamber B-4
tidy_regression(lmer(average_speed~chamber + (1|ID), data=dB), is_color=output_col)

####### No effect of test date
tidy_regression(lmer(average_speed~days_from_start + (1|ID), data=dB), is_color=output_col)

####### No effect of test time
tidy_regression(lmer(average_speed~min_from_IncStart + (1|trial_type), data=dB), is_color=output_col)
```

<a id="wo.massb"></a>

**Without Mass**

```{r}
data<-data.frame(R=dB$average_speed,
                 A=dB$host_c, 
                 B=dB$sex_c, 
                 C=dB$sym_dist, 
                 X=dB$ID, Y=dB$chamber) 

source("src/compare_models.R")
errors <- withWarnings(model_comparisonsAIC("src/generic models-gaussian glmer 2-RF + 3-FF.R"))
cat("Number of models that did not converge: ", length(errors$warnings))
```

```{r}
anova(m18, m0, test="Chisq") # Replacing Y with X improves the fit
anova(m0, m36, test="Chisq") # Adding Y Marginally improves the fit
```

```{r}
burster_model <- lmer(average_speed ~ (1|ID), data=dB) # null model the best fit model
tidy_regression(burster_model, is_color=output_col)
```

```{r}
s.test <- paste("pval: ", shapiro.test(residuals(burster_model))$p.value)
qqnorm(resid(burster_model))
qqline(resid(burster_model))
text(-1, 0.1, s.test)
```


<a id="w.massb"></a>

**With Mass**

```{r}
data<-data.frame(R=dB$average_speed,
                 A=dB$host_c, 
                 B=dB$sex_c, 
                 C=dB$sym_dist,
                 D=dB$mass_c,
                 X=dB$trial_type, Y=dB$chamber)

source("src/compare_models.R")
errors <- withWarnings(model_comparisonsAIC("src/generic models-gaussian glmer 2-RF + 4-FF.R"))
cat("Number of models that did not converge: ", length(errors$warnings))
```

```{r}
anova(m113, m117, test="Chisq") # Adding D does not improve fit
anova(m113, m226, test="Chisq") # Adidng X does not improve fit
anova(m0, m4, test="Chisq") # Adding D does not improve fit
anova(m226, m230, test="Chisq") # Adding D does not improve fit
```

```{r}
burster_model <- lmer(average_speed ~ (1|chamber), data=dB) # null model the best fit model
tidy_regression(burster_model, is_color=output_col)
```

```{r}
s.test <- paste("pval: ", shapiro.test(residuals(burster_model))$p.value)
qqnorm(resid(burster_model))
qqline(resid(burster_model))
text(-1, 0.1, s.test)
```

<a id="morphb"></a>

**Morphology**

```{r}
dB<-data_flew[data_flew$flight_type=="B",] 
dB <- dB %>%
  filter(!is.na(body))
dB <- center_data(dB)

data<-data.frame(R=dB$average_speed,
                 A=dB$thorax_c, 
                 B=dB$body_c, 
                 C=dB$wing_c,
                 X=dB$trial_type, Y=dB$chamber) 

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian glmer 2-RF + 3-FF.R")
```

```{r}
anova(m18, m19, test="Chisq") # Adding A does marginally improve fit
anova(m0, m36, test="Chisq") # Adding Y does not impove fit
```

```{r}
burster_morph_model <- lmer(average_speed ~ thorax_c + (1|chamber), data=dB)
tidy_regression(burster_morph_model, is_color=output_col) # thorax not significant - null model best fit 
burster_morph_model <- lmer(average_speed ~ (1|chamber), data=dB)
tidy_regression(burster_morph_model, is_color=output_col) 
```

```{r}
s.test <- paste("pval: ", shapiro.test(residuals(burster_morph_model))$p.value)
qqnorm(resid(burster_morph_model))
qqline(resid(burster_morph_model))
text(-1, 0.1, s.test)
```



### Burster Plots

```{r}
gf_histogram(~average_speed,data=dB)
```


```{r}
source("src/plotting-lm.R")
source("src/plotting-lm2.R")
```

```{r echo=FALSE, warning=FALSE}

#, fig.width=2.8, fig.height=2.3}

lm_plot(dB, "sym_dist", "average_speed", "Distance From Sympatric Zone", "Average Speed (m/s)", "Average Speed by Distance From Sympatric Zone")

lm_plot2(dB, "sex_c", "average_speed", "Sex", "Average Speed (m/s)", "Average Speed by Sex")
axis(1, at = seq(-1,1,.5),labels=c("M", " ", " ", " ", "F"))

lm_plot2(dB, "host_c", "average_speed", "Host Plant", "Average Speed (m/s)", "Average Speed by Host Plant")
axis(1, at = seq(-1,1,.5),labels=c("BV", " ", " ", " ", "GRT"))

dB <- dB %>%
  filter(!is.na(mass_c))
lm_plot(dB, "mass_c", "average_speed", "Mass", "Average Speed (m/s)", "Average Speed by Mass")
```

```{r echo=FALSE, warning=FALSE}

#, fig.width=2.8, fig.height=2.3}

lm_plot(dB, "thorax_c", "average_speed", "Thorax Length (mm)",  "Average Speed (m/s)", "Average Speed by Thorax Length") # best model
lm_plot(dB, "body_c", "average_speed", "Body Length (mm)", "Average Speed (m/s)", "Average Speed by Body Length")
lm_plot(dB, "wing_c", "average_speed", "Wing Length (mm)", "Average Speed (m/s)", "Average Speed by Wing Length")
lm_plot(dB, "beak_c", "average_speed", "Beak Length (mm)", "Average Speed (m/s)", "Average Speed by Beak Length")
```