---
title: 'Modeling Flight Speed'
author: "Anastasia Bernat"
date: "5/4/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
rm(list=ls())
setwd("~/Desktop/Rstats-winter2020/")

library(lme4)
library(rethinking)

library(lubridate)
library(chron)
library(dplyr)
library(tidyselect)
library(stringr)

library(glmnet)
library(ggplotify)
library(gridExtra)
library(ggformula)
library(randomcoloR)

knitr::opts_chunk$set(echo = TRUE)
```

## Winter 2020 Flight Trials: Spped Flight Modeling {.tabset}

Flight Trials Winter 2020 Dataset was conducted from 2/17/2020 - 3/10/2020. Soapberry bugs were flight tested twice for multiple hours in the flight mill and observed from 8 AM to (5-8 PM) each day. Used multivariate (glm) and mixed effect modeling (glmer) to analyze the flight results.

### All Flyers

[Cleaning Data](#clean)

[Testing Covariates](#test)

[Without Mass](#wo.mass)

[With Mass](#w.mass)

[Morphology](#morph)

<a id="clean"></a>

**Cleaning the Data**

```{r}
rm(list=ls())
output_col = FALSE # Recommend changing this to TRUE if working in Base R or RStudio, and FALSE if generating an html
source("clean_flight_data.R") # Script that loads and cleans up the data
source("regression_output.R") # A script that cleans up regression outputs and prints in color or black and white
source("center_flight_data.R")

data <- read_flight_data("all_flight_data-Winter2020.csv")
data_all <- data[[1]]
data_tested <- data[[2]]

### Remove everyone who didn't fly (then remove distances = 0, if that didn't work fully)
data_flew_all <- data_tested[data_tested$flew_b == 1, ] 
```

```{r}
### Check for low speeds
low_speeds <- data_flew_all %>%
  filter(average_speed <0.05)

### Check for high speeds
high_speeds <- data_flew_all %>%
  filter(average_speed >0.65)

low_speeds$flight_type# have 7 bugs with average_speed = 0 but were marked as bursters (this could be just something very short (second burst) - not enough to grant a calculation) - I decided to remove them. But one bug was continuous and had 0 distance and 0 speeds - that was bug 196 T2 set011-3-03-2020-A3_196.txt
high_speeds$flight_type # 3 bugs - also bursters. Could also be short explosive bursts but not true to the biology of these bugs (more like us blowing on them).
```

```{r}
### Remove outliers
data_flew <- data_flew_all %>%
  filter(average_speed > 0.05) %>%
  filter(average_speed < 0.65)
  
data_flew <- center_data(data_flew)
```

<a id="test"></a>

**Testing Models and Covariates**

```{r}
test_model<-lmer(average_speed~host_c*sex_c*sym_dist + (1|ID), data=data_flew) #it converges - Run one set of models with ID as a random factor.
getME(test_model, "lower")
```

```{r echo=FALSE}
#######testing some covariates:
data_flew$chamber <- relevel(data_flew$chamber, ref="A-2")

####### Effect of chamber B-4
tidy_regression(lmer(average_speed~chamber + (1|ID), data=data_flew), is_color=output_col) ###Possibly reductions in speed 

####### No effect of test date
tidy_regression(lmer(average_speed~days_from_start + (1|chamber), data=data_flew), is_color=output_col)

####### No effect of test time
tidy_regression(lmer(average_speed~min_from_IncStart + (1|trial_type), data=data_flew), is_color=output_col)

```

<a id="wo.mass"></a>

**Without Mass**

```{r}
data<-data.frame(R=data_flew$average_speed,
                 A=data_flew$host_c, 
                 B=data_flew$sex_c, 
                 C=data_flew$sym_dist, 
                 X=data_flew$trial_type)

source("AICprobabilities.R")
source("generic models-gaussian glmer 1-RF + 3-FF.R") 
AICs <- sort(summary$AIC)
models_init <- sort(P, decreasing=TRUE, index.return=TRUE)
top <- length(models_init$x[which(models_init$x>0.05)])

AICs <- AICs[1:top]
models <- lapply(models_init$ix[1:top],1,FUN=as.integer)
probs <- models_init$x[1:top]
rbind(AICs, models, probs)
```
Top model (holding a 95% probability) contains only the random effects of trial type; no effects of host, sex, or distance from the sympatric zone on speed

- m18 <- lmer(R ~ (1 | Y), data=data)

```{r}
speed_model <- lmer(average_speed ~ (1 | trial_type), data=data_flew)
tidy_regression(speed_model, is_color=output_col)
```

<a id="w.mass"></a>

**With Mass**

```{r}
data_flew <- data_flew %>%
  filter(!is.na(mass))
data_flew <- center_data(data_flew)

data<-data.frame(R=data_flew$average_speed,
                 A=data_flew$host_c, 
                 B=data_flew$sex_c, 
                 C=data_flew$sym_dist, 
                 D=data_flew$mass_c,
                 X=data_flew$chamber) 

source("AICprobabilities.R")
source("generic models-gaussian glmer 1-RF + 4-FF.R") 
AICs <- sort(summary$AIC)
models_init <- sort(P, decreasing=TRUE, index.return=TRUE)
top <- length(models_init$x[which(models_init$x>0.05)])

AICs <- AICs[1:top]
models <- lapply(models_init$ix[1:top],1,FUN=as.integer)
probs <- models_init$x[1:top]
rbind(AICs, models, probs)
```

- m4 <- R ~ D + (1 | X)
- m0 <- R ~ 1 + (1 | X)

```{r}
anova(m0, m4, test="Chisq") # Adding D does improve fit
# note: if use data_flew_all you get the null model as the best fit model
```


```{r}
speed_model <- lmer(average_speed ~ mass_c + (1 | chamber), data=data_flew)
tidy_regression(speed_model, is_color=output_col)
```

* possibly marginal? effect of mass...

<a id="morph"></a>

**Morphology**

```{r}
data_flew <- data_flew %>%
  filter(!is.na(mass)) %>%
  filter(!is.na(body))
data_flew <- center_data(data_flew)

data<-data.frame(R=data_flew$average_speed,
                 A=data_flew$beak_c, 
                 B=data_flew$thorax_c, 
                 C=data_flew$body_c, 
                 D=data_flew$wing_c,
                 X=data_flew$chamber) # 5 models failed to converge
source("AICprobabilities.R")
source("generic models-gaussian glmer 1-RF + 4-FF.R") 
AICs <- sort(summary$AIC)
models_init <- sort(P, decreasing=TRUE, index.return=TRUE)
top <- length(models_init$x[which(models_init$x>0.05)])

AICs <- AICs[1:top]
models <- lapply(models_init$ix[1:top],1,FUN=as.integer)
probs <- models_init$x[1:top]
rbind(AICs, models, probs)
```

- m2 <- R ~ B + (1 | X)
- m4 <- R ~ D + (1 | X)
- m0 <- R ~ 1 + (1 | X)
- m3 <-  R ~ C + (1 | X)

- m115 <- R ~ B + (1 | Y)
- m113 <- R ~ 1 + (1 | Y)
- m228 <- R ~ B + (1 | X) + (1 | Y)
- m226 <- R ~ 1 + (1 | X) + (1 | Y)

```{r}
anova(m0, m2, test="Chisq") # Adding B does improve fit
anova(m0, m4, test="Chisq") # Adding D does improve fit
anova(m0, m3, test="Chisq") # Adding C does improve fit
anova(m2, m4, test="Chisq")   
anova(m2, m3, test="Chisq") # m2 the better fit model 
```

```{r}
speed_morph_model <- lmer(average_speed ~ thorax_c + (1|chamber), data=data_flew)
tidy_regression(speed_morph_model, is_color=output_col)
```

* positive effect of thorax, where the longer the thorax, the faster the speed

### All Flyers Plots

```{r}
gf_histogram(~average_speed,data=data_flew_all, col=~flight_type) # before filtering outliers
gf_histogram(~average_speed,data=data_flew, col=~flight_type) # after filtering outliers
```

```{r}
source("plotting-lm.R")
source("plotting-lm2.R")
```

```{r echo=FALSE, warning=FALSE}
#, fig.width=2.8, fig.height=2.3}

lm_plot(data_flew, "sym_dist", "average_speed", "Distance From Sympatric Zone", "Average Speed (m/s)", "Average Speed by Distance From Sympatric Zone (°)")

lm_plot2(data_flew, "sex_c", "average_speed", "Sex", "Average Speed (m/s)", "Average Speed by Sex")
axis(1, at = seq(-1,1,.5),labels=c("M", " ", " ", " ", "F"))

lm_plot2(data_flew, "host_c", "average_speed", "Host Plant", "Average Speed (m/s)", "Average Speed by Host Plant")
axis(1, at = seq(-1,1,.5),labels=c("BV", " ", " ", " ", "GRT"))

data_flew <- data_flew %>%
  filter(!is.na(mass_c))
lm_plot(data_flew, "mass_c", "average_speed", "Mass",  "Average Speed (m/s)", "Average Speed by Mass")
```

```{r echo=FALSE, warning=FALSE}

#, fig.width=2.8, fig.height=2.3}

lm_plot(data_flew, "thorax_c", "average_speed", "Thorax Length (mm)",  "Average Speed (m/s)", "Average Speed by Thorax Length") # best model
lm_plot(data_flew, "body_c", "average_speed", "Body Length (mm)", "Average Speed (m/s)", "Average Speed by Body Length")
lm_plot(data_flew, "wing_c", "average_speed", "Wing Length (mm)", "Average Speed (m/s)", "Average Speed by Wing Length")
lm_plot(data_flew, "beak_c", "average_speed", "Beak Length (mm)", "Average Speed (m/s)", "Average Speed by Beak Length")
```

### Continuous Flyers

[Cleaning Data](#cleandata)

[Testing Covariates](#testing)

[Without Mass](#wo.massc)

[With Mass](#w.massc)

[Morphology](#morphc)

<a id="cleandata"></a>

**Cleaning the data**

```{r}
rm(list=ls())
output_col = FALSE 
source("clean_flight_data.R") 
source("regression_output.R")
source("center_flight_data.R")

data <- read_flight_data("all_flight_data-Winter2020.csv")
data_all <- data[[1]]
data_tested <- data[[2]]

### Remove everyone who didn't fly (then remove distances = 0, if that didn't work fully)
data_flew <- data_tested[data_tested$flew_b == 1, ]
data_flew <- data_flew[data_flew$average_speed > 0, ]
data_flew <- center_data(data_flew)

### Break up by flight type
dC<-data_flew[data_flew$flight_type=="C",] # this exclusdes BC or CB flyers 
dC <- center_data(dC)
```

<a id="testing"></a>

**Testing Models and Covariates**

```{r}
test_model<-lmer(average_speed~host_c*sex_c*sym_dist + (1|ID), data=dC) #it converges - Run one set of models with ID as a random factor.
getME(test_model, "lower")
```

```{r echo=FALSE}
#######testing some covariates:
dC$chamber <- relevel(dC$chamber, ref="A-2")

####### Effect of chamber B-4
tidy_regression(lmer(average_speed~chamber + (1|ID), data=dC), is_color=output_col) ###Possibly reductions in speed 

####### No effect of test date
tidy_regression(lmer(average_speed~days_from_start + (1|ID), data=dC), is_color=output_col)

####### No effect of test time
tidy_regression(lmer(average_speed~min_from_IncStart + (1|ID), data=dC), is_color=output_col)
```

<a id="wo.massc"></a>

**Without Mass**

```{r}
data<-data.frame(R=dC$average_speed,
                 A=dC$host_c, 
                 B=dC$sex_c, 
                 C=dC$sym_dist, 
                 X=dC$ID, Y=dC$chamber)

source("AICprobabilities.R")
source("generic models-gaussian glmer 2-RF + 3-FF.R") 
AICs <- sort(summary$AIC)
models_init <- sort(P, decreasing=TRUE, index.return=TRUE)
top <- length(models_init$x[which(models_init$x>0.05)])

AICs <- AICs[1:top]
models <- lapply(models_init$ix[1:top],1,FUN=as.integer)
probs <- models_init$x[1:top]
rbind(AICs, models, probs)
```

- m18 <- R ~ (1 | Y)
- m36 <- R ~ (1 | X) + (1 | Y)
- m20 <- R ~ B + (1 | Y)
- m38 <- R ~ B + (1 | X) + (1 | Y)

```{r}
anova(m18, m36, test="Chisq") # Adding (1|X) does not improve fit
anova(m18, m20, test="Chisq") # Adding B does improve fit!
anova(m20, m38, test="Chisq") # Adding (1|X) does not improve fit
```

```{r}
continuous_model <- lmer(average_speed ~ sex_c + (1|chamber), data=dC)
tidy_regression(continuous_model, is_color=output_col)
```

*marginal? effect of sex_c, where if a female continuous-flyer bug then more likely to fly faster on average 

<a id="w.massc"></a>

**With Mass**

```{r}
data<-data.frame(R=dC$average_speed,
                 A=dC$host_c, 
                 B=dC$sex_c, 
                 C=dC$sym_dist,
                 D=dC$mass_c,
                 X=dC$ID, Y=dC$chamber)

source("AICprobabilities.R")
source("generic models-gaussian glmer 2-RF + 4-FF.R") # 4 models failed
AICs <- sort(summary$AIC)
models_init <- sort(P, decreasing=TRUE, index.return=TRUE)
top <- length(models_init$x[which(models_init$x>0.05)])

AICs <- AICs[1:top]
models <- lapply(models_init$ix[1:top],1,FUN=as.integer)
probs <- models_init$x[1:top]
rbind(AICs, models, probs) # top two models have combined prob of 89%
```

- m133 <- R ~ B * D + (1 | Y)
- m246 <- R ~ B * D + (1 | X) + (1 | Y)

```{r}
anova(m133, m246, test="Chisq") # Adding (1|X) does not improve fit
```

```{r}
continuous_model <- lmer(average_speed ~ sex_c*mass_c + (1|chamber), data=dC)
tidy_regression(continuous_model, is_color=output_col)
```
* no effect of sex_c
* no effect of mass_c
* negative effect of sex_c*mass_c, where if female and heavier, then slower avg speed (make sense because probably due to egg-laying)

<a id="morphc"></a>

**Morphology**

```{r}
dC<-data_flew[data_flew$flight_type=="C",] 
dC <- dC %>%
  filter(!is.na(body))
dC <- center_data(dC)

data<-data.frame(R=dC$average_speed,
                 A=dC$beak_c, 
                 B=dC$thorax_c, 
                 C=dC$body_c, 
                 D=dC$wing_c,
                 X=dC$ID, Y=dC$chamber) # 8 models failed to converge

source("AICprobabilities.R")
source("generic models-gaussian glmer 2-RF + 4-FF.R") 
AICs <- sort(summary$AIC)
models_init <- sort(P, decreasing=TRUE, index.return=TRUE)
top <- length(models_init$x[which(models_init$x>0.05)])

AICs <- AICs[1:top]
models <- lapply(models_init$ix[1:top],1,FUN=as.integer)
probs <- models_init$x[1:top]
rbind(AICs, models, probs)
```

- m115 <- R ~ B + (1 | Y)
- m228 <- R ~ B + (1 | X) + (1 | Y)
- m113 <- R ~ 1 + (1 | Y)
- m116 <- R ~ C + (1 | Y)
- m117 <- R ~ D + (1 | Y)
- m226 <- R ~ 1 + (1 | X) + (1 | Y)
- m229 <- R ~ C + (1 | X) + (1 | Y)
- m230 <- R ~ D + (1 | X) + (1 | Y)

```{r}
anova(m115, m228, test="Chisq") # Adding (1|X) does not improve fit
anova(m113, m115, test="Chisq") # Adding B does improve fit
anova(m113, m116, test="Chisq") # Adding C does improve fit
anova(m113, m117, test="Chisq") # Adding D does improve fit

anova(m115, m116, test="Chisq") # Replacing B with C improves fit
anova(m115, m117, test="Chisq") # Replacing B with D improves fit
anova(m117, m116, test="Chisq") # Replacing D with C improves fit

# m115 best fit model?
```

```{r}
continuous_morph_model <- lmer(average_speed ~ thorax_c + (1|chamber), data=dC)
tidy_regression(continuous_morph_model, is_color=output_col)
```

* positive effect of thorax length, where the longer the thorax the faster the avg speed

### Continuous Flyers Plots

```{r}
gf_histogram(~average_speed,data=dC)
```

```{r}
source("plotting-lm.R")
source("plotting-lm2.R")
```

```{r echo=FALSE, warning=FALSE}

#, fig.width=2.8, fig.height=2.3}

lm_plot(dC, "sym_dist", "average_speed", "Distance From Sympatric Zone", "Average Speed (m/s)", "Average Speed by Distance From Sympatric Zone")

lm_plot2(dC, "sex_c", "average_speed", "Sex", "Average Speed (m/s)", "Average Speed by Sex")
axis(1, at = seq(-1,1,.5),labels=c("M", " ", " ", " ", "F"))

lm_plot2(dC, "host_c", "average_speed", "Host Plant", "Average Speed (m/s)", "Average Speed by Host Plant")
axis(1, at = seq(-1,1,.5),labels=c("BV", " ", " ", " ", "GRT"))

dC <- dC %>%
  filter(!is.na(mass_c))
lm_plot(dC, "mass_c", "average_speed", "Mass", "Average Speed (m/s)", "Average Speed by Mass")
```

```{r echo=FALSE, warning=FALSE}

#, fig.width=2.8, fig.height=2.3}

lm_plot(dC, "thorax_c", "average_speed", "Thorax Length (mm)",  "Average Speed (m/s)", "Average Speed by Thorax Length") # best model
lm_plot(dC, "body_c", "average_speed", "Body Length (mm)", "Average Speed (m/s)", "Average Speed by Body Length")
lm_plot(dC, "wing_c", "average_speed", "Wing Length (mm)", "Average Speed (m/s)", "Average Speed by Wing Length")
lm_plot(dC, "beak_c", "average_speed", "Beak Length (mm)", "Average Speed (m/s)", "Average Speed by Beak Length")
```

### Bursters

[Cleaning Data](#cleanthedata)

[Testing Covariates](#testingcov)

[Without Mass](#wo.massb)

[With Mass](#w.massb)

[Morphology](#morphb)

<a id="cleanthedata"></a>

**Cleaning the data**

```{r}
rm(list=ls())
output_col = FALSE 
source("clean_flight_data.R") 
source("regression_output.R")
source("center_flight_data.R")

data <- read_flight_data("all_flight_data-Winter2020.csv")
data_all <- data[[1]]
data_tested <- data[[2]]

### Remove everyone who didn't fly (then remove distances = 0, if that didn't work fully)
data_flew <- data_tested[data_tested$flew_b == 1, ]
data_flew <- data_flew %>%
  filter(average_speed > 0) %>%
  filter(average_speed < 0.65)
data_flew <- center_data(data_flew)

### Break up by flight type
dB<-data_flew[data_flew$flight_type=="B",] # this exclusdes BC or CB flyers 
dB <- center_data(dB)
```

<a id="testingcov"></a>

**Testing Models and Covariates**

```{r}
test_model<-lmer(average_speed~host_c*sex_c*sym_dist + (1|ID), data=dB) #it converges - Run one set of models with ID as a random factor.
getME(test_model, "lower")
```

```{r echo=FALSE}
#######testing some covariates:
dB$chamber <- relevel(dB$chamber, ref="A-2")

####### No effect of chamber
tidy_regression(lmer(average_speed~chamber + (1|ID), data=dB), is_color=output_col)

####### No effect of test date
tidy_regression(lmer(average_speed~days_from_start + (1|ID), data=dB), is_color=output_col)

####### No effect of test time
tidy_regression(lmer(average_speed~min_from_IncStart + (1|trial_type), data=dB), is_color=output_col)
```

<a id="wo.massb"></a>

**Without Mass**

```{r}
data<-data.frame(R=dB$average_speed,
                 A=dB$host_c, 
                 B=dB$sex_c, 
                 C=dB$sym_dist, 
                 X=dB$ID, Y=dB$trial_type) 

source("AICprobabilities.R")
source("generic models-gaussian glmer 2-RF + 3-FF.R") # 1 model failed to converge
AICs <- sort(summary$AIC)
models_init <- sort(P, decreasing=TRUE, index.return=TRUE)
top <- length(models_init$x[which(models_init$x>0.05)])

AICs <- AICs[1:top]
models <- lapply(models_init$ix[1:top],1,FUN=as.integer)
probs <- models_init$x[1:top]
rbind(AICs, models, probs)
```

- m18 <- R ~ (1 | Y)
- m0 <- R ~ (1 | X)
- m36 <- R ~ (1 | X) + (1 | Y)

```{r}
anova(m18, m0, test="Chisq") # Replacing Y with X improves the fit
anova(m0, m36, test="Chisq") # Adding Y does not improve the fit
```

```{r}
burster_model <- lmer(average_speed ~ (1|ID), data=dB) # null model the best fit model
tidy_regression(burster_model, is_color=output_col)
```

<a id="w.massb"></a>

**With Mass**

```{r}
data<-data.frame(R=dB$average_speed,
                 A=dB$host_c, 
                 B=dB$sex_c, 
                 C=dB$sym_dist,
                 D=dB$mass_c,
                 X=dB$trial_type, Y=dB$chamber)

source("AICprobabilities.R")
source("generic models-gaussian glmer 2-RF + 4-FF.R") 
AICs <- sort(summary$AIC)
models_init <- sort(P, decreasing=TRUE, index.return=TRUE)
top <- length(models_init$x[which(models_init$x>0.05)])

AICs <- AICs[1:top]
models <- lapply(models_init$ix[1:top],1,FUN=as.integer)
probs <- models_init$x[1:top]
rbind(AICs, models, probs) # top two models have combined prob of 92%
```

- m113 <- R ~ 1 + (1 | Y)
- m117 <- R ~ D + (1 | Y)
- m226 <- R ~ 1 + (1 | X) + (1 | Y)
- m0 <-  R ~ 1 + (1 | X)
- m230 <- R ~ D + (1 | X) + (1 | Y)
- m4 <- R ~ D + (1 | X)

```{r}
anova(m113, m117, test="Chisq") # Adding D does not improve fit
anova(m113, m226, test="Chisq") # Adidng X does not improve fit
anova(m0, m4, test="Chisq") # Adding D does not improve fit
anova(m226, m230, test="Chisq") # Adding D does not improve fit
```

```{r}
burster_model <- lmer(average_speed ~ (1|chamber), data=dB) # null model the best fit model
tidy_regression(burster_model, is_color=output_col)
```

<a id="morphb"></a>

**Morphology**

```{r}
dB<-data_flew[data_flew$flight_type=="B",] 
dB <- dB %>%
  filter(!is.na(body))
dB <- center_data(dB)

data<-data.frame(R=dB$average_speed,
                 A=dB$beak_c, 
                 B=dB$thorax_c, 
                 C=dB$body_c, 
                 D=dB$wing_c,
                 X=dB$trial_type, Y=dB$chamber) 

source("AICprobabilities.R")
source("generic models-gaussian glmer 2-RF + 4-FF.R") 
AICs <- sort(summary$AIC)
models_init <- sort(P, decreasing=TRUE, index.return=TRUE)
top <- length(models_init$x[which(models_init$x>0.05)])

AICs <- AICs[1:top]
models <- lapply(models_init$ix[1:top],1,FUN=as.integer)
probs <- models_init$x[1:top]
rbind(AICs, models, probs)
```

- m113 <- R ~ 1 + (1 | Y)
- m226 <- R ~ 1 + (1 | X) + (1 | Y)
- m0 <-  R ~ 1 + (1 | X)
- m115 <- R ~ B + (1 | Y)

```{r}
anova(m113, m115, test="Chisq") # Adding B marginally improves fit
anova(m0, m226, test="Chisq") # Adding Y does not impove fit
```

```{r}
burster_morph_model <- lmer(average_speed ~ thorax_c + (1|chamber), data=dB)
tidy_regression(burster_morph_model, is_color=output_col) # thorax not significant - null model best fit 
burster_morph_model <- lmer(average_speed ~ (1|chamber), data=dB)
tidy_regression(burster_morph_model, is_color=output_col) 
```

### Burster Plots

```{r}
gf_histogram(~average_speed,data=dB)
```


```{r}
source("plotting-lm.R")
source("plotting-lm2.R")
```

```{r echo=FALSE, warning=FALSE}

#, fig.width=2.8, fig.height=2.3}

lm_plot(dB, "sym_dist", "average_speed", "Distance From Sympatric Zone", "Average Speed (m/s)", "Average Speed by Distance From Sympatric Zone")

lm_plot2(dB, "sex_c", "average_speed", "Sex", "Average Speed (m/s)", "Average Speed by Sex")
axis(1, at = seq(-1,1,.5),labels=c("M", " ", " ", " ", "F"))

lm_plot2(dB, "host_c", "average_speed", "Host Plant", "Average Speed (m/s)", "Average Speed by Host Plant")
axis(1, at = seq(-1,1,.5),labels=c("BV", " ", " ", " ", "GRT"))

dB <- dB %>%
  filter(!is.na(mass_c))
lm_plot(dB, "mass_c", "average_speed", "Mass", "Average Speed (m/s)", "Average Speed by Mass")
```

```{r echo=FALSE, warning=FALSE}

#, fig.width=2.8, fig.height=2.3}

lm_plot(dB, "thorax_c", "average_speed", "Thorax Length (mm)",  "Average Speed (m/s)", "Average Speed by Thorax Length") # best model
lm_plot(dB, "body_c", "average_speed", "Body Length (mm)", "Average Speed (m/s)", "Average Speed by Body Length")
lm_plot(dB, "wing_c", "average_speed", "Wing Length (mm)", "Average Speed (m/s)", "Average Speed by Wing Length")
lm_plot(dB, "beak_c", "average_speed", "Beak Length (mm)", "Average Speed (m/s)", "Average Speed by Beak Length")
```