---
title: "Modeling Morphology"
author: "Anastasia Bernat"
date: "6/24/2020"
output: html_document
---

```{r setup, include=FALSE}
rm(list=ls())
setwd("~/Desktop/Rstats-winter2020/")

# For Modeling
library(lme4)

# For Data Summaries and Manipulation
library(dplyr)
library(mosaic)

# For Plotting
library(ggplotify)
library(gridExtra) # "grid" graphics

knitr::opts_chunk$set(echo = TRUE)
ggplot2::theme_set(ggplot2::theme_bw())
ggplot2::theme_update(panel.grid.major = ggplot2::element_line(colour = "grey75"))
```

## Winter 2020 Flight Trials: Morphology Modeling {.tabset}

Flight Trials Winter 2020 Dataset was conducted from 2/17/2020 - 3/10/2020. Soapberry bugs were flight tested twice for multiple hours in the flight mill and observed from 8 AM to (5-8 PM) each day. No morphological covariates were included for thorax width or body length (just looked at sex, host, and sym_dist). For wing length and beak length, sex, host, sym_dist, and thorax width were all included.

### All Data

[Reading Data](#data_clean)

[Testing Covariates](#testing)

Comparing Models:

* [Thorax](#thorax)

* [Body](#body)

* [Wing](#wing)

* [Beak](#beak)

<a id="data_clean"></a>

#### Reading the data

```{r}
rm(list=ls())
output_col = FALSE # Recommend changing this to TRUE if working in Base R or RStudio, and FALSE if generating an html
source("src/clean_flight_data.R") # Script that loads and cleans up the data
source("src/regression_output.R") # A script that cleans up regression outputs and prints in color or black and white
source("src/center_flight_data.R")

data <- read_flight_data("data/all_flight_data-Winter2020.csv")
data_all <- data[[1]]
data_tested <- data[[2]]
```

<a id="testing"></a>

#### Testing Models and Covariates

**Experimental Set-Up Effects**

```{r echo=FALSE}
###### No effect of chamber
tidy_regression(lmer(thorax_c~chamber + (1|trial_type), data=data_tested), is_color=output_col) 

####### No effect of test date
tidy_regression(lmer(thorax_c~days_from_start_c+ (1|trial_type), data=data_tested), is_color=output_col)

####### No effect of test time
tidy_regression(lmer(thorax_c~min_from_IncStart_c + (1|trial_type), data=data_tested), is_color=output_col) 
```

<a id="thorax"></a>

#### Comparing Models

**Thorax**

```{r, warning=FALSE, message=FALSE}
data<-data.frame(R=data_tested$thorax_c, 
                 A=data_tested$host_c, 
                 B=data_tested$sex_c, 
                 C=data_tested$sym_dist, 
                 X=data_tested$trial_type, Y=data_tested$chamber)

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian glmer 2-RF + 3-FF.R")
```

```{r}
anova(m7, m43, test="Chisq") # Adding (1|Y) does not improve fit
```

```{r}
best.fit.thorax <- lmer(thorax_c ~ host_c + sex_c + sym_dist + (1|trial_type), data=data_tested)
tidy_regression(best.fit.thorax, is_color=output_col)
```

* negative effect of host plant where if from GRT then smaller thorax length
* strong positive effect of sex where if Female then larger thorax length
* positive effect of sym_dist where if farther from Homestead, then longer thorax length

<a id="body"></a>

**Body Length**

```{r, warning=FALSE, message=FALSE}
data<-data.frame(R=data_tested$body_c, 
                 A=data_tested$host_c, 
                 B=data_tested$sex_c, 
                 C=data_tested$sym_dist, 
                 X=data_tested$trial_type, Y=data_tested$chamber)

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian glmer 2-RF + 3-FF.R")
```

```{r}
anova(m11, m29, test="Chisq") # Replacing X with Y does improve fit
anova(m29, m33, test="Chisq") # Adding B*C marginally improves fit
anova(m11, m15, test="Chisq") # Adding B*C marginally improves fit
anova(m15, m33, test="Chisq") # Replacing X with Y does improve fit
```

```{r}
best.fit.body <- lmer(body_c ~ host_c * sex_c + sym_dist + (1 | chamber), data=data_tested)
tidy_regression(best.fit.body, is_color=output_col)
```

* negative effect of host where if from GRT then smaller body
* strong positve effect of sex, where if Female then larger body
* positve effect of sym_dist where if farther from Homestead, then larger body

* positive host*sex interaction where if Female and from GRT then larger body

<a id="wing"></a>

**Wing Length**

```{r, warning=FALSE, message=FALSE}
data<-data.frame(R=data_tested$wing_c, 
                 A=data_tested$host_c, 
                 B=data_tested$sex_c, 
                 C=data_tested$sym_dist, 
                 D=data_tested$thorax_c,
                 X=data_tested$trial_type, Y=data_tested$chamber)

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian glmer 2-RF + 4-FF.R")
```

```{r}
anova(m72, m185, test="Chisq") # Replacing X with Y does not improve fit
anova(m72, m298, test="Chisq") # Adding Y does not improve fit
```

```{r}
best.fit.wing <- lmer(body_c ~ host_c * sex_c + host_c * thorax_c + sex_c * thorax_c + (1 | trial_type), data=data_tested)
tidy_regression(best.fit.wing, is_color=output_col)
```

* no effect of host
* positive effect of sex where if Female then have longer wings
* strong positive effect of thorax where the longer the thorax, the longer the wings

* positive effect of host*sex where if Female and from GRT then have longer wings
* negative effect of host*thorax where if from GRT and have logner thorax, then have longer wings
* no effect of sex*thorax

<a id="beak"></a>

**Beak Length**

```{r, warning=FALSE, message=FALSE}
data<-data.frame(R=data_tested$beak_c, 
                 A=data_tested$host_c, 
                 B=data_tested$sex_c, 
                 C=data_tested$sym_dist, 
                 D=data_tested$thorax_c,
                 X=data_tested$trial_type, Y=data_tested$chamber)

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian glmer 2-RF + 4-FF.R")
```

```{r}
anova(m80, m193, test="Chisq") # Replacing X with Y does not improve fit
anova(m80, m306, test="Chisq") # Adding Y does not improve fit
```

```{r}
best.fit.beak <- lmer(body_c ~ host_c * sym_dist + sex_c * sym_dist + sex_c * thorax_c + (1 | trial_type), data=data_tested)
tidy_regression(best.fit.beak, is_color=output_col)
```

* positive effect of host where if from GRT then have longer beak
* no effect of sym_dist
* no effect of sex
* strong positive effect of thorax where the longer the thorax the longer the beak

* no effect of host*sym_dist
* no effect of sym_dist*sex
* no effect of sex*thorax

### All Data Plots

```{r}
par(mfrow=c(2,3))
plot(thorax_c ~ host_c, data=data_tested, main="data_tested")
plot(thorax_c ~ sym_dist, data=data_tested, main="data_tested")
plot(thorax_c ~ sex_c, data=data_tested, main="data_tested")

plot(thorax_c ~ host_c, data=data_all, main="data_all")
plot(thorax_c ~ sym_dist, data=data_all, main="data_all")
plot(thorax_c ~ sex_c, data=data_all, main="data_all")
```

```{r}
gf_point(thorax_c ~ sym_dist, col=~host_c, alpha=~sex_c, data=data_tested)
gf_point(thorax_c ~ sym_dist, col=~host_c, alpha=~sex_c, data=data_all)
```


### Female Data

[Reading Data](#data_cleanF)

[Testing Covariates](#testingF)

Comparing Models:

* [Thorax](#thoraxF)

* [Body](#bodyF)

* [Wing](#wingF)

* [Beak](#beakF)

<a id="data_cleanF"></a>

```{r}
rm(list=ls())
output_col = FALSE 
source("src/clean_flight_data.R") 
source("src/regression_output.R")
source("src/center_flight_data.R")

data <- read_flight_data("data/all_flight_data-Winter2020.csv")
data_all <- data[[1]]
data_tested <- data[[2]]

data_fem <- data_tested[data_tested$sex=="F",]
data_fem <- center_data(data_fem)
```

<a id="testingF"></a>

#### Testing Models and Covariates

**Experimental Set-Up Effects**

```{r echo=FALSE}
###### Marginal effect of chambers B-2 and B-3
tidy_regression(lmer(thorax_c~chamber + (1|trial_type), data=data_fem), is_color=output_col) 

####### No effect of test date
tidy_regression(lmer(thorax_c~days_from_start_c+ (1|trial_type), data=data_fem), is_color=output_col)

####### No effect of test time
tidy_regression(lmer(thorax_c~min_from_IncStart_c + (1|trial_type), data=data_fem), is_color=output_col) 
```

<a id="thoraxF"></a>

#### Comparing Models

**Thorax**

```{r, warning=FALSE, message=FALSE}
data<-data.frame(R=data_fem$thorax_c, 
                 A=data_fem$host_c, 
                 B=data_fem$sym_dist, 
                 X=data_fem$trial_type, Y=data_fem$chamber)

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian glmer 2-RF + 2-FF.R")
```

```{r}
best.fem.thorax <- lmer(thorax_c ~ (1|chamber), data=data_fem)
tidy_regression(best.fem.thorax, is_color=output_col)
```

<a id="bodyF"></a>

**Body**

```{r, warning=FALSE, message=FALSE}
data<-data.frame(R=data_fem$body_c, 
                 A=data_fem$host_c, 
                 B=data_fem$sym_dist,
                 X=data_fem$trial_type, Y=data_fem$chamber)

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian glmer 2-RF + 2-FF.R")
```

```{r}
anova(m0, m10, test="Chisq") # Adding Y marginally improves fit
anova(m0, m6, test="Chisq") # Adding A does improve fit
anova(m0, m7, test="Chisq") # Adding B does improve fit
anova(m5, m6, test="Chisq") # Adding A does not improve fit
anova(m5, m7, test="Chisq") # Adding B does not improve fit
anova(m0, m5, test="Chisq") # Replacing Y with X improves fit
```

```{r}
best.fem.body <- lmer(body_c ~ (1 | trial_type), data=data_fem)
tidy_regression(best.fem.body, is_color=output_col)
```

<a id="wingF"></a>

**Wing**

```{r, warning=FALSE, message=FALSE}
data<-data.frame(R=data_fem$wing_c, 
                 A=data_fem$host_c, 
                 B=data_fem$sym_dist,
                 C=data_fem$thorax_c,
                 X=data_fem$trial_type)

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian glmer 1-RF + 3-FF.R")
```

```{r}
best.fem.wing <- lmer(wing_c ~ host_c * thorax_c + (1 | trial_type), data=data_fem)
tidy_regression(best.fem.wing, is_color=output_col)
```

* positive effect of host where if from GRT then longer wing
* strong positive effect of thorax where the longer the thorax the longer the wing
* negative effect of host*thorax where the longer the thorax and from GRT then the shorter the wing (conflicting interaction.)

<a id="beakF"></a>

**Beak**

```{r, warning=FALSE, message=FALSE}
data<-data.frame(R=data_fem$beak_c, 
                 A=data_fem$host_c, 
                 B=data_fem$sym_dist,
                 C=data_fem$thorax_c,
                 X=data_fem$trial_type)

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian glmer 1-RF + 3-FF.R")
```

```{r}
anova(m5, m7, test="Chisq") # Adding B does improve fit
anova(m7, m11, test="Chisq") # Adding A*B does not improve fit
```

```{r}
best.fem.beak <- lmer(beak_c ~ host_c + sym_dist + thorax_c + (1 | trial_type), data=data_fem)
tidy_regression(best.fem.beak, is_color=output_col)
```

* negative effect of host where if from GRT then the shorter the beak
* marginal negative effect/no effect of sym_dist
* strong positive effect of host where the longer the thorax, the longer the beak


### Female Plots

```{r}
plot(thorax ~ host_c, data=data_fem)
plot(thorax ~ sym_dist, data=data_fem)
gf_point(thorax_c ~ sym_dist, col=~host_c, data=data_fem)
```


### Male Data 

[Reading Data](#data_cleanM)

[Testing Covariates](#testingM)

Comparing Models:

* [Thorax](#thoraxM)

* [Body](#bodyM)

* [Wing](#wingM)

* [Beak](#beakM)

<a id="data_cleanM"></a>

```{r}
rm(list=ls())
output_col = FALSE #
source("src/clean_flight_data.R") 
source("src/regression_output.R") 
source("src/center_flight_data.R")

data <- read_flight_data("data/all_flight_data-Winter2020.csv")
data_all <- data[[1]]
data_tested <- data[[2]]

data_male <- data_tested[data_tested$sex=="M",]
data_male <- center_data(data_male)
```

<a id="testingM"></a>

#### Testing Models and Covariates

**Experimental Set-Up Effects**

```{r echo=FALSE}
###### No effect of chamber
tidy_regression(lmer(thorax_c~chamber + (1|trial_type), data=data_male), is_color=output_col) 

####### No effect of test date
tidy_regression(lmer(thorax_c~days_from_start_c+ (1|trial_type), data=data_male), is_color=output_col)

####### No effect of test time
tidy_regression(lmer(thorax_c~min_from_IncStart_c + (1|trial_type), data=data_male), is_color=output_col) 
```

<a id="thoraxM"></a>

#### Comparing Models

**Thorax**

```{r, warning=FALSE , message=FALSE}
data<-data.frame(R=data_male$thorax_c, 
                 A=data_male$host_c, 
                 B=data_male$sym_dist, 
                 X=data_male$chamber, Y=data_male$trial_type)

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian glmer 2-RF + 2-FF.R")
```

```{r}
anova(m3, m13, test="Chisq") # adding (1|Y) does not improve fit
anova(m3, m8, test="Chisq") # replacing X with Y does improve fit
anova(m8, m13, test="Chisq") # Adding (1|X) does not improve fit
```

```{r}
best.male.thorax <- lmer(thorax_c ~ host_c + sym_dist + (1|trial_type), data=data_male)
tidy_regression(best.male.thorax, is_color=output_col)
```

* negative effect of host where if form GRT, the shorter the thorax
* positve effect of sym_dist where the farther from Homestead, the longer the thorax

<a id="bodyM"></a>

**Body**

```{r, warning=FALSE, message=FALSE}
data<-data.frame(R=data_male$body_c, 
                 A=data_male$host_c, 
                 B=data_male$sym_dist,
                 X=data_male$chamber, Y=data_male$trial_type)

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian glmer 2-RF + 2-FF.R")
```

```{r}
anova(m3, m8, test="Chisq") # Replacing X with Y does not improve fit
anova(m3, m4, test="Chisq") # Adding A*B does not improve fit
anova(m3, m13, test="Chisq") # Adding Y does not improve fit
```

```{r}
best.male.body <- lmer(body_c ~ host_c + sym_dist + (1 | chamber), data=data_male)
tidy_regression(best.male.body, is_color=output_col)
```

* negative effect of host where if from GRT then smaller body
* positive effect of sym_dist where if farther from Homestead then larger body

<a id="wingM"></a>

**Wing**

```{r, warning=FALSE, message=FALSE}
data<-data.frame(R=data_male$wing_c, 
                 A=data_male$host_c, 
                 B=data_male$sym_dist,
                 C=data_male$thorax_c,
                 X=data_male$chamber, Y=data_male$trial_type)

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian glmer 2-RF + 3-FF.R")
```

```{r}
anova(m3, m21, test="Chisq") # Replacing X with Y does not improve fit
anova(m3, m39, test="Chisq") # Adding Y does not improve fit
```

```{r}
best.male.wing <- lmer(wing_c ~ thorax_c + (1 | chamber), data=data_male)
tidy_regression(best.male.wing, is_color=output_col)
```

* strong positive effect of thorax where the longer the thorax, the longer the wing

<a id="beakM"></a>

**Beak**

```{r, warning=FALSE, message=FALSE}
data<-data.frame(R=data_male$beak_c, 
                 A=data_male$host_c, 
                 B=data_male$sym_dist,
                 C=data_male$thorax_c,
                 X=data_male$chamber, Y=data_male$trial_type)

source("src/compare_models.R")
model_comparisonsAIC("src/generic models-gaussian glmer 2-RF + 3-FF.R")
```

```{r}
anova(m5, m23, test="Chisq") # replacing X with Y does not improve fit
anova(m5, m41, test="Chisq") # adding Y does not improve fit
```

```{r}
best.male.beak <- lmer(wing_c ~ host_c + thorax_c + (1 | chamber), data=data_male)
tidy_regression(best.male.beak, is_color=output_col)
```

* no effect of hsot
* strong positive effect of thorax where the larger the thorax, the longer the beak

### Male Plots

```{r}
plot(thorax ~ host_c, data=data_male)
plot(thorax ~ sym_dist, data=data_male)
gf_point(thorax_c ~ sym_dist, col=~host_c, data=data_male)
```
