---
title: "Preliminary Extreme Trial Analyses"
author: "Anastasia Bernat"
date: "10/5/2021"
output: html_document
---

```{r setup, include=FALSE}
rm(list=ls())
dir = "/Users/anastasiabernat/Desktop/git_repositories/SBB-dispersal/avbernat_working_on/Dispersal/Summer_2021/windaq_processing/data/"
setwd(dir)

knitr::opts_chunk$set(echo = TRUE)
```

```{r}
prelims = read.csv(paste0(dir,"prelim_extreme_data.csv"), header=TRUE, stringsAsFactors = TRUE)
```

```{r}
library(dplyr)
prelims = prelims %>%
  filter(died. != "?", died. != "Y", ID != "ID", set_number!="")
prelims = prelims[,-c(2,13)]
```


```{r}
colnames(prelims) = c("ID1", "chamber", "set_number", "test_date", "time_start", "time_end", 
                      "temp", "RH", "population", "flew", "flight_type", "mass", "EWM", "NOTES",
                      "acclimated")
```

```{r}
prelims$acc_b[prelims$acclimated=="Y"] = 1
prelims$acc_b[prelims$acclimated=="N"] = -1

prelims$flew_b[prelims$flew=="Y"] = 1
prelims$flew_b[prelims$flew=="N"] = -1
```

```{r}
library(ggplot2)

ggplot(prelims[prelims$temp == "12",], aes(x=acc_b, fill=flew)) +
  geom_histogram() # 70 % RH 
ggplot(prelims[prelims$temp == "20",], aes(x=acc_b, fill=flew)) +
  geom_histogram() # 70 % RH 
# ggplot(prelims[prelims$temp == "28",], aes(x=acc_b, fill=flew)) +
#   geom_histogram() # 40 % RH 
# ggplot(prelims[prelims$temp == "36",], aes(x=acc_b, fill=flew)) +
#   geom_histogram() # 70 % RH # all not acclimated
ggplot(prelims[prelims$temp == "38",], aes(x=acc_b, fill=flew)) +
  geom_histogram() # 70 % RH 
```

```{r}
# how does acclimation affect whether bugs flew or not?
m = glm(flew ~ acc_b, data=prelims[prelims$temp != "36",], family = "binomial")
summary(m) # it doesn't

#  does acclimating matter more for continuous flyers?

contin = prelims %>%
  filter(temp != "36", flight_type != "B")

contin$temp = as.integer(as.character(contin$temp))

m = glm(flew ~ acc_b, data=contin, family = "binomial")
summary(m) # it doesn't


# how does temperature affect whether flew or not?

m = glm(flew ~ temp, data=contin, family = "binomial")
summary(m) # it doesn't
```


```{r}
ggplot(contin, aes(x=temp, fill=flew)) +
  geom_histogram(bins=20) 


ggplot(contin, aes(x=temp, fill=flew)) +
  geom_histogram(bins=20)  +
  facet_grid(acclimated ~ ., scales = "free")  +
  ylim(0, 12)

# but see declines and that parabolic curve...distance and speed 
# would be more accurate measurement for this 
```

It's hard to tell what the effect of temperature is if these bugs still flew at 38 C and 70%. More visible may be the effect of temperature on distance or speed for those that flew continuously? Acclimation doesn't seem to matter here in a consistently obvious way either.


Temporary conclusion, SBB have a **wide thermal tolerance and high dispersal drive.** Still need to check distance and speeds. Quick proxi for distance is duration spent flying:


```{r}
prelims$temp = as.integer(as.character(prelims$temp))
```

```{r}
prelims
```

```{r}

#times2decimals <- as.character(prelims$time_start)

times2decimals = function(times) {
  decimals = sapply(strsplit(times,":"),
  function(x) {
    x <- as.numeric(x)
    x[1]+x[2]/60
    }
)
  return(decimals)
}


prelims[prelims$ID1 == 229,] # row 81
prelims[prelims$ID1 == 191,] # row 119
prelims[prelims$ID1 == 27,] # row 139
prelims[119,] 

prelims$time_start_dec = times2decimals(as.character(prelims$time_start))
prelims$time_end_dec = times2decimals(as.character(prelims$time_end))

duration = prelims$time_end_dec - prelims$time_start_dec

# add 24 hrs for the bugs that flew overnight 
duration[duration<0] + 24
overnight_i = which(duration<0)
duration[overnight_i] = duration[overnight_i] + 24

prelims$duration = duration
```

```{r}
contin2 = prelims %>%
  filter(temp != "36", flight_type != "B", flew != "N")

plot(contin2$temp, contin2$duration, xlab="temp (C)", ylab="duration (hrs)")

m = lm(duration ~ temp, data=contin2)
summary(m)
```


```{r}
m = lm(duration ~ acc_b, data=contin2)
summary(m) # acclimation does not matter for duration
```

```{r}
ggplot(contin2, aes(x=duration, fill=flew)) +
  geom_histogram(bins=30)  +
  facet_grid(acclimated ~ ., scales = "free")  +
  ylim(0, 12)
```






