---
title: "All Morphology: Testing for Spatial Dependencies"
author: "Anastasia Bernat"
date: "4/23/2021"
output: html_document
---

```{r setup, include=FALSE}
rm(list=ls())
dir = "~/Desktop/git_repositories/SBB-dispersal/avbernat_working_on/All_Morphology/stats/"
setwd(dir)

library(dplyr)
library(zoo)

library(vegan)
library(lattice)
library(MASS)
library(sp)
library(raster)
library(dismo)
library(splancs)
library(INLA)
library(reshape)
library(gstat)
library(ggplot2)
library(ggmap)
library(rworldmap)
library(rgdal)
library(fields)
library(rgeos)

library(sf)

knitr::opts_chunk$set(echo = TRUE)
```

# Barrier Model


## Sourcing Scripts and Reading the Data

```{r}
source_path = "~/Desktop/git_repositories/SBB-dispersal/avbernat_working_on/Rsrc/"

script_names = c("compare_models.R",
                 "regression_output.R",
                 "clean_morph_data2.R", # two functions: read_morph_data and remove_torn_wings
                 "AICprobabilities.R",
                 "HighstatLibV13.R")

for (script in script_names) { 
  path = paste0(source_path, script)
  source(path) 
}

source("~/Desktop/git_repositories/SBB-dispersal/avbernat_working_on/RTsrc/ts_auxiliary_funs.R") # time
source("~/Desktop/git_repositories/SBB-dispersal/avbernat_working_on/RSsrc/spatial_dependencies.R") # space
```

### Clean the Data

```{r}
data_list <- read_morph_data("data/allmorphology04.28.21-coors.csv")
raw_data = data_list[[1]]
all_bugs = nrow(raw_data)
data_long = data_list[[2]] 

# Remove individuals with torn wings first
raw_data = remove_torn_wings(raw_data) # **don't need to remove torn wings for wing morph analysis
data_long = remove_torn_wings(data_long) # **don't need to remove torn wings for wing morph analysis
clean_bugs = nrow(raw_data)

cat("number of bugs with torn wings:", all_bugs - clean_bugs, "\n\n")
```

```{r}
# remove NA dates
d = data_long %>%
  filter(!is.na(wing2body), !is.na(datetime))

range(d$wing2body) 
```

```{r}
names(data_long)
```

```{r}
dotplot = function(var, data) {
  plot(data[,var], seq(1:nrow(data)),
       col=data$color, ylab="Order of the data from text file", xlab=paste0("Value of ", var))
}
```


```{r}
MyVar <- c("wing2body", "body", "wing", "beak", "thorax", "lat", "long") 
d$color = "black"
d$color[d$year == "2019" & d$month =="May"] = "red"

par(mfrow=c(3,3))
for (v in MyVar) {
  dotplot(v, d)
}

# remove outliers?
d2 = d %>% 
  filter(!beak<4.5)

par(mfrow=c(3,3))
for (v in MyVar) {
  dotplot(v, d2)
}
```

```{r}
coors = unique(d2[,c("population", "lat","long")])
coors[order(-coors$lat),]
```
August	2017...is it Key Largo or is it North Key Largo? Or are the coordinates seem a little off?

K.elegans	Homestead	M	5.59	2.84	7.5	10.55	February	2020	81	winter	L	25.1755702	-80.36780089	non-diapausing	2.5.2020 - Either this is Homestead or North Key Largo.


Here is the block

K.elegans	Gainesville	M	5.51	3.08	8.28	11.13	February	2020	81	winter	L	29.6620224	-82.347313
K.elegans	Homestead	M	5.59	2.84	7.5	10.55	February	2020	81	winter	L	25.1755702	-80.36780089 ***
K.elegans	Homestead	F	7.47	3.82	9.66	13.31	February	2020	81	winter	L	25.491362	-80.485821

Data cleaning is not linear, it's incrementally non-linear. 

We need to start adding site to the allmorph datasheet

This one too...lat is a little too hight to make sense: 	Key_Largo	25.17563	-80.36775	(search for the lon and you'll find it in the dataset)

August 2017, May 2019

```{r}
spatial_data <- st_as_sf(d2,
                         coords = c("long", "lat"),
                         crs = 4326,
                         agr = "constant")

spatial_data <- st_transform(spatial_data, 2236)
plot(spatial_data) 
```

Make a model and then create a variogram for it to check for spatial dependencies.

What effects wing2body?

```{r}
d2
```

```{r}
d2$wing2body_s = (d2$wing2body - mean(d2$wing2body)) / sd(d2$wing2body)
```

```{r}
# And this is the Poisson GLMM

M1 <- inla(wing2body ~ sex_binom + pophost_binom + month_of_year + months_since_start, 
                    control.compute = list(dic = TRUE), 
                    family = "gaussian", 
                    data = d2)

M2 <- inla(wing2body ~ sex + pophost + month_of_year + months_since_start +
                            f(population, model = "iid"), 
                    control.compute = list(dic = TRUE),          
                    family = "gaussian", 
                    data = d2)
# Compare them
dic  <- c(M1$dic$dic, M2$dic$dic) # M2 better
dic
```

```{r}
temp = d2 %>% 
  filter(!is.na(wing2body))

check_spatial_dependencies(M1, temp, temp$long, temp$lat, zone = 16, cutoff=10000, is_inla=TRUE)
```

```{r}
keys = c("North_Key_Largo", "Key_Largo", "Plantation_Key", "Homestead", "HomesteadBV", "HomesteadGRT")

d3 = d2 %>%
  filter(population == "North_Key_Largo" | population =="Key_Largo" | population == "Plantation_Key" | population == "Homestead" | population == "HomesteadBV" | population == "HomesteadGRT")
```

```{r}
spatial_data <- st_as_sf(d3,
                         coords = c("long", "lat"),
                         crs = 4326,
                         agr = "constant")

spatial_data <- st_transform(spatial_data, 2236)
plot(spatial_data)
```

```{r}
M1 <- inla(wing2body ~ sex_binom + pophost + month_of_year + months_since_start +
                    f(population, model = "iid"),  
                    control.compute = list(dic = TRUE), 
                    family = "gaussian", 
                    data = d3)
```

```{r}
temp = d3 %>% 
  filter(!is.na(wing2body))

check_spatial_dependencies(M1, temp, temp$long, temp$lat, zone = 16, cutoff=10000, is_inla=TRUE)
```


Remove homestead

```{r}
d4 = d2 %>%
  filter(population == "North_Key_Largo" | population =="Key_Largo" | population == "Plantation_Key")
```

```{r}
spatial_data <- st_as_sf(d4,
                         coords = c("long", "lat"),
                         crs = 4326,
                         agr = "constant")

spatial_data <- st_transform(spatial_data, 2236)
plot(spatial_data)
```

```{r}
M1 <- inla(wing2body ~ sex_binom + month_of_year + months_since_start +
                    f(population, model = "iid"),  
                    control.compute = list(dic = TRUE), 
                    family = "gaussian", 
                    data = d4)
```

```{r}
temp = d4 %>% 
  filter(!is.na(wing2body))

check_spatial_dependencies(M1, temp, temp$long, temp$lat, zone = 16, cutoff=10000, is_inla=TRUE)
```

Will probably need to take means because the layering is obscuring what is happening per site...
