---
title: "Modeling Wing Morphology Summary File"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
rm(list=ls())

# Set Working Directory
dir = "~/Desktop/git_repositories/SBB-dispersal/avbernat_working_on/All_Morphology/stats/"
setwd(dir)

knitr::opts_chunk$set(echo = TRUE)
```

# Data Cleaning And Exploration

## Read Libraries

```{r message=FALSE}
library(lme4) # fit regressions 
library(dplyr) # data manipulation 
library(ggformula) # ggplot plotting
library(cowplot) # ggplot helper functions to arrange multi-panel figures
```

## Read Source Files

```{r message=FALSE}
source_path = "~/Desktop/git_repositories/SBB-dispersal/avbernat_working_on/Rsrc/"

script_names = c("compare_models.R", # 1 function: model_comparisonsAIC()
                 "regression_output.R", # 1 function: tidy_regression()
                 "clean_morph_data3.R", # 2 functions: read_morph_data(), remove_torn_wings()
                 "AICprobabilities.R") # 1 function: AICprobs()

for (script in script_names) { 
  path = paste0(source_path, script)
  source(path) 
}

source("~/Desktop/git_repositories/SBB-dispersal/avbernat_working_on/RTsrc/vartests.R")
```

## Read the Data

```{r}
data_list <- read_morph_data("data/allmorphology05.18.21.csv")
raw_data = data_list[[1]]
data_long = data_list[[2]] # long-wing bugs only

data_long = remove_torn_wings(data_long)
```

## Histograms of Wing Morph Data 

&nbsp;

```{r fig.height=5.7, fig.width=6.6, echo=FALSE}
raw_data_missing = raw_data %>% 
  filter(w_morph=="" | is.na(w_morph)) 

par(mfrow=c(3,1))
hist(raw_data$wing[raw_data$w_morph=="L"]/raw_data$thorax[raw_data$w_morph=="L"],
     main="Histogram of wing length/thorax length for long winged SBB",
     xlab="wing length/thorax length",
     breaks=seq(0.5, 3.8, by=0.05))
hist(raw_data_missing$wing/raw_data_missing$thorax, 
     main="Histogram of wing length/thorax length for SBB w/o recorded wing morph",
     xlab="wing length/thorax length",
     breaks=seq(0.5, 3.8, by=0.05))
hist(raw_data$wing[raw_data$w_morph=="S"]/raw_data$thorax[raw_data$w_morph=="S"],
     main="Histogram of wing length/thorax length for short winged SBB",
     xlab="wing length/thorax length",
     breaks=seq(0.5,3.8,by=0.05))
```

Notice that there are 30 bugs that are hard to identify as either S or L. 

\newpage

## Barplots 

Bugs were collected on different years and months. These barplots show the bugs collected per **population**, **host plant**, and **sex** across the years and months.

```{r echo=FALSE}
# repeating plot features
customPlot = list(geom_bar(position='stack', color="black", width=0.7),
                  theme_bw(),
                  theme_classic(),
                  theme(axis.text.y = element_text(size=11),
                        axis.text.x=element_text(size=11, angle = 45, hjust = 1.1),
                        axis.title=element_text(size=17,face="bold")),
                  theme(axis.title.x = element_text(vjust = -3)),
                  theme(axis.title.y = element_text(vjust = 4)),
                  theme(plot.margin=unit(c(1,1,1.1,1.2),"cm"))
                  )
```

### Plot 1: Collection numbers grouped by Population and Datetime

```{r message=FALSE, fig.height=4.1, fig.width=7.2, echo=FALSE}
pop_colors = c("#787a87", "#E69F00", "#56B4E9", "royalblue", 
               "grey", "gold", "#409973", "#9bc969", "ivory2") 

p1 = ggplot(data=subset(raw_data, !is.na(datetime)), aes(x=datetime, fill=population)) + 
  labs(title=" ", fill="Population", 
       x="Field Collection Month", y = "Number of Bugs Collected") +
  scale_fill_manual(values=pop_colors) + customPlot

groups = raw_data %>% 
  group_by(sex, datetime) %>% 
  summarize(count = n())

numF = paste0("F(", as.character(unlist(groups[1:10,3])), ")")
numM = paste0("M(", as.character(unlist(groups[11:20,3])), ")")
text = paste0(numF, "\n", numM)
p1 = p1 + annotate(geom="text", x=1, y=400, label=text[1],
              color="black", size=3) +
  annotate(geom="text", x=2, y=470, label=text[2],
           color="black", size=3) +
  annotate(geom="text", x=3, y=270, label=text[3],
           color="black", size=3) +
  annotate(geom="text", x=4, y=220, label=text[4],
           color="black", size=3) +
  annotate(geom="text", x=5, y=320, label=text[5],
           color="black", size=3) +
  annotate(geom="text", x=6, y=500, label=text[6],
           color="black", size=3) +
  annotate(geom="text", x=7, y=240, label=text[7],
           color="black", size=3) +
  annotate(geom="text", x=8, y=730, label=paste0("\n", text[8]),
           color="black", size=3) +
  annotate(geom="text", x=9, y=440, label=text[9],
           color="black", size=3) +
  annotate(geom="text", x=10, y=650, label=text[10],
           color="black", size=3)
p1
```

### Plot 2: Collection numbers grouped by Host Plant and Datetime

```{r fig.height=4.1, fig.width=7.2, echo=FALSE}
p2 = ggplot(data=subset(raw_data, !is.na(datetime)), aes(x=datetime, fill=pophost)) + 
  labs(title=" ", fill="Host Plant",
              x="Field Collection Month", y = "Number of Bugs Collected") +
  scale_fill_manual(values=c("#56B4E9", "chartreuse4")) + customPlot
p2
```

### Plot 3: Collection numbers grouped by Sex and Datetime

```{r fig.height=4.1, fig.width=7.2, echo=FALSE}
p3 = ggplot(data=subset(raw_data, !is.na(datetime)), aes(x=datetime, fill=sex)) + 
  labs(title=" ", fill="Sex",
              x="Field Collection Month", y = "Number of Bugs Collected") +
  scale_fill_manual(values=c("salmon1", "darkslategray3")) + customPlot
p3
```

&nbsp;

# Regression Modeling

## Long-Wing Morph Frequency

How does sex, host plant, month, and/or year effect whether a soapberry bug is long-winged (wing_morph_binom=1) or short-winged (wing_morph_binom=0)? 

```{r}
data<-data.frame(R=raw_data$wing_morph_binom, 
                 A=raw_data$sex_binom, 
                 B=raw_data$pophost_binom, 
                 C=(raw_data$month_of_year),
                 D=raw_data$months_since_start)

model_script = paste0(source_path,"generic models-binomial glm 4-FF.R")
model_comparisonsAIC(model_script)
```

&nbsp;

```{r results = "hold"}
anova(m98, m110, test="Chisq") # adding B*D does not improve fit
anova(m84, m98, test="Chisq") # adding A*B improves fit
anova(m63, m84, test="Chisq") # Adding C*D improves fit
anova(m51, m63, test="Chisq") # Adding B improves fit
```

\newpage

## Best Fit

```{r}
M1 = glm(wing_morph_binom ~ sex_binom * pophost_binom + sex_binom * months_since_start +
           pophost_binom * month_of_year + month_of_year * months_since_start, 
         data=raw_data, family="binomial") 
summary(M1)
```

## Modeling Variance

In addition to modeling how sex, host plant, month, or year effects whether a soapberry bug is long-winged or short-winged, we modeled how those same factors affected the variance of wing morph.

```{r}
SE = function(x){sd(x)/sqrt(length(x))}
```

```{r}
wmorph_table<-aggregate(wing_morph_binom~sex_binom*pophost_binom*month_of_year*months_since_start, 
                           data=raw_data, FUN=mean)
wmorph_table$sd<-aggregate(wing_morph_binom~sex_binom*pophost_binom*month_of_year*months_since_start, 
                              data=raw_data,FUN=sd)$wing_morph_binom
wmorph_table$se<-aggregate(wing_morph_binom~sex_binom*pophost_binom*month_of_year*months_since_start, 
                              data=raw_data,FUN=SE)$wing_morph_binom
```

```{r}
data = wmorph_table
data<-data.frame(R=data$sd, 
                 A=data$sex_binom, 
                 B=data$pophost_binom, 
                 C=(data$month_of_year),
                 D=data$months_since_start)

model_script = paste0(source_path,"generic models-gaussian glm 4-FF.R")
model_comparisonsAIC(model_script)
```

&nbsp;

```{r results = "hold"}
anova(m2, m5, test="Chisq") # Adding A does not improve fit
anova(m2, m8, test="Chisq") # Adding C does not improve fit
anova(m2, m9, test="Chisq") # Adding D does not improve fit
anova(m0, m2, test="Chisq") # Adding B improves fit
```

## Best Fit

```{r}
M2 = glm(sd ~ pophost_binom, data=wmorph_table, family="gaussian")
summary(M2) 
```

&nbsp;

## Wing-to-body Ratio 

How does sex, host plant, month, and/or year effect the wing-to-body ratio of long-winged soapberry bugs?

```{r}
data<-data.frame(R=data_long$wing2body_c, # centered
                 A=data_long$sex_binom,
                 B=data_long$pophost_binom,
                 C=data_long$month_of_year_c, # centered
                 D=data_long$months_since_start_c) # centered

model_script = paste0(source_path,"generic models-gaussian glm 4-FF.R")
model_comparisonsAIC(model_script)
```

&nbsp;

```{r results = "hold"}
anova(m22, m42, test="Chisq") # adding B*C does not improve fit
anova(m16, m22, test="Chisq") # adding C does improve fit  
anova(m22, m34, test="Chisq") # adding D does not improve fit
anova(m34, m58, test="Chisq") # Adding B*D marginally improves fit
```

&nbsp;

## Best Fit

```{r}
M3 = glm(wing2body_c ~ sex_binom*pophost_binom + month_of_year_c, data=data_long, family=gaussian) 
summary(M3) 
```

&nbsp;

## Modeling Variance 

In addition to modeling how sex, host plant, month, or year effects the wing-to-body ratio of a soapberry bug, we modeled how those same factors affected the variance of wing-to-body ratio.

```{r}
w2b_table<-aggregate(wing2body~sex_binom*pophost_binom*month_of_year*months_since_start, 
                     data=data_long, FUN=mean)
w2b_table$sd<-aggregate(wing2body~sex_binom*pophost_binom*month_of_year*months_since_start, 
                           data=data_long, FUN=sd)$wing2body
w2b_table$se<-aggregate(wing2body~sex_binom*pophost_binom*month_of_year*months_since_start, 
                           data=data_long, FUN=SE)$wing2body
```

```{r}
data = w2b_table
data<-data.frame(R=data$sd, 
                 A=data$sex_binom, 
                 B=data$pophost_binom, 
                 C=(data$month_of_year),
                 D=data$months_since_start)

model_script = paste0(source_path,"generic models-gaussian glm 4-FF.R")
model_comparisonsAIC(model_script)
```

&nbsp;

```{r results = "hold"}
anova(m8, m19, test="Chisq") # Adding B*C does not improve fit
anova(m2, m8, test="Chisq") # Adding C does not improve fit
anova(m0, m2, test="Chisq") # Adding B improves fit
```

## Best Fit

```{r}
M4 = glm(sd ~ pophost_binom, data=w2b_table, family=gaussian) 
summary(M4)
```

&nbsp;

# LOESS & Linear Regression Plots

```{r echo=FALSE}
# repeating plotting features
customPlot = list( theme_classic(),
                   theme(axis.text=element_text(size=13),
                         axis.title=element_text(size=16), 
                         plot.title=element_text(size=20),),
                   theme(legend.position = c(0.2, 0.9)),
                   theme(legend.title = element_text(size=14, face="italic"),
                         legend.text = element_text(size = 13, face="italic"))
)

xlab_years = na.omit(sort(unique(data_long$dates))[-2])

xlab_allmonths = na.omit(sort(unique(data_long$month_of_year)))
xlab_months = xlab_allmonths[c(-2,-5)]
month_labs <- c("Feb", "May", "Aug", "Oct", "Dec")

hp_col_pts = c("turquoise3", "springgreen4")
hp_col_shade = c("turquoise3", "green")
sex_col_pts = c("brown3", "black")
sex_col_shade = c("brown1", "sienna4")
```

## Wing Morph Frequency

### Group significant elements 

```{r}
w_morph_summary<-aggregate(wing_morph_binom~sex*pophost*month_of_year*months_since_start
                           , data=raw_data, FUN=mean)
w_morph_summary$se<-aggregate(wing_morph_binom~sex*pophost*month_of_year*months_since_start
                              , data=raw_data, FUN=SE)$wing_morph_binom

jitter = runif(n=nrow(w_morph_summary), min=-0.5, max=0.5) # jitter points slightly
w_morph_summary$dates <- w_morph_summary$month_of_year + jitter

dd = w_morph_summary
```

### Check for LOESS Residuals 

```{r}
plot_lowess_residuals = function(lfit, x, y) {
  lfun <- approxfun(lfit)
  fitted <- lfun(x)
  resid <- y-fitted
  plot(fitted,resid)
  abline(h=0,col=8)
}
```

```{r fig.height=3.3}
l1 = lowess(dd$dates, dd$wing_morph_binom, f=0.4) # f = alpha, the smoother span

par(mfrow=c(1,2))
plot(dd$dates, dd$wing_morph_binom)
lines(l1, type = "l")
plot_lowess_residuals(l1, dd$dates, dd$wing_morph_binom)
```

### Figure: Panels A, B, C, D (long-wing morph freq with month) & E (long-wing morph freq with year)

```{r echo=FALSE}
# relevel factors for cleaner plotting
dd$pophost = factor(dd$pophost, levels = c("K. elegans", "C. corindum") )
dd$`Host Plant` = dd$pophost

dd$sex[dd$sex=="F"]<-"Females"
dd$sex[dd$sex=="M"]<-"Males"
dd$sex = factor(dd$sex, levels = c("Males", "Females") )
dd$`Sex` = dd$sex 
```

```{r}
summary(M1)
```

```{r warning=FALSE, message=FALSE, echo=FALSE}
# single-variate model of month predicting wing morph
# AB: best fit model, M1, has multiple interaction terms with month, so a single-variate model
# was substituted 
#fit = glm(wing_morph_binom ~ month_of_year, family = "binomial", data = raw_data)
xmonth <- seq(2,12, 0.01234)
bhost = sample(c(1), replace=TRUE, size=length(xmonth))
bsex = sample(c(1), replace=TRUE, size=length(xmonth))
zyear = sample(c(81), replace=TRUE, size=length(xmonth))
#zyear = seq(min(raw_data$months_since_start),max(raw_data$months_since_start), 0.1)


wing_probs <- predict(M1, list( month_of_year=xmonth, 
                                pophost_binom=bhost, 
                                sex_binom=bsex, 
                                months_since_start=zyear), type="response")

#wing_probs <- predict(fit, list(month_of_year = xmonth),type="response")
fit_pvalue = round(summary(fit)$coeff[,"Pr(>|z|)"][2],10)
fit_pvalue = round(summary(M1)$coeff[,"Pr(>|z|)"][5],5)
pvalue = paste0("italic(p)[glm]==", fit_pvalue)
```

```{r}
ggplot() + 
  ggtitle("A") + xlab("Month") + ylab("Long-Wing Morph Frequency") +
  geom_vline(xintercept = xlab_allmonths, color="gainsboro") + 
  geom_line(aes(x=xmonth, y=wing_probs), linetype = "dashed") 
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
p4 = ggplot() + 
  ggtitle("A") + xlab("Month") + ylab("Long-Wing Morph Frequency") +
  geom_vline(xintercept = xlab_allmonths, color="gainsboro") + 
  geom_line(aes(x=xmonth, y=wing_probs), linetype = "dashed") +
  geom_smooth(data=dd, method="loess", 
              mapping = aes(x = month_of_year, y = wing_morph_binom, 
                            colour=`Host Plant`, fill=`Host Plant`)) + 
  geom_point(data=dd, mapping = aes(x = month_of_year, y = wing_morph_binom, colour=`Host Plant`)) +
  customPlot + 
  scale_color_manual(values=c("C. corindum" = hp_col_pts[1], "K. elegans" = hp_col_pts[2])) +
  scale_fill_manual(values = c("C. corindum" = hp_col_shade[1], "K. elegans" = hp_col_shade[2])) +
  scale_x_continuous(breaks=xlab_months, labels= month_labs) + 
  ylim(0.3,1.2) 

p5 = ggplot() + 
  ggtitle("B") + xlab("Month") + ylab("Long-Wing Morph Frequency") +
  geom_vline(xintercept = xlab_allmonths, color="gainsboro") + 
  geom_line(aes(x=xmonth, y=wing_probs), linetype = "dashed") +
  geom_smooth(data=dd, method="loess", 
              mapping = aes(x = month_of_year, y = wing_morph_binom, colour=Sex, fill=Sex)) + 
  geom_point(data=dd, mapping = aes(x = month_of_year, y = wing_morph_binom, colour=Sex)) +
  customPlot +
  scale_color_manual(values=c("Females" = sex_col_pts[1], "Males" = sex_col_pts[2])) +
  scale_fill_manual(values = c("Females" = sex_col_shade[1], "Males" = sex_col_shade[2])) +
  scale_x_continuous(breaks=xlab_months, labels= month_labs) + 
  ylim(0.3,1.2) + 
  theme(axis.title.y = element_blank()) 
# theme(legend.position = c(0.23, 0.95))

# extract alpha, smoothing parameter, from LOESS regression(s))
alpha = paste("alpha[loess]==", ggplot_build(p5)$data[[3]]$alpha[1])
degree="lambda[loess]==0"

p4 = p4 + 
  annotate(geom="text", x=7.4, y=1.0, label=alpha, parse=TRUE) +
  annotate(geom="text", x=7.4, y=1.04, label=degree, parse=TRUE) +
  annotate(geom="text", x=5, y=0.718, label=pvalue, parse=TRUE)

p5 = p5 + 
  annotate(geom="text", x=7.4, y=1.1, label=alpha, parse=TRUE) +
  annotate(geom="text", x=7.4, y=1.15, label=degree, parse=TRUE) +
  annotate(geom="text", x=5.5, y=1, label=pvalue, parse=TRUE)
```

```{r}
p4
```

```{r echo=FALSE, message=FALSE}
# create groups by sex
dfF = dd[dd$sex=="Females",]
dfM = dd[dd$sex=="Males",]

fit = glm(wing_morph_binom ~ month_of_year + sex_binom * pophost_binom + 
          pophost_binom * month_of_year, family = "binomial", data = raw_data)
xmonth <- seq(2,12, 0.01)
set.seed(194842)
bsex = sample(c(-1,1), replace=TRUE, size=length(xmonth))
bhost = sample(c(-1,1), replace=TRUE, size=length(xmonth))
wing_probs <- predict(fit, list(sex_binom = bsex,
                                pophost_binom = bhost,
                                month_of_year = xmonth),type="response")
fit_pvalue = round(summary(M1)$coeff[,"Pr(>|z|)"][6],4)
pvalue = paste0("italic(p)[glm]==", fit_pvalue)


pred = cbind(xmonth, bsex, bhost, wing_probs)
pred = as.data.frame(pred)

predFK = pred[pred$bhost==1 & pred$bsex==1,]
predFC = pred[pred$bhost==-1 & pred$bsex==1,]

predMK = pred[pred$bhost==1 & pred$bsex==-1,]
predMC = pred[pred$bhost==-1 & pred$bsex==-1,]
```

```{r echo=FALSE, message=FALSE}
### females
p8 = ggplot() + 
  ggtitle("C") + xlab("Month") + ylab("Long-Wing Morph Frequency") +
  geom_vline(xintercept = xlab_allmonths, color="gainsboro") + 
  geom_line(aes(x=predFK$xmonth, y=predFK$wing_probs), linetype = "dashed", color="darkgreen") +
  geom_line(aes(x=predFC$xmonth, y=predFC$wing_probs), linetype = "dashed", color="skyblue2") +
  geom_smooth(data=dfF, method="loess", span=0.8,
              mapping = aes(x = month_of_year, y = wing_morph_binom, colour=`Host Plant`, fill=`Host Plant`)) + 
  geom_point(data=dfF, mapping = aes(x = month_of_year, y = wing_morph_binom, colour=`Host Plant`)) +
  customPlot + 
  scale_color_manual(values=c("C. corindum" = hp_col_pts[1], "K. elegans" = hp_col_pts[2])) +
  scale_fill_manual(values = c("C. corindum" = hp_col_shade[1], "K. elegans" = hp_col_shade[2])) +
  scale_x_continuous(breaks=xlab_months, labels= month_labs) +
  ylim(0,1.3)

### males

p9 = ggplot() + 
  ggtitle("D") + xlab("Month") + ylab("Long-Wing Morph Frequency") +
  geom_vline(xintercept = xlab_allmonths, color="gainsboro") + 
  geom_line(aes(x=predMK$xmonth, y=predMK$wing_probs), linetype = "dashed", color="darkgreen") +
  geom_line(aes(x=predMC$xmonth, y=predMC$wing_probs), linetype = "dashed", color="skyblue2") +
  geom_smooth(data=dfM, method="loess", span=0.8,
              mapping = aes(x = month_of_year, y = wing_morph_binom, colour=`Host Plant`, fill=`Host Plant`)) + 
  geom_point(data=dfM, mapping = aes(x = month_of_year, y = wing_morph_binom, colour=`Host Plant`)) +
  customPlot + 
  scale_color_manual(values=c("C. corindum" = hp_col_pts[1], "K. elegans" = hp_col_pts[2])) +
  scale_fill_manual(values = c("C. corindum" = hp_col_shade[1], "K. elegans" = hp_col_shade[2])) +
  scale_x_continuous(breaks=xlab_months, labels= month_labs) +
  ylim(0,1.3)

alpha = paste("alpha[loess]==", ggplot_build(p8)$data[[4]]$alpha[1])
degree="lambda[loess]==0"

p8 = p8 + 
  annotate(geom="text", x=10.7, y=1.3, label="Females", color="black", size=6, fontface = 'italic') + 
  annotate(geom="text", x=10, y=1.1, label=alpha, color="black", parse=TRUE) +
  annotate(geom="text", x=10, y=1.17, label=degree, color="black", parse=TRUE) +
  annotate(geom="text", x=10, y=0.25, label=pvalue, color="black", parse=TRUE)
#+ theme(legend.title = element_blank())

p9 = p9 + 
  annotate(geom="text", x=11, y=1.3, label="Males", color="black", size=6, fontface = 'italic') + 
  annotate(geom="text", x=10, y=1.1, label=alpha, color="black", parse=TRUE) +
  annotate(geom="text", x=10, y=1.17, label=degree, color="black",parse=TRUE) +
  annotate(geom="text", x=10, y=0.25, label=pvalue, color="black", parse=TRUE) +
  theme(legend.position = "none") + theme(axis.title.y = element_blank()) 
```

```{r echo=FALSE, message=FALSE}
threshold = sort(unique(raw_data$dates))[5]
wm_table<-aggregate(wing_morph_binom~dates, data=raw_data, FUN=mean)
wm_table$date_b = "2013-2015"
wm_table$date_b[wm_table$dates >= threshold] = "2016-2020"
wm_table$date_b = as.factor(wm_table$date_b)

xlab_years = sort(unique(raw_data$dates))
```

```{r echo=FALSE, message=FALSE}
p7 = ggplot() + 
  ggtitle("C") + xlab("Year") + ylab("Long-Wing Morph Frequency") +
  geom_vline(xintercept = xlab_years, color="gainsboro") + 
  geom_smooth(data=wm_table, method="lm", se=FALSE, linetype = "dashed",
              mapping = aes(x = dates, y = wing_morph_binom), colour="black", lwd=0.5) +
  geom_smooth(data=wm_table, method="loess", 
              mapping = aes(x = dates, y = wing_morph_binom), colour="black") + 
  geom_point(data=wm_table, mapping = aes(x = dates, y = wing_morph_binom)) +
  guides(fill = guide_legend(reverse = TRUE)) + theme_classic() +
  theme(axis.text=element_text(size=13),
        axis.title=element_text(size=16), 
        plot.title=element_text(size=20)) + 
  guides(color = FALSE) + 
  scale_linetype(guide = FALSE) +
  theme(legend.key = element_rect(fill = "white", color = NA), 
        legend.key.size = unit(0.5, "cm"),
        legend.key.width = unit(0.5,"cm")) + 
  scale_fill_manual(values = "blue", labels=c("C. corindum", "K. elegans")) + 
  labs(fill = "Host Plant") + 
  theme(legend.title = element_text(size = 14),
        legend.text = element_text(size = 13, face="italic")) +
  scale_color_manual(values="blue") + 
  theme(legend.position = c(0.2, 0.88)) + 
  theme(legend.title = element_blank()) + ylim(0.3,1.2)
```


```{r echo=FALSE, message=FALSE}
#### loess and linear regressions
alpha = paste("alpha[loess]==", ggplot_build(p7)$data[[3]]$alpha[1])
degree="lambda[loess]==0"
mlinear = glm(wing_morph_binom ~ dates, data=dd, family=gaussian)
pvalue = paste0("italic(p)[glm]==",round(summary(mlinear)$coeff[[8]],2))

p7 = p7 + annotate(geom="text", x=unique(raw_data$dates)[6], y=1.05, label=alpha, color="black", parse=TRUE, size=6) +
  annotate(geom="text", x=unique(raw_data$dates)[6], y=1.12, label=degree, color="black",parse=TRUE, size=6) +
  annotate(geom="text", x=unique(raw_data$dates)[5], y=0.50, label=pvalue, color="black", parse=TRUE, size=6)
```



```{r echo=FALSE, message=FALSE}
p7 = p7 + ggtitle("E")
p8 = p8 + theme(legend.position = "none") +
  annotate(geom="text", x=2.9, y=1.3, label="Females", color="black", size=6, fontface = 'italic')

p9 = p9 + theme(legend.position = "none") + 
  annotate(geom="text", x=2.7, y=1.3, label="Males", color="black", size=6, fontface = 'italic') +
  theme(axis.title.y = element_blank())
```

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=13.5, fig.width=10}
figure4 = ggdraw() +
  draw_plot(p5, 0, .66, .5, .33) +
  draw_plot(p6, 0.5, .66, .5, .33) +
  draw_plot(p8, 0, .33, .5, .33) +
  draw_plot(p9, 0.5, .33, .5, .33) +
  draw_plot(p7, 0, 0, 1, .33)
figure4
```



## Wing-to-body Ratio

```{r}
w2b_summary<-aggregate(wing2body~sex*pophost*dates, data=data_long, FUN=mean)
w2b_summary$se<-aggregate(wing2body~sex*pophost*dates, data=data_long,
                          FUN=SE)$wing2body

jitter = runif(n=nrow(w2b_summary), min=-0.5, max=0.5) # jitter slightly
w2b_summary$dates <- w2b_summary$dates + jitter
d = w2b_summary
```

## Check for LOESS Residuals 

```{r fig.height=3.3}
l1 = lowess(d$dates, d$wing2body, f=0.4)

par(mfrow=c(1,2))
plot(d$dates, d$wing2body)
lines(l1, type = "l")
plot_lowess_residuals(l1, d$dates, d$wing2body)
```

## Figure: Panel A & B (wing-to-body with month)

```{r}
w2b_summary<-aggregate(wing2body~sex*pophost*month_of_year, data=data_long, FUN=mean)
df = w2b_summary

df$pophost = factor(df$pophost, levels = c("K. elegans", "C. corindum") )
df$`Host Plant` = df$pophost
df$sex[df$sex=="F"]<-"Females"
df$sex[df$sex=="M"]<-"Males"
df$sex = factor(df$sex, levels = c("Males", "Females") )
df$`Sex` = df$sex 
```

```{r}
p1 = ggplot() + 
  ggtitle("A") + xlab("Month") + ylab("Wing-to-Body Ratio") +
  geom_vline(xintercept = xlab_allmonths, color="gainsboro") + 
  geom_smooth(data=data_long, method="glm", se=FALSE, linetype = "dashed", 
              mapping = aes(x = month_of_year, y = wing2body), colour="black", lwd=0.5) + 
  geom_smooth(data=df, method="loess", 
              mapping = aes(x = month_of_year, y = wing2body, colour=`Host Plant`, fill=`Host Plant`)) + 
  geom_point(data=df, mapping = aes(x = month_of_year, y = wing2body, colour=`Host Plant`)) + 
  ylim(0.71, 0.75) +
  customPlot +
  scale_color_manual(values=c("C. corindum" = hp_col_pts[1], "K. elegans" = hp_col_pts[2])) +
  scale_fill_manual(values = c("C. corindum" = hp_col_shade[1], "K. elegans" = hp_col_shade[2])) +
  scale_x_continuous(breaks=xlab_months, labels=month_labs)

p2 = ggplot() + 
  ggtitle("B") + xlab("Month") + ylab("Wing-to-Body Ratio") + 
  geom_vline(xintercept = xlab_allmonths, color="gainsboro") + 
  geom_smooth(data=data_long, method="glm", se=FALSE, linetype = "dashed",  
              mapping = aes(x = month_of_year, y = wing2body), colour="black", lwd=0.5) + 
  geom_smooth(data=df, method="loess",
              mapping = aes(x = month_of_year, y = wing2body, colour=Sex, fill=Sex)) + 
  geom_point(data=df, mapping = aes(x = month_of_year, y = wing2body, colour=Sex)) +
  ylim(0.71, 0.75) +
  customPlot +
  scale_color_manual(values=c("Females" = sex_col_pts[1], "Males" = sex_col_pts[2])) +
  scale_fill_manual(values = c("Females" = sex_col_shade[1], "Males" = sex_col_shade[2])) +
  scale_x_continuous(breaks=xlab_months, labels= month_labs) + theme(axis.title.y = element_blank())
```

```{r}
# Extract from LOESS and linear regressions
alpha = paste("alpha[loess]==", ggplot_build(p1)$data[[3]]$alpha[1])
degree="lambda[loess]==0"
mlinear = glm(wing2body ~ month_of_year, data=df, family=gaussian) # aggregated df
summary(M3) # individual bugs df (best fit model)
M3_beta = summary(M3)$coeff[,"Estimate"][4]
M3_pvalue = round(summary(M3)$coeff[,"Pr(>|t|)"][4],10)
pvalue = paste0("italic(p)[glm]==", M3_pvalue)
```


```{r}
  
# sex_col_shade = c("brown1", "sienna4")
# sex_col_pts = c("brown1", "grey27")
```

