---
title: "All Morphology: Testing for Spatial Dependencies"
author: "Anastasia Bernat"
date: "4/23/2021"
output: html_document
---

```{r setup, include=FALSE}
rm(list=ls())
dir = "~/Desktop/git_repositories/SBB-dispersal/avbernat_working_on/All_Morphology/stats/"
setwd(dir)

library(dplyr)
library(zoo)

library(vegan)
library(lattice)
library(MASS)
library(sp)
library(raster)
library(dismo)
library(splancs)
library(INLA)
library(reshape)
library(gstat)
library(ggplot2)
library(ggmap)
library(rworldmap)
library(rgdal)
library(fields)
library(rgeos)

library(sf)

knitr::opts_chunk$set(echo = TRUE)
```

# Barrier Model


## Sourcing Scripts and Reading the Data

```{r}
source_path = "~/Desktop/git_repositories/SBB-dispersal/avbernat_working_on/Rsrc/"

script_names = c("compare_models.R",
                 "regression_output.R",
                 "clean_morph_data2.R", # two functions: read_morph_data and remove_torn_wings
                 "AICprobabilities.R",
                 "HighstatLibV13.R")

for (script in script_names) { 
  path = paste0(source_path, script)
  source(path) 
}

source("~/Desktop/git_repositories/SBB-dispersal/avbernat_working_on/RTsrc/ts_auxiliary_funs.R") # time
source("~/Desktop/git_repositories/SBB-dispersal/avbernat_working_on/RSsrc/spatial_dependencies.R") # space
```

### Clean the Data

```{r}
data_list <- read_morph_data("data/allmorphology04.26.21-coors2.csv")
raw_data = data_list[[1]]
all_bugs = nrow(raw_data)
data_long = data_list[[2]] 

# Remove individuals with torn wings first
raw_data = remove_torn_wings(raw_data) # **don't need to remove torn wings for wing morph analysis
data_long = remove_torn_wings(data_long) # **don't need to remove torn wings for wing morph analysis
clean_bugs = nrow(raw_data)

cat("number of bugs with torn wings:", all_bugs - clean_bugs, "\n\n")
```

```{r}
# remove NA dates
d = data_long %>%
  filter(!is.na(wing2body), !is.na(datetime))

range(d$wing2body) 
```

```{r}
names(data_long)
```

```{r}
dotplot = function(var, data) {
  plot(data[,var], seq(1:nrow(data)),
       col=data$color, ylab="Order of the data from text file", xlab=paste0("Value of ", var))
}
```


```{r}
MyVar <- c("wing2body", "body", "wing", "beak", "thorax", "lat", "long") 
d$color = "black"
d$color[d$year == "2019" & d$month =="May"] = "red"

par(mfrow=c(3,3))
for (v in MyVar) {
  dotplot(v, d)
}

# remove outliers?
d2 = d %>% 
  filter(!beak<4.5)

par(mfrow=c(3,3))
for (v in MyVar) {
  dotplot(v, d2)
}
```

```{r}
coors = unique(d2[,c("population", "lat","long")])
coors[order(-coors$lat),]
```
August	2017...is it Key Largo or is it North Key Largo? Or are the coordinates seem a little off?

K.elegans	Homestead	M	5.59	2.84	7.5	10.55	February	2020	81	winter	L	25.1755702	-80.36780089	non-diapausing	2.5.2020 - Either this is Homestead or North Key Largo.


Here is the block

K.elegans	Gainesville	M	5.51	3.08	8.28	11.13	February	2020	81	winter	L	29.6620224	-82.347313
K.elegans	Homestead	M	5.59	2.84	7.5	10.55	February	2020	81	winter	L	25.1755702	-80.36780089 ***
K.elegans	Homestead	F	7.47	3.82	9.66	13.31	February	2020	81	winter	L	25.491362	-80.485821

Data cleaning is not linear, it's incrementally non-linear. 

We need to start adding site to the allmorph datasheet

This one too...lat is a little too hight to make sense: 	Key_Largo	25.17563	-80.36775	(search for the lon and you'll find it in the dataset)

August 2017, May 2019

```{r}
spatial_data <- st_as_sf(d2,
                         coords = c("long", "lat"),
                         crs = 4326,
                         agr = "constant")

spatial_data <- st_transform(spatial_data, 2236)
plot(spatial_data) 
```

Make a model and then create a variogram for it to check for spatial dependencies.

What effects wing2body?

```{r}
d2
```

```{r}
d2$wing2body_s = (d2$wing2body - mean(d2$wing2body)) / sd(d2$wing2body)
```

```{r}
# And this is the Poisson GLMM

M1 <- inla(wing2body ~ sex_binom + pophost_binom + month_of_year + months_since_start, 
                    control.compute = list(dic = TRUE), 
                    family = "gaussian", 
                    data = d2)

M2 <- inla(wing2body ~ sex + pophost + month_of_year + months_since_start +
                            f(population, model = "iid"), 
                    control.compute = list(dic = TRUE),          
                    family = "gaussian", 
                    data = d2)
# Compare them
dic  <- c(M1$dic$dic, M2$dic$dic) # M2 better
dic
```

```{r}
# Pearson residuals vs. fitted values
mu2 <- M2$summary.fitted.values[,"mean"]
E2  <- (d2$wing2body - mu2) / sqrt(mu2)

par(mfrow = c(1,1), mar = c(5,5,2,2), cex.lab = 1.5)
plot(x = mu2, 
     y = E2,
     xlab = "Fitted values",
    ylab = "Pearson residuals")
abline(h = 0, lty = 2)  
```

```{r}
# Plot residuals vs each covariate in the model
# And vs each covariate not in the model.
d2$E2 <- E2
MyVar <- c("sex", "pophost", "month_of_year", 
            "months_since_start")
MyMultipanel.ggp2(Z = d2, 
                  varx = MyVar, 
                  vary = "E2", 
                  ylab = "Pearson residuals gaussian GLMM",
                  addSmoother = FALSE,
                  addRegressionLine = FALSE,
                  addHorizontalLine = TRUE)
```

```{r}
LongLatToUTM <- function(x,y,zone){
  xy <- data.frame(ID = 1:length(x), X = x, Y = y)
  coordinates(xy) <- c("X", "Y")
  proj4string(xy) <- CRS("+proj=longlat +datum=WGS84")  
  res <- spTransform(xy, CRS(paste("+proj=utm +zone=",zone," ellps=WGS84",sep='')))
  return(as.data.frame(res))
}

LL <- LongLatToUTM(x = d2$long, y = d2$lat, zone = 16)
d2$X.utm <- LL$X
d2$Y.utm <- LL$Y
```

```{r}
xyplot(Y.utm / 1000 ~ X.utm / 1000,
       data = d2,
       aspect = "iso",
       pch = 16)

# What are the distances between the points?
Loc <- cbind(d2$X.utm, d2$Y.utm)
D <- dist(Loc)
par(mfrow = c(1,1), mar = c(5,5,2,2), cex.lab = 1.5)
hist(D, 
     freq = TRUE,
     main = "", 
     xlab = "Distance between sites (m)",
     ylab = "Frequency")
```

```{r}
# 1. Plot the Pearson versus their spatial coordinates.
MyData <- data.frame(E1  = E2,
                     Xkm = d2$X.utm,
                     Ykm = d2$Y.utm)

MyData$MySize <- 3 * abs(MyData$E1) / max( MyData$E1)
MyData$MyCol <- ifelse(MyData$E1> 0, 1, 2)
xyplot(Ykm ~ Xkm,
       data = MyData,
       cex = MyData$MySize,
       col = MyData$MyCol,
       pch = 1)

# We seem to have some spatial correlation in the Pearson residuals! We have groups 
# of red dots close to each other. And also groups of black dots.
```


```{r}
# Sample-variogram with distances up to 1200000 m
MyData  <- data.frame(E2 = E2, 
                      X  = d2$X.utm, 
                      Y  = d2$Y.utm)
coordinates(MyData)    <- c("X", "Y")
V12 <- variogram(E2 ~ 1, MyData, cressie = TRUE, cutoff = (1200000))
plot(V12, 
     main = "", 
     xlab = list(label = "Distance (m)", cex = 1.5), 
     ylab = list(label = "Semi-variogram", cex = 1.5),
       pch = 16,
       col = 1,
       cex = 1.5
       )

# Sample-variogram with distances up to 400000 m
MyData  <- data.frame(E2 = E2, 
                      X  = d2$X.utm, 
                      Y  = d2$Y.utm)
coordinates(MyData)    <- c("X", "Y")
V12 <- variogram(E2 ~ 1, MyData, cressie = TRUE, cutoff = (400000))
plot(V12, 
     main = "", 
     xlab = list(label = "Distance (m)", cex = 1.5), 
     ylab = list(label = "Semi-variogram", cex = 1.5),
       pch = 16,
       col = 1,
       cex = 1.5
       )

# Sample-variogram with distances up to 100000 m
MyData  <- data.frame(E2 = E2, 
                      X  = d2$X.utm, 
                      Y  = d2$Y.utm)
coordinates(MyData)    <- c("X", "Y")
V12 <- variogram(E2 ~ 1, MyData, cressie = TRUE, cutoff = (100000))
plot(V12, 
     main = "", 
     xlab = list(label = "Distance (m)", cex = 1.5), 
     ylab = list(label = "Semi-variogram", cex = 1.5),
       pch = 16,
       col = 1,
       cex = 1.5
       )

# Sample-variogram with distances up to 10000 m
MyData  <- data.frame(E2 = E2, 
                      X  = d2$X.utm, 
                      Y  = d2$Y.utm)
coordinates(MyData)    <- c("X", "Y")
V12 <- variogram(E2 ~ 1, MyData, cressie = TRUE, cutoff = (10000))
plot(V12, 
     main = "", 
     xlab = list(label = "Distance (m)", cex = 1.5), 
     ylab = list(label = "Semi-variogram", cex = 1.5),
       pch = 16,
       col = 1,
       cex = 1.5
       )

# Sample-variogram with distances up to 500 m
MyData  <- data.frame(E2 = E2, 
                      X  = d2$X.utm, 
                      Y  = d2$Y.utm)
coordinates(MyData)    <- c("X", "Y")
V12 <- variogram(E2 ~ 1, MyData, cressie = TRUE, cutoff = (500))
plot(V12, 
     main = "", 
     xlab = list(label = "Distance (m)", cex = 1.5), 
     ylab = list(label = "Semi-variogram", cex = 1.5),
       pch = 16,
       col = 1,
       cex = 1.5
       )
```
```{r}
# What is the distance for which we would expect dependency? (Since a bug can fly up to 14 km...that would be the max and then it would tapper off)
# Sample-variogram with distances up to 14000 m
MyData  <- data.frame(E2 = E2, 
                      X  = d2$X.utm, 
                      Y  = d2$Y.utm)
coordinates(MyData)    <- c("X", "Y")
V12 <- variogram(E2 ~ 1, MyData, cressie = TRUE, cutoff = (14000))
plot(V12, 
     main = "", 
     xlab = list(label = "Distance (m)", cex = 1.5), 
     ylab = list(label = "Semi-variogram", cex = 1.5),
       pch = 16,
       col = 1,
       cex = 1.5
       ) # or more realistically 7 km because it tappers around that

# But maybe only the keys and homestead, let's model only the keys
```

```{r}
keys = c("North_Key_Largo", "Key_Largo", "Plantation_Key", "Homestead", "HomesteadBV", "HomesteadGRT")

d3 = d2 %>%
  filter(population == "North_Key_Largo" | population =="Key_Largo" | population == "Plantation_Key" | population == "Homestead" | population == "HomesteadBV" | population == "HomesteadGRT")
```

```{r}
spatial_data <- st_as_sf(d3,
                         coords = c("long", "lat"),
                         crs = 4326,
                         agr = "constant")

spatial_data <- st_transform(spatial_data, 2236)
plot(spatial_data)
```

```{r}
M1 <- inla(wing2body ~ sex_binom + pophost + month_of_year + months_since_start +
                    f(population, model = "iid"),  
                    control.compute = list(dic = TRUE), 
                    family = "gaussian", 
                    data = d3)
```

```{r}
# Pearson residuals vs. fitted values
mu1 <- M1$summary.fitted.values[,"mean"]
E1  <- (d3$wing2body - mu1) / sqrt(mu1)

par(mfrow = c(1,1), mar = c(5,5,2,2), cex.lab = 1.5)
plot(x = mu1, 
     y = E1,
     xlab = "Fitted values",
    ylab = "Pearson residuals")
abline(h = 0, lty = 2)  
```

```{r}
# Plot residuals vs each covariate in the model
# And vs each covariate not in the model.
d3$E1 <- E1
MyVar <- c("sex", "month_of_year", "pophost",
            "months_since_start")
MyMultipanel.ggp2(Z = d3, 
                  varx = MyVar, 
                  vary = "E1", 
                  ylab = "Pearson residuals gaussian GLMM",
                  addSmoother = FALSE,
                  addRegressionLine = FALSE,
                  addHorizontalLine = TRUE)
```

```{r}
LL <- LongLatToUTM(x = d3$long, y = d3$lat, zone = 16)
d3$X.utm <- LL$X
d3$Y.utm <- LL$Y

xyplot(Y.utm / 1000 ~ X.utm / 1000,
       data = d3,
       aspect = "iso",
       pch = 16)

# What are the distances between the points?
Loc <- cbind(d3$X.utm, d3$Y.utm)
D <- dist(Loc)
par(mfrow = c(1,1), mar = c(5,5,2,2), cex.lab = 1.5)
hist(D, 
     freq = TRUE,
     main = "", 
     xlab = "Distance between sites (m)",
     ylab = "Frequency")
```

```{r}
# Sample-variogram with distances up to 250000 m
MyData  <- data.frame(E1 = E1, 
                      X  = d3$X.utm, 
                      Y  = d3$Y.utm)
coordinates(MyData)    <- c("X", "Y")
V12 <- variogram(E1 ~ 1, MyData, cressie = TRUE, cutoff = (250000))
plot(V12, 
     main = "", 
     xlab = list(label = "Distance (m)", cex = 1.5), 
     ylab = list(label = "Semi-variogram", cex = 1.5),
       pch = 16,
       col = 1,
       cex = 1.5
       )
with(V12, scatter.smooth(x = dist, 
                        y = gamma,
                        xlab = "Distance",
                        ylab = "Sample variogram"))

# Sample-variogram with distances up to 14000 m
MyData  <- data.frame(E1 = E1, 
                      X  = d3$X.utm, 
                      Y  = d3$Y.utm)
coordinates(MyData)    <- c("X", "Y")
V12 <- variogram(E1 ~ 1, MyData, cressie = TRUE, cutoff = (14000))
plot(V12, 
     main = "", 
     xlab = list(label = "Distance (m)", cex = 1.5), 
     ylab = list(label = "Semi-variogram", cex = 1.5),
       pch = 16,
       col = 1,
       cex = 1.5
       )
with(V12, scatter.smooth(x = dist, 
                        y = gamma,
                        xlab = "Distance",
                        ylab = "Sample variogram"))

# Sample-variogram with distances up to 1000 m
MyData  <- data.frame(E1 = E1, 
                      X  = d3$X.utm, 
                      Y  = d3$Y.utm)
coordinates(MyData)    <- c("X", "Y")
V12 <- variogram(E1 ~ 1, MyData, cressie = TRUE, cutoff = (1000))
plot(V12, 
     main = "", 
     xlab = list(label = "Distance (m)", cex = 1.5), 
     ylab = list(label = "Semi-variogram", cex = 1.5),
       pch = 16,
       col = 1,
       cex = 1.5
       )
with(V12, scatter.smooth(x = dist, 
                        y = gamma,
                        xlab = "Distance",
                        ylab = "Sample variogram"))
```


```{r}
# 1. Plot the Pearson versus their spatial coordinates.
MyData <- data.frame(E1  = E1,
                     Xkm = d3$X.utm,
                     Ykm = d3$Y.utm)

MyData$MySize <- 3 * abs(MyData$E1) / max( MyData$E1)
MyData$MyCol <- ifelse(MyData$E1> 0, 1, 2)
xyplot(Ykm ~ Xkm,
       data = MyData,
       cex = MyData$MySize,
       col = MyData$MyCol,
       pch = 1)

# We seem to have some spatial correlation in the Pearson residuals! We have groups 
# of red dots close to each other. And also groups of black dots.
```


Remove homestead

```{r}
d4 = d2 %>%
  filter(population == "North_Key_Largo" | population =="Key_Largo" | population == "Plantation_Key")
```
```{r}
spatial_data <- st_as_sf(d4,
                         coords = c("long", "lat"),
                         crs = 4326,
                         agr = "constant")

spatial_data <- st_transform(spatial_data, 2236)
plot(spatial_data)
```

```{r}
M1 <- inla(wing2body ~ sex_binom + month_of_year + months_since_start +
                    f(population, model = "iid"),  
                    control.compute = list(dic = TRUE), 
                    family = "gaussian", 
                    data = d4)
```

```{r}
# Pearson residuals vs. fitted values
mu1 <- M1$summary.fitted.values[,"mean"]
E1  <- (d4$wing2body - mu1) / sqrt(mu1)

par(mfrow = c(1,1), mar = c(5,5,2,2), cex.lab = 1.5)
plot(x = mu1, 
     y = E1,
     xlab = "Fitted values",
    ylab = "Pearson residuals")
abline(h = 0, lty = 2)  
```

```{r}
LL <- LongLatToUTM(x = d4$long, y = d4$lat, zone = 16)
d4$X.utm <- LL$X
d4$Y.utm <- LL$Y

xyplot(Y.utm / 1000 ~ X.utm / 1000,
       data = d4,
       aspect = "iso",
       pch = 16)

# What are the distances between the points?
Loc <- cbind(d4$X.utm, d4$Y.utm)
D <- dist(Loc)
par(mfrow = c(1,1), mar = c(5,5,2,2), cex.lab = 1.5)
hist(D, 
     freq = TRUE,
     main = "", 
     xlab = "Distance between sites (m)",
     ylab = "Frequency")
```



```{r}
# 1. Plot the Pearson versus their spatial coordinates.
MyData <- data.frame(E1  = E1,
                     Xkm = d4$X.utm,
                     Ykm = d4$Y.utm)

MyData$MySize <- 3 * abs(MyData$E1) / max( MyData$E1)
MyData$MyCol <- ifelse(MyData$E1> 0, 1, 2)
xyplot(Ykm ~ Xkm,
       data = MyData,
       cex = MyData$MySize,
       col = MyData$MyCol,
       pch = 1)

# We seem to have some spatial correlation in the Pearson residuals! We have groups 
# of red dots close to each other. And also groups of black dots.
```

Will probably need to take means because the layering is obscuring what is happening per site...
