---
title: 'Appendix: Modeling Wing Morphology'
author: "for 'paper_title'; Bernat, AV, Cenzer, ML"
geometry: margin=2cm
output:
  pdf_document:
    includes: null
    toc: yes
    toc_depth: 3
  html_document:
    toc: yes
    toc_depth: '3'
    df_print: paged
line-height: 1.5
fontsize: 11pt
classoption: a4paper
language: en
---

```{r setup, include=FALSE}
rm(list = ls())

# set working directory
dir = "~/Desktop/git_repositories/SBB-dispersal/avbernat_working_on/"
#dir = "~/Documents/GitHub/SBB-dispersal/avbernat_working_on/"
setwd(dir)

knitr::opts_knit$set(root.dir = dir)
knitr::opts_chunk$set(echo = TRUE)
```

\newpage

# Details of the Analyses

This document was generated by R Markdown on `r format(Sys.time(), '%Y-%m-%d')` using R version 4.0.5. The document provides the step-by-step analytical methods used in the manuscript by Anastasia Bernat (AVB) and Meredith Cenzer (MLC). Multiple draft scripts were written by AVB between 2021-03-01 and 2021-07-26 until being distilled and complied by AVB and code reviewed by MLC at the University of Chicago into this comprehensive script. All draft scripts can be viewed in the GitHub repository, SBB-dispersal (https://github.com/mlcenzer/SBB-dispersal), within the directory **avbernat_working_on > All_Morphology > stats**. 

All code and output from the statistical analyses is shown. Code for data cleaning and the generation of plots is not displayed, but can be viewed in the wing_summary.Rmd file and its accompanying three sourced scripts. To repeat analyses and the generation of plots, all data files and sourced scripts should follow the directory structure presented in the SBB-dispersal repository.

## Description of the Data

This document analyzes two main datasets, **raw_data** and **data_long**. The raw_data set provides morphology measurements for each soapberry bug, *Jadera haematoloma*, collected and measured between the years 2013 and 2020. There are four morphology measurements: beak length, thorax width, wing length, and body length. The sex, wing morph (long-winged, shot-winged, or ambiguously-winged), and host plant the bug was collected from as well as the month and year each bug was collected in was recorded. The data_long set provides the same recordings as the raw_data set, but it has been filtered for only long-winged soapberry bugs.

## Abbreviations Used in the Data and Code

* **sd** - standard deviation
* **se** - standard error
* **binom** - a column name that ends in 'binom' is a column that has been recodified into binary data (0's and 1's). Example columns: sex_binom, pophost_binom, and wing_morph_binom
* **pophost** - the host plant soapberry bugs were collected from, which was either *Koelreuteria elegans* or *Cardiospermum corindum*, occasionally called (and abbreviated) as goldenrain tree (GRT) or balloon vine (BV), respectively
* **months_since_start** - proxy for year where the first collection occurred on May 2013
* **month_of_year** - proxy for season where collections occurred only in months February, April, May, August, September, October, and September
* **wing2body** - a computed and unitless column that calculates the wing length divided by the body length of a soapberry bug
* **wing2thorax** - a computed and unitless column that calculates the wing length divided by the thorax width of a soapberry bug
* **S** - short-winged morph
* **L** - long-winged morph
* **LS** or **SL** - ambiguous wing morph

# Data Cleaning And Exploration

## Read Libraries

The occurrence of long-wing morphology and the wing-to-body ratio of *J. haematoloma* were analyzed using multivariate, generalized linear modeling (GLM) as implemented in the R packages `lme4` and `binom`. The `dplyr` package helped speed data manipulation processes by grouping data quickly. All plots, except the histograms, were generated using `ggplot` libraries and helper functions found in R packages `ggformula` and `cowplot`.

Additional R packages not shown below, but embedded in the sourced scripts are `zoo` and `lubridate`, which aid in data manipulation and datetime manipulation, respectively. 

```{r message=FALSE}
library(lme4) # fit regressions 
library(dplyr) # data manipulation 
library(ggformula) # ggplot plotting
library(cowplot) # ggplot helper functions to arrange multi-panel figures
library(binom) # binomial confidence intervals
```

## Read Source Files

Need to explain this: 

model_script = paste0(source_path,"generic models-binomial glm 4-FF.R")
model_comparisonsAIC(model_script)

```{r message=FALSE}
source_path = paste0(dir,"/Rsrc/")

script_names = c("clean_morph_data3.R",  # 1 function: read_morph_data() 
                 "remove_torn_wings.R",  # 1 function: remove_torn_wings()
                 "compare_models.R",     # 1 function: model_comparisonsAIC()
                 "get_Akaike_weights.R") # 1 function: get_model_probs()

for (script in script_names) { 
  path = paste0(source_path, script)
  source(path) 
}
```

## Read the Data

```{r}
data_list = read_morph_data(paste0(dir,"All_Morphology/stats/data/allmorphology05.18.21.csv"))
raw_data = data_list[[1]]
data_long = data_list[[2]] # long-wing bugs only

data_long = remove_torn_wings(data_long)
```
\newpage

## Histograms of Wing Morph Data 

&nbsp;

```{r fig.height=5.7, fig.width=6.6, echo=FALSE}
raw_data_missing = raw_data %>% 
  filter(w_morph=="" | is.na(w_morph)) 

par(mfrow=c(3,1))
hist(raw_data$wing[raw_data$w_morph=="L"]/raw_data$thorax[raw_data$w_morph=="L"],
     main="Histogram of wing length/thorax length for long winged SBB",
     xlab="wing length/thorax length",
     breaks=seq(0.5, 3.8, by=0.05))
hist(raw_data_missing$wing/raw_data_missing$thorax, 
     main="Histogram of wing length/thorax length for SBB w/o recorded wing morph",
     xlab="wing length/thorax length",
     breaks=seq(0.5, 3.8, by=0.05))
hist(raw_data$wing[raw_data$w_morph=="S"]/raw_data$thorax[raw_data$w_morph=="S"],
     main="Histogram of wing length/thorax length for short winged SBB",
     xlab="wing length/thorax length",
     breaks=seq(0.5,3.8,by=0.05))
```

Notice that there are 30 bugs that are hard to identify as either S or L. 

\newpage

## Barplots 

Bugs were collected on different years and months. These barplots show the bugs collected per **population**, **host plant**, and **sex** across the years and months.

```{r echo=FALSE}
# repeating plot features
customPlot = list(geom_bar(position='stack', color="black", width=0.7),
                  theme_classic(),
                  theme(axis.text.y = element_text(size=11),
                        axis.text.x = element_text(size=11, angle = 45, hjust = 1.1),
                        axis.title = element_text(size=17,face="bold")),
                  theme(axis.title.x = element_text(vjust = -3)),
                  theme(axis.title.y = element_text(vjust = 4)),
                  theme(plot.margin=unit(c(1,1,1.1,1.2),"cm"))
                  )
```

### Plot 1: Collection numbers grouped by Population and Datetime

```{r message=FALSE, fig.height=4.1, fig.width=7.2, echo=FALSE}
pop_colors = c("#787a87", "#E69F00", "#56B4E9", "royalblue", 
               "grey", "gold", "#409973", "#9bc969", "ivory2") 

p1 = ggplot(data=subset(raw_data, !is.na(datetime)), aes(x=datetime, fill=population)) + 
  labs(title=" ", fill="Population", 
       x="Field Collection Month", y = "Number of Bugs Collected") +
  scale_fill_manual(values=pop_colors) + customPlot

groups = raw_data %>% 
  group_by(sex, datetime) %>% 
  summarize(count = n())

numF = paste0("F(", as.character(unlist(groups[1:10,3])), ")")
numM = paste0("M(", as.character(unlist(groups[11:20,3])), ")")
text = paste0(numF, "\n", numM)
p1 = p1 + annotate(geom="text", x=1, y=400, label=text[1],
              color="black", size=3) +
  annotate(geom="text", x=2, y=470, label=text[2],
           color="black", size=3) +
  annotate(geom="text", x=3, y=270, label=text[3],
           color="black", size=3) +
  annotate(geom="text", x=4, y=220, label=text[4],
           color="black", size=3) +
  annotate(geom="text", x=5, y=320, label=text[5],
           color="black", size=3) +
  annotate(geom="text", x=6, y=500, label=text[6],
           color="black", size=3) +
  annotate(geom="text", x=7, y=240, label=text[7],
           color="black", size=3) +
  annotate(geom="text", x=8, y=730, label=paste0("\n", text[8]),
           color="black", size=3) +
  annotate(geom="text", x=9, y=440, label=text[9],
           color="black", size=3) +
  annotate(geom="text", x=10, y=650, label=text[10],
           color="black", size=3)
p1
```

### Plot 2: Collection numbers grouped by Host Plant and Datetime

```{r fig.height=4.1, fig.width=7.2, echo=FALSE}
p2 = ggplot(data=subset(raw_data, !is.na(datetime)), aes(x=datetime, fill=pophost)) + 
  labs(title=" ", fill="Host Plant",
              x="Field Collection Month", y = "Number of Bugs Collected") +
  scale_fill_manual(values=c("#56B4E9", "chartreuse4")) + customPlot
p2
```

### Plot 3: Collection numbers grouped by Sex and Datetime

```{r fig.height=4.1, fig.width=7.2, echo=FALSE}
p3 = ggplot(data=subset(raw_data, !is.na(datetime)), aes(x=datetime, fill=sex)) + 
  labs(title=" ", fill="Sex",
              x="Field Collection Month", y = "Number of Bugs Collected") +
  scale_fill_manual(values=c("salmon1", "darkslategray3")) + customPlot
p3
```

&nbsp;

# Regression Modeling

## Long-Wing Morph Frequency

How does sex, host plant, month, and/or year effect whether a soapberry bug is long-winged (wing_morph_binom=1) or short-winged (wing_morph_binom=0)? 

```{r}
data = data.frame(R=raw_data$wing_morph_binom, 
                 A=raw_data$sex_binom, 
                 B=raw_data$pophost_binom, 
                 C=(raw_data$month_of_year),
                 D=raw_data$months_since_start)

model_script = paste0(source_path,"generic models-binomial glm 4-FF.R")
model_comparisonsAIC(model_script)
```

&nbsp;

```{r results = "hold"}
anova(m98, m110, test="Chisq") # adding B*D does not improve fit
anova(m84, m98, test="Chisq") # adding A*B improves fit
anova(m63, m84, test="Chisq") # Adding C*D improves fit
anova(m51, m63, test="Chisq") # Adding B improves fit
```

\newpage

### Best Fit

```{r}
M1 = glm(wing_morph_binom ~ sex_binom * pophost_binom + sex_binom * months_since_start +
           pophost_binom * month_of_year + month_of_year * months_since_start, 
         data=raw_data, family="binomial") 
summary(M1)
```

## Long-Wing Morph Variance

In addition to modeling how sex, host plant, month, or year effects whether a soapberry bug is long-winged or short-winged, we modeled how those same factors affected the variance of wing morph.

```{r}
SE = function(x){sd(x)/sqrt(length(x))}
```

```{r}
wmorph_table = aggregate(wing_morph_binom~sex_binom*pophost_binom*month_of_year*months_since_start, 
                           data=raw_data, FUN=mean)
wmorph_table$sd = aggregate(wing_morph_binom~sex_binom*pophost_binom*month_of_year*months_since_start, 
                              data=raw_data,FUN=sd)$wing_morph_binom
wmorph_table$se = aggregate(wing_morph_binom~sex_binom*pophost_binom*month_of_year*months_since_start, 
                              data=raw_data,FUN=SE)$wing_morph_binom
wmorph_table$n = aggregate(wing_morph_binom~sex_binom*pophost_binom*month_of_year*months_since_start, 
                              data=raw_data,FUN=length)$wing_morph_binom
```

```{r}
data = wmorph_table
data = data.frame(R=data$sd, 
                 A=data$sex_binom, 
                 B=data$pophost_binom, 
                 C=(data$month_of_year),
                 D=data$months_since_start)

model_script = paste0(source_path,"generic models-gaussian glm 4-FF.R")
model_comparisonsAIC(model_script)
```

&nbsp;

```{r results = "hold"}
anova(m2, m5, test="Chisq") # Adding A does not improve fit
anova(m2, m8, test="Chisq") # Adding C does not improve fit
anova(m2, m9, test="Chisq") # Adding D does not improve fit
anova(m0, m2, test="Chisq") # Adding B improves fit
```

### Best Fit

```{r}
M2 = glm(sd ~ pophost_binom, data=wmorph_table, family="gaussian")
summary(M2) 
```

&nbsp;

## Wing-to-Body Ratio 

How does sex, host plant, month, and/or year effect the wing-to-body ratio of long-winged soapberry bugs?

```{r}
data = data.frame(R=data_long$wing2body_c, # centered
                 A=data_long$sex_binom,
                 B=data_long$pophost_binom,
                 C=data_long$month_of_year_c, # centered
                 D=data_long$months_since_start_c) # centered

model_script = paste0(source_path,"generic models-gaussian glm 4-FF.R")
model_comparisonsAIC(model_script)
```

&nbsp;

```{r results = "hold"}
anova(m88, m58, test="Chisq") # Adding A*D marginally improves fit
anova(m76, m58, test="Chisq") # Adding C*D does not improve fit
anova(m88, m99, test="Chisq") # adding C*D does not improve fit
anova(m34, m58, test="Chisq") # Adding B*D improves fit
```

&nbsp;

### Best Fit

```{r}
M3 = glm(wing2body_c ~ sex_binom*pophost_binom + pophost_binom*months_since_start_c 
         + month_of_year_c, data=data_long, family=gaussian) 
summary(M3) 
```

&nbsp;

## Wing-to-Body Ratio Variance 

In addition to modeling how sex, host plant, month, or year effects the wing-to-body ratio of a soapberry bug, we modeled how those same factors affected the variance of wing-to-body ratio.

```{r}
w2b_table = aggregate(wing2body~sex_binom*pophost_binom*month_of_year*months_since_start, 
                     data=data_long, FUN=mean)
w2b_table$sd = aggregate(wing2body~sex_binom*pophost_binom*month_of_year*months_since_start, 
                           data=data_long, FUN=sd)$wing2body
w2b_table$se = aggregate(wing2body~sex_binom*pophost_binom*month_of_year*months_since_start, 
                           data=data_long, FUN=SE)$wing2body
```

```{r}
data = w2b_table
data = data.frame(R=data$sd, 
                 A=data$sex_binom, 
                 B=data$pophost_binom, 
                 C=(data$month_of_year),
                 D=data$months_since_start)

model_script = paste0(source_path,"generic models-gaussian glm 4-FF.R")
model_comparisonsAIC(model_script)
```

&nbsp;

```{r results = "hold"}
anova(m8, m19, test="Chisq") # Adding B*C does not improve fit
anova(m2, m8, test="Chisq") # Adding C does not improve fit
anova(m0, m2, test="Chisq") # Adding B improves fit
```

### Best Fit

```{r}
M4 = glm(sd ~ pophost_binom, data=w2b_table, family=gaussian) 
summary(M4)
```

&nbsp;

# LOESS & Linear Regression Plots

```{r echo=FALSE}
# repeating plotting features
customPlot = list( theme_classic(),
                   theme(axis.text=element_text(size=13),
                         axis.title=element_text(size=16), 
                         plot.title=element_text(size=20),),
                   theme(legend.position = c(0.2, 0.9)),
                   theme(legend.title = element_text(size=14, face="italic"),
                         legend.text = element_text(size = 13, face="italic"))
)

xlab_years = na.omit(sort(unique(data_long$dates))[-2])
xlab_yrs = sort(unique(raw_data$dates))

xlab_allmonths = na.omit(sort(unique(data_long$month_of_year)))
xlab_months = xlab_allmonths[c(-2,-5)]
month_labs = c("Feb", "May", "Aug", "Oct", "Dec")

hp_col_pts = c("turquoise3", "springgreen4")
hp_col_shade = c("turquoise3", "green")
hp_col_lines = c("skyblue2", "darkgreen")
sex_col_pts = c("brown3", "black")
sex_col_shade = c("brown1", "sienna4")
sex_col_lines = c("brown3", "black")

font_size = 4.5
```

## Wing Morph Frequency

### Group significant elements 

```{r}
# function to calculate 95% confidence interval (CI). 
CI_95 = function(x){qnorm(0.975)*sd(x)/sqrt(length(x))}
CI_95_binom_upper = function(y) {binom.confint(x=sum(y, na.rm=TRUE), n=length(y[!is.na(y)]), 
                                                conf.level=0.95, methods='exact')$upper}
CI_95_binom_lower = function(y) {binom.confint(x=sum(y, na.rm=TRUE), n=length(y[!is.na(y)]), 
                                                conf.level=0.95, methods='exact')$lower}
```

```{r}
# aggregate the full data and compute standard error (SE), upper and lower CI, & sample size (n)
w_morph_summary = aggregate(wing_morph_binom~sex*pophost*month_of_year*months_since_start,
                           data=raw_data, FUN=mean)
w_morph_summary$se = aggregate(wing_morph_binom~sex*pophost*month_of_year*months_since_start,
                              data=raw_data, FUN=SE)$wing_morph_binom
w_morph_summary$upper = aggregate(wing_morph_binom~sex*pophost*month_of_year*months_since_start,
                                 data=raw_data, FUN=CI_95_binom_upper)$wing_morph_binom
w_morph_summary$lower = aggregate(wing_morph_binom~sex*pophost*month_of_year*months_since_start,
                                 data=raw_data, FUN=CI_95_binom_lower)$wing_morph_binom
w_morph_summary$n = aggregate(wing_morph_binom~sex*pophost*month_of_year*months_since_start,
                             data=raw_data, FUN=length)$wing_morph_binom

jitter = runif(n=nrow(w_morph_summary), min=-0.1, max=0.1) # jitter points slightly
w_morph_summary$dates = w_morph_summary$month_of_year + jitter

dd = w_morph_summary
```

### Check for LOESS Residuals 

```{r}
plot_lowess_residuals = function(lfit, x, y) {
  lfun = approxfun(lfit)
  fitted = lfun(x)
  resid = y-fitted
  plot(fitted,resid)
  abline(h=0,col=8)
}
```

```{r fig.height=4.5}
l1 = lowess(dd$dates, dd$wing_morph_binom, f=0.4) # f = alpha, the smoother span

par(mfrow=c(1,2))
plot(dd$dates, dd$wing_morph_binom)
lines(l1, type = "l")
plot_lowess_residuals(l1, dd$dates, dd$wing_morph_binom)
```

### Figure: Panels A, B, C, D (long-wing morph freq with month) & E (long-wing morph freq with year)

```{r echo=FALSE}
# relevel factors for cleaner plotting
dd$pophost = factor(dd$pophost, levels = c("K. elegans", "C. corindum") )
dd$`Host Plant` = dd$pophost

# rename columns 
dd$sex[dd$sex=="F"] = "Females"
dd$sex[dd$sex=="M"] = "Males"
dd$sex = factor(dd$sex, levels = c("Males", "Females") )
dd$`Sex` = dd$sex 

raw_data$'Host Plant'=raw_data$pophost
raw_data$Sex=NA
raw_data$Sex[raw_data$sex=="F"] = "Females"
raw_data$Sex[raw_data$sex=="M"] = "Males"
```

Note: Single-variate models or simpler models substituted best fit slope calculations below in order to avoid multiple interaction terms found in the best fit model. This led to cleaner glm line plotting.

\begin{center}
 \textbf{Panels A and B Regression Computations:}
\end{center}

```{r}
# single-variate model of month predicting wing morph
fit = glm(wing_morph_binom ~ month_of_year, family="binomial", data=raw_data)
xmonth = seq(2,12, 0.01)
wing_probs = predict(fit, list(month_of_year=xmonth), type="response")

# extract pvalue from best fit regression model
fit_pvalue = round(summary(M1)$coeff[,"Pr(>|z|)"][5],5)
pvalue = paste0("italic(p)[glm]==", fit_pvalue)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
### Panels A and B

### host plant
p4 = ggplot() + 
  ggtitle("A\n") + xlab("Month") + ylab("Long-Wing Morph Frequency") +
  geom_vline(xintercept = xlab_allmonths, color="gainsboro") + 
  geom_line(aes(x=xmonth, y=wing_probs), linetype = "dashed") +
  geom_smooth(data=raw_data[!is.na(raw_data$wing_morph_binom),], method="loess",
              mapping = aes(x = month_of_year, y = wing_morph_binom, colour=`Host Plant`, fill=`Host Plant`)) +
  geom_point(data=dd, 
             mapping = aes(x = month_of_year, y = wing_morph_binom, colour=`Host Plant`, size=n), show.legend=F) +
  customPlot + 
  scale_color_manual(values=c("C. corindum" = hp_col_pts[1], "K. elegans" = hp_col_pts[2])) +
  scale_fill_manual(values = c("C. corindum" = hp_col_shade[1], "K. elegans" = hp_col_shade[2])) +
  scale_x_continuous(breaks=xlab_months, labels= month_labs) + 
  ylim(0,1.2) 

### sex
p5 = ggplot() + 
  ggtitle("B\n") + xlab("Month") + ylab("Long-Wing Morph Frequency") +
  geom_vline(xintercept = xlab_allmonths, color="gainsboro") + 
  geom_line(aes(x=xmonth, y=wing_probs), linetype = "dashed") +
  geom_smooth(data=raw_data[!is.na(raw_data$wing_morph_binom),], method="loess",
              mapping = aes(x = month_of_year, y = wing_morph_binom, colour=`Sex`, fill=`Sex`)) + 
  geom_point(data=dd, 
             mapping = aes(x = month_of_year, y = wing_morph_binom, colour=Sex, size=n), show.legend=F) +
  customPlot +
  scale_color_manual(values=c("Females" = sex_col_pts[1], "Males" = sex_col_pts[2])) +
  scale_fill_manual(values = c("Females" = sex_col_shade[1], "Males" = sex_col_shade[2])) +
  scale_x_continuous(breaks=xlab_months, labels= month_labs) + 
  ylim(0,1.2) + 
  theme(axis.title.y = element_blank()) 

# extract alpha, smoothing parameter, from LOESS regression(s)
alpha = paste("alpha[loess]==", ggplot_build(p5)$data[[3]]$alpha[1])
degree="lambda[loess]==0"

p4 = p4 + 
  annotate(geom="text", x=7.2, y=0.92, label=alpha, parse=TRUE, size=font_size) +
  annotate(geom="text", x=7.2, y=0.99, label=degree, parse=TRUE, size=font_size) +
  annotate(geom="text", x=5.5, y=0.718, label=pvalue, parse=TRUE, size=font_size) + 
  theme(legend.position = c(0.18, 1.02)) + 
  labs(x = " ", y = " ") +
  coord_cartesian(ylim=c(0.30,1.1))

p5 = p5 + 
  annotate(geom="text", x=7.2, y=0.98, label=alpha, parse=TRUE, size=font_size) +
  annotate(geom="text", x=7.2, y=1.05, label=degree, parse=TRUE, size=font_size) +
  annotate(geom="text", x=6.5, y=0.85, label=pvalue, parse=TRUE, size=font_size) + 
  theme(legend.position = c(0.15, 1.02)) + 
  labs(x = " ", y = " ") +
  coord_cartesian(ylim=c(0.30,1.1))
```

```{r echo=FALSE}
# group data by sex 
dfF = dd[dd$sex=="Females",]
dfM = dd[dd$sex=="Males",]
```

\begin{center}
  \textbf{Panels C and D Regression Computations:}
\end{center}

```{r message=FALSE}
# multi-variate model with month, sex, and host plant predicting wing morph
fit2 = glm(wing_morph_binom ~ sex_binom * pophost_binom + 
          pophost_binom * month_of_year, family = "binomial", data = raw_data)
xmon = seq(2,12, 0.01)
set.seed(194842)
bsex = sample(c(-1,1), replace=TRUE, size=length(xmon))
bhost = sample(c(-1,1), replace=TRUE, size=length(xmon))
wprobs = predict(fit2, list(sex_binom = bsex,
                                pophost_binom = bhost,
                                month_of_year = xmon), type="response")

pred = cbind(xmon, bsex, bhost, wprobs)
pred = as.data.frame(pred)

predFK = pred[pred$bhost==1 & pred$bsex==1,]
predFC = pred[pred$bhost==-1 & pred$bsex==1,]

predMK = pred[pred$bhost==1 & pred$bsex==-1,]
predMC = pred[pred$bhost==-1 & pred$bsex==-1,]

# extract pvalue from best fit regression model
fit_pvalue = round(summary(M1)$coeff[,"Pr(>|z|)"][6],4)
pvalue = paste0("italic(p)[glm]==", fit_pvalue)
```

```{r echo=FALSE, message=FALSE}
### Panels C and D

### females & host plant
p6 = ggplot() + 
  ggtitle("C") + xlab("Month") + ylab("Long-Wing Morph Frequency") +
  geom_vline(xintercept = xlab_allmonths, color="gainsboro") + 
  geom_line(aes(x=predFK$xmon, y=predFK$wprobs), linetype = "dashed", color=hp_col_lines[2]) +
  geom_line(aes(x=predFC$xmon, y=predFC$wprobs), linetype = "dashed", color=hp_col_lines[1]) +
  geom_smooth(data=raw_data[raw_data$sex=="F" & !is.na(raw_data$wing_morph_binom),], method="loess", span=0.8,
              mapping = aes(x = month_of_year, y = wing_morph_binom, colour=`Host Plant`, fill=`Host Plant`)) +
  geom_point(data=dfF, mapping = aes(x = month_of_year, y = wing_morph_binom, colour=`Host Plant`, size=n),
             show.legend=F) +
  customPlot + 
  scale_color_manual(values=c("C. corindum" = hp_col_pts[1], "K. elegans" = hp_col_pts[2])) +
  scale_fill_manual(values = c("C. corindum" = hp_col_shade[1], "K. elegans" = hp_col_shade[2])) +
  scale_x_continuous(breaks=xlab_months, labels= month_labs) +
  ylim(0,1.3)

### males & host plant
p7 = ggplot() + 
  ggtitle("D") + xlab("Month") + ylab("Long-Wing Morph Frequency") +
  geom_vline(xintercept = xlab_allmonths, color="gainsboro") + 
  geom_line(aes(x=predMK$xmon, y=predMK$wprobs), linetype = "dashed", color=hp_col_lines[2]) +
  geom_line(aes(x=predMC$xmon, y=predMC$wprobs), linetype = "dashed", color=hp_col_lines[1]) +
  geom_smooth(data=raw_data[raw_data$sex=="M" & !is.na(raw_data$wing_morph_binom),], method="loess", span=0.8,
              mapping = aes(x = month_of_year, y = wing_morph_binom, colour=`pophost`, fill=`pophost`),
              show.legend=F) + 
  geom_point(data=dfM, mapping = aes(x = month_of_year, y = wing_morph_binom, colour=`Host Plant`, size=n)) +
  guides(color=F) +
  customPlot + 
  scale_color_manual(values=c("C. corindum" = hp_col_pts[1], "K. elegans" = hp_col_pts[2])) +
  scale_fill_manual(values = c("C. corindum" = hp_col_shade[1], "K. elegans" = hp_col_shade[2])) +
  scale_x_continuous(breaks=xlab_months, labels= month_labs) +
  ylim(0,1.3)

# extract alpha, smoothing parameter from LOESS regression(s)
alpha = paste("alpha[loess]==", ggplot_build(p7)$data[[4]]$alpha[1])
degree="lambda[loess]==0"

p6 = p6 + 
  annotate(geom="text", x=3, y=1.07, label="Females", size=6, fontface = 'italic') + 
  annotate(geom="text", x=6.5, y=0.99, label=alpha, parse=TRUE, size=font_size) +
  annotate(geom="text", x=6.5, y=1.06, label=degree, parse=TRUE, size=font_size) +
  annotate(geom="text", x=7.5, y=0.40, label=pvalue, parse=TRUE, size=font_size) + 
  theme(legend.position = "none") + 
  labs(x = " ", y = " ") +
  coord_cartesian(ylim=c(0.30,1.1))

p7 = p7 + 
  annotate(geom="text", x=3, y=1.07, label="Males", size=6, fontface = 'italic') + 
  annotate(geom="text", x=6.5, y=0.99, label=alpha, parse=TRUE, size=font_size) +
  annotate(geom="text", x=6.5, y=1.06, label=degree,parse=TRUE, size=font_size) +
  annotate(geom="text", x=7.5, y=0.40, label=pvalue, parse=TRUE, size=font_size) +
  #theme(legend.position = "none") + 
  theme(legend.position = c(0.87, 0.25)) + 
  theme(axis.title.y = element_blank()) + 
  labs(x = " ", y = " ") +
  coord_cartesian(ylim=c(0.30,1.1))
```

```{r echo=FALSE}
# group data by sex 
wm_table = aggregate(wing_morph_binom~dates*sex, data=raw_data, FUN=mean)
wm_table$n = aggregate(wing_morph_binom~dates*sex, data=raw_data, FUN=length)$wing_morph_binom
dd = wm_table

dd$sex[dd$sex=="F"] = "Females"
dd$sex[dd$sex=="M"] = "Males"
dd$sex = factor(dd$sex, levels = c("Males", "Females") )
dd$`Sex` = dd$sex 

dF = dd[dd$sex=="Females",]
dM = dd[dd$sex=="Males",]
```

\begin{center}
  \textbf{Panel F Regression Computations:}
\end{center}

```{r}
# multi-variate model with year, sex, and host plant predicting wing morph
fit3 = glm(wing_morph_binom ~ sex_binom * dates, family = "binomial", data = raw_data)
xyr = seq(sort(unique(dd$dates))[1],sort(unique(dd$dates))[10], 1)
set.seed(194842)
bsex = sample(c(-1,1), replace=TRUE, size=length(xyr))
bhost = sample(c(-1,1), replace=TRUE, size=length(xyr))
wprobs = predict(fit3, list(sex_binom = bsex,
                                pophost_binom = bhost,
                                dates = xyr), type="response")

pred = cbind(xyr, bsex, bhost, wprobs)
pred = as.data.frame(pred)
pred$xyr = as.Date.numeric(pred$xyr)

predF = pred[pred$bsex==1,]
predM = pred[pred$bsex==-1,]

# extract pvalue from best fit regression model
fit_pvalue = round(summary(M1)$coeff[,"Pr(>|z|)"][7],3)
pvalue = paste0("italic(p)[glm]==", fit_pvalue)
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
### Panel F

p9 = ggplot() + 
  ggtitle("F") + xlab("Year") + ylab("Long-Wing Morph Frequency") +
  geom_vline(xintercept = xlab_yrs, color="gainsboro") + 
  geom_line(aes(x=predF$xyr, y=predF$wprobs), linetype = "dashed", color=sex_col_lines[1], lwd=0.5) +
  geom_line(aes(x=predM$xyr, y=predM$wprobs), linetype = "dashed", color=sex_col_lines[2], lwd=0.5) +
  geom_smooth(data=raw_data[!is.na(raw_data$wing_morph_binom),], method="loess", span=0.8,
              mapping = aes(x = dates, y = wing_morph_binom, colour=`Sex`, fill=`Sex`)) + 
  geom_point(data=dd, mapping = aes(x = dates, y = wing_morph_binom, colour=`Sex`, size=n),
             show.legend=F) +
  customPlot + 
  scale_color_manual(values=c("Females" = sex_col_pts[1], "Males" = sex_col_pts[2])) +
  scale_fill_manual(values = c("Females" = sex_col_shade[1], "Males" = sex_col_shade[2])) +
  ylim(0,1.3) + # ylim bounds must include must all data points or points be marked as 'non-finite'
  coord_cartesian(ylim=c(0.30,1.1)) +
  theme(legend.position = "none") 

p9 = p9 +
  annotate(geom="text", x=xlab_yrs[5], y=0.95, label=alpha, parse=TRUE, size=font_size) +
  annotate(geom="text", x=xlab_yrs[5], y=1.02, label=degree, parse=TRUE, size=font_size) +
  annotate(geom="text", x=xlab_yrs[4], y=0.59, label=pvalue, parse=TRUE, size=font_size)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=13.5, fig.width=11}
customGRID1 = list(draw_plot(p4, 0, .66, .5, .33),
                   draw_plot(p5, 0.5, .66, .5, .33),
                   draw_plot(p6, 0, .33, .5, .33),
                   draw_plot(p7, 0.5, .33, .5, .33),
                   draw_plot(p9, 0, 0, 1, .33),
                   draw_plot_label("Month", 0.47, 0.338, fontface="plain"),
                   draw_plot_label("Long-Wing Morph Frequency", 0, 0.46, fontface="plain", angle=90)
                  )

paper_figure1 = ggdraw() +
  customGRID1
paper_figure1
```

```{r eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE, fig.height=10, fig.width=10}

p6B = p6 + theme(legend.position = c(0.83, 0.96)) + ggtitle("A")
p7B = p7 + ggtitle("B")
p9B = p9 + theme(legend.position = c(0.83, 0.96)) + ggtitle("C")

customGRID2 = list(draw_plot(p6B, 0, .5, .5, .5),
                  draw_plot(p7B, 0.5, .5, .5, .5),
                  draw_plot(p9B, 0, 0, 1, .5),
                  draw_plot_label("Month", 0.47, 0.52, fontface="plain"),
                  draw_plot_label("Long-Wing Morph Frequency", 0, 0.46, fontface="plain", angle=90)
                  )

paper_figure1 = ggdraw() +
  customGRID2
paper_figure1
```


&nbsp;

\newpage

## Wing-to-Body Ratio

### Group significant elements 

```{r}
w2b_summary = aggregate(wing2body~sex*pophost*dates*month_of_year, data=data_long, FUN=mean)
w2b_summary$se = aggregate(wing2body~sex*pophost*dates, data=data_long,
                          FUN=SE)$wing2body
w2b_summary$n = aggregate(wing2body~sex*pophost*dates, data=data_long,
                          FUN=length)$wing2body

jitter = runif(n=nrow(w2b_summary), min=-0.1, max=0.1) # jitter slightly
w2b_summary$dates = w2b_summary$dates + jitter
d = w2b_summary
```

### Check for LOESS Residuals 

```{r fig.height=4.5}
l1 = lowess(d$dates, d$wing2body, f=0.4)

par(mfrow=c(1,2))
plot(d$dates, d$wing2body)
lines(l1, type = "l")
plot_lowess_residuals(l1, d$dates, d$wing2body)
```

### Figure: Panels A, B, (wing-to-body ratio with month) & C (wing-to-body ratio with year)

```{r echo=FALSE, message=FALSE, warning=FALSE}
# relevel and rename factors for cleaner plotting
d$pophost = factor(d$pophost, levels = c("K. elegans", "C. corindum") )
d$`Host Plant` = d$pophost

d$sex[d$sex=="F"] = "Females"
d$sex[d$sex=="M"] = "Males"
d$sex = factor(d$sex, levels = c("Males", "Females") )
d$`Sex` = d$sex 

data_long$'Host Plant' = factor(data_long$pophost, levels = c("K. elegans", "C. corindum") )
data_long$Sex[data_long$sex=="F"] = "Females"
data_long$Sex[data_long$sex=="M"] = "Males"
data_long$Sex = factor(data_long$Sex, levels = c("Males", "Females") )
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
### Panels A and B

### host plant
p10 = ggplot() + 
  ggtitle("A") + xlab("Month") + ylab(" ") +
  geom_vline(xintercept = xlab_allmonths, color="gainsboro") + 
  geom_smooth(data=data_long, method="glm", se=FALSE, linetype = "dashed", 
              mapping = aes(x = month_of_year, y = wing2body), colour="black", lwd=0.5) + 
  geom_smooth(data=data_long, method="loess", 
              mapping = aes(x = month_of_year, y = wing2body, colour=`Host Plant`, fill=`Host Plant`)) + 
  geom_point(data=d, mapping = aes(x = month_of_year, y = wing2body, colour=`Host Plant`, size=n), show.legend=F) + 
  ylim(0.65, 0.85) + # data outside this range will be dropped
  coord_cartesian(ylim=c(0.705,0.75)) + 
  customPlot +
  scale_color_manual(values=c("C. corindum" = hp_col_pts[1], "K. elegans" = hp_col_pts[2])) +
  scale_fill_manual(values = c("C. corindum" = hp_col_shade[1], "K. elegans" = hp_col_shade[2])) +
  scale_x_continuous(breaks=xlab_months, labels=month_labs) + 
  theme(axis.title.x = element_blank())

### sex
p11 = ggplot() + 
  ggtitle("B") + xlab("Month") + ylab(" ") + 
  geom_vline(xintercept = xlab_allmonths, color="gainsboro") + 
  geom_smooth(data=data_long, method="glm", se=FALSE, linetype = "dashed",  
              mapping = aes(x = month_of_year, y = wing2body), colour="black", lwd=0.5) + 
  geom_smooth(data=data_long, method="loess",
              mapping = aes(x = month_of_year, y = wing2body, colour=Sex, fill=Sex)) + 
  geom_point(data=d, mapping = aes(x = month_of_year, y = wing2body, colour=Sex, size=n), show.legend=F) +
  ylim(0.65, 0.85) +
  coord_cartesian(ylim=c(0.705,0.75)) + 
  customPlot +
  scale_color_manual(values=c("Females" = sex_col_pts[1], "Males" = sex_col_pts[2])) +
  scale_fill_manual(values = c("Females" = sex_col_shade[1], "Males" = sex_col_shade[2])) +
  scale_x_continuous(breaks=xlab_months, labels= month_labs) + 
  theme(axis.title.x = element_blank())

# extract alpha, smoothing parameter from LOESS regression(s) and pvalue from linear model
alpha = paste("alpha[loess]==", ggplot_build(p10)$data[[3]]$alpha[1])
degree="lambda[loess]==0"
fit_pvalue = round(summary(M3)$coeff[,"Pr(>|t|)"][5],7)
pvalue = paste0("italic(p)[glm]==", fit_pvalue)

p10 = p10 +
  annotate(geom="text", x=10, y=0.745, label=alpha, parse=TRUE, size=font_size) +
  annotate(geom="text", x=10, y=0.749, label=degree, parse=TRUE, size=font_size) +
  annotate(geom="text", x=5, y=0.715, label=pvalue, parse=TRUE, size=font_size)

p11 = p11 +
  annotate(geom="text", x=10, y=0.745, label=alpha, parse=TRUE, size=font_size) +
  annotate(geom="text", x=10, y=0.749, label=degree, parse=TRUE, size=font_size) +
  annotate(geom="text", x=5, y=0.715, label=pvalue, parse=TRUE, size=font_size)
```

\begin{center}
 \textbf{Panel C Regression Computations:}
\end{center}

```{r}
# multi-variate model with year, sex, and host plant predicting wing2body ratio
fit4 = glm(wing2body ~ pophost_binom * dates, data = data_long)
xyr = seq(sort(unique(dd$dates))[1],sort(unique(dd$dates))[10], 1)
set.seed(194842)
bhost = sample(c(-1,1), replace=TRUE, size=length(xyr))
wprobs = predict(fit4, list(pophost_binom = bhost,
                                dates = xyr), type="response")

pred = cbind(xyr, bhost, wprobs)
pred = as.data.frame(pred)
pred$xyr = as.Date.numeric(pred$xyr)

predK = pred[pred$bhost==1,]
predC = pred[pred$bhost==-1,]

# extract pvalue from best fit regression model
fit_pvalue = round(summary(M3)$coeff[,"Pr(>|t|)"][7],3)
pvalue = paste0("italic(p)[glm]==", fit_pvalue)
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
### Panel C

p12 = ggplot() + 
  ggtitle("C") + xlab("Year") + ylab(" ") +
  geom_vline(xintercept = xlab_years, color="gainsboro") + 
  geom_line(aes(x=predK$xyr, y=predK$wprobs), linetype = "dashed", color=hp_col_lines[2], lwd=0.5) +
  geom_line(aes(x=predC$xyr, y=predC$wprobs), linetype = "dashed", color=hp_col_lines[1], lwd=0.5) +
  geom_smooth(data=data_long, method="loess", span=0.8,
              mapping = aes(x = dates, y = wing2body, colour=`Host Plant`, fill=`Host Plant`),
              show.legend=F) + 
  geom_point(data=d, mapping = aes(x = dates, y = wing2body, colour=`Host Plant`, size=n)) +
  guides(color=F) + 
  customPlot +
  scale_color_manual(values=c("C. corindum" = hp_col_pts[1], "K. elegans" = hp_col_pts[2])) +
  scale_fill_manual(values = c("C. corindum" = hp_col_shade[1], "K. elegans" = hp_col_shade[2])) +
  ylim(0.65, 0.85) +
  theme(legend.position = c(0.1, 0.9)) +
  #theme(legend.position = "none") +
  coord_cartesian(ylim=c(0.705,0.75))

#### loess and linear regressions
alpha = paste("alpha[loess]==", ggplot_build(p12)$data[[4]]$alpha[1])
degree="lambda[loess]==0"

p12 = p12 +
  annotate(geom="text", x=unique(d$dates)[17]-330, y=0.743, label=alpha, parse=TRUE, size=font_size) +
  annotate(geom="text", x=unique(d$dates)[17]-330, y=0.747, label=degree, parse=TRUE, size=font_size) +
  annotate(geom="text", x=unique(d$dates)[15], y=0.716, label=pvalue, parse=TRUE, size=font_size)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=10, fig.width=10}
paper_figure2 = ggdraw() +
  draw_plot(p10, 0, .5, .5, .5) +
  draw_plot(p11, .5, .5, .5, .5) +
  draw_plot(p12, 0, 0, 1, .5) +  
  draw_plot_label("Month", 0.47, 0.51, fontface="plain") +
  draw_plot_label("Wing-to-Body Ratio", 0, 0.35, fontface="plain", angle=90)
paper_figure2
```

