---
title: "All Morphology: Wing LOESS"
author: "Anastasia Bernat"
date: "4/8/2021"
output:
  html_document:
    theme: united
---

```{r setup, include=FALSE}
rm(list=ls())
library(ggformula)
library(splines)
library(gridExtra) #grid.arrange()
library(ggpmisc) #stat_fit_glance
library(cowplot)

dir = "~/Desktop/git_repositories/SBB-dispersal/avbernat_working_on/All_Morphology/stats/"
setwd(dir)

knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#knitr::purl("wing_loess.Rmd", output = "wing_loess.R") # convert Rmd to R script
```

# {.tabset}

## Cleaning Data and Sourcing Scripts

```{r}
source_path = "~/Desktop/git_repositories/SBB-dispersal/avbernat_working_on/Rsrc/"

script_names = c("compare_models.R","regression_output.R", "clean_morph_data3.R", "AICprobabilities.R")

for (script in script_names) { 
  path = paste0(source_path, script)
  source(path) 
}
```

### Read the data

```{r warning=FALSE}
data_list <- read_morph_data("data/allmorphology05.10.21-coors.csv")
raw_data = data_list[[1]]
data_long = data_list[[2]]
all_bugs = nrow(data_long)

data_long = remove_torn_wings(data_long)
```

## Wing2Body 

```{r}
SE = function(x){sd(x)/sqrt(length(x))}
w2b_summary<-aggregate(wing2body~sex*pophost*dates, data=data_long, FUN=mean)
w2b_summary$se<-aggregate(wing2body~sex*pophost*dates, data=data_long,
                          FUN=SE)$wing2body

#jitter slightly
jitter = runif(n=nrow(w2b_summary), min=-0.5, max=0.5)
w2b_summary$dates <- w2b_summary$dates + jitter
```

```{r}
# check for residuals for loess
l1 = lowess(w2b_summary$dates, w2b_summary$wing2body, f=0.4)
plot(w2b_summary$dates, w2b_summary$wing2body)
lines(l1, type = "l")

plot_lowess_residuals = function(lfit, x, y) {
  lfun <- approxfun(lfit)
  fitted <- lfun(x)
  resid <- y-fitted
  plot(fitted,resid)
  abline(h=0,col=8)
}

plot_lowess_residuals(l1, w2b_summary$dates, w2b_summary$wing2body)
```

```{r}
d = w2b_summary

# general plotting features
xlab_years = na.omit(sort(unique(data_long$dates))[-2])
host_colors_shade = c("turquoise3", "green")
host_colors_pts = c("turquoise3", "springgreen4")

# date plotting features
threshold = unique(d$dates)[13]
threshold
d$date_b = "2013-2014"
d$date_b[d$dates >= threshold] = "2015-2020"

d$date_b = as.factor(d$date_b)

w2b_table<-aggregate(wing2body~dates, data=data_long, FUN=mean)
w2b_table$date_b = "2013-2014"
w2b_table$date_b[w2b_table$dates >= threshold] = "2015-2020"
w2b_table$date_b = as.factor(w2b_table$date_b)

last_n_years = w2b_table[w2b_table$date_b=="2015-2020",]
last_n_years
```

```{r}
dfF = d[d$sex=="F",]
dfM = d[d$sex=="M",]

dfF$pophost[dfF$pophost=="C.corindum"]<-"C. corindum"
dfF$pophost[dfF$pophost=="K.elegans"]<-"K. elegans"
dfM$pophost[dfM$pophost=="C.corindum"]<-"C. corindum"
dfM$pophost[dfM$pophost=="K.elegans"]<-"K. elegans"

dfF$pophost = factor(dfF$pophost, levels = c("K. elegans", "C. corindum") )
dfF$`Host Plant` = dfF$pophost
dfM$pophost = factor(dfM$pophost, levels = c("K. elegans", "C. corindum") )
dfM$`Host Plant` = dfM$pophost
```


```{r fig.width=4.7, fig.height=2.15}
p0 = ggplot() + 
  ggtitle("C") + xlab("Year") + ylab("Wing-to-Body Ratio") +
  geom_vline(xintercept = xlab_years, color="gainsboro") + 
   geom_smooth(data=w2b_table, method="lm", se=FALSE, linetype = "dashed",
              mapping = aes(x = dates, y = wing2body), colour="black", lwd=0.5) +
  geom_smooth(data=w2b_table, method="loess", 
              mapping = aes(x = dates, y = wing2body), colour="black") + 
  geom_point(data=w2b_table, mapping = aes(x = dates, y = wing2body)) +
  ylim(0.71, 0.75) +
  stat_smooth(data=last_n_years, mapping = aes(x = dates, y = wing2body), method = "lm", formula = y ~ x, se = FALSE,
               colour="blue") # linetype = "longdash",

p0 = p0 + guides(fill = guide_legend(reverse = TRUE)) + theme_classic() +
  theme(axis.text=element_text(size=13),
        axis.title=element_text(size=16), 
        plot.title=element_text(size=20)) + guides(color = FALSE)  + scale_linetype(guide = FALSE) +
  theme(legend.key = element_rect(fill = "white", color = NA), 
        legend.key.size = unit(0.5, "cm"),
        legend.key.width = unit(0.5,"cm")) + 
  scale_fill_manual(values = "blue", labels=c("C. corindum", "K. elegans")) + 
  labs(fill = "Host Plant") + theme(legend.title = element_text(size = 14),
                                    legend.text = element_text(size = 13, face="italic")) +
  scale_color_manual(values="blue") + theme(legend.position = c(0.2, 0.88)) + theme(legend.title = element_blank())

# loess 
alpha = paste("alpha[loess]==", ggplot_build(p0)$data[[2]]$alpha[1])
degree="lambda[loess]==0"
mlinear = glm(wing2body ~ dates, data=d, family=gaussian)
pvalue = paste0("italic(p)[glm]==",round(summary(mlinear)$coeff[[8]],2))

p0 = p0 + annotate(geom="text", x=unique(d$dates)[10], y=0.74, label=alpha, color="black", parse=TRUE, size=6) +
          annotate(geom="text", x=unique(d$dates)[10], y=0.748, label=degree, color="black",parse=TRUE, size=6) +
          annotate(geom="text", x=unique(d$dates)[15], y=0.718, label=pvalue, color="black", parse=TRUE, size=6)

# last 5 years
mllinear = glm(wing2body ~ dates, data=w2b_table[w2b_table$date_b=="2015-2020",], family=gaussian)
pval = paste0("italic(p)[glm]==",round(summary(mllinear)$coeff[[8]],2))
summary(mllinear)

p0 = p0 + annotate(geom="text", x=unique(d$dates)[24], y=0.743, label=pval, color="blue", parse=TRUE, size=6)
p0
```
```{r}
head(ggplot_build(p0)$data)
```


```{r}
w2b_summary<-aggregate(wing2body~sex*pophost*month_of_year, data=data_long, FUN=mean)
df = w2b_summary

xlab_dates = na.omit(sort(unique(data_long$month_of_year)))
xlab_months = xlab_dates[c(-2,-5)]
month_labs <- c("Feb", "May", "Aug", "Oct", "Dec")

sex_colors_shade = c("brown1", "sienna4")
sex_colors_pts = c("brown1", "grey27")
```

```{r}
df$pophost[df$pophost=="C.corindum"]<-"C. corindum"
df$pophost[df$pophost=="K.elegans"]<-"K. elegans"
df$pophost = factor(df$pophost, levels = c("K. elegans", "C. corindum") )

df$`Host Plant` = df$pophost

df$sex[df$sex=="F"]<-"Females"
df$sex[df$sex=="M"]<-"Males"
df$sex = factor(df$sex, levels = c("Males", "Females") )
df$`Sex` = df$sex 
```

```{r}
customPlot = list( theme_classic(),
                   theme(axis.text=element_text(size=13),
                      axis.title=element_text(size=16), 
                      plot.title=element_text(size=20),),
                   theme(legend.position = c(0.2, 0.9)),
                   theme(legend.title = element_text(size=14, face="italic"),
                         legend.text = element_text(size = 13, face="italic"))
                   )
```

```{r fig.width=4.7, fig.height=2.3}
# Females
p1 = ggplot() + 
  ggtitle("Females") + xlab("Year") + ylab("Wing-to-Body Ratio") +                          # title and labels
  geom_vline(xintercept = xlab_years, color="gainsboro") +                                  # add grey lines for the each collection date
  geom_smooth(data=dfF, method="glm", se=FALSE, linetype = "dashed",                        # GLM regression and pvalue
              mapping = aes(x = dates, y = wing2body), colour="black", lwd=0.5) + 
  geom_smooth(data=dfF, method="loess",                                                     # LOESS regression
              mapping = aes(x = dates, y = wing2body, colour=`Host Plant`, fill=`Host Plant`)) + 
  geom_point(data=dfF, mapping = aes(x = dates, y = wing2body, colour=`Host Plant`)) +   # data points and yaxis value range
  ylim(0.70, 0.76) +
  customPlot +
  scale_color_manual(values=c("C. corindum" = "turquoise3", "K. elegans" = "springgreen4")) +
  scale_fill_manual(values = c("C. corindum" = "turquoise3", "K. elegans" = "green"))

# Males
p2 = ggplot() + 
  ggtitle("Males") + xlab("Year") + ylab("Wing-to-Body Ratio") +                            # title and labels
  geom_vline(xintercept = xlab_years, color="gainsboro") +                                  # add grey lines for the each collection date
  geom_smooth(data=dfM, method="glm", se=FALSE, linetype = "dashed",                        # GLM regression and pvalue
              mapping = aes(x = dates, y = wing2body), colour="black", lwd=0.5) + 
  geom_smooth(data=dfM, method="loess",                                                     # LOESS regression
              mapping = aes(x = dates, y = wing2body, colour=`Host Plant`, fill=`Host Plant`)) + 
  geom_point(data=dfM, mapping = aes(x = dates, y = wing2body, colour=`Host Plant`)) +   # data points and yaxis value range
  ylim(0.70, 0.76) +
  customPlot +
  scale_color_manual(values=c("C. corindum" = "turquoise3", "K. elegans" = "springgreen4")) +
  scale_fill_manual(values = c("C. corindum" = "turquoise3", "K. elegans" = "green"))

grid.arrange(p1, p2, ncol=2)
```

Now do it by season because that's more interesting...

```{r fig.width=4.7, fig.height=2.3}
p5 = ggplot() + 
  ggtitle("A") + xlab("Month") + ylab("Wing-to-Body Ratio") +                              # title and labels
  geom_vline(xintercept = xlab_dates, color="gainsboro") +                                          # add grey lines for the each collection date
  geom_smooth(data=df, method="glm", se=FALSE, linetype = "dashed",                        # GLM regression and pvalue
              mapping = aes(x = month_of_year, y = wing2body), colour="black", lwd=0.5) + 
  geom_smooth(data=df, method="loess",                                                     # LOESS regression
              mapping = aes(x = month_of_year, y = wing2body, colour=`Host Plant`, fill=`Host Plant`)) + 
  geom_point(data=df, mapping = aes(x = month_of_year, y = wing2body, colour=`Host Plant`)) +   # data points and yaxis value range
  ylim(0.71, 0.75) +
  customPlot +
  scale_color_manual(values=c("C. corindum" = "turquoise3", "K. elegans" = "springgreen4")) +
  scale_fill_manual(values = c("C. corindum" = "turquoise3", "K. elegans" = "green")) +
  scale_x_continuous(breaks=xlab_months, labels= month_labs)

p6 = ggplot() + 
  ggtitle("B") + xlab("Month") + ylab("Wing-to-Body Ratio") +                             # title and labels
  geom_vline(xintercept = xlab_dates, color="gainsboro") +                                # add grey lines for the each collection date
  geom_smooth(data=df, method="glm", se=FALSE, linetype = "dashed",                        # GLM regression and pvalue
              mapping = aes(x = month_of_year, y = wing2body), colour="black", lwd=0.5) + 
  geom_smooth(data=df, method="loess",                                                     # LOESS regression
              mapping = aes(x = month_of_year, y = wing2body, colour=Sex, fill=Sex)) + 
  geom_point(data=df, mapping = aes(x = month_of_year, y = wing2body, colour=Sex)) +   # data points and yaxis value range
  ylim(0.71, 0.75) +
  customPlot +
  scale_color_manual(values=c("Females" = "brown3", "Males" = "black")) +
  scale_fill_manual(values = c("Females" = "brown1", "Males" = "sienna4")) +
  scale_x_continuous(breaks=xlab_months, labels= month_labs) + theme(axis.title.y = element_blank())


# extract the LOESS
alpha = paste("alpha[loess]==", ggplot_build(p5)$data[[2]]$alpha[1])
degree="lambda[loess]==0"

mlinear = glm(wing2body ~ month_of_year, data=df, family=gaussian)
#round(summary(mlinear)$coeff[[8]],3) 
pvalue = paste0("italic(p)[glm]==",round(summary(mlinear)$coeff[[8]],3))

p5 = p5 + annotate(geom="text", x=10, y=0.744, label=alpha, color="black", parse=TRUE) +
          annotate(geom="text", x=10, y=0.747, label=degree, color="black", parse=TRUE) +
          annotate(geom="text", x=6.3, y=0.715, label=pvalue, color="black", parse=TRUE) 
p6 = p6 + annotate(geom="text", x=10, y=0.744, label=alpha, color="black", parse=TRUE) +
          annotate(geom="text", x=10, y=0.747, label=degree, color="black", parse=TRUE) +
          annotate(geom="text", x=6.3, y=0.715, label=pvalue, color="black", parse=TRUE) 
```

```{r fig.width=4.7*1.1, fig.height=4.4*1.1}
figure = ggdraw() +
  draw_plot(p5, 0, .5, .5, .5) +
  draw_plot(p6, 0.5, .5, .5, .5) +
  draw_plot(p0, 0, 0, 1, .5)
figure
#ggsave("w2b_over_time.pdf", plot=figure)
```

Loess: a nonparametric, graphical tool for depicting relationships between variables. http://www.jtrive.com/docs/Jacoby_LOESS.pdf

```{r}
# extract the LOESS
head(ggplot_build(p5)$data[[2]]) # alpha = 0.4 (smoothing parameter)
head(ggplot_build(p6)$data[[2]]) # alpha = 0.4 (smoothing parameter); lambda = degree of the local polynomial 
# degree is 0 (estimated below)
# Using a zero degree polynomial turns LOESS into a weighted moving average. 
# degrees: https://www.itl.nist.gov/div898/handbook/pmd/section1/pmd144.htm
```

Try ggdraw() http://www.sthda.com/english/wiki/wiki.php?id_contents=7930
Map separate lines: https://stats.idre.ucla.edu/r/faq/how-can-i-explore-different-smooths-in-ggplot2/


```{r}
dfF = df[df$sex=="Females",]
dfM = df[df$sex=="Males",]

dfF$pophost = factor(dfF$pophost, levels = c("K. elegans", "C. corindum") )
dfF$`Host Plant` = dfF$pophost
dfM$pophost = factor(dfM$pophost, levels = c("K. elegans", "C. corindum") )
dfM$`Host Plant` = dfM$pophost
```


```{r fig.width=4.7, fig.height=2.3}
# females
p3 = ggplot() + 
  ggtitle("Females") + xlab("Month") + ylab("Wing-to-Body Ratio") +
  geom_vline(xintercept = xlab_dates, color="gainsboro") + 
  geom_smooth(data=dfF, method="lm", se=FALSE, linetype = "dashed", lwd=0.5,
              mapping = aes(x = month_of_year, y = wing2body, colour = `Host Plant`)) +
  geom_smooth(data=dfF, method="loess", 
              mapping = aes(x = month_of_year, y = wing2body, colour=`Host Plant`, fill=`Host Plant`)) + 
  geom_point(data=dfF, mapping = aes(x = month_of_year, y = wing2body, colour=`Host Plant`)) +
  ylim(0.70, 0.765) +
  customPlot + 
  scale_color_manual(values=c("C. corindum" = "turquoise3", "K. elegans" = "springgreen4")) +
  scale_fill_manual(values = c("C. corindum" = "turquoise3", "K. elegans" = "green")) +
  scale_x_continuous(breaks=xlab_months, labels= month_labs)

# males
p4 = ggplot() + 
  ggtitle("Males") + xlab("Month") + ylab("Wing-to-Body Ratio") +
  geom_vline(xintercept = xlab_dates, color="gainsboro") + 
  geom_smooth(data=dfM, method="lm", se=FALSE, linetype = "dashed", lwd=0.5,
              mapping = aes(x = month_of_year, y = wing2body, colour = `Host Plant`)) +
  geom_smooth(data=dfM, method="loess", 
              mapping = aes(x = month_of_year, y = wing2body, colour=`Host Plant`, fill=`Host Plant`)) + 
  geom_point(data=dfM, mapping = aes(x = month_of_year, y = wing2body, colour=`Host Plant`)) +
  ylim(0.70, 0.765) +
  customPlot + 
  scale_color_manual(values=c("C. corindum" = "turquoise3", "K. elegans" = "springgreen4")) +
  scale_fill_manual(values = c("C. corindum" = "turquoise3", "K. elegans" = "green")) +
  scale_x_continuous(breaks=xlab_months, labels= month_labs)

grid.arrange(p3, p4, ncol=2)
```

```{r}
# extract the LOESS
head(ggplot_build(p3)$data[[2]])
head(ggplot_build(p4)$data[[2]])

# these calculations give you similar tables. Here is an example:
mloess <- loess(wing2body ~ month_of_year, data=dfF, degree=0)
# Using a zero degree polynomial turns LOESS into a weighted moving average. 
xrange <- range(dfF$month_of_year)
xseq <- seq(from=xrange[1], to=xrange[2], length=80)
pred <- predict(mloess, newdata = data.frame(month_of_year = xseq), se=TRUE)
y = pred$fit
ci <- pred$se.fit * qt(0.95 / 2 + .5, pred$df)
ymin = y - ci
ymax = y + ci
loess.DF <- data.frame(x = xseq, y, ymin, ymax, se = pred$se.fit)
head(loess.DF)
summary(loess.DF)
summary(mloess)
```

```{r}
# compare models
mlinear = glm(wing2body ~ month_of_year, data=df, family=gaussian)
mlinearF = glm(wing2body ~ month_of_year, data=dfF, family=gaussian)
mlinearM = glm(wing2body ~ month_of_year, data=dfM, family=gaussian)

summary(mlinear)$coeff[[8]] # significant for all (but only with K.elegans)
summary(mlinearF)$coeff # not significant for females (or when split by host plant)
summary(mlinearM)$coeff # significant for males (for K.elegans, not for C.corindum)
```

scale_colour_discrete(name ="Host Plant", labels=c("C. corindum", "K. elegans")) +

## Wing Morph Freq

```{r}
#############wing morph plots

##Results to visualize: sex, host, months_since_start.

w_morph_summary<-aggregate(wing_morph_binom~sex*pophost*month_of_year, data=raw_data, FUN=mean)
w_morph_summary$se<-aggregate(wing_morph_binom~sex*pophost*month_of_year, data=raw_data, FUN=SE)$wing_morph_binom

#jitter slightly
jitter = runif(n=nrow(w_morph_summary), min=-0.5, max=0.5)
w_morph_summary$dates <- w_morph_summary$month_of_year + jitter

# create groups
dd = w_morph_summary

dd$pophost[dd$pophost=="C.corindum"]<-"C. corindum"
dd$pophost[dd$pophost=="K.elegans"]<-"K. elegans"
dd$pophost = factor(dd$pophost, levels = c("K. elegans", "C. corindum") )

dd$`Host Plant` = dd$pophost

dd$sex[dd$sex=="F"]<-"Females"
dd$sex[dd$sex=="M"]<-"Males"
dd$sex = factor(dd$sex, levels = c("Males", "Females") )
dd$`Sex` = dd$sex 

dfF = dd[dd$sex=="Females",]
dfM = dd[dd$sex=="Males",]
```

```{r fig.width=4.7, fig.height=2.3}
p7 = ggplot() + 
  ggtitle("Females") + xlab("Month") + ylab("Long-Wing Morph Frequency") +
  geom_vline(xintercept = xlab_dates, color="gainsboro") + 
  geom_smooth(data=dfF, method="lm", se=FALSE, linetype = "dashed", lwd=0.5,
              mapping = aes(x = month_of_year, y = wing_morph_binom, colour = `Host Plant`)) +
  geom_smooth(data=dfF, method="loess", 
              mapping = aes(x = month_of_year, y = wing_morph_binom, colour=`Host Plant`, fill=`Host Plant`)) + 
  geom_point(data=dfF, mapping = aes(x = month_of_year, y = wing_morph_binom, colour=`Host Plant`)) +
  customPlot + 
  scale_color_manual(values=c("C. corindum" = "turquoise3", "K. elegans" = "springgreen4")) +
  scale_fill_manual(values = c("C. corindum" = "turquoise3", "K. elegans" = "green")) +
  scale_x_continuous(breaks=xlab_months, labels= month_labs) +
  ylim(-0.75,2.25)

p8 = ggplot() + 
  ggtitle("Males") + xlab("Month") + ylab("Long-Wing Morph Frequency") +
  geom_vline(xintercept = xlab_dates, color="gainsboro") + 
  geom_smooth(data=dfM, method="lm", se=FALSE, linetype = "dashed", lwd=0.5,
              mapping = aes(x = month_of_year, y = wing_morph_binom, colour = `Host Plant`)) +
  geom_smooth(data=dfM, method="loess", 
              mapping = aes(x = month_of_year, y = wing_morph_binom, colour=`Host Plant`, fill=`Host Plant`)) + 
  geom_point(data=dfM, mapping = aes(x = month_of_year, y = wing_morph_binom, colour=`Host Plant`)) +
  customPlot + 
  scale_color_manual(values=c("C. corindum" = "turquoise3", "K. elegans" = "springgreen4")) +
  scale_fill_manual(values = c("C. corindum" = "turquoise3", "K. elegans" = "green")) +
  scale_x_continuous(breaks=xlab_months, labels= month_labs) +
  ylim(-0.75,2.25)

grid.arrange(p7,p8, ncol=2)
```


```{r fig.width=4.7, fig.height=2.3*2}
extrafigure = grid.arrange(p3,p4,p7,p8, ncol=2)
#ggsave("extra_figure.pdf", plot=extrafigure)
```


```{r}
customPlot = list( theme_classic(),
                   theme(axis.text=element_text(size=13),
                      axis.title=element_text(size=16), 
                      plot.title=element_text(size=20),),
                   theme(legend.position = c(0.23, 0.95)),
                   theme(legend.title = element_text(size=14, face="italic"), 
                         legend.text = element_text(size = 13, face="italic"))
                   )
```

```{r fig.width=4.7, fig.height=2.15}
#########
p9 = ggplot() + 
  ggtitle("A") + xlab("Month") + ylab("Long-Wing Morph Frequency") +
  geom_vline(xintercept = xlab_dates, color="gainsboro") + 
  geom_smooth(data=dd, method="lm", se=FALSE, linetype = "dashed", lwd=0.5,
              mapping = aes(x = month_of_year, y = wing_morph_binom), colour="black") +
  geom_smooth(data=dd, method="loess", 
              mapping = aes(x = month_of_year, y = wing_morph_binom, colour=`Host Plant`, fill=`Host Plant`)) + 
  geom_point(data=dd, mapping = aes(x = month_of_year, y = wing_morph_binom, colour=`Host Plant`)) +
  customPlot + 
  scale_color_manual(values=c("C. corindum" = "turquoise3", "K. elegans" = "springgreen4")) +
  scale_fill_manual(values = c("C. corindum" = "turquoise3", "K. elegans" = "green")) +
  scale_x_continuous(breaks=xlab_months, labels= month_labs) + ylim(0.3,1.2)

p10 = ggplot() + 
  ggtitle("B") + xlab("Month") + ylab("Long-Wing Morph Frequency") +
  geom_vline(xintercept = xlab_dates, color="gainsboro") + 
  geom_smooth(data=dd, method="lm", se=FALSE, linetype = "dashed", lwd=0.5,
              mapping = aes(x = month_of_year, y = wing_morph_binom), colour="black") +
  geom_smooth(data=dd, method="loess", 
              mapping = aes(x = month_of_year, y = wing_morph_binom, colour=Sex, fill=Sex)) + 
  geom_point(data=dd, mapping = aes(x = month_of_year, y = wing_morph_binom, colour=Sex)) +
  customPlot + 
  scale_color_manual(values=c("Females" = "brown3", "Males" = "black")) +
  scale_fill_manual(values = c("Females" = "brown1", "Males" = "sienna4")) +
  scale_x_continuous(breaks=xlab_months, labels= month_labs) + theme(axis.title.y = element_blank()) + ylim(0.3,1.2)

# extract the LOESS
alpha = paste("alpha[loess]==", ggplot_build(p9)$data[[2]]$alpha[1])
degree="lambda[loess]==0"

mlinear = glm(wing_morph_binom ~ month_of_year, data=dd, family=gaussian)
round(summary(mlinear)$coeff[[8]],3) # significant for all (but only with K.elegans)
pvalue = paste0("italic(p)[glm]==",round(summary(mlinear)$coeff[[8]],2))

p9 = p9 + annotate(geom="text", x=7.4, y=1.0, label=alpha, color="black", parse=TRUE) +
          annotate(geom="text", x=7.4, y=1.04, label=degree, color="black", parse=TRUE) +
          annotate(geom="text", x=5, y=0.718, label=pvalue, color="black", parse=TRUE)

p10 = p10 + annotate(geom="text", x=7.4, y=1.1, label=alpha, color="black", parse=TRUE) +
          annotate(geom="text", x=7.4, y=1.15, label=degree, color="black",parse=TRUE) +
          annotate(geom="text", x=5, y=1, label=pvalue, color="black", parse=TRUE)

grid.arrange(p9,p10, ncol=2)
```


```{r}
threshold = sort(unique(raw_data$dates))[5]
df$date_b = "2013-2015"
df$date_b[df$dates >= threshold] = "2016-2020"

df$date_b = as.factor(df$date_b)
```

```{r}
wm_table<-aggregate(wing_morph_binom~dates, data=raw_data, FUN=mean)
wm_table$date_b = "2013-2015"
wm_table$date_b[wm_table$dates >= threshold] = "2016-2020"
wm_table$date_b = as.factor(wm_table$date_b)

last_4_years = wm_table[wm_table$date_b=="2016-2020",]
first_3_years = wm_table[wm_table$date_b=="2013-2015",]

xlab_years = sort(unique(raw_data$dates))
```

```{r fig.width=4.7, fig.height=2.15}

p11 = ggplot() + 
  ggtitle("C") + xlab("Year") + ylab("Long-Wing Morph Frequency") +
  geom_vline(xintercept = xlab_years, color="gainsboro") + 
   geom_smooth(data=wm_table, method="lm", se=FALSE, linetype = "dashed",
              mapping = aes(x = dates, y = wing_morph_binom), colour="black", lwd=0.5) +
  geom_smooth(data=wm_table, method="loess", 
              mapping = aes(x = dates, y = wing_morph_binom), colour="black") + 
  geom_point(data=wm_table, mapping = aes(x = dates, y = wing_morph_binom))
  #ylim(0.71, 0.75) +
 #+ stat_smooth(data=first_3_years, mapping = aes(x = dates, y = wing_morph_binom), method = "lm", formula = y ~ x, se = FALSE,
 #              colour="blue") # linetype = "longdash",

p11 = p11 + guides(fill = guide_legend(reverse = TRUE)) + theme_classic() +
  theme(axis.text=element_text(size=13),
        axis.title=element_text(size=16), 
        plot.title=element_text(size=20)) + guides(color = FALSE)  + scale_linetype(guide = FALSE) +
  theme(legend.key = element_rect(fill = "white", color = NA), 
        legend.key.size = unit(0.5, "cm"),
        legend.key.width = unit(0.5,"cm")) + 
  scale_fill_manual(values = "blue", labels=c("C. corindum", "K. elegans")) + 
  labs(fill = "Host Plant") + theme(legend.title = element_text(size = 14),
                                    legend.text = element_text(size = 13, face="italic")) +
  scale_color_manual(values="blue") + theme(legend.position = c(0.2, 0.88)) + theme(legend.title = element_blank()) + ylim(0.3,1.2)

alpha = paste("alpha[loess]==", ggplot_build(p0)$data[[2]]$alpha[1])
degree="lambda[loess]==0"
mlinear = glm(wing_morph_binom ~ dates, data=dd, family=gaussian)
pvalue = paste0("italic(p)[glm]==",round(summary(mlinear)$coeff[[8]],2))


p11 = p11 + annotate(geom="text", x=unique(raw_data$dates)[7], y=1, label=alpha, color="black", parse=TRUE, size=6) +
          annotate(geom="text", x=unique(raw_data$dates)[7], y=1.05, label=degree, color="black",parse=TRUE, size=6) +
          annotate(geom="text", x=unique(raw_data$dates)[5], y=0.59, label=pvalue, color="black", parse=TRUE, size=6)
# first 3 years
#mlinear = glm(wing_morph_binom ~ dates, data=first_3_years, family=gaussian)
#pvalue = paste0("italic(p)[glm]==",round(summary(mlinear)$coeff[[8]],2))

#p11 = p11 + annotate(geom="text", x=unique(raw_data$dates)[3], y=1, label=pvalue, color="blue", parse=TRUE, size=6)

p11
```

```{r fig.width=4.7*1.1, fig.height=4.4*1.1}
figure2 = ggdraw() +
  draw_plot(p9, 0, .5, .5, .5) +
  draw_plot(p10, 0.5, .5, .5, .5) +
  draw_plot(p11, 0, 0, 1, .5)
figure2
#ggsave("wmorph_over_time.pdf", plot=figure2)
```

MLC: Something is going wrong here calculating variance for a binomial response - a frequency or probability plot should have a y-axis with a strict minimum of 0 and maximum of 1, but that is not happening with the variance plotted here. So, the estimated variance is probably not accurate, and the expanded y-axis range makes it difficult to see changes that might be visible along the month axis. I have historically used binom.confint() in the binom() library to calculate binomial confidence intervals; I don't know what the smoothing options are, but I do think we need a different function for this one. In principal, I like this idea.

MLC: Actually, this is cuing me into something being off about the way loess is visualizing variance in general - the error bars on many of these plots are way wider (x10?) than the standard error on the data itself. I think it may be related to using a summary file instead of a complete data file. 


AB:

The prediction interval predicts in what range a future individual observation will fall, while a confidence interval shows the likely range of values associated with some statistical parameter of the data, such as the population mean.

What is being plotted is a CI but a CI will converge to a single value as the sample size increases. That is why the shaded regions change so much as you keep adding more data points.

What we want then is a prediction interval. Even though the "y" column is the predicted value, the "ymin" and "ymax" are the lower and upper pointwise confidence interval around the mean, note the predicted interval.

```{r}
library(msir)
data(cars)
# Calculates and plots a 1.96 * SD prediction band, that is,
# a 95% prediction band
l <- loess.sd(cars, nsigma = 1.96)
plot(cars, main = "loess.sd(cars)", col="red", pch=19)
lines(l$x, l$y)
lines(l$x, l$upper, lty=2)
lines(l$x, l$lower, lty=2)
```


