---
title: 'Data Visualizations: flight_yes_no'
author: "Anastasia Bernat"
date: "5/4/2020"
output:
  html_document: default
  pdf_document: default
---

# Winter 2020 Flight Trials 

```{r setup, include=FALSE}
rm(list=ls())
dir = "~/Desktop/git_repositories/SBB-dispersal/avbernat/Dispersal/Winter_2020/stats/"
setwd(dir) 

library(lme4)
library(rethinking)

library(lubridate)
library(chron)
library(dplyr)
library(tidyselect)

library(glmnet)
library(ggplotify)
library(gridExtra)
library(ggformula)
library(randomcoloR)

knitr::opts_chunk$set(echo = TRUE)
```

Flight Trials Winter 2020 Dataset was conducted from 2/17/2020 - 3/10/2020. Soapberry bugs were flight tested twice in the flight mill and observed from 8 AM to (5-8 PM) each day.

# Loading, cleaning, and recoding the data

```{r}
source_path = "~/Desktop/git_repositories/SBB-dispersal/avbernat/Rsrc/"

script_names = c("center_flight_data.R", # Re-centers data 
                 "clean_flight_data.R", # Loads and cleans data
                 # groups explanatory values into categories and calculates 
                 # the sample of proportion of flight for each category
                 "categorize_binomial_data.R",
                 "categorize_binomial_data-MF.R", 
                 "categorize_binomial_data3.R",
                 "categorize_binomial_data-MF2.R")

for (script in script_names) { 
  path = paste0(source_path, script)
  source(path) 
}

output_col = FALSE # Change to TRUE if working in Base R or RStudio; FALSE if generating an HTML

data <- read_flight_data("data/all_flight_data-Winter2020.csv")
data_all <- data[[1]]
data_tested <- data[[2]]
```

# ALL DATA PLOTS

```{r fig.width=7, fig.height=2.5}
cat_data1 <- categorize_data(data_tested, all_of("min_from_IncStart"), all_of("flew_b"), 30, 0, 30)
cat_data2 <- categorize_data(data_tested, all_of("minute_duration"), all_of("flew_b"), 30, 0, 30)

# Plotting observed proportions of yes flew by minutes from incubation start 
p1 <- as.grob(expression(plot(sample_prop ~ min_from_IncStart, data=cat_data1,
     ylab="Sample Proportion of Yes Flew", xlab="Minutes From Incubtation Start (min) ", 
     main="Observed proportions of yes flew by minutes from incubation start")
))

# Plotting observed proportions of yes flew by minutes from trial start
p2 <- as.grob(expression(
  plot(sample_prop ~ minute_duration, data=cat_data2,
     ylab="Sample Proportion of Yes Flew", xlab="Minutes From Trial Start (min)", 
     main="Observed proportions of yes flew by minutes from trial start") %>%
  abline(v=120, col="red")
))

grid.arrange(p1,p2, ncol=2)
```


```{r}
cat_data1$min_from_IncStart
fit <- lm(`sample_prop`~ min_from_IncStart, data=cat_data1)
summary(fit) # NO significant effect

cf <- round(coef(fit), 2)  ## rounded coefficients for better output
eq <- paste0("portion_flew = ", cf[1],
             ifelse(sign(cf[2])==1, " + ", " - "), 
             abs(cf[2]), "*time_start ") ## sign check to avoid having plus followed by minus for negative coefficients

plot(cat_data1$min_from_IncStart,cat_data1$sample_prop, 
     xlab= "Starting Time", ylab="Proportion of 24-H Day Spent Flying")
  abline(coef(fit)[1:2], col="blue")
mtext(eq, 3, line=-2, side=3)
text(300,0.4, c("not significant"), col= "red")
```

```{r}
cat_data2$log_min_duration <- log(cat_data2$minute_duration)
m <- quap(
  alist(
    sample_prop ~ dnorm(mu, sigma),
    mu <- alpha + beta*(log_min_duration),
    alpha ~ dnorm(5.792217, 5), # this should be in terms of height not weight
    beta ~ dlnorm(0,1),
    sigma ~ dunif(0,10)
  ),
  data=cat_data2
)
precis(m)

#x_seq <- log(min(cat_data2$minute_duration):max(cat_data2$minute_duration))
x_seq <- log(seq(from=0, to=840, length.out=30))
mu <- sim(m,data=list(log_min_duration=x_seq)) 

mu_mean <- apply(mu,2,mean) 
mu_ci <- apply(mu,2,PI,prob=0.89) 

# Plot Log
plot( cat_data2$log_min_duration , cat_data2$sample_prop , col=col.alpha(rangi2,0.7),
      ylab="Sample Proportion of Yes Flew", 
      xlab="Log(Minutes From Trial Start) log(min)", 
      main="Observed proportions of yes flew by log(minutes) from trial start" )
lines(x_seq, mu_mean) 
shade(mu_ci, lim = x_seq)
text(6,0.3, c("significant"), col= "darkolivegreen4")

# Plot Non-log
plot( cat_data2$minute_duration , cat_data2$sample_prop, col=col.alpha(rangi2,0.7), 
      ylab="Sample Proportion of Yes Flew", 
      xlab="Minutes From Trial Start (min)", 
      main="Observed proportions of yes flew by minutes from trial start" )
lines( exp(x_seq), mu_mean) 
shade(mu_ci, lim = exp(x_seq))
abline(v=120, col="red", lty=2)
```

```{r}
par(mfrow=c(1,2))
# What is happening for those who flew?
x = chron(times = data_tested$time_start)
y = data_tested$recording_duration / (60*60) # hours 

fit <- lm(y~x, data=data_tested)
summary(fit) # significant effect

cf <- round(coef(fit), 2)  ## rounded coefficients for better output
eq <- paste0("portion_flew = ", cf[1],
             ifelse(sign(cf[2])==1, " + ", " - "), 
             abs(cf[2]), "*time_start ") ## sign check to avoid having plus followed by minus for negative coefficients

plot(x,y, xlab= "Starting Time", ylab="Duration (hr)")
  abline(coef(fit)[1:2], col="blue") # Alternative: lines(x, fitted(fit), col="blue")
#mtext(eq, 3, line=-2, side=3)

# Filter out those that didn't fly.
all_flew <- filter(data_all, flew_b == 1)

# What is happening for those who flew?
x = chron(times = all_flew$time_start)
y = all_flew$recording_duration / (60*60) # hours 

fit <- lm(y~x, data=all_flew)
summary(fit) # significant effect

cf <- round(coef(fit), 2)  ## rounded coefficients for better output
eq <- paste0("portion_flew = ", cf[1],
             ifelse(sign(cf[2])==1, " + ", " - "), 
             abs(cf[2]), "*time_start ") ## sign check to avoid having plus followed by minus for negative coefficients

plot(x,y, xlab= "Starting Time", ylab="Duration (hr)")
  abline(coef(fit)[1:2], col="blue") # Alternative: lines(x, fitted(fit), col="blue")
#mtext(eq, 3, line=-2, side=3)
```

```{r}
# Missing mass for some (3 NA)
missing_mass <- subset(data_all, is.na(data_all$mass))
# IDs 339, 48, and 342 have no mass and were tested on the same date
# Remove missing mass
data_mass <- data_all %>%
  filter(!is.na(mass))

## ALL DATA: Observed proportions of yes flew by mass
cat_data3 <- categorize_data(data_mass, all_of("mass"), all_of("flew_b"),  0.010, 0.015, 0.025) 

## TRIAL 1 DATA: Observed proportions of yes flew by mass
data_T1 <-data_all[data_all$trial_type=="T1",]
data_mass_T1 <- data_T1 %>%
  filter(!is.na(mass))

cat_data4 <- categorize_data(data_mass_T1, all_of("mass"), all_of("flew_b"),  0.010, 0.015, 0.025) 

```

```{r}
plot(data_T1$mass, data_T1$flew_b)
```

```{r}
# ALL DATA 
#all_fit <-lm(flew_b~mass, data=data_mass) 
all_fit <-lm(sample_prop~ mass, data=cat_data3) 
coeff <- coefficients(summary(all_fit))
coeff

mass_seq <- seq(from=0.02, to=0.20, length.out=30)
mu <- link(all_fit, data=data.frame(mass=mass_seq)) # gives warning
mu.PI <- apply(mu,2,PI, prob=0.89)

eq <- paste0("portion_flew = ", round(coeff[1],3),
             ifelse(sign(coeff[2])==1, " + ", " - "), 
             abs(round(coeff[2],3)), "*mass")

plot(as.matrix(cat_data3[1]), 
     as.matrix(cat_data3[3]),
     ylab="Sample Proportion of Yes Flew", 
     xlab="Mass (g)", 
     main="All Data: Observed proportions of yes flew by mass")
shade(mu.PI, mass_seq)
abline(coeff[1], coeff[2], col="blue")
mtext(eq, side=4)
```

```{r}
# TRIAL 1 - FIX THIS TOO

fit_mass<-lm(flew_b~mass, data=data_mass_T1) 
coeff <- coefficients(summary(fit_mass))

mass_seq <- seq(from=0.01, to=0.180, length.out=30)
mu <- link(fit_mass, data=data.frame(mass=mass_seq)) # gives warning
mu.PI <- apply(mu,2,PI, prob=0.89)

eq <- paste0("portion_flew = ", round(coeff[1],3),
             ifelse(sign(coeff[2])==1, " + ", " - "), 
             abs(round(coeff[2],3)), "*mass")

plot(as.matrix(cat_data4[1]), 
     as.matrix(cat_data4[3]),
     ylab="Sample Proportion of Yes Flew", 
     xlab="Mass (g)", 
     main="Trial 1: Observed proportions of yes flew by mass")
abline(coeff[1], coeff[2], col="blue")
mtext(eq, side=4)
shade(mu.PI, mass_seq)
```

```{r fig.width=5, fig.height=3}
summary <-aggregate(flew_b~sex*host_plant*sym_dist, data=data_all, FUN=mean)
summary

par(mar=c(6.5, 5.5, 5.5, 9.5), xpd=TRUE) # Add extra space to right of plot area; change clipping to figure
plot(summary$flew_b~summary$sym_dist, 
     col=c(1,2)[as.factor(summary$sex)], # Female = Black, Male = Red
     pch=c(19,22)[as.factor(summary$host_plant)],
     main="Observed Data",
     xlab = "Distance from Sympatric Zone (°)",
     ylab= "Proportion Flew" # K. elegans = Squares C.corindum = circles
     #sub=eq_glmer
     ) 
legend("topright",
       legend = c("C. corindum and F", "C. corindum and M", "K. elegans and F" , "K. elegans and M"),
       inset=c(-0.23,0.2),
       col=c(1,2),
       pch = c(19,19, 22,22),
       title="Groups")

```

```{r}
missing_mass <- subset(data_all, is.na(data_all$mass))
data_mass <- data_all %>%
  filter(!is.na(mass))

cat_data5 <- categorize_data_MF(data_mass, all_of("mass"), all_of("sex_c"), 
                                all_of("host_c"), all_of("flew_b"), 0.010, 0.015, 0.025)
cat_data5
```

```{r fig.width=5, fig.height=3}

gf_point(sample_prop ~ mass, data=cat_data5, col=~sex_c, size=~host_c,
         xlab = "Mass (g)", ylab= "Sample Proportion that Flew",
         title= "Observed Data")

par(mar=c(6.5, 5.5, 5.5, 9.5), xpd=TRUE) 
plot(cat_data5$sample_prop~cat_data5$mass, 
     col=c(1,2)[as.factor(cat_data5$sex_c)], # Female = Black, Male = Red
     pch=c(19,22)[as.factor(cat_data5$host_c)],
     main="Observed Data",
     xlab = "Mass (g)",
     ylab= "Proportion Flew", # K. elegans = Squares, C.corindum = circles
     #sub=eq_glmer
     ) 
legend("topright",
       legend = c("C. corindum and F", "C. corindum and M", "K. elegans and F" , "K. elegans and M"),
       inset=c(-0.23,0.2),
       col=c(1,2),
        pch = c(19,19, 22,22),
       title="Groups")
```

# TRIAL 1 PLOTS



# TRIAL 2 PLOTS

```{r}
data_T2 <-data_all[data_all$trial_type=="T2",]
plot(data_T2$mass, data_T2$flew_b)
```

```{r}
##### TRIAL 2 DATA: Observed proportions of yes flew by mass #############
cat_data6 <- categorize_data(data_T2, all_of("mass"), all_of("flew_b"),  0.010, 0.015, 0.025)
```

```{r}
fit<-lm(flew_b~mass, data=data_T2) 
coeff <- coefficients(summary(fit))

mass_seq <- seq(from=0.0, to=0.18, length.out=30)
mu <- link(fit, data=data.frame(mass=mass_seq)) # gives warning
mu.PI <- apply(mu,2,PI, prob=0.89)

eq <- paste0("portion_flew = ", round(coeff[1],3),
             ifelse(sign(coeff[2])==1, " + ", " - "), 
             abs(round(coeff[2],3)), "*mass")

plot(as.matrix(cat_data6[1]), 
     as.matrix(cat_data6[3]),
     ylab="Sample Proportion of Yes Flew", 
     xlab="Mass (g)", 
     main="Trial 2: Observed proportions of yes flew by mass")
shade(mu.PI, mass_seq)
abline(coeff[1], coeff[2], col="blue")
mtext(eq, side=4)
```

## FEMALE DATA PLOTS

```{r}
data_fem <- data_all[data_all$sex=="F",]
```

```{r fig.width=5, fig.height=3}
plot(data_fem$sym_dist, data_fem$flew_b)

fem_sum <-aggregate(flew_b~host_plant*sym_dist, data=data_fem, FUN=mean)
fem_sum

par(mar=c(6.5, 5.5, 5.5, 9.5), xpd=TRUE) # Add extra space to right of plot area; change clipping to figure
plot(fem_sum$flew_b~fem_sum$sym_dist, 
     pch=c(19,22)[as.factor(fem_sum$host_plant)],
     main="Observed Data",
     xlab = "Distance from Sympatric Zone (°)",
     ylab= "Proportion Flew", # K. elegans = Squares C.corindum = circles
     #sub=eq_glmer
     ) 
legend("topright",
       legend = c("C.corindum","K.elegans"),
       inset=c(-0.23,0.2),
       pch = c(19,22),
       title="Groups")

```


```{r}
cat_data7 <- categorize_data(data_fem, all_of("sym_dist"), all_of("flew_b"),  0.1, 0, 0.05) 

fit_fem <- lm(flew_b~sym_dist, data=data_fem) 
coefff <- coefficients(summary(fit_fem))

eq <- paste0("portion_flew = ", round(coefff[1],3),
             ifelse(sign(coefff[2])==1, " + ", " - "), 
             abs(round(coefff[2],3)), "*sym_dist")

plot(as.matrix(cat_data7[1]), 
     as.matrix(cat_data7[3]),
     ylab="Sample Proportion of Yes Flew", 
     xlab="Distance From Sympatric Zone (°)", 
     main="Observed proportions of yes flew by distance from sympatric zone")
abline(coefff[1], coefff[2], col="blue")
mtext(eq, side=4)
```


```{r fig.width=7, fig.height=2.5}
pf1 <- as.grob(expression(
  plot(data_fem$flew_b, data_fem$total_eggs)))
# seems like if you generally laid less eggs you flew, but a lot of females laid few and also didn't fly
pf2 <- as.grob(expression(
  plot(data_fem$mass, data_fem$total_eggs)))
pf3 <- as.grob(expression(
  plot(data_fem$sym_dist,data_fem$total_eggs)))
grid.arrange(pf1,pf2,pf3, ncol=3)
```

```{r}
####### Sample proportion that laid eggs by mass ####### AND
####### Total eggs laid  by mass #######

data <- read_flight_data("data/all_flight_data-Winter2020.csv")
data_all <- data[[1]]
data_tested <- data[[2]]
data_fem <- data_tested[data_tested$sex=="F",]

# Data Manipulation - filtering, removing duplicates, and recentering mass
data_fem_filtered <- data_fem %>%
  filter(!is.na(total_eggs)) %>%
  filter(!is.na(mass))
data_fem_filtered <-data_fem_filtered[!duplicated(data_fem_filtered$ID, 
                                                  data_fem_filtered$total_eggs),]

data_fem_filtered$mass_c <- data_fem_filtered$mass - mean(data_fem_filtered$mass)

cat_data8 <- categorize_data(data_fem_filtered, all_of("mass"), all_of("total_eggs"),  0.010, 0.015, 0.025) 
```


```{r}
####### Sample proportion that laid eggs by mass ####### AND
####### Total eggs laid  by mass #######

# Plotting - lm, PI

par(mfrow=c(1,2)) 

# Plot 1 ---------------------------------------------------

fit_prop <- lm(sample_prop~mass, data=cat_data8) 
coeff <- coefficients(summary(fit_prop))

eq <- paste0("portion_total_eggs = ", round(coeff[1],3),
             ifelse(sign(coeff[2])==1, " + ", " - "), 
             abs(round(coeff[2],3)), "*mass")

mass.seq <- seq(from=0, to=.25, length.out=30)
mu <- link(fit_prop, data=data.frame(mass=mass.seq)) # gives warning
mu.PI <- apply(mu,2,PI, prob=0.89)

plot(as.matrix(cat_data8[1]),
     as.matrix(cat_data8[3]),
     ylab="Proportion of Total Eggs Laid", 
     xlab="Mass (g)", 
     main="Observed proportions of eggs laid by mass",
     cex.main=0.8)
abline(coeff[1], coeff[2], col="blue")
mtext(eq, side=4, cex=0.9)
shade(mu.PI, mass.seq)

# Plot 2 ---------------------------------------------------
fit_eggs <- lm(total_eggs~mass_c, data=data_fem_filtered) 
coeff <- coefficients(summary(fit_eggs))

eq <- paste0("total_eggs = ", round(coeff[1],3),
             ifelse(sign(coeff[2])==1, " + ", " - "), 
             abs(round(coeff[2],3)), "*mass_c")

mass_c.seq <- seq(from=-.1, to=.15, length.out=30)
mu <- link(fit_eggs, data=data.frame(mass_c=mass_c.seq)) # gives warning
mu.PI <- apply(mu,2,PI, prob=0.89)

plot(total_eggs ~ mass_c, data= data_fem_filtered,
     ylab="Total Eggs Laid", 
     xlab="Mass Centered (g)", 
     main="Female Data: Total eggs laid by mass",
     cex.main=0.9)
abline(coeff[1], coeff[2], col="blue")
mtext(eq, side=4, cex=0.9)
shade(mu.PI, mass_c.seq)
```


```{r}
##### Egg laid day of flight by variations/change in mass ########

# Data Manipulation - filtering, removing duplicates, recentering mass, and calculating changes in mass

data_fem <- data_tested[data_tested$sex=="F",] # 1.) Reselect the data (for integrity purposes)
data_f <- select(data_fem, ID, eggs_b)
duplicates <- data_f[duplicated(data_f),] # 2.) This will give you duplicate rows:
dim(data_f[duplicated(data_f),])[1] # 3.) This will give you the number of duplicates:
omit <- duplicates$ID # 4.) # Pull out IDs that are duplicate and put them in an omit vector

# 5.) Remove those IDs from the original dataframe and return dataframe of unique values
data_fem_u <- data_fem[!data_fem$ID %in% omit, ]

# 6.) Filter and recenter
data_eggs <- data_fem_u %>%
  filter(!is.na(mass))

data_eggs$mass_c <- data_eggs$mass - mean(data_eggs$mass)

# 7.) Determine in diff dataframes which female bugs had egg switch responses (on one day they laid eggs and were tested and on another they were testedthey did not lay eggs)

data_eggs_f <- select(data_eggs, ID, eggs_b)
dupCheck<- data_eggs_f[duplicated(data_eggs_f),] # check that there are no duplicates of ID and eggs_b
dupCheck # if there are zero rows then good

dupIDs <- data_eggs[duplicated(data_eggs$ID),] # duplicated IDs in this set (aka. T2 bugs)
data_eggs_compare <- data_eggs[data_eggs$ID %in% dupIDs$ID, ] # keep all duplicated IDs (aka. T1 and T2 bugs who switch egg repsonse)
#ID_match <- data_eggs_compare[duplicated(data_eggs_compare$ID),] 

data_eggs_compare2 <- data_eggs_compare %>% # filter for only T1 bugs
  filter(trial_type == "T1")

# NOTE: one bug: ID number 109 had two trial 1's so we'll say the second trial 1 is 'trial 2'
rows_to_keep <- data_eggs_compare2$filename != "set005-2-21-2020-B3_109.txt"
egg_data = data_eggs_compare2[rows_to_keep,]

egg_data_s <- egg_data[order(egg_data$ID),] # Sort the data
ID_match_s <- dupIDs[order(dupIDs$ID),] # Sort the data

```

By this point we have filtered the data for female bugs that have an egg switch response, meaning, on one day they laid eggs and were tested and on another they were testedthey did not lay eggs. But how did the flight response change? Let's calculate flight responses and reinscribe those as well.

First, calculate variances/change in mass between trial 1 and trial 2 (between trials). Second, reinscribe flew_b to be 0 if flew in T1 but not T2 and 1 if flew in T2 but not T1 

```{r}
# 8.) Calcualte mass differences and flight response differences for fem bugs who had yes-no egg laying differences.

#######################################################################################
# KEY
#
# if mass diff is + , then mass was gained from T1 to T2
# if mass diff is -, then mass was lost from T1 to T2
#
# if flight diff = 0, then no flight reponse change (flew or did not fly both times))
# if flight diff = -1, then flew in T1 but not T2
# if flight diff = 1, then flew in T2 but not T1
#
#######################################################################################


egg_data_s <- egg_data[order(egg_data$ID),] # Sort the data
egg_data_s
if (egg_data$ID %in% ID_match_s$ID) {
  egg_data_s$mass_diff <- ID_match_s$mass - egg_data_s$mass # T2 - T1 masses differences
  egg_data_s$flew_delta <- ID_match_s$flew_b - egg_data_s$flew_b # T2 - T1 flight response differences
} 

# 9.) Calcualte the number of times the bug yes flew for all trials

for (row in seq(1:nrow(egg_data_s))) {
  egg_data_s$flew_count[row] <- 1 # flew once
  if (ID_match_s$flew_b[row] == 1 && egg_data_s$flew_b[row] == 1) {
    egg_data_s$flew_count[row] <- 2 # flew twice
  }
  if (ID_match_s$flew_b[row] == 0 && egg_data_s$flew_b[row] == 0) {
    egg_data_s$flew_count[row] <- 0 # didn't fly for either 
  }
} # 12 of the 34 rows flew twice

egg_data_s2 <- egg_data_s[order(egg_data_s$mass_diff),]
egg_data_compact <- select(egg_data_s2, ID, flew_delta, mass_diff, flew_count)
egg_data_compact 
```

Three things noticed:

(i) 12 of the 34 rows flew twice (in both trials).

(ii) 12 of the 34 did not flew at all (in either trial).

(iii) there were no cases in which the flight diff = 1, so no female bugs flew in T2 but not T1. But 10 cases of the 34 that flew in T1 but not T2.

```{r}
# Plot flight proportions by negative | positive mass_change. 
cat_data9 <- categorize_data3(egg_data_s2, all_of("mass_diff"), all_of("flew_delta"),  0.005, 0.010, -0.040) 
```


```{r}
##### Proportion flew for females who laid eggs day of flight in one trial but not the other COMPARED to their variations/change in mass ########

fit <- lm(sample_prop~mass_diff, data=cat_data9) 
coeff <- coefficients(summary(fit))

eq <- paste0("portion_flew = ", round(coeff[1],3),
             ifelse(sign(coeff[2])==1, " + ", " - "), 
             abs(round(coeff[2],3)), "*mass_diff")

mass_diff.seq <- seq(from=-0.04, to=0.08, length.out=30)
mu <- link(fit, data=data.frame(mass_diff=mass_diff.seq)) # gives warning
mu.PI <- apply(mu,2,PI, prob=0.89)

plot(as.matrix(cat_data9[1]),
     as.matrix(cat_data9[3]),
     ylim = c(-1,1),
     ylab="Proportion Flew Over Trials", 
     xlab="Mass Differences between T1 to T2 (g)", 
     main="Observed proportions of yes flew by mass differences between trials",
     cex=0.9)
abline(coeff[1], coeff[2], col="blue")
mtext(eq, side=4, cex=0.9)
shade(mu.PI, mass_diff.seq)
abline(v=0, col="mediumpurple3", type="l", lty=2)
text(0.01, 0.95, c("gained mass"),col="mediumpurple4")
text(-0.009, 0.963, c("lost mass"), col="mediumpurple4")
abline(h=0, col="darkolivegreen4", type="l", lty=2)
text(0.060, 0.087, c("flew in T2 but not T1"),col="darkolivegreen4")
text(0.060, -0.087, c("flew in T1 but not T2"),col="darkolivegreen4")
```

```{r}
egg_data_compact 
```

```{r}
cat_data10 <- categorize_data_MF2(egg_data_s2, all_of("mass_diff"), all_of("flew_count"), all_of("flew_delta"), 0.005, 0.010, -0.040) 
```

```{r}
gf_point(sample_prop ~ mass_diff, data=cat_data10, col=~flew_count,
         xlab = "Mass Differences between T1 and T2 (g)", ylab= "Proportion Flew Over Trials",
         title= "Observed proportions of yes flew by mass differences between trials")

fit <- lm(sample_prop~mass_diff, data=cat_data10) 
coeff <- coefficients(summary(fit))

eq <- paste0("portion_flew = ", round(coeff[1],3),
             ifelse(sign(coeff[2])==1, " + ", " - "), 
             abs(round(coeff[2],3)), "*mass_diff")

mass_diff.seq <- seq(from=-0.04, to=0.08, length.out=30)
mu <- link(fit, data=data.frame(mass_diff=mass_diff.seq)) # gives warning
mu.PI <- apply(mu,2,PI, prob=0.89)
```

```{r}
plot(as.matrix(cat_data10[1]),
     as.matrix(cat_data10[3]),
     #col=c(3,2,1)[as.factor(cat_data10$flew_count)], # green = 2, black = -1
     #col=c(1,2)[as.factor(summary$sex)], # Female = Black, Male = Red
     pch=c(19,22,21)[as.factor(cat_data10$flew_count)],
     ylim = c(-1,1),
     ylab="Proportion Flew Over Trials", 
     xlab="Mass Differences between T1 to T2 (g)", 
     main="Observed proportions of yes flew by mass differences between trials",
     cex=0.9)
abline(coeff[1], coeff[2], col="blue")
mtext(eq, side=4, cex=0.9)
shade(mu.PI, mass_diff.seq)
abline(v=0, col="mediumpurple3", type="l", lty=2)
text(0.01, 0.95, c("gained mass"),col="mediumpurple4")
text(-0.009, 0.963, c("lost mass"), col="mediumpurple4")
abline(h=0, col="darkolivegreen4", type="l", lty=2)
text(0.060, 0.087, c("flew in T2 but not T1"),col="darkolivegreen4")
text(0.060, -0.087, c("flew in T1 but not T2"),col="darkolivegreen4")
legend("topright",
       #legend = c("Both Trials","One Trial", "No Trial"),
       legend = c("2", "1", "0"),
       #inset=c(-0.03,0.02),
       pch = c(19,22,21),
       title="Flight Count")
```

```{r}
data <- read_flight_data("data/all_flight_data-Winter2020.csv")
data_all <- data[[1]]
data_tested <- data[[2]]
#  Eggs laid by week graph
eggs_laid <- read.csv("data/egg_data-Winter2020.csv") # go merge ID with population in Python - could also transpose this data into the data_all sheet

# Check which ID's flew to see which female bugs were tested and only keep those
highest_ID <- max(as.integer(data_tested$ID))
cat("Highest ID tested:", highest_ID)
eggs_laid <- filter(eggs_laid, ID <= 333)

# Convert date_collected to a Date object.
eggs_laid$date <- as_date(eggs_laid$date_collected) 
eggs_laid

# Make sure all the eggs are summed up for that week
df <- eggs_laid %>% 
  group_by(ID, date,pop) %>% 
  summarise(egg_n = sum(eggs))

# Cumulative sum of eggs through time 
df <- df %>%
  group_by(c(ID)) %>%
  mutate(c_eggs = cumsum(egg_n))

# For simplicity, add a new column - days_from_EggStart (days from when first collected eggs)
df$days_from_EggStart <- df$date - df$date[1]
```

```{r}
# Calculate mean eggs each day laid by all females:

d <- c()
m <- c()
for (i in 1:max(df$days_from_EggStart)) {
  df_day <- filter(df, days_from_EggStart == i)
  if (nrow(df_day) == 0 | nrow(df_day) ==1) { # those that only said eggs once removed.
    next
  }
  d <- c(d, i)
  mean <- round(mean(df_day$c_eggs))
  m <- c(m,mean)
}

# Plot all lines - could color by population or distance from the sympatric zone or do the mean of each population
cols <- terrain.colors(max(df$ID))

plot(NA, ylim = c(0, 360), xlim = c(0, 24), 
     xlab="Days From First Egg Collection", 
     ylab="Cumulative Eggs Laid",
     main="Cumulative Eggs Laid Over Time",
     bty="n")
df_m <- c()
for(i in 1:max(df$ID)) {
  df_ID <- filter(df, ID == i)
  if (nrow(df_ID) == 0 | nrow(df_ID) == 1) { # those that only said eggs once removed.
    next
  }
  lines(c_eggs ~ days_from_EggStart, data=df_ID, type="o", col=cols[i])
}
lines(d, m, type="o")
text(23.5,110, labels = c("mean"))
```

```{r}
rand_colors <- randomColor(df$ID, luminosity="bright")
```

```{r}
# Group and calcualte mean eggs laid each day by population:

plot(NA, ylim = c(0, 250), xlim = c(0, 28), 
     xlab="Days From First Egg Collection", 
     ylab="Mean Eggs Laid",
     main="Mean Eggs Laid Over Time By Population",
     bty="n")
pops <- unique(unlist(df$pop)) 
e <- 0
o <- -1
for (p in 1:length(pops)) {
  df_pop <- filter(df, pop == pops[p])
  days <- c()
  pm <- c()
  for (d in 1:max(df_pop$days_from_EggStart)) { # want the mean of each day for each population
    df_pop_day <- filter(df_pop,days_from_EggStart == d )  
    if (nrow(df_pop_day) == 0 ) { 
      next }
    days <- c(days, d)
    pop_m <- round(mean(df_pop_day$c_eggs))
    pm <- c(pm, pop_m)
  }
  lines(days, pm, type="o", col=rand_colors[p]) # why won't the color change??
  
  if (p %% 2 == 0) {
    e <- e + 2
    text(x=25.5, y=pop_m, labels=pops[e],cex=0.8)
  } else { 
    o <- o + 2
    text(x=25.5, y=(pop_m+10), labels=pops[o],cex=0.8)
    } 
}
```

- could do cumulative eggs laid over time
- plot change in mass over time (lines?)

## MALE DATA PLOTS

```{r}
data <- read_flight_data("data/all_flight_data-Winter2020.csv")
data_all <- data[[1]]
data_tested <- data[[2]]
data_male <- data_tested[data_tested$sex=="M",]
```

```{r fig.width=5, fig.height=3}
plot(data_male$sym_dist, data_male$flew_b)

male_sum <-aggregate(flew_b~host_plant*sym_dist, data=data_male, FUN="mean")
male_sum

par(mar=c(6.5, 5.5, 5.5, 9.5), xpd=TRUE) # Add extra space to right of plot area; change clipping to figure
plot(male_sum$flew_b~male_sum$sym_dist, 
     pch=c(19,22)[as.factor(male_sum$host_plant)],
     main="Observed Data",
     xlab = "Distance from Sympatric Zone (°)",
     ylab= "Proportion Flew", # K. elegans = Squares C.corindum = circles
     #sub=eq_glmer
     ) 
legend("topright",
       legend = c("C.corindum","K.elegans"),
       inset=c(-0.23,0.2),
       pch = c(19,22),
       title="Groups")

```


```{r}
cat_data7 <- categorize_data(data_male, all_of("sym_dist"), all_of("flew_b"),  0.1, 0, 0.05) 

fit_male <- lm(flew_b~sym_dist, data=data_male) 
coeffm <- coefficients(summary(fit_male))
coeffm

eq <- paste0("portion_flew = ", round(coeffm[1],3),
             ifelse(sign(coeffm[2])==1, " + ", " - "), 
             abs(round(coeffm[2],3)), "*sym_dist")

plot(as.matrix(cat_data7[1]), 
     as.matrix(cat_data7[3]),
     ylab="Sample Proportion of Yes Flew", 
     xlab="Distance From Sympatric Zone (°)", 
     main="Observed proportions of yes flew by distance from sympatric zone")
abline(coeffm[1], coeffm[2], col="blue")
mtext(eq, side=4)
```

```{r}
plot(data_male$mass, data_male$flew_b) # only a few large ones...really driving this but need to go back and check in the morphology if it's a typo or a a big male, but beside that looks pretty 50-50
```



