---
title: "Flight Summary Stats"
date: "9/24/2020"
output: html_document
---

```{r setup, include=FALSE}
rm(list=ls())
dir = "~/Desktop/git_repositories/SBB-dispersal/avbernat_working_on/Dispersal/Winter_2020/stats/"
setwd(dir) 

library(lme4)
library(dplyr)
library(tidyselect)
library(zoo)
library(faraway)
library(rstatix)

library(ggplot2)
library(ggpubr)
library(ggformula)
library(ggplotify)
library(gridExtra)
library(glmnet)

library(randomcoloR)

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = dir)
getwd()
```

# Winter 2020 Flight Trials Modeling {.tabset}

## Y/N Flight Reponse {.tabset}

### The Data

#### Source Scripts and Clean the Data

```{r}
source_path = "~/Desktop/git_repositories/SBB-dispersal/avbernat_working_on/Rsrc/"

script_names = c("center_flight_data.R", # Re-centers data 
                 "clean_flight_data.R", # Loads and cleans data
                 "unique_flight_data.R",
                 "get_warnings.R", 
                 "compare_models.R",
                 "regression_output.R", # Cleans regression outputs; prints in color or B&W
                 "AICprobabilities.R")

for (script in script_names) { 
  path = paste0(source_path, script)
  source(path) 
}

output_col = FALSE # Change to TRUE if working in Base R or RStudio; FALSE if generating an HTML
```

```{r}
data <- read_flight_data("data/all_flight_data-Winter2020.csv")
data_all <- data[[1]]
data_tested <- data[[2]]

# keep bugs only tested once then recenter:
d <- create_delta_data(data_tested, remove_bugs_tested_once = FALSE) 
d_all <- center_data(d, is_not_unique_data = FALSE)
```

#### Exploratory Plots

```{r}
# Compare Number of Trials a Bug Flew in to Average Days
d_all$num_trials <- 0
d_all$num_trials <- rowSums(d_all[,c("num_flew", "num_notflew")])

# Scatter Plot
# AB: some bugs we couldn't find early on in the bins and were tested into trial 2 time. Some bugs were also accidentally tested twice in trial 1. 
s <- ggplot(d_all, aes(x=avg_days, y=num_trials))  +
  geom_point(size=2, shape=23)
  
# Histogram
d_all$num_trials <- as.factor(rowSums(d_all[,c("num_flew", "num_notflew")]))
h <- ggplot(d_all, aes(x=avg_days, color=num_trials)) +
  geom_histogram(binwidth=1, fill="white")+ # removed argument position="dodge"
  theme(legend.position="top") +
  scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9")) +
  scale_fill_manual(values=c("#999999", "#E69F00", "#56B4E9"))  

grid.arrange(s, h, nrow=1)
```

```{r}
d <- center_data(d_all, is_not_unique_data = FALSE)
```

### All

```{r}
days_model<-glm(cbind(num_flew,num_notflew)~avg_days_c, data=d, family=binomial)
summary(days_model)
```

While average days does not show the same strong effect as days_from_start, this should still control for the fact that some individuals were only tested early, and some by chance had two earlier or later days within a trial. This allows us to retain the structure with R1 and R2, which controls for ID, but still converges.

```{r}
R1 = d$num_flew
R2 = d$num_notflew
A = d$host_c
B = d$sex_c
C = d$sym_dist_s
D = d$avg_mass_logsqrt
E = d$avg_days_c

data<-data.frame(R1, R2, A, B, C, D, E)

model_script = paste0(source_path,"generic models-binomial glm 2R ~ 4-FF + E.R")
errors = withWarnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m63, m85, test="Chisq") # Adding B*D does not improve fit
anova(m26, m36, test="Chisq") # Adding C does not improve fit 
anova(m12, m26, test="Chisq") # Adding A*D does improve fit
anova(m63, m36, test="Chisq") # Adding C*D does improve fit | m63 is the best fit
```

```{r}
mass_model_all <- glm(cbind(num_flew, num_notflew) ~ host_c * average_mass_trans 
                      + sym_dist_s * average_mass_trans + sex + avg_days_c, data=d, family=binomial)

summary(mass_model_all) # matches with earlier top models but earlier top model was simplier (host*avg_mass + sex_c)
```

MLC: Because mass and morphology are so dimorphic between sexes, and sex itself has a strong direct effect, let's go now to the split-by-sex results. It may be that in wing2body and mass, we have pinpointed the reasons that sexes are different; but they differ in so many ways, that is a strong inference to try and make.

AB: significant host_c:average_mass_trans relationship where if from GRT and heavy then fly more times. (It could be a category dominated by females). This was also noticed in the top model performed without transformations and avg_days_c.

### Females

```{r}
data_fem <- d_all[d_all$sex=="F",]
data_fem <- center_data(data_fem, is_not_binded = FALSE)
```

```{r}
R1 = data_fem$num_flew
R2 = data_fem$num_notflew
A = data_fem$host_c
B = data_fem$sym_dist
C = data_fem$average_mass_trans
D = data_fem$wing2body_trans
E = data_fem$avg_days_c

data<-data.frame(R1, R2, A, B, C, D, E)

model_script = paste0(source_path,"generic models-binomial glm 2R ~ 4-FF + E.R")
errors = withWarnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m25, m13, test='Chisq') #adding A*C improves fit
anova(m25, m17, test="Chisq") #adding D improves fit
anova(m10, m13, test="Chisq") #adding A alone does not improve fit
anova(m10, m3, test="Chisq") #adding D improves fit
anova(m10, m4, test="Chisq") #adding C improves fit
anova(m25, m45, test='Chisq') #adding A*D does not improve fit | top model is m25
```

```{r}
mass_model_fem <- glm(cbind(num_flew, num_notflew) ~ host_c * average_mass_trans +  wing2body_trans + avg_days_c, data=data_fem, family=binomial) ### sym_dist does not matter for females; AB: this matches with earlier top models (host * avg_mass + wing) 

summary(mass_model_fem)
```

```{r}
# AB: conflation for females between days from start and eggs laid? Check total eggs vs. days from start (this is a rough sketch)
gf_point(total_eggs ~ days_from_start, data=data_tested) + 
  geom_smooth(method='lm')
# not a strong enough relationship
summary(lm(total_eggs ~ days_from_start, data=data_tested)) # not significant
```

### Males

```{r}
data_male <- d_all[d_all$sex=="M",]
data_male <- center_data(data_male, is_not_binded = FALSE)
```

```{r}
R1 = data_male$num_flew
R2 = data_male$num_notflew
A = data_male$host_c
B = data_male$sym_dist_s
C = data_male$average_mass_trans 
D = data_male$wing2body_trans  
E = data_male$avg_days_c

data<-data.frame(R1, R2, A, B, C, D, E)

model_script = paste0(source_path,"generic models-binomial glm 2R ~ 4-FF + E.R")
errors = withWarnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m83, m105, test="Chisq") #marginal improvement from adding C*D
anova(m50, m62, test="Chisq") #no improvement from adding C | m50 is the top model
anova(m83, m62, test="Chisq") #marginal improvement from B*C
```

```{r}
mass_model_male<-glm(cbind(num_flew, num_notflew)~host_c*wing2body_trans + sym_dist_s*wing2body_trans + avg_days_c, family=binomial, data=data_male) ## AB: host showing up matches top (simpler) model from original dataset (host + mass + days + wing but the nuance is that when used ID host was not significant and when use trial_type it was). Also strange to not see mass show up. It was significant in all the top models but not when it was the only predictor variable in a model.

summary(mass_model_male)
# Tests
summary(glm(cbind(num_flew, num_notflew)~wing2body_trans + sym_dist_s*wing2body_trans + avg_days_c, family=binomial, data=data_male))
plot(host_c ~ sym_dist_s, data=data_male)
```

* mass does not matter but morphology does by host or sym_dist

### Plots 

MLC: I just made this a function so it's easy to collapse. Probably won't stay in the final summary script, but I found it helpful for looking at these interactions. In making the data_temp summaries, data=d can be swapped for data=d[d$sex=="M",] (or F) to look at host effects within one sex only.


```{r}
six_plots<-function(){
##quick plots for y/n flight
##plot-specific grouping variables
d$mass_block<-round(d$average_mass/0.005)*0.005
d$f_prob<-d$num_flew/(d$num_flew+d$num_notflew)
d$wing2body_block<-round(d$wing2body, digits=2)
d$days_block<-round(d$avg_days, digits=0)


par(mfrow=c(2,3), mai=c(0.6, 0.6, 0.02, 0.02))
##mass by sex
data_temp<-aggregate(f_prob~sex*mass_block, data=d, FUN=mean)
data_temp$n<-aggregate(f_prob~sex*mass_block, data=d, FUN=length)$f_prob
plot(data_temp$f_prob~data_temp$mass_block, pch=c(2,19)[as.factor(data_temp$sex)], ylab="Flight probability", xlab="Mass")
##Indeed, we can see that the effect of mass is very pronounced in females but essentially absent in males.
##Really clear to see - great image.
## AB: However, based on the delta mass analyses, it might be more nuanced (but it's definitely weaker). Mass does effect males, it might just be in a smaller window because their mass doesn't change as much? Consider the male group point at 1 flight probability - it might be an outlier? If you print the data temp you see that there is only 1 male bug in that group:

print(data_temp) # I don't think that their masses play a huge role but it does follow the trend in the overall data. I guess one way to parse this out experimentally would be to see how virgin female mass impacts flight probability. Would would have a smaller mass window and maybe more dispersed data points.


##wing2body by sex
data_temp<-aggregate(f_prob~sex*wing2body_block, data=d, FUN=mean)
data_temp$n<-aggregate(f_prob~sex*wing2body_block, data=d, FUN=length)$f_prob
plot(data_temp$f_prob~data_temp$wing2body_block, pch=c(2,19)[as.factor(data_temp$sex)], ylab="Flight probability", xlab="wing-to-body ratio")
##Here we can see that as wing2body ratio increases, flight probability increases, but that the effect is much more pronounced in males.  


##average days from start by sex
data_temp<-aggregate(f_prob~sex*days_block, data=d, FUN=mean)
data_temp$n<-aggregate(f_prob~sex*days_block, data=d, FUN=length)$f_prob
plot(data_temp$f_prob~data_temp$days_block, pch=c(2,19)[as.factor(data_temp$sex)], ylab="Flight probability", xlab="Days from start")
##Here we can see that as the average days from start increases, flight probability increases, but that the effect is really only visible in females. 
##This does not mean days from start didn't impact males - only that our experimental design successfully stopped 
##that from being confounding (eg, males may have been less likely to die, or had less biased mortality by host).
## AB: Or does this mean this is a pool size issue? Also, I might argue that for both males and females they are similarly weak? Not entirely visible for females either.


##mass by host
data_temp<-aggregate(f_prob~host_c*mass_block, data=d, FUN=mean)
data_temp$n<-aggregate(f_prob~host_c*mass_block, data=d, FUN=length)$f_prob
plot(data_temp$f_prob~data_temp$mass_block, pch=19, col=c(rgb(1,0.1,0,0.8),rgb(0,1,0.8,0.8))[as.factor(data_temp$host_c)], ylab="Flight probability", xlab="Mass")
##Here, we can see that the effect of mass is clear on GRT (red) but weak on BV (blue)

print(data_temp)

##wing2body by host
data_temp<-aggregate(f_prob~host_c*wing2body_block, data=d, FUN=mean)
data_temp$n<-aggregate(f_prob~host_c*wing2body_block, data=d, FUN=length)$f_prob
plot(data_temp$f_prob~data_temp$wing2body_block, pch=19, col=c(rgb(1,0.1,0,0.8),rgb(0,1,0.8,0.8))[as.factor(data_temp$host_c)], , ylab="Flight probability", xlab="wing-to-body ratio")
##Here, we can see that the positive effect of wing2body ratio is clear on GRT (red) and on BV (blue)


##average days from start by host
data_temp<-aggregate(f_prob~host_c*days_block, data=d, FUN=mean)
data_temp$n<-aggregate(f_prob~host_c*days_block, data=d, FUN=length)$f_prob
plot(data_temp$f_prob~data_temp$days_block, pch=19, col=c(rgb(1,0.1,0,0.8),rgb(0,1,0.8,0.8))[as.factor(data_temp$host_c)], , ylab="Flight probability", xlab="Days from start")
##Here we can see no clear impact of days from start when separated by host
}
six_plots()
```

AB: Nuances in mass, host, sex, and flight probability goes back to those bivariate plots I made. You can more subtly see how each responds as they change. I know you wnat to see things more continuous, but I'm going to plot all points because it doesn't fall into those uneven group pitfalls/limitations. 

```{r}
gf_point(num_flew ~ average_mass, data=d) + 
  geom_smooth(method='lm')  +
  ggtitle("All Data: Number Times Flew vs. Average Mass (g)") # as average mass increases, the number of times flew decreases
```

Then break it between host and average mass:

```{r}
p1 <- gf_point(num_flew ~ average_mass, col=~host_plant, data=d[d$average_mass_c < 0,], title="(a)   avgmass_c <0") + 
  geom_smooth(method='lm')
p2 <- gf_point(num_flew ~ average_mass, col=~host_plant, data=d[d$average_mass_c > 0,], title="(b)   avgmass_c >0") + 
  geom_smooth(method='lm')
grid.arrange(p1,p2, nrow=2)
```

Consistent with the model and graphs: (a) If GRT and had an average mass below the mean of the population, then you flew less times. (b) But if were GRT and had an average mass above the mean of the population, then you flew more times.

```{r}
p3 <- gf_point(num_notflew ~ average_mass, col=~host_plant, data=d[d$average_mass_c < 0,], title="(a)   avgmass_c <0") + 
  geom_smooth(method='lm')
p4 <- gf_point(num_notflew ~ average_mass, col=~host_plant, data=d[d$average_mass_c > 0,], title="(b)   avgmass_c >0") + 
  geom_smooth(method='lm')
grid.arrange(p3,p4, nrow=2)
```

```{r}
p5 <- gf_point(num_flew ~ average_mass, col=~sex, data=d[d$average_mass_c < 0,], title="(a)   avgmass_c <0") + 
  geom_smooth(method='lm')
p6 <- gf_point(num_flew ~ average_mass, col=~sex, data=d[d$average_mass_c > 0,], title="(b)   avgmass_c >0") + 
  geom_smooth(method='lm')
grid.arrange(p5,p6, nrow=2)
```
```{r}
p7 <- gf_point(num_notflew ~ average_mass, col=~sex, data=d[d$average_mass_c < 0,], title="(a)   avgmass_c <0") + 
  geom_smooth(method='lm')
p8 <- gf_point(num_notflew ~ average_mass, col=~sex, data=d[d$average_mass_c > 0,], title="(b)   avgmass_c >0") + 
  geom_smooth(method='lm')
grid.arrange(p7,p8, nrow=2)
```

Most males' average mass is below the average mass of the population


```{r}
gf_point(num_flew ~ average_mass, col=~sex, data=d) + 
  geom_smooth(method='lm')
```


Note: I did not take out bugs that only flew once. That is important in order to do delta analyses and important to consider on certain limitations in these graphs (e.g. bugs would never reach 2 num_flew or 2 num_notflew). 

### Questions


Coefficients between the interactions host * average_mass_trans and sym_dist * average_mass_trans for the Males top model are conflicting. Problematic?

Coefficients between the interactions host * wing2body and sym_dist * wing2body for the Males top model are conflicting. Problematic?

noE vs. E source script? Why use the + E source script?


## Flight Speed {.tabset}

### The Data

```{r}
data <- read_flight_data("data/all_flight_data-Winter2020.csv")
data_all <- data[[1]]
data_tested <- data[[2]]

### Remove everyone who didn't fly (then remove distances = 0, if that didn't work fully)
data_flew_all <- data_tested[data_tested$flew_b == 1, ] 

### Check for low speeds
low_speeds <- data_flew_all %>%
  filter(average_speed < 0.05)

### Check for high speeds
high_speeds <- data_flew_all %>%
  filter(average_speed > 0.65)

low_speeds$flight_type# have 7 bugs with average_speed = 0 but were marked as bursters (this could be just something very short (second burst) - not enough to grant a calculation) - I decided to remove them. But one bug was continuous and had 0 distance and 0 speeds - that was bug 196 T2 set011-3-03-2020-A3_196.txt
high_speeds$flight_type # 3 bugs - also bursters. Could also be short explosive bursts but not true to the biology of these bugs (more like us blowing on them).

### Remove outliers
data_flew <- data_flew_all %>%
  filter(average_speed > 0.05) 
  
data_flew <- center_data(data_flew)

## transform mass & speed
data_flew$mass_trans<-log(data_flew$mass)-mean(log(data_flew$mass), na.rm=TRUE)

data_flew$speed_trans<-log(data_flew$average_speed)-mean(log(data_flew$average_speed), na.rm=TRUE)
```


```{r}
#######do flight types differ?
data_flew$flight_type <- relevel(data_flew$flight_type, ref="B")
tidy_regression(lmer(speed_trans~flight_type + (1|chamber) + (1|ID), data=data_flew), is_color=output_col) 
# yes, B and C differ distinctly in average speed; BC and CB are not different from C, so let's keep those in the continuous flight analyses.

####### Effect of chamber B-1, B-4, A-2, A-1 
data_flew$chamber <- relevel(data_flew$chamber, ref="A-4")
tidy_regression(lmer(speed_trans~chamber + (1|ID), data=data_flew), is_color=output_col) ###Possibly reductions in speed 

####### No effect of test date
tidy_regression(lmer(speed_trans~days_from_start + (1|chamber), data=data_flew), is_color=output_col)

####### No effect of test time
tidy_regression(lmer(average_speed~min_from_IncStart + (1|chamber), data=data_flew), is_color=output_col)
```

MLC: bursters are likely to be much less reliable than continuous flyers; so, let's exclude them.

### Non-Bursters 

C, BC, and CB data for speed

```{r}
data_flew <- data_flew %>%
  filter(!is.na(mass))
data_flew <- center_data(data_flew)

dC<-data_flew[data_flew$flight_type=="C" | data_flew$flight_type=="BC" | data_flew$flight_type=="CB" ,] 
dC <- dC %>%
  filter(!is.na(body))
dC <- center_data(dC)
```

```{r}
data<-data.frame(R=dC$speed_trans,
                 A=dC$host_c, 
                 B=dC$sex_c, 
                 C=dC$mass_trans,
                 D=dC$sym_dist_s, #I note it does not matter whether this is sym_dist or wing2body 
                 X=dC$chamber,
                 Y=dC$ID) 

model_script = paste0(source_path,"generic models-gaussian glmer 2-RF + 4-FF REMLF.R")
errors = withWarnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m19, m28, test="Chisq") #adding A does not improve fit
anova(m19, m29, test="Chisq") #adding D does not improve fit
anova(m19, m8, test="Chisq") #adding B*C interaction does improve fit | m19 best fit
```

```{r}
continuous_model<-lmer(speed_trans~sex*mass_trans + (1|ID) + (1|chamber), data=dC)

summary(continuous_model) 
```


```{r}
# QQ plot of residuals 
s.test <- paste("pval: ", shapiro.test(residuals(continuous_model))$p.value)
qqnorm(resid(continuous_model))
qqline(resid(continuous_model))
text(-2, 0.1, s.test) #  normal post-transform

# Diagnostic plot
dC['fitted_values'] = fitted.values(continuous_model)
dC['fitted_residuals'] = residuals(continuous_model)

ggplot(data = dC ) + geom_point(aes(x = fitted_values, y = fitted_residuals)) +
  geom_hline(yintercept = 0) +
  labs(title = "Plot of Fitted Residuals against Fitted Values",
       x = "Fitted Values", y = "Fitted Residuals")

# Plot of the leverage points
halfnorm(hatvalues(continuous_model))
hatvalues(continuous_model)[138]
hatvalues(continuous_model)[1]
```


### Plots

```{r}
speed_summary<-aggregate(average_speed~sex, data=dC, FUN=mean)

speed_summary$se<-aggregate(average_speed~sex, data=dC, FUN=function(x)
  sd(x)/sqrt(length(x)))$average_speed
```

MLC: There is not enough replication to break this down by sex, so we'll leave it as it is. This could be generalized, although it's probably not worth the time?

```{r}
four_plots<-function(){
##quick plots for speed
##plot-specific grouping variables
data_flew$mass_block<-round(data_flew$mass/0.005)*0.005 # AB (self-note): rounded up so the range, or block, begins 0.05 g less than the recorded mass_block value in data_tmep
data_flew$wing2body_block<-round(data_flew$wing2body, digits=2)

par(mfrow=c(2,2), mai=c(0.8, 0.8, 0.02, 0.02))
##mass by sex
data_temp<-aggregate(average_speed~sex*mass_block, data=data_flew, FUN=mean)
data_temp$n<-aggregate(average_speed~sex*mass_block, data=data_flew, FUN=length)$average_speed
plot(data_temp$average_speed~data_temp$mass_block, pch=c(2,19)[as.factor(data_temp$sex)], ylab="Flight speed", xlab="Mass")
##Unlike for flight probability, here we see a clear effect of mass on male flight speed, but not female flight speed.


##wing2body by sex
data_temp<-aggregate(average_speed~sex*wing2body_block, data=data_flew, FUN=mean)
data_temp$n<-aggregate(average_speed~sex*wing2body_block, data=data_flew, FUN=length)$average_speed
plot(data_temp$average_speed~data_temp$wing2body_block, pch=c(2,19)[as.factor(data_temp$sex)], ylab="Flight speed", xlab="wing-to-body ratio")
##Here we can see that as wing2body ratio increases, flight speed increases marginally in males, but not noticeably in females

#mass by host
data_temp<-aggregate(average_speed~host_c*mass_block, data=data_flew, FUN=mean)
data_temp$n<-aggregate(average_speed~host_c*mass_block, data=data_flew, FUN=length)$average_speed
plot(data_temp$average_speed~data_temp$mass_block, pch=19, col=c(rgb(1,0.1,0,0.8),rgb(0,1,0.8,0.8))[as.factor(data_temp$host_c)], ylab="Flight speed", xlab="Mass")
##Here we can see a weak positive effect of mass on flight speed, that does not appear to differ between hosts


##wing2body by host
data_temp<-aggregate(average_speed~host_c*wing2body_block, data=data_flew, FUN=mean)
data_temp$n<-aggregate(average_speed~host_c*wing2body_block, data=data_flew, FUN=length)$average_speed
plot(data_temp$average_speed~data_temp$wing2body_block, pch=19, col=c(rgb(1,0.1,0,0.8),rgb(0,1,0.8,0.8))[as.factor(data_temp$host_c)], ylab="Flight speed", xlab="wing-to-body ratio")
##Here we can see a potential weak positive effect of wing2body ratio on flight speed that does not differ between hosts.

}
four_plots()
```

AB: 2 females at 0.41 m/s average speed at mass 0.35 to 0.40 g and 1 male at 0.42 m/s average speed at mass 0.60 to 0.65 g

```{r}
# pwc = "Pairwise Wilcoxon test" between groups
pwc <- dC %>% 
  dunn_test(average_speed ~ sex, p.adjust.method = "bonferroni") %>%
  add_xy_position(x = "sex")
res.kruskal <- dC %>% kruskal_test(average_speed ~ sex)

ggplot(dC, aes(sex, average_speed)) + 
  geom_violin() +
    labs(
    subtitle = get_test_label(res.kruskal, detailed = TRUE), 
    caption = get_pwc_label(pwc)
    ) + 
  theme(legend.position="none") +
  xlab("sex") +
  ylab("average speed (m/s)") +
  geom_boxplot(width=0.1) +
  stat_pvalue_manual(pwc, hide.ns = TRUE)
```

Close to being significantly different between the sexes - mostly marginal 

```{r}
# check out the outliers
select(dC[dC$sex == "M" & dC$average_speed > 0.5, ], average_speed, distance, chamber) # really fast moving male bugs
```

```{r}
dC_ro <- dC[dC$average_speed < 0.51, ]

pwc <- dC_ro %>% 
  dunn_test(average_speed ~ sex, p.adjust.method = "bonferroni") %>%
  add_xy_position(x = "sex")
res.kruskal <- dC_ro %>% kruskal_test(average_speed ~ sex)
pwc
ggplot(dC_ro, aes(sex, average_speed)) + 
  geom_violin() +
    labs(
    subtitle = get_test_label(res.kruskal, detailed = TRUE), 
    caption = get_pwc_label(pwc)
    ) + 
  theme(legend.position="none") +
  xlab("sex") +
  ylab("average speed (m/s)") +
  geom_boxplot(width=0.1) +
  stat_pvalue_manual(pwc, hide.ns = TRUE)
```


AB: relationship significant if you remove 'outliers' but I wouldn't consider them outliers. Also important to remember the very different sex pool sizes: there are over 3.5 times as many males as females. Generally, across the literature/across species, speed doesn't show much difference between sexes within a species but max_speed has been used as a parameter. This parameter is tricky because I would need to go back and make sure I remove any of those first fast outliers from us blowing on them. It's possible and there are many flight parameters that we aren't using that may show us more about how these insects perform (Jones 2015) really takes advantage of many flight parameters

### Questions

Difference in pool sizes between sexes causes this hard-to-parse-out comparison?

What to do about potential "outliers"? I don't see them as outliers, just some powerful male flyers. 

Use max speed instead of average speed?

## Flight Distance {.tabset}

### The Data

```{r}
data <- read_flight_data("data/all_flight_data-Winter2020.csv")
data_all <- data[[1]]
data_tested <- data[[2]]

### Remove everyone who didn't fly (then remove distances = 0, if that didn't work fully)
data_flew <- data_tested[data_tested$flew_b == 1, ]
data_flew <- data_flew[data_flew$distance > 0, ]
data_flew <- center_data(data_flew) ##bursters are likely to be much less reliable than continuous flyers; so, let's exclude them.

data_flew <- data_flew %>%
  filter(!is.na(mass))
data_flew <- center_data(data_flew)
```

```{r}
#######do flight types differ?
data_flew$flight_type <- relevel(data_flew$flight_type, ref="B")
tidy_regression(lmer(distance~flight_type + (1|chamber) + (1|ID), data=data_flew), is_color=output_col) ###This is intensely influenced by flight_type - similar to speed, I think these are much more meaningful for continuous fliers.
```

### Non-Bursters

C, BC, and CB data for distance

```{r}
dC<-data_flew[data_flew$flight_type=="C" | data_flew$flight_type=="BC" | data_flew$flight_type=="CB" ,] 
dC <- dC %>%
  filter(!is.na(body))
dC <- center_data(dC)

##transform mass
dC$mass_trans<-log(dC$mass)-mean(log(dC$mass), na.rm=TRUE)

##transform distance
dC$distance_trans<-log(sqrt(100+dC$distance))-mean(log(sqrt(100+dC$distance)), na.rm=TRUE)
```

```{r}
####### Effect of chamber B-1 and A-1
dC$chamber <- relevel(dC$chamber, ref="B-4")
tidy_regression(lmer(distance_trans~chamber + (1|ID), data=dC), is_color=output_col) ###yes, this one's an issue 

####### No effect of test date
tidy_regression(lmer(distance_trans~days_from_start + (1|chamber), data=dC), is_color=output_col)

####### no effect of start time
tidy_regression(lmer(distance_trans~min_from_IncStart + (1|chamber), data=dC), is_color=output_col)
```

```{r}
### model transformed distance
data<-data.frame(R=dC$distance_trans,
                 A=dC$host_c, 
                 B=dC$sex_c, 
                 C=dC$mass_trans,
                 D=dC$wing2body, #Sym_dist showed up in nothing, reran with wing2body
                 X=dC$chamber,
                 Y=dC$ID) 

model_script = paste0(source_path,"generic models-gaussian glmer 2-RF + 4-FF REMLF.R")
errors = withWarnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m2, m9, test="Chisq") #adding D does not improve things
anova(m3, m8, test="Chisq") #adding B does not improve things
anova(m2, m8, test="Chisq") #adding C does not improve things
anova(m3, m10, test="Chisq") # Adding D does not improve fit
```

MLC: Once again, unclear if it's mass or sex driving the increase in distance traveled. Unfortunately, we still don't have the replication to split this by sex (at least, not for females). So, we just have to accept the uncertainty here - but we can use the delta stats to help dig into this for mass!

```{r}
distance_model_all<-lmer(distance_trans~mass_trans + (1|chamber) + (1|ID), data=dC)
summary(distance_model_all)
```

```{r}
alt_distance_model<-lmer(distance_trans~sex + (1|chamber) + (1|ID), data=dC)
summary(alt_distance_model)
```


```{r}
# QQ plot of residuals 
s.test <- paste("pval: ", shapiro.test(residuals(distance_model_all))$p.value)
qqnorm(resid(distance_model_all))
qqline(resid(distance_model_all))
text(-2, 0.1, s.test) #  normal post-transform

# Diagnostic plot
dC['fitted_values'] = fitted.values(distance_model_all)
dC['fitted_residuals'] = residuals(distance_model_all)

ggplot(data = dC ) + geom_point(aes(x = fitted_values, y = fitted_residuals)) +
  geom_hline(yintercept = 0) +
  labs(title = "Plot of Fitted Residuals against Fitted Values",
       x = "Fitted Values", y = "Fitted Residuals")

# AB: Diagnostic plot not looking so good, there's a striking linear relationship. It might not be best to actually let it pass

# Plot of the leverage points
halfnorm(hatvalues(distance_model_all))
hatvalues(distance_model_all)[138]
hatvalues(distance_model_all)[127]
```

```{r}
# QQ plot of residuals 
s.test <- paste("pval: ", shapiro.test(residuals(alt_distance_model))$p.value)
qqnorm(resid(alt_distance_model))
qqline(resid(alt_distance_model))
text(-2, 0.1, s.test) #  normal post-transform

# Diagnostic plot
dC['fitted_values'] = fitted.values(alt_distance_model)
dC['fitted_residuals'] = residuals(alt_distance_model)

ggplot(data = dC ) + geom_point(aes(x = fitted_values, y = fitted_residuals)) +
  geom_hline(yintercept = 0) +
  labs(title = "Plot of Fitted Residuals against Fitted Values",
       x = "Fitted Values", y = "Fitted Residuals")

# AB: this plot isn't looking any better

# Plot of the leverage points
halfnorm(hatvalues(alt_distance_model))
hatvalues(alt_distance_model)[167]
hatvalues(alt_distance_model)[168]
```

### Plots 

```{r}
four_plots<-function(){
##quick plots for distance
##plot-specific grouping variables
dC$mass_block<-round(dC$mass/0.005)*0.005
dC$wing2body_block<-round(dC$wing2body, digits=2)


par(mfrow=c(2,2), mai=c(0.8, 0.8, 0.02, 0.02))
##mass by sex
data_temp<-aggregate(distance~sex*mass_block, data=dC, FUN=mean)
data_temp$n<-aggregate(distance~sex*mass_block, data=dC, FUN=length)$distance
plot(data_temp$distance~data_temp$mass_block, pch=c(2,19)[as.factor(data_temp$sex)], ylab="Flight distance", xlab="Mass")
##Like flight speed, here we again see what looks like a positive effect of mass for males, but not for females


##wing2body by sex
data_temp<-aggregate(distance~sex*wing2body_block, data=dC, FUN=mean)
data_temp$n<-aggregate(distance~sex*wing2body_block, data=dC, FUN=length)$distance
plot(data_temp$distance~data_temp$wing2body_block, pch=c(2,19)[as.factor(data_temp$sex)], ylab="Flight distance", xlab="wing-to-body ratio")
##There's not much going on here.

#mass by host
data_temp<-aggregate(distance~host_c*mass_block, data=dC, FUN=mean)
data_temp$n<-aggregate(distance~host_c*mass_block, data=dC, FUN=length)$distance
plot(data_temp$distance~data_temp$mass_block, pch=19, col=c(rgb(1,0.1,0,0.8),rgb(0,1,0.8,0.8))[as.factor(data_temp$host_c)], ylab="Flight distance", xlab="Mass")
##Not much here either


##wing2body by host
data_temp<-aggregate(distance~host_c*wing2body_block, data=dC, FUN=mean)
data_temp$n<-aggregate(distance~host_c*wing2body_block, data=dC, FUN=length)$distance
plot(data_temp$distance~data_temp$wing2body_block, pch=19, col=c(rgb(1,0.1,0,0.8),rgb(0,1,0.8,0.8))[as.factor(data_temp$host_c)], ylab="Flight distance", xlab="wing-to-body ratio")
##Not much here either

}
four_plots()
```

### Questions

MLC pointed out that distance is not normal even after transforming it, but I'm really not sure if it's ok to let it pass. Seems like it's better not. We might have to use nonparametric tests like the Kruskal-Wallis.












