---
title: 'Binomial Modeling: flight_yes_no'
author: "Anastasia Bernat"
date: "3/30/2020 - 8/5/2020"
output:
  html_document: default
  pdf_document: default
---
```{r setup, include=FALSE}
rm(list=ls())
setwd("~/Desktop/git_repositories/SBB-dispersal/avbernat/Dispersal/Winter_2020/stats/")

library(lme4)
library(dplyr)
library(ggplotify)
library(gridExtra)
library(ggformula)
library(tidyselect)
library(zoo)
library(nnet) # multinom package

library(rethinking)

knitr::opts_chunk$set(echo = TRUE)
```

## Winter 2020 Flight Trials: Binomial Flight Modeling {.tabset}

Flight Trials Winter 2020 Dataset was conducted from 2/17/2020 - 3/10/2020. Soapberry bugs were flight tested twice for multiple hours in the flight mill and observed from 8 AM to (5-8 PM) each day. Used multivariate (glm) and mixed effect modeling (glmer) to analyze the flight results.

### All Trials

[Cleaning Data](#data_clean)

[Testing glmer() and Covariates](#testing)

[Without Mass](#all_womass)

[With Mass](#all_wmass)

[Morphology](#all_morph)

[Morphology + Mass + Sex](#all_mms)

<a id="data_clean"></a>

**Cleaning the Data**

```{r}
source_path = "~/Desktop/git_repositories/SBB-dispersal/avbernat/Rsrc/"

script_names = c("center_flight_data.R", # Re-centers data 
                 "clean_flight_data.R", # Loads and cleans data
                 "unique_flight_data.R",
                 "get_warnings.R", 
                 "compare_models.R",
                 "regression_output.R", # Cleans regression outputs; prints in color or B&W
                 "AICprobabilities.R",
                 "get_Akaike_weights.R")

for (script in script_names) { 
  path = paste0(source_path, script)
  source(path) 
}

output_col = FALSE # Change to TRUE if working in Base R or RStudio; FALSE if generating an HTML

data <- read_flight_data("data/all_flight_data-Winter2020.csv")
data_all <- data[[1]]
data_tested <- data[[2]]
d <- create_delta_data(data_tested)
```

```{r}
d <- center_data(d, is_not_unique_data = FALSE)
d
```

<a id="testing"></a>

**Note**

Modeling the dataset as it was before led to lots of converging issues that couldn't be easily resolved. Now we are using a modified dataset, d, to model the data.

**Testing Models and Covariates**


```{r}
# test mixed effect model
model <- glmer(cbind(num_flew, num_notflew)~sex_c*host_c + (1|population), data=d, family=binomial)
tidy_regression(model, is_color=output_col) ####MLC: updated model, but won't converge with population as a RF 
model <- glmer(cbind(num_flew, num_notflew)~sex_c*host_c + (1|site), data=d, family=binomial)
tidy_regression(model, is_color=output_col) 
getME(model, "lower")

## AB: What is a random factor here exactly? B/c it seems to matter which we use.
## AB: I tried this and it didn't work but is there a way to cbind() random factors so I can put in chamber, test date, or test time?
```

**Experimental, Biological, and Morphological Effects**

```{r, echo=FALSE}
####### Can't test chamber, test date, or test time any more
####### Can't test eggs_b b/c it's date specific

####### (Strong) Effect of average mass
tidy_regression(glm(cbind(num_flew, num_notflew)~avg_mass_c, data=d, family=binomial), is_color=output_col) 

####### Effect of number of eggs laid
tidy_regression(glm(cbind(num_flew, num_notflew)~total_eggs, data=d, family=binomial), is_color=output_col) 

####### Effect of beak length
tidy_regression(glm(cbind(num_flew, num_notflew)~beak_c, data=d, family=binomial), is_color=output_col) 

####### Effect of thorax length
tidy_regression(glm(cbind(num_flew, num_notflew)~thorax_c, data=d, family=binomial), is_color=output_col) 

####### Effect of body length
tidy_regression(glm(cbind(num_flew, num_notflew)~body_c, data=d, family=binomial), is_color=output_col) 

####### No effect of wing length
tidy_regression(glm(cbind(num_flew, num_notflew)~wing_c, data=d, family=binomial), is_color=output_col)
```

<a id="all_womass"></a>

#### Without Mass

```{r message=FALSE}
R1 = d$num_flew
R2 = d$num_notflew
A = d$host_c
B = d$sex_c
C = d$sym_dist
D = d$avg_mass_c 
X = d$population # or population get the same top model. Variance is 0 if use pop, not 0 if use site. However, variance is 0 if use site and avg_mass (down below). 

data<-data.frame(R1, R2, A, B, C, D, X)

model_script = paste0(source_path,"generic models-binomial glmer 2R ~ 1 RF + 3-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m4, m8, test="Chisq") # Adding A*B marginally improves fit if use pop as RF | marginally improves fit if use site as RF
anova(m2, m4, test="Chisq") # Adding A does not improve fit
anova(m2, m6, test="Chisq") # Adding C does not improve fit
```

```{r}
nomass_model_all <- glmer(cbind(num_flew, num_notflew) ~ sex_c + (1|population), data=d, family=binomial)
tidy_regression(nomass_model_all, is_color=output_col)

nomass_model_all <- glmer(cbind(num_flew, num_notflew) ~ sex_c + (1|site), data=d, family=binomial)
tidy_regression(nomass_model_all, is_color=output_col)
```

* strong negative effect of sex where if F then less number of times will fly

<a id="all_wmass"></a>

#### With Mass 

```{r message=FALSE}
model_script = paste0(source_path,"generic models-binomial glmer 2R ~ 1-RF + 4-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m63, m85, test="Chisq") # Adding B*D does not improve fit

anova(m26, m36, test="Chisq") # Adding C does not improve fit | is sym dist important?
anova(m12, m26, test="Chisq") # Adding A*D does improve fit
```

Wanted to check whether sym_dist is making much of an impact overall and you can see it's far from significant.

```{r}
tidy_regression(glmer(cbind(num_flew, num_notflew) ~ sym_dist + (1|population), data=d, family=binomial), is_color=output_col)
```

```{r}
# model m26 might be the better model than this because even adding C did not improve fit
# and the difference in AIC between the two isn't huge.
mass_model_all <- glmer(cbind(num_flew, num_notflew) ~ host_c * avg_mass_c +  sex_c + (1|population), data=d, family=binomial)
tidy_regression(mass_model_all, is_color=output_col)
mass_model_all <- glmer(cbind(num_flew, num_notflew) ~ host_c * avg_mass_c + sym_dist * avg_mass_c + sex_c + (1|population), data=d, family=binomial)
tidy_regression(mass_model_all, is_color=output_col)
```

Now there's a conflicting interaction between average mass and host*avg_mass...and there's 

* no effect of host
* no effect of avg_mass_c
* no effect of sym_dist

but effect of sex and interactions but the interactions are conflicting...

How do you graph these models? How would you graph cbind(num_flew, num_notflew)?

```{r}
plot( cbind(num_flew, num_notflew) ~ cbind(host_c, avg_mass), data=d , col=rangi2)
```

How does num_flew or num_notflew vary with host plant?

```{r}
coplot(num_flew ~ avg_mass_c | host_c, data=d)
coplot(num_notflew~ avg_mass_c | host_c, data=d)
```

```{r, echo=FALSE}
# error with these
m7.8 <- map( alist(
  num_flew ~ dnorm( mu , sigma ) ,
  mu <- a + bM * avg_mass_c + bH*host_c + bMH*avg_mass_c*host_c ,
  a ~ dnorm( 0 , 10 ) ,
  bM ~ dnorm( 0 , 1 ) ,
  bH ~ dnorm( 0 , 1 ) ,
  bMH ~ dnorm( 0 , 1 ) ,
  sigma ~ dunif( 0 , 1 ) 
),
data=d)

m7.9 <- map( alist(
  num_flew ~ dnorm( mu , sigma ) ,
  mu <- a + bM * avg_mass_c + bH*host_c + bMH*avg_mass_c*host_c + bS*sex_c,
  a ~ dnorm( 0 , 10 ) ,
  bM ~ dnorm( 0 , 1 ) ,
  bH ~ dnorm( 0 , 1 ) ,
  bMH ~ dnorm( 0 , 1 ) ,
  bS ~ dnorm( 0 , 1),
  sigma ~ dunif( 0 , 1 ) 
),
data=d)

#precis(m7.9) # look at how high the sd's are for bM and bMH

m7.10 <- map( alist(
  num_notflew ~ dnorm( mu , sigma ) ,
  mu <- a + bM * avg_mass_c + bH*host_c + bMH*avg_mass_c*host_c,
  a ~ dnorm( 0 , 10 ) ,
  bM ~ dnorm( 0 , 1 ) ,
  bH ~ dnorm( 0 , 1 ) ,
  bMH ~ dnorm( 0 , 1 ) ,
  sigma ~ dunif( 0 , 1 ) 
),
data=d)

#precis(m7.10)
```

Plot the interaction between host*avg_mass:

```{r echo=FALSE, fig.height=10}
par(mfrow=c(4,2)) 

avgmass.seq <- c(-1,1) # -0.3, .12
for ( m in c(-1,1) ) {
  dt <- d[d$host_c==m,]
  plot( num_flew ~ avg_mass_c, data=dt , col=rangi2 ,
        main=paste("host_c =", m) , xaxp=c(-1,1,2) ,
        xlim=c(-0.03,0.13),
        ylim=c(-1,3) ,
        xlab="average mass (centered)" )
  mu <- link( m7.8 , data=data.frame(host_c=m,avg_mass_c=avgmass.seq) ) 
  mu.mean <- apply( mu , 2 , mean )
  mu.PI <- apply( mu , 2 , PI , prob=0.97 )
  lines( avgmass.seq , mu.mean )
  lines( avgmass.seq , mu.PI[1,] , lty=2 )
  lines( avgmass.seq , mu.PI[2,] , lty=2 )
}

host.seq <- c(-1,1)
for ( h in c(-1,1)) {
  if (h == -1) {
    dt <- d[d$avg_mass_c < 0,]
    text <- paste("avg_mass_c < 0 ")
  }
    if (h == 1) {
    dt <- d[d$avg_mass_c > 0,]
    text <- paste("avg_mass_c > 0")
  }
  plot( num_flew ~ host_c, data=dt , col=rangi2 ,
        main=text , xaxp=c(-1,1,2) ,
        xlim=c(-1,1),
        ylim=c(-4,6) ,
        xlab="host (centered)" )
  mu <- link( m7.8 , data=data.frame(avg_mass_c=h,host_c=host.seq) ) 
  mu.mean <- apply( mu , 2 , mean )
  mu.PI <- apply( mu , 2 , PI , prob=0.97 )
  lines( host.seq , mu.mean )
  lines( host.seq , mu.PI[1,] , lty=2 )
  lines( host.seq , mu.PI[2,] , lty=2 )
}

avgmass.seq <- c(-1,1) # -0.3, .12
for ( m in c(-1,1) ) {
  dt <- d[d$host_c==m,]
  plot( num_notflew ~ avg_mass_c, data=dt , col=rangi2 ,
        main=paste("host_c =", m) , xaxp=c(-1,1,2) ,
        xlim=c(-0.03,0.13),
        ylim=c(-1,3) ,
        xlab="average mass (centered)" )
  mu <- link( m7.10 , data=data.frame(host_c=m,avg_mass_c=avgmass.seq) ) 
  mu.mean <- apply( mu , 2 , mean )
  mu.PI <- apply( mu , 2 , PI , prob=0.97 )
  lines( avgmass.seq , mu.mean )
  lines( avgmass.seq , mu.PI[1,] , lty=2 )
  lines( avgmass.seq , mu.PI[2,] , lty=2 )
}

host.seq <- c(-1,1)
for ( h in c(-1,1)) {
  if (h == -1) {
    dt <- d[d$avg_mass_c < 0,]
    text <- paste("avg_mass_c < 0 ")
  }
    if (h == 1) {
    dt <- d[d$avg_mass_c > 0,]
    text <- paste("avg_mass_c > 0")
  }
  plot( num_notflew ~ host_c, data=dt , col=rangi2 ,
        main=text , xaxp=c(-1,1,2) ,
        xlim=c(-1,1),
        ylim=c(-4,6) ,
        xlab="host (centered)" )
  mu <- link( m7.10 , data=data.frame(avg_mass_c=h,host_c=host.seq) ) 
  mu.mean <- apply( mu , 2 , mean )
  mu.PI <- apply( mu , 2 , PI , prob=0.97 )
  lines( host.seq , mu.mean )
  lines( host.seq , mu.PI[1,] , lty=2 )
  lines( host.seq , mu.PI[2,] , lty=2 )
}
```

Consistent with the model and graphs: If GRT and had an average mass below the mean of the population, then you flew less times. But if were GRT and had an average mass above the mean of the population, then you flew more times.

Overall mass vs. num_flew trend:

```{r}
gf_point(num_flew ~ avg_mass, data=d) + 
  geom_smooth(method='lm')
```

Then break it between host and average mass:

```{r}
p1 <- gf_point(num_flew ~ avg_mass, col=~host_plant, data=d[d$avg_mass_c < 0,], title="(a)   avgmass_c <0") + 
  geom_smooth(method='lm')
p2 <- gf_point(num_flew ~ avg_mass, col=~host_plant, data=d[d$avg_mass_c > 0,], title="(b)   avgmass_c >0") + 
  geom_smooth(method='lm')
grid.arrange(p1,p2, nrow=2)

p3 <- gf_point(num_notflew ~ avg_mass, col=~host_plant, data=d[d$avg_mass_c < 0,], title="(a)   avgmass_c <0") + 
  geom_smooth(method='lm')
p4 <- gf_point(num_notflew ~ avg_mass, col=~host_plant, data=d[d$avg_mass_c > 0,], title="(b)   avgmass_c >0") + 
  geom_smooth(method='lm')
grid.arrange(p3,p4, nrow=2)
```

GRT flyers: Even when they weighed more than the average mass of the population, they flew more times, which breaks the overall trend between number times bug flew vs. average mass which is typically negative. 

<a id="all_morph"></a>

#### Morphology 

```{r}
d <- d %>%
  filter(!is.na(body))
d$thorax_c <- d$thorax - mean(d$thorax)
d$wing_c <- d$wing - mean(d$wing)
d$body_c <- d$body - mean(d$body)

R1 = d$num_flew
R2 = d$num_notflew
A = d$thorax_c
B = d$body_c
C = d$wing_c 
X = d$population

data<-data.frame(R1, R2, A, B, C, X)

model_script = paste0(source_path,"generic models-binomial glmer 2R ~ 1 RF + 3-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m15, m17, test="Chisq") # Adding A*C does not improve fit
anova(m13, m15, test="Chisq") # Adding A*B does improve fit
```

```{r}
morph_model_all <- glmer(cbind(num_flew, num_notflew) ~ thorax_c * body_c + body_c * wing_c + (1|population), data=d, family=binomial)
tidy_regression(morph_model_all, is_color=output_col)
```
* strong negative effect of thorax where the wider the thorax, the less times the bug flies
* strong negative effect of body where the longer the body, the less times a bug flies
* strong positive effect of wing where the longer the wing, the more times the bug flies

* strong positive effect of thorax*body where the longer the body and wider the thorax, the more times a bug flies (hm seems conflicting to the single effects)
* negative effect of of body*wing where the longer the body and wing, the less times the bug flies

```{r}
R1 = d$num_flew
R2 = d$num_notflew
A = d$thorax_c
B = d$wing2body_c
X = d$population

data<-data.frame(R1, R2, A, B, X)

model_script = paste0(source_path,"generic models-binomial glmer 2R ~ 1-RF + 2-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m3, m4, test="Chisq") # Adding A*B improves fit
```

```{r}
morph_model_all2 <- glmer(cbind(num_flew, num_notflew) ~ wing2body_c * thorax_c + (1|population), data=d, family=binomial)
tidy_regression(morph_model_all2, is_color=output_col)
```

* strong positive effect of wing2body where the greater the ratio the more times a bug flies
* negative effect of thorax where the the wider the thorax the less times a bug flies
* marginal strong negative effect of wing2body*thorax where the wider the thorax and large the ratio, the less times a bug flies

```{r}
# check for collinearity
covariates <- data.frame(d$thorax, d$wing2body, d$avg_mass)
pairs(covariates)
```

Exponential relationships between mass and thorax. Thorax can only get so big - mass can increase but thorax caps out (top right). 

```{r}
covariates <- data.frame(d$thorax, d$wing, d$body, d$avg_mass)
pairs(covariates)
```

Seems like once morphology caps out so does mass not stretch too far.

<a id="all_mms"></a>

#### Morphology + Mass + Sex

```{r}
R1 = d$num_flew
R2 = d$num_notflew
A = d$thorax_c
B = d$wing2body_c
C = d$avg_mass_c
X = d$population

data<-data.frame(R1, R2, A, B, C, X)

model_script = paste0(source_path,"generic models-binomial glmer 2R ~ 1 RF + 3-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m11, m15, test="Chisq") # Adding B*C improves fit
anova(m15, m17, test="Chisq") # Adding A*C does not improve fit
```

```{r}
multi_model_all <- glmer(cbind(num_flew, num_notflew) ~ thorax_c * wing2body_c + wing2body_c * avg_mass + (1|population), data=d, family=binomial) 
tidy_regression(multi_model_all, is_color=output_col)
```

coefficients are huge. Probably a collinearity issue between morph measurements * mass.

```{r}
R1 = d$num_flew
R2 = d$num_notflew
A = d$thorax_c
B = d$wing2body_c
C = d$sex_c
X = d$population

data<-data.frame(R1, R2, A, B, C, X)

model_script = paste0(source_path,"generic models-binomial glmer 2R ~ 1 RF + 3-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m11, m15, test="Chisq") # Adding B*C does not improve fit
anova(m11, m14, test="Chisq") # Adding A*C does not improve fit
anova(m7, m11, test="Chisq") # Adding A*B does improve fit
```

```{r}
multi_model_all2 <- glmer(cbind(num_flew, num_notflew) ~ thorax_c * wing2body_c + sex_c + (1|population), data=d, family=binomial) 
tidy_regression(multi_model_all2, is_color=output_col)
```

* no effect of thorax
* strong positve effect of ratio where the larger the ratio the more times a bug flew
* negative effect of sex where if F then less times bug will fly

* strong negative effect of thorax*ratio where the larger the ratio and wider the thorax, the less times a bug flew.

```{r}
# weak positive relationship between thorax vs. wing2body 
p1 <- gf_point(thorax ~ wing2body, col=~num_flew, data=d) + 
  geom_smooth(method='lm')
p2 <- gf_point(thorax ~ wing2body, data=d[d$num_flew == 2,], title="num_flew=2") + 
  geom_smooth(method='lm')
p3 <- gf_point(thorax ~ wing2body, data=d[d$num_flew == 1,], title="num_flew=1") + 
  geom_smooth(method='lm')
p4 <- gf_point(thorax ~ wing2body, data=d[d$num_flew == 0,], title="num_flew=0") + 
  geom_smooth(method='lm')

grid.arrange(p1,p2,p3,p4, ncol=2)
```

As wing2body and thorax increases, num_flew decreases too, but its really the thorax shrinking that seems to be driving this.

```{r}
gf_point(num_flew ~ thorax, data=d)  + 
  geom_smooth(method='lm')

gf_point(num_flew ~ wing2body, data=d)  + 
  geom_smooth(method='lm')
```


```{r}
R1 = d$num_flew
R2 = d$num_notflew
A = d$thorax_s
B = d$wing2body_s
C = d$sex_c
D = d$avg_mass_s
X = d$population

data<-data.frame(R1, R2, A, B, C, D, X)

model_script = paste0(source_path,"generic models-binomial glmer 2R ~ 1-RF + 4-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
# Less fail if use standardized variables. Also, all the top models have multiple interactions b/c I'm guessing, they're correlated in some ways.
```

```{r}
anova(m58, m74, test="Chisq") # Adding B*C marginally improves fit
anova(m74, m94, test="Chisq") # Adding A*C does not improve fit
```

```{r}
multi_model <- glmer(cbind(num_flew, num_notflew) ~ thorax_s * wing2body_s + wing2body_s * avg_mass_s + sex_c + (1|population), data=d, family=binomial) 
tidy_regression(multi_model, is_color=output_col)
```

### Summary

```{r}
tidy_regression(nomass_model_all, is_color=output_col)
```

```{r}
tidy_regression(mass_model_all, is_color=output_col)
```

Why this host*avg_mass interaction? GRT flyers: Even when they weighed more than the average mass of the population, they flew more times, which breaks the overall trend between number times bug flew vs. average mass which is typically negative. (Refer to plots in the "All Trials" tab)
```{r}
tidy_regression(morph_model_all, is_color=output_col)
tidy_regression(morph_model_all2, is_color=output_col)
```

Why this wing2body*thorax interaction? As wing2body and thorax increases, num_flew decreases too, but its really the thorax shrinking that seems to be driving this.

```{r}
tidy_regression(multi_model_all, is_color=output_col)
tidy_regression(multi_model_all2, is_color=output_col)
```

Huge coefficients! I'm sure this is all due to collinearity. Thoughts?

### All Trials - Old Method

[Cleaning Data](#data_clean_old)

[Testing glmer() and Covariates](#testing_old)

[Without Mass](#old_womass)

[With Mass](#old_wmass)

[Morphology](#old_morph)

<a id="data_clean_old"></a>

**Cleaning the Data**

```{r}
data <- read_flight_data("data/all_flight_data-Winter2020.csv")
data_all <- data[[1]]
data_tested <- data[[2]]
```

<a id="testing_old"></a>

**Testing Models and Covariates**

```{r}
# test mixed effect model
model <- glmer(flew_b~sex_c*host_c + (1|ID) + (1|trial_type), data=data_tested, family=binomial)
summary(model)
getME(model, "lower")
```

**Experimental Set-Up Effects**

```{r, echo=FALSE}
####### No effect of chamber
tidy_regression(glm(flew_b~chamber, data=data_tested, family=binomial), is_color=output_col)

####### Effect of test date (but no effect of test date when you split between T1 and T2 - which makes sense)
tidy_regression(glm(flew_b~days_from_start_c, data=data_tested, family=binomial), is_color=output_col)

####### No effect of test time
tidy_regression(glm(flew_b~min_from_IncStart_c, data=data_tested, family=binomial), is_color=output_col)
```

**Biological Effects**

```{r, echo=FALSE}
####### (Strong) Effect of mass
tidy_regression(glm(flew_b~mass_c, data=data_tested, family=binomial), is_color=output_col)

####### Effect of number of eggs laid
tidy_regression(glm(flew_b~total_eggs, data=data_tested, family=binomial), is_color=output_col)

####### Effect of whether eggs were laid or not
tidy_regression(glm(flew_b~eggs_b, data=data_tested, family=binomial), is_color=output_col)
```

**Morphology Effects**

```{r, echo=FALSE}
####### Effect of beak length
tidy_regression(glm(flew_b~beak_c, data=data_tested, family=binomial), is_color=output_col)

####### Effect of thorax length
tidy_regression(glm(flew_b~thorax_c, data=data_tested, family=binomial), is_color=output_col)

####### Effect of body length
tidy_regression(glm(flew_b~body_c, data=data_tested, family=binomial), is_color=output_col)

####### No effect of wing length
tidy_regression(glm(flew_b~wing_c, data=data_tested, family=binomial), is_color=output_col)
```

<a id="old_womass"></a>

#### Without Mass (Test)

```{r}
# Remove any missing masses
data_tested <- data_tested %>%
  filter(!is.na(mass))
data_tested <- center_data(data_tested)

R = data_tested$flew_b
A = data_tested$host_c
B = data_tested$sex_c
C = data_tested$sym_dist
D = data_tested$mass_c 
X = data_tested$ID
Y = data_tested$trial_type

data<-data.frame(R, A, B, C, D, X, Y)

model_script = paste0(source_path,"generic models-binomial glmer 2 RF + 3-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m42, m46, test="Chisq") # Adding A*B marginally improves fit
anova(m40, m42, test="Chisq") # Adding A does not improve fit
```

```{r}
nomass_model_all_old<-glmer(flew_b~sex_c*host_c + (1|ID) + (1|trial_type), family=binomial, data=data_tested)
tidy_regression(nomass_model_all_old, is_color=output_col)
```

* strong negative effect of sex where if F then less likely to fly
* no effect of host
* marginal effect of host*sex 

Probbaly females from GRT are flyers becuse there are the colonizers of the population.

<a id="old_wmass"></a>

#### With Mass (Test)

```{r}
# model_script = paste0(source_path,"generic models-binomial glmer 2-RF + 4-FF.R")
# errors = catch_warnings(model_comparisonsAIC(model_script))
# cat("Number of models that failed to converge: ", length(errors$warnings))
# too many models failed to even keep running
```

<a id="old_morph"></a>

#### Morphology 

```{r}
data_tested <- data_tested %>%
  filter(!is.na(body))
data_tested <- center_data(data_tested)

R = data_tested$flew_b
A = data_tested$thorax_c
B = data_tested$body_c
C = data_tested$wing_c 
X = data_tested$ID
Y = data_tested$trial_type

data<-data.frame(R, A, B, C, X, Y)

model_script = paste0(source_path,"generic models-binomial glmer 2 RF + 3-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m51, m53, test="Chisq") # Adding A*B improves fit
```

```{r}
morph_model_all_old<-glmer(flew_b~ thorax_c * body_c + body_c * wing_c + (1|ID), family=binomial, data=data_tested) 
tidy_regression(morph_model_all_old, is_color=output_col) 
```

```{r}
R = data_tested$flew_b
A = data_tested$thorax_c
B = data_tested$wing2body_c
X = data_tested$ID
Y = data_tested$trial_type

data<-data.frame(R, A, B, C, X, Y)

model_script = paste0(source_path,"generic models-binomial glmer 2-RF + 2-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m13, m14, test="Chisq") # Adding A*B does not improve fit
```

```{r}
morph_model_all_old2<-glmer(flew_b~ thorax_c + wing2body_c + (1|trial_type) + (1|ID), family=binomial, data=data_tested)
tidy_regression(morph_model_all_old2, is_color=output_col) # coefficients are huge...
```

### Summary

```{r}
tidy_regression(nomass_model_all_old, is_color=output_col)
```

It's possible that females from GRT are more likley to fly because they are colonizers.

```{r}
# There isn't a mass_model_all_old because too many models failed to converge 
```

```{r}
tidy_regression(morph_model_all_old, is_color=output_col) # same morph model as the morph_model_all
tidy_regression(morph_model_all_old2, is_color=output_col) 
```

This morph model is a better fit model than the morph_model_all_old model, but HUGE coefficient for wing2body_c. Why?

### Trial 1

[Cleaning Data](#clean_data)

[Testing Covariates](#exp_effects)

[Testing glmer() with pop and chamber](#glmertesting1)

[Without Mass Modeling](#without_mass)

[With Mass Modeling](#with_mass)

[Morphology Modeling](#morph1)

<a id="clean_data"></a>

**Cleaning the Data**

```{r}
data <- read_flight_data("data/all_flight_data-Winter2020.csv")
data_all <- data[[1]]
data_tested <- data[[2]]

data_T1 <-data_tested[data_tested$trial_type=="T1",]
data_T1 <- center_data(data_T1)
```

<a id="exp_effects"></a>

**Experimental Set-Up Effects**

```{r echo=FALSE}
####### No effect of chamber
tidy_regression(glm(flew_b~chamber, data=data_T1, family=binomial), is_color=output_col)

####### No effect of test date
tidy_regression(glm(flew_b~days_from_start_c, data=data_T1, family=binomial), is_color=output_col)

####### No effect of test time
tidy_regression(glm(flew_b~min_from_IncStart, data=data_T1, family=binomial), is_color=output_col)
```

**Biological Effects**

```{r, echo=FALSE}
####### (Strong) Effect of mass
tidy_regression(glm(flew_b~mass_c, data=data_T1, family=binomial), is_color=output_col)
```

**Morphology Effects**

```{r, echo=FALSE}
####### Effect of beak length
tidy_regression(glm(flew_b~beak_c, data=data_T1, family=binomial), is_color=output_col)

####### Effect of thorax length
tidy_regression(glm(flew_b~thorax_c, data=data_T1, family=binomial), is_color=output_col)

####### No effect of body length
tidy_regression(glm(flew_b~body_c, data=data_T1, family=binomial), is_color=output_col)

####### No effect of wing length
tidy_regression(glm(flew_b~wing_c, data=data_T1, family=binomial), is_color=output_col)
```

<a id="glmertesting1"></a>

**Glmer() Testing Notes**

(1|chamber) as a random effect does not need to be included in the Trial 1 glmer() analyses because in all cases the variance is essentially 0. What pops up if you do try glmer on the best fit model? A singular fit error is raised: “When you obtain a singular fit, this is often indicating that the model is overfitted – that is, the random effects structure is too complex to be supported by the data, which naturally leads to the advice to remove the most complex part of the random effects structure (usually random slopes). The benefit of this approach is that it leads to a more parsimonious (conservative) model that is not over-fitted.” Source: https://stats.stackexchange.com/questions/378939/dealing-with-singular-fit-in-mixed-models

However, (1|population) and (1|days_from_start_c) does matter when doing the morphology analyses in Trial 1, and will not throw an error because the variance is large enough to change the fit of a model.

<a id="without_mass"></a>

#### Without Mass

```{r}
# Remove any missing masses
data_T1 <- data_T1 %>%
  filter(!is.na(mass))
data_T1 <- center_data(data_T1)

R = data_T1$flew_b
A = data_T1$host_c
B = data_T1$sex_c
C = data_T1$sym_dist
D = data_T1$mass_c 

data<-data.frame(R, A, B, C, D)

model_script = paste0(source_path,"generic models-binomial glm 3-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m8, m11, test="Chisq") ## Adding C does not improve fit
anova(m11, m15, test="Chisq") ## Adding B*C interaction does not improve fit
anova(m11, m14, test="Chisq") ## Adding A*C interaction does not improve fit 
```

```{r}
nomass_T1<-glm(flew_b~sex_c*host_c, family=binomial, data=data_T1)
tidy_regression(nomass_T1, is_color=output_col)
```
* strong negative effect of sex, where if you are female you are less likely to fly.
* No direct effect of host plant
* strong positive effect of sex and host where if female and from GRT, then more likely to fly.

<a id="with_mass"></a>

#### With Mass

```{r}
model_script = paste0(source_path,"generic models-binomial glm 4-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))

R = data_T1$flew_b
A = data_T1$host_c
B = data_T1$sex_c
C = data_T1$mass_c 

data<-data.frame(R, A, B, C)

model_script = paste0(source_path,"generic models-binomial glm 3-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m9, m12, test="Chisq") # Adding B does not improve fit
```

```{r}
anova(m51, m63, test="Chisq") # Adding B does not improve fit
anova(m27, m51, test="Chisq") # Adding C*D does improve fit
```

```{r}
# top model if don't include sym_dist
mass_T1_nosym <- glm(flew_b~host_c*mass_c, family=binomial, data=data_T1)
tidy_regression(mass_T1_nosym, is_color=output_col)

# top model if you do include sym_dist
mass_T1_sym<- glm(flew_b~host_c*mass_c + sym_dist*mass_c, family=binomial, data=data_T1)
tidy_regression(mass_T1_sym, is_color=output_col)
```

<a id="morph1"></a>

#### Morphology 

```{r}
data_T1 <-data_tested[data_tested$trial_type=="T1",]
data_T1 <- data_T1 %>%
  filter(!is.na(body))
data_T1 <- center_data(data_T1)

R = data_T1$flew_b
A = data_T1$thorax_c
B = data_T1$body_c
C = data_T1$wing_c 

data<-data.frame(R, A, B, C)

model_script = paste0(source_path,"generic models-binomial glm 3-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m7, m13, test="Chisq") # Adding B*C marginally improves fit
anova(m13, m16, test="Chisq") # Adding A*C does not improve fit
```


```{r}
morph_T1 <- glm(flew_b ~ body_c * wing_c + thorax_c, family = binomial, data = data_T1)
tidy_regression(morph_T1, is_color=output_col)
```

* maringal negative effect of body wher ethe larger the body the less likely to fly
* strong positive effect of wing length where the longer the wing, the more likely to fly
* storng negative effect of thorax where the larger the thorasxs, the less likely to fly

* no effect of body*wing (we know body and wing are closely related so they could be canceling each other out (b/c body is a proxi for mass))

```{r}
morph_T1glmer <- glmer(formula = flew_b ~  body_c * wing_c + thorax_c + (1| population), family = binomial, data = data_T1) 
tidy_regression(morph_T1glmer, is_color=output_col) # did change the coefficients slightly, but not a better fit
```

```{r}
R = data_T1$flew_b
A = data_T1$thorax_c
B = data_T1$wing2body_c

data<-data.frame(R, A, B)

model_script = paste0(source_path,"generic models-binomial glm 2-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m3, m4, test="Chisq") # Adding A*B does not improve fit
```

```{r}
morph_T1_ratio <- glm(flew_b ~ wing2body_c + thorax_c, family = binomial, data = data_T1)
tidy_regression(morph_T1_ratio, is_color=output_col)
```

### Summary

```{r}
tidy_regression(nomass_T1, is_color=output_col)
tidy_regression(mass_T1_sym, is_color=output_col)
tidy_regression(mass_T1_nosym, is_color=output_col)
tidy_regression(morph_T1, is_color=output_col) # glmer did not improve fit
tidy_regression(morph_T1_ratio, is_color=output_col)
```

### Trial 2

[Cleaning Data](#clean)

[Testing Covariates](#experimental_effects)

[Testing glmer() with pop and chamber](#glmertesting2)

[Without Mass Modeling](#wo_mass)

[With Mass Modeling](#w_mass)

[Morphology Modeling](#morph2)

<a id="clean"></a>

**Cleaning the Data**

```{r}
data <- read_flight_data("data/all_flight_data-Winter2020.csv")
data_all <- data[[1]]
data_tested <- data[[2]]

data_T2 <-data_tested[data_tested$trial_type=="T2",]
data_T2 <- center_data(data_T2)
```

<a id="experimental_effects"></a>

**Experimental Set-Up Effects**

```{r echo=FALSE}
####### No effect of chamber
tidy_regression(glm(flew_b~chamber, data=data_T2, family=binomial), is_color=output_col)

####### No effect of test date
tidy_regression(glm(flew_b~days_from_start_c, data=data_T2, family=binomial), is_color=output_col)

####### No effect of test time
tidy_regression(glm(flew_b~min_from_IncStart, data=data_T2, family=binomial), is_color=output_col)
```

**Biological Effects**

```{r, echo=FALSE}
####### (Strong) Effect of mass
tidy_regression(glm(flew_b~mass_c, data=data_T2, family=binomial), is_color=output_col)
```

**Morphology Effects**

```{r, echo=FALSE}
####### Effect of beak length
tidy_regression(glm(flew_b~beak_c, data=data_T2, family=binomial), is_color=output_col)

####### Effect of thorax length
tidy_regression(glm(flew_b~thorax_c, data=data_T2, family=binomial), is_color=output_col)

####### Effect of body length
tidy_regression(glm(flew_b~body_c, data=data_T2, family=binomial), is_color=output_col)

####### Marginal effect of wing length
tidy_regression(glm(flew_b~wing_c, data=data_T2, family=binomial), is_color=output_col)
```

<a id="wo_mass"></a>

<a id="glmertesting2"></a>

**Glmer() Testing Notes**

(1|population) + (1|chamber) as random effects do not need to be included in the Trial 2 glmer() mass and no mass analyses because in all cases the variance is essentially 0. What pops up if you do try glmer on the best fit model? A singular fit error is raised. Refer to the reasoning in Trial 1.

However, no error is thrown when included in the morphology analyses.

#### Without Mass

```{r}
R = data_T2$flew_b
A = data_T2$host_c
B = data_T2$sex_c
C = data_T2$sym_dist
D = data_T2$mass_c # as a covariate 

data<-data.frame(R, A, B, C, D)

model_script = paste0(source_path,"generic models-binomial glm 3-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m2, m4, test="Chisq") # Adding A does not improve fit
anova(m2, m6, test="Chisq") # Adding C does not improve fit
```

```{r}
nomass_T2 <-glm(flew_b~sex_c, family=binomial, data=data_T2)
tidy_regression(nomass_T2, is_color=output_col)
```

* sex is the only significant and strong effect, so that if you are female you are much less likely to disperse/fly

<a id="w_mass"></a>

#### With Mass

```{r}
model_script = paste0(source_path,"generic models-binomial glm 4-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m4, m7, test="Chisq") # Adding A does not improve fit
anova(m7, m12, test="Chisq") # Adding B does not improve fit
anova(m4, m9, test="Chisq") # Adding B does not improve fit
```

```{r}
mass_T2 <-glm(flew_b~mass_c, family=binomial, data=data_T2)
tidy_regression(mass_T2, is_color=output_col)
```

* mass is the only significant and extremely strong effect, so that if you are heavy you are much less likely to disperse/fly

Mass is really dominating everything, so let's split by sex after looking at Trial 2.

<a id="morph2"></a>

#### Morphology 

```{r}
# Remove any missing masses and morphology measurements
data_T2 <- data_T2 %>%
  filter(!is.na(mass))  %>%
  filter(!is.na(body))

data_T2 <- center_data(data_T2)

R = data_T2$flew_b
A = data_T2$thorax_c
B = data_T2$body_c
C = data_T2$wing_c 

data<-data.frame(R, A, B, C)

model_script = paste0(source_path,"generic models-binomial glm 3-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m13, m16, test="Chisq") # Adding A*C improves fit
```

```{r}
morph_T2 <- glm(flew_b ~ thorax_c * wing_c + body_c * wing_c, family = binomial, data = data_T2)
tidy_regression(morph_T2, is_color=output_col)
```

* negative marginal effect of thorax, where the wider the thorax, the then bug less likely to fly
* positive effect of wing length, where if wing longer then bug more likely to fly
* negative effect of body, where the longer the body the less likely the bug is to fly

* positive effect of thorax:wing where the wider the thorax and wing the more likely the bug is to fly *(this encapsulates overall flight power in a way b/c thorax muscles and wing length)*
* negative effect of wing and body where the longer the wing and body the less likely the bug is to fly *(this encapsulates mass limitations)*


```{r}
R = data_T2$flew_b
A = data_T2$thorax_c
B = data_T2$wing2body_c

data<-data.frame(R, A, B)

model_script = paste0(source_path,"generic models-binomial glm 2-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m3, m4, test="Chisq") # Adding A*B does not improve fit
```

```{r}
morph_T2_ratio <- glm(flew_b ~ thorax_c + wing2body_c, family = binomial, data = data_T2)
tidy_regression(morph_T2_ratio, is_color=output_col)
```

### Summary

```{r}
tidy_regression(nomass_T2, is_color=output_col)
tidy_regression(mass_T2, is_color=output_col)
tidy_regression(morph_T2, is_color=output_col) # glmer did not improve fit
tidy_regression(morph_T2_ratio, is_color=output_col)
```

### Females 

[Cleaning Data](#cleanup_data_fem)

[Testing Covariates](#fem_exp)

[Without Mass Modeling](#without.mass.fem)

[With Mass Modeling](#with.mass.fem)

[Eggs Modeling](#egg.fem)

[Morphology Modeling](#morph.fem)

<a id="cleanup_data_fem"></a>

**Cleaning the Data**

```{r}
data <- read_flight_data("data/all_flight_data-Winter2020.csv")
data_all <- data[[1]]
data_tested <- data[[2]]
d = create_delta_data(data_tested)
```

```{r}
data_fem <- d[d$sex=="F",]
data_fem <- center_data(data_fem, is_not_unique_data = FALSE)
data_fem
```
<a id="fem_exp"></a>

**Experimental, Biological, and Morphological Effects**

```{r, echo=FALSE}
####### Can't test chamber, test date, or test time any more
####### Can't specific mass anymore - only average mass. (Could do mass diff?)

####### (Strong) Effect of average mass
tidy_regression(glm(cbind(num_flew, num_notflew)~avg_mass, data=data_fem, family=binomial), is_color=output_col) 

####### Effect of number of eggs laid
tidy_regression(glm(cbind(num_flew, num_notflew)~total_eggs, data=data_fem, family=binomial), is_color=output_col) 

####### Effect of beak length
tidy_regression(glm(cbind(num_flew, num_notflew)~beak_c, data=data_fem, family=binomial), is_color=output_col) 

####### No effect of thorax length
tidy_regression(glm(cbind(num_flew, num_notflew)~thorax_c, data=data_fem, family=binomial), is_color=output_col) 

####### Effect of body length
tidy_regression(glm(cbind(num_flew, num_notflew)~body_c, data=data_fem, family=binomial), is_color=output_col) 

####### Effect of wing length
tidy_regression(glm(cbind(num_flew, num_notflew)~wing_c, data=data_fem, family=binomial), is_color=output_col)
```

<a id="without.mass.fem"></a>

#### Without Mass

```{r message=FALSE}
R1 = data_fem$num_flew
R2 = data_fem$num_notflew
A = data_fem$host_c
B = data_fem$wing_c
C = data_fem$avg_mass 
X = data_fem$population # can't do trial type anymore

data<-data.frame(R1, R2, A, B, C, X)

model_script = paste0(source_path,"generic models-binomial glmer 2R ~ 1 RF + 2-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m0, m2, test="Chisq") # Adding B improves fit
anova(m2, m3, test="Chisq") # Adding A does not improve fit
```

```{r}
nomass_fem <- glmer(cbind(num_flew, num_notflew) ~ wing_c + (1|population), data=data_fem, family=binomial)
tidy_regression(nomass_fem, is_color=output_col)
```

* positive effect of wing length where the longer the wing the more times the bug will fly

<a id="with.mass.fem"></a>

#### With Mass 

```{r message=FALSE}
model_script = paste0(source_path,"generic models-binomial glmer 2R ~ 1 RF + 3-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m7, m12, test="Chisq") # Adding A*C improves fit
anova(m12, m14, test="Chisq") # Adding A*B does not improve fit
```

```{r}
mass_fem <- glmer(cbind(num_flew, num_notflew) ~ host_c * avg_mass + wing_c + (1|population), data=data_fem, family=binomial)
tidy_regression(mass_fem, is_color=output_col)
```

* negative effect of host where if from GRT the less times a bug will fly
* strong negative effect of average mass where the heavier the less times a bug will fly
* positive effect of wing length where the longer the wing the more times the bug will fly

* strong positive effect of host*average mass where the heavier the bug and from GRT the more times the bug will fly (I've seen these strong conflicting interactions a lot with host and something else)

```{r}
# check for collinearity between mass and wing length
data<-data.frame(data_fem$host_c, data_fem$wing_c, data_fem$avg_mass )
colnames(data) <- c("Host Plant", "Wing Length", "Average Mass")
pairs(data)
```

GRT bugs weigh less.

<a id="egg.fem"></a>

#### Eggs

```{r message=FALSE} 
R1 = data_fem$num_flew
R2 = data_fem$num_notflew
A = data_fem$host_c
B = data_fem$wing_c
C = data_fem$avg_mass 
D = data_fem$total_eggs_c # can't use eggs_b anymore
X = data_fem$population 

data<-data.frame(R1, R2, A, B, C, D, X)

model_script = paste0(source_path,"generic models-binomial glmer 2R ~ 1-RF + 4-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m9, m20, test="Chisq") # Adding B*D does not improve fit
```

```{r}
egg_model <- glmer(cbind(num_flew, num_notflew) ~ wing_c + total_eggs_c + (1|population), data=data_fem, family=binomial)
tidy_regression(egg_model, is_color=output_col)
```

* positive effect of wing length
* negative effect of total eggs

<a id="morph.fem"></a>

#### Morphology 

```{r message=FALSE}
d <- data_fem %>%
  filter(!is.na(body))
d$thorax_c <- d$thorax - mean(d$thorax)
d$wing_c <- d$wing - mean(d$wing)
d$body_c <- d$body - mean(d$body)

R1 = d$num_flew
R2 = d$num_notflew
A = d$thorax_c
B = d$body_c
C = d$wing_c 
X = d$population

data<-data.frame(R1, R2, A, B, C, X)

model_script = paste0(source_path,"generic models-binomial glmer 2R ~ 1 RF + 3-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m15, m16, test="Chisq") # replacing A*B with A*C improves fit
anova(m13, m16, test="Chisq") # Adding B*C improves fit
```

```{r}
morph_fem <- glmer(cbind(num_flew, num_notflew) ~ thorax_c * wing_c + body_c * wing_c + (1 | population), data=d, family=binomial)
tidy_regression(morph_fem, is_color=output_col)
```
* strong negative effect of thorax where the wider the thorax, the less times the bug flies
* no effect of wing length
* no effect of body length

* strong positive (marginal) effect of thorax*body where the longer the body and wider the thorax, the more times a bug flies (seems to conflict the single effects)
* negative effect of of body*wing where the longer the body and wing, the less times the bug flies

```{r message=FALSE}
R1 = d$num_flew
R2 = d$num_notflew
A = d$thorax_c
B = d$wing2body_c
X = d$population

data<-data.frame(R1, R2, A, B, X)

model_script = paste0(source_path,"generic models-binomial glmer 2R ~ 1 RF + 2-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m0, m2, test="Chisq") # Adding B marginally improves fit
anova(m2, m3, test="Chisq") # Adding B does not improve fit
```

```{r}
morph_fem2 <- glmer(cbind(num_flew, num_notflew) ~ wing2body_c + (1 | population), data=d, family=binomial)
tidy_regression(morph_fem2, is_color=output_col)
```

* marginal strong positive effect of wing:body where the longer the body and heavier the body, the more times a female bug flew. (hm)

### Summary

```{r}
tidy_regression(nomass_fem, is_color=output_col)
tidy_regression(mass_fem, is_color=output_col)
tidy_regression(egg_model, is_color=output_col)
tidy_regression(morph_fem, is_color=output_col)
tidy_regression(morph_fem2, is_color=output_col)
```

### Females - Old Method

[Cleaning Data](#cleanup_data)

[Testing Covariates](#expset-up_effects)

[Without Mass Modeling](#without.mass)

[With Mass Modeling](#with.mass)

* [glmer() A=host, B=mass, C=wing, X=ID](#glmer)

[Eggs Modeling](#eggs)

* [glmer() A=host, B=wing_c, C=mass, D=eggs_b, X=ID](#glmer_eggs)

[Morphology Modeling](#morph_fem)

* [glmer() A=thorax, B=body, C=wing X=trial_type](#glmer_morph)

<a id="cleanup_data"></a>

**Cleaning the Data**

```{r}
data <- read_flight_data("data/all_flight_data-Winter2020.csv")
data_all <- data[[1]]
data_tested <- data[[2]]

data_fem <- data_tested[data_tested$sex=="F",]
data_fem <- center_data(data_fem)
```

<a id="expset-up_effects"></a>

**Experimental Set-Up Effects**

```{r echo=FALSE}
####### Marginal effect of chamber B-4 (ah this again!)
tidy_regression(glm(flew_b~chamber, data=data_fem, family=binomial), is_color=output_col)

####### No effect of test date (almost marginal)
tidy_regression(glm(flew_b~days_from_start, data=data_fem, family=binomial), is_color=output_col)

####### No effect of test time
tidy_regression(glm(flew_b~min_from_IncStart, data=data_fem, family=binomial), is_color=output_col)
```

**Biological Effects**

```{r echo=FALSE}
####### (Strong) Effect of mass  
tidy_regression(glm(flew_b~mass_c, data=data_fem, family=binomial), is_color=output_col)

####### Effect of number of eggs laid
tidy_regression(glm(flew_b~total_eggs_c, data=data_fem, family=binomial), is_color=output_col) 

####### Effect of whether eggs were laid or not
tidy_regression(glm(flew_b~eggs_b, data=data_fem, family=binomial), is_color=output_col) 
```

**Morphology Effects**

```{r echo=FALSE}
####### Effect of beak length
tidy_regression(glm(flew_b~beak_c, data=data_fem, family=binomial), is_color=output_col)  

####### No effect of thorax length
tidy_regression(glm(flew_b~thorax_c, data=data_fem, family=binomial), is_color=output_col) 

####### Effect of body length
tidy_regression(glm(flew_b~body_c, data=data_fem, family=binomial), is_color=output_col) 

####### Effect of wing length
tidy_regression(glm(flew_b~wing_c, data=data_fem, family=binomial), is_color=output_col) 
```

<a id="without.mass"></a>

#### Without Mass

```{r}
R = data_fem$flew_b
A = data_fem$host_c
B = data_fem$wing_c # before had sym_dist but now leave out because did not have significant effects or interactions within the total model dataset.
C = data_fem$mass_c

data<-data.frame(R, A, B, C)

model_script = paste0(source_path,"generic models-binomial glm 2-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m0, m2, test="Chisq") # Adding B does improve fit
anova(m2, m3, test="Chisq") # Adding A does not improve fit
```

```{r echo=FALSE}
nomass_fem2 <-glm(flew_b~wing_c, family=binomial, data=data_fem)
tidy_regression(nomass_fem2, is_color=output_col) 
```

* positive effect of wing length where the longer the wing then female more likely to fly.

<a id="with.mass"></a>

#### With Mass

```{r}
data_fem <- data_fem %>%
  filter(!is.na(mass))
data_fem <- center_data(data_fem)

R = data_fem$flew_b
A = data_fem$host_c
B = data_fem$wing_c # used to be sym_dis
C = data_fem$mass_c

data<-data.frame(R, A, B, C)

model_script = paste0(source_path,"generic models-binomial glm 3-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m3, m6, test="Chisq") # Adding B does improve fit
anova(m3, m5, test="Chisq") # Adding A does not improve fit
anova(m6, m7, test="Chisq") # Adding A does not improve fit
anova(m6, m10, test="Chisq") # Adding B*C does not improve fit
```

```{r echo=FALSE}
mass_fem2 <-glm(flew_b~wing_c + mass_c, family=binomial, data=data_fem)
tidy_regression(mass_fem2, is_color=output_col) 
```

* large negative effect of mass, where the heavier the bug is the less likely the bus will fly/disperse.
* positve effect of wing morph where the longer the wing then more likely to disperse.

<a id="glmer"></a>

#### Consider for females: A=host, B=mass, C=wing, X=ID

```{r}
R = data_fem$flew_b
A = data_fem$host_c
B = data_fem$mass_c
C = data_fem$wing_c # replaced sym_dist
X = data_fem$ID  # variance is zero if use trial_type | 3 models fail to converge if use ID

data<-data.frame(R, A, B, C, X)

model_script = paste0(source_path,"generic models-binomial glmer 1 RF + 3-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
# top model if use ID:
multi_glmer_fem <-glmer(flew_b~host_c * mass_c + host_c * wing_c + (1 | ID), family=binomial, data=data_fem)
tidy_regression(multi_glmer_fem, is_color=output_col) 

# top model if use trial_type:
multi_glmer_fem2 <-glmer(flew_b~mass_c + wing_c + (1|trial_type), family=binomial, data=data_fem)
# if you replace trial_type with ID then you'll see the coefficients are way too huge and at different scales. Why?
tidy_regression(multi_glmer_fem2, is_color=output_col) 
```

<a id="eggs"></a>

#### Eggs

```{r}
R = data_fem$flew_b
A = data_fem$host_c
B = data_fem$wing_c
C = data_fem$mass_c
D = data_fem$eggs_b
  
data<-data.frame(R, A, B, C, D)

model_script = paste0(source_path,"generic models-binomial glm 4-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
egg_model2 <-glm(flew_b~wing_c + mass_c + eggs_b, family=binomial, data=data_fem)
tidy_regression(egg_model2, is_color=output_col)
```

* Strong negative effect if laid eggs that day
* strong marginal negative effect of mass
* positive effect of wing length

```{r}
egg_glmer <-glmer(flew_b~wing_c + mass_c + eggs_b + (1|ID), family=binomial, data=data_fem) 
tidy_regression(egg_glmer, is_color=output_col) # model fits better
```

still a strong negative effect if laid eggs that day but mass and wing length are no longer significant...

<a id="glmer_eggs"></a>

#### glmer() A=host, B=wing_c, C=mass, D=eggs_b, X=ID

```{r}
R = data_fem$flew_b
A = data_fem$host_c
B = data_fem$wing_c
C = data_fem$mass_c
D = data_fem$eggs_b
X = data_fem$ID
  
data<-data.frame(R, A, B, C, D, X)

model_script = paste0(source_path,"generic models-gaussian glmer 1-RF + 4-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m21, m33, test="Chisq") # Adding B improves fits
anova(m33, m53, test="Chisq") # Adding B*C does not improve fit
```

```{r}
egg_glmer2<-glmer(flew_b~mass_c *eggs_b + wing_c + (1|ID), family=binomial, data=data_fem) 
tidy_regression(egg_glmer2, is_color=output_col)
```

coefficients are extremely out of scale.
* negative effect of mass
* negative effect of eggs
*positive effect of wing
* positivie effect mass*eggs? doesn't really make sense. --> collinearity

<a id="morph_fem"></a>

#### Morphology

```{r}
R = data_fem$flew_b
A = data_fem$thorax_c
B = data_fem$body_c
C = data_fem$wing_c 

data<-data.frame(R, A, B, C)

model_script = paste0(source_path,"generic models-binomial glm 3-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m13, m16, test="Chisq") # Adding A*C improves fit
```

```{r}
morph_fem2 <- glm(flew_b ~ thorax_c * wing_c + body_c * wing_c, data=data_fem)
tidy_regression(morph_fem2, is_color=output_col)
```
* no significant effects 

```{r}
morph_fem_glmer <- glmer(flew_b ~ thorax_c * wing_c + body_c * wing_c + (1|ID), data=data_fem, family=binomial)
tidy_regression(morph_fem_glmer, is_color=output_col) #even better fitted model; effects no longer significant and coefficients changed. 
```

**Trying wing2body**

```{r}
R = data_fem$flew_b
A = data_fem$thorax_c
B = data_fem$wing2body_c

data<-data.frame(R, A, B)

model_script = paste0(source_path,"generic models-binomial glm 2-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m0, m2, test="Chisq") # Adding B improves fit
anova(m2, m3, test="Chisq") # Adding A does not improve fit
```

```{r}
morph_fem3 <- glm(flew_b ~ wing2body_c, data=data_fem, family=binomial) # variance of population and days_from_start_c = 0
tidy_regression(morph_fem3, is_color=output_col)
```

* strong positive effect of wing2body where the larger the ration (the more wing than body), the more likely the female bug disperses

Worth running a glmer() script.

<a id="glmer_morph"></a>

#### Consider for females morph: A=thorax, B=body, C=wing X=trial_type

```{r}
R = data_fem$flew_b
A = data_fem$thorax_c
B = data_fem$body_c
C = data_fem$wing_c
X = data_fem$ID # you get a different set of top models if you use trial_type or ID, you get the null model if ID and the usual model if trial_type

data<-data.frame(R, A, B, C, X)

model_script = paste0(source_path,"generic models-binomial glmer 1 RF + 3-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m0, m3, test="Chisq") # Adding C does not improve fit
anova(m0, m2, test="Chisq") # Adding B does not improve fit
anova(m0, m1, test="Chisq") # Adding A does not improve fit
```

```{r}
# when use ID and trial type as RFs get the null model as the top model
morph_fem_glmer3 <- glmer(flew_b ~  1 + (1|ID), data=data_fem, family=binomial)
tidy_regression(morph_fem_glmer3 , is_color=output_col) 
# when use only trial type as RF:
morph_fem_glmer4 <- glmer(flew_b ~  thorax_c * wing_c + body_c * wing_c + (1|trial_type), data=data_fem, family=binomial) 
tidy_regression(morph_fem_glmer4 , is_color=output_col) 
```

**Trying wing2body**

```{r}
R = data_fem$flew_b
A = data_fem$thorax_c
B = data_fem$wing2body_c
X = data_fem$trial_type # you get a different set of top models if you use trial_type or ID

data<-data.frame(R, A, B, X)

model_script = paste0(source_path,"generic models-binomial glmer 1-RF + 2-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m0, m2, test="Chisq") # Adding A marginally imprvoes fit
anova(m0, m1, test="Chisq")
```

```{r}
# best fit model if only use trial type as RF:
morph_fem_glmer5 <- glmer(flew_b ~  wing2body_c + (1|trial_type), data=data_fem, family=binomial) 
tidy_regression(morph_fem_glmer5 , is_color=output_col) 
# best fit model is the null if use both trial type and ID as RFs: 
morph_fem_glmer6 <- glmer(flew_b ~  1 + (1|ID), data=data_fem, family=binomial)
tidy_regression(morph_fem_glmer6, is_color=output_col) 
```

### Summary

```{r}
tidy_regression(nomass_fem2, is_color=output_col) 
tidy_regression(mass_fem2, is_color=output_col) 
tidy_regression(multi_glmer_fem, is_color=output_col) 
tidy_regression(multi_glmer_fem2, is_color=output_col) 
tidy_regression(egg_model2, is_color=output_col)
tidy_regression(egg_glmer, is_color=output_col)
tidy_regression(egg_glmer2, is_color=output_col)
```

All the morph models had an issue between using ID vs. trial type as the RF. When I used trial_type I got reasonable results. When I used ID I always got the null model. Why is this?

```{r}
tidy_regression(morph_fem2, is_color=output_col)
tidy_regression(morph_fem_glmer, is_color=output_col) 
tidy_regression(morph_fem3, is_color=output_col)
tidy_regression(morph_fem_glmer3 , is_color=output_col) 
tidy_regression(morph_fem_glmer4 , is_color=output_col) 
tidy_regression(morph_fem_glmer5 , is_color=output_col) 
tidy_regression(morph_fem_glmer6, is_color=output_col)
```

### Males

[Cleaning Data](#cleanup_male)

[Testing Covariates](#male_exp)

[Without Mass Modeling](#without.mass.m)

[With Mass Modeling](#with.mass.m)

[Morphology Modeling](#morph.m)

<a id="cleanup_male"></a>

**Cleaning the Data**

```{r}
data <- read_flight_data("data/all_flight_data-Winter2020.csv")
data_all <- data[[1]]
data_tested <- data[[2]]
d = create_delta_data(data_tested)
data_male <- d[d$sex=="M",]
data_male <- center_data(data_male, is_not_unique_data = FALSE) 
data_male
```

<a id="male_exp"></a>

**Experimental, Biological, and Morphological Effects**

```{r, echo=FALSE}
####### Can't test chamber, test date, or test time any more
####### Can't specific mass anymore - only average mass. (Could do mass diff?)

####### No effect of average mass (not ideal - using this cbind() method seems to be muting mass)
tidy_regression(glm(cbind(num_flew, num_notflew)~avg_mass, data=data_male, family=binomial), is_color=output_col) 

####### Effect of beak length
tidy_regression(glm(cbind(num_flew, num_notflew)~beak_c, data=data_male, family=binomial), is_color=output_col) 

####### No effect of thorax length
tidy_regression(glm(cbind(num_flew, num_notflew)~thorax_c, data=data_male, family=binomial), is_color=output_col) 

####### Effect of body length
tidy_regression(glm(cbind(num_flew, num_notflew)~body_c, data=data_male, family=binomial), is_color=output_col) 

####### Effect of wing length
tidy_regression(glm(cbind(num_flew, num_notflew)~wing_c, data=data_male, family=binomial), is_color=output_col)
```

<a id="without.mass.m"></a>

#### Without Mass

```{r message=FALSE}
R1 = data_male$num_flew
R2 = data_male$num_notflew
A = data_male$host_c
B = data_male$wing_c
C = data_male$avg_mass 
X = data_male$population # can't do trial type anymore

data<-data.frame(R1, R2, A, B, C, X)

model_script = paste0(source_path,"generic models-binomial glmer 2R ~ 1 RF + 2-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m2, m3, test="Chisq") # Adding A does not improve fit
```

```{r}
nomass_male <- glmer(cbind(num_flew, num_notflew) ~ wing_c + (1|population), data=data_male, family=binomial)
tidy_regression(nomass_male, is_color=output_col)
```

* positive effect of wing

<a id="with.mass.m"></a>

#### With Mass 

```{r message=FALSE}
model_script = paste0(source_path,"generic models-binomial glmer 2R ~ 1 RF + 3-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m6, m10, test="Chisq") # Adding B*C does not improve fit
anova(m6, m7, test="Chisq") # Adding A does not improve fot
```

```{r}
mass_male <- glmer(cbind(num_flew, num_notflew) ~ wing_c + avg_mass + (1|population), data=data_male, family=binomial)
tidy_regression(mass_male, is_color=output_col)
```

* strong positive effect of wing
* strong negative effect of average mass 

```{r}
# check for collinearity between mass and wing length
data<-data.frame(data_male$host_c, data_male$wing_c, data_male$avg_mass)
colnames(data) <- c("Host Plant", "Wing Length", "Average Mass")
pairs(data)
```

<a id="morph.m"></a>

#### Morphology 

```{r message=FALSE}
d <- data_male %>%
  filter(!is.na(body))
d$thorax_c <- d$thorax - mean(d$thorax)
d$wing_c <- d$wing - mean(d$wing)
d$body_c <- d$body - mean(d$body)

R1 = d$num_flew
R2 = d$num_notflew
A = d$thorax_c
B = d$body_c
C = d$wing_c 
X = d$population

data<-data.frame(R1, R2, A, B, C, X)

model_script = paste0(source_path,"generic models-binomial glmer 2R ~ 1 RF + 3-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m13, m16, test="Chisq") # Adding B*C does not improve fit
anova(m10, m13, test="Chisq") # Adding A does not improve fit
anova(m6, m10, test="Chisq") # Adding B*C improves fit
```

```{r}
morph_male <- glmer(cbind(num_flew, num_notflew) ~ body_c * wing_c + (1 | population), data=d, family=binomial)
tidy_regression(morph_male, is_color=output_col)
```

* marginal negative effect of body
* positive effect of wing
* negative wing*body where the longer the wing and body the less times a male bug flies.

```{r message=FALSE}
R1 = d$num_flew
R2 = d$num_notflew
A = d$thorax_c
B = d$wing2body_c
X = d$population

data<-data.frame(R1, R2, A, B, X)

model_script = paste0(source_path,"generic models-binomial glmer 2R ~ 1 RF + 2-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m2, m3, test="Chisq") # Adding B does not improve fit
```

```{r}
morph_male2 <- glmer(cbind(num_flew, num_notflew) ~ wing2body_c + (1 | population), data=d, family=binomial)
tidy_regression(morph_male2, is_color=output_col)
```

* strong positive effect of wing2body where the more wing2body the more times a male bug flies.

```{r}
# check for collinearity between mass and morhpology
data<-data.frame(data_male$thorax_c, data_male$wing_c, data_male$body_c, data_male$avg_mass)
colnames(data) <- c("Thorax Length", "Wing Length", "Body Length", "Average Mass")
pairs(data)
```


```{r}
data_male <- data_tested[data_tested$sex=="M",]
data_male <- center_data(data_male) 

data<-data.frame(data_male$thorax_c, data_male$wing_c, data_male$body_c, data_male$mass_c, data_male$days_from_start_c)
colnames(data) <- c("Thorax Length", "Wing Length", "Body Length", "Average Mass", "Days from Start")
pairs(data)

m <- lm(mass ~ days_from_start, data=data_male)
summary(m)$coefficients
plot(mass ~ days_from_start, data=data_male)
abline(m, col="red")
#plot(body ~ days_from_start, data=data_male)
text(mass ~ days_from_start, labels=data_male$ID, data=data_male, cex=0.5, font=2, pos=4) # maybe the smaller males died faster? that's why it looks like.
```

### Summary

```{r}
tidy_regression(nomass_male, is_color=output_col)
tidy_regression(mass_male, is_color=output_col)
tidy_regression(morph_male, is_color=output_col)
tidy_regression(morph_male2, is_color=output_col)
```


### Males - Old Method

[Cleaning Data](#cleanup_d)

[Testing Covariates](#expsetup_effects)

[Without Mass Modeling + Days From Start](#no.mass)

[With Mass Modeling + Days From Start](#mass)

[Morphology Modeling](#male_morph)

* [glmer() A=thorax, B=body, C=wing, D=days_from_start, X=ID, Y=trial_type](#morph_glmer)

* [glmer() A=wing2body, B=thorax, C=days_from_start, X=ID, Y=trial_type](#morph_glmer3)

<a id="cleanup_d"></a>

**Cleaning Data**

```{r}
data <- read_flight_data("data/all_flight_data-Winter2020.csv")
data_all <- data[[1]]
data_tested <- data[[2]]

data_male <- data_tested[data_tested$sex=="M",]
data_male <- center_data(data_male)
```

<a id="expsetup_effects"></a>

**Experimental Set-Up Effects**

```{r echo=FALSE}
####### No effect of chamber
tidy_regression(glm(flew_b~chamber, data=data_male, family=binomial), is_color=output_col)

####### Effect of test date 
tidy_regression(glm(flew_b~days_from_start_c, data=data_male, family=binomial), is_color=output_col)

####### Maringal effect of test time (p = 0.081)
tidy_regression(glm(flew_b~min_from_IncStart, data=data_male, family=binomial), is_color=output_col)

# seems like males are very time sensitive
```

**Biological Effects**

```{r echo=FALSE}
####### No effect of mass
tidy_regression(glm(flew_b~mass_c, data=data_male, family=binomial), is_color=output_col) 
```

**Morphology Effects**

```{r echo=FALSE}
####### Effect of beak length
tidy_regression(glm(flew_b~beak_c, data=data_male, family=binomial), is_color=output_col)

####### No effect of thorax length
tidy_regression(glm(flew_b~thorax_c, data=data_male, family=binomial), is_color=output_col) 

####### Effect of body length
tidy_regression(glm(flew_b~body_c, data=data_male, family=binomial), is_color=output_col) 

####### Effect of wing length
tidy_regression(glm(flew_b~wing_c, data=data_male, family=binomial), is_color=output_col) 
```

<a id="no.mass"></a>

#### Without Mass + Days From Start

```{r}
# Remove any missing masses
data_male <- data_male %>%
  filter(!is.na(mass), !is.na(wing))
data_male <- center_data(data_male)

R = data_male$flew_b
A = data_male$host_c
B = data_male$days_from_start_c
C = data_male$wing_c
D = data_male$mass_c 

data<-data.frame(R, A, B, C, D)

model_script = paste0(source_path,"generic models-binomial glm 3-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m7, m12, test="Chisq") # Adding A*C does not improve fit
anova(m6, m7, test="Chisq") # Adding A improves fit
```

```{r echo=FALSE}
nomass_male<-glm(flew_b~host_c + days_from_start_c + wing_c, family=binomial, data=data_male)
tidy_regression(nomass_male, is_color=output_col)
```

* Negative effect if from GRT
* Negative effect of days from start where the farther out the day from the start of the experiment (e.g. maybe a proxi for age), the less likely the male bug is to fly
* Positive effect wing length where the longer the wing the more likely the male bug is to fly

```{r}
## Consider covariates
model_male_final <-glmer(flew_b~host_c + days_from_start_c + wing_c +  (1|ID), family=binomial, data=data_male) 
tidy_regression(model_male_final, is_color=output_col)
## Changed the effect slightly and improved the model
```

* Marginal negative effect if from GRT
* Negative effect of days from start where the farther out the day from the start of the experiment (e.g. maybe a proxi for age), the less likely the male bug is to fly
* Positive effect wing length where the longer the wing the more likely the male bug is to fly

<a id="mass"></a>

#### With Mass + Days From Start

```{r}
model_script = paste0(source_path,"generic models-binomial glm 4-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m15, m35, test="Chisq") # Adding A*C does not improve fit
```

```{r echo=FALSE}
mass_male <-glm(flew_b~host_c + mass_c + days_from_start_c + wing_c, family=binomial, data=data_male)
tidy_regression(mass_male, is_color=output_col)
```

* Strong negative effect if from GRT
* Strong negative effect of mass, that if weigh more then less likely to disperse
* negative effect of days from start where the farther out the day from the start of the experiment (e.g. maybe a proxi for age), the less likely the male bug is to fly
* positive effect of wing length where the longer th wing, the more likely the male bug is to fly

```{r}
## Consider covariates
mass_male_glmer <-glmer(flew_b~host_c + mass_c + days_from_start_c + wing_c + (1|population) + (1|trial_type), family=binomial, data=data_male) # no error, but singular fit error 

tidy_regression(mass_male_glmer, is_color=output_col)
## Changed the effects slightly, a better fitting model, and made host not significant
```

<a id="male_morph"></a>

#### Morphology

```{r}
data_male <- data_tested[data_tested$sex=="M",]
data_male <- data_male %>%
  filter(!is.na(mass)) %>%
  filter(!is.na(body))
data_male <- center_data(data_male)

R = data_male$flew_b
A = data_male$thorax_c
B = data_male$body_c
C = data_male$wing_c 

data<-data.frame(R, A, B, C)

model_script = paste0(source_path,"generic models-binomial glm 3-FF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m10, m13, test="Chisq") # Adding A does not improve fit
```

```{r}
morph_male <-glm(flew_b~body_c* wing_c, family=binomial, data=data_male) 
tidy_regression(morph_male, is_color=output_col) 
```

* no effect of body length
* marginal positive effect of wing length, where if longer wing then more likely to fly

* negative effect of body*wing where if longer body and wing then less likely to fly

```{r}
morph_male_glmer <- glmer(flew_b ~ body_c* wing_c + (1|ID), family=binomial, data=data_male) 
tidy_regression(morph_male_glmer, is_color=output_col) # model improves fit and it converges
```
* no effect of body length
* no effect of wing length

* negative effect of body*wing where if longer body and wing then less likely to fly

<a id="morph_glmer"></a>

#### glmer() A=thorax, B=body, C=wing, D=days_from_start, X=ID, Y=trial_type

```{r message=FALSE}
R = data_male$flew_b
A = data_male$thorax_c 
B = data_male$body_c
C = data_male$wing_c 
D = data_male$days_from_start_c 
X = data_male$ID # 3 models failed to converge when ID as RF
#Y = data_male$trial_type # more models fail if both ID and trial as RF

data<-data.frame(R, A, B, C, D, X)

model_script = paste0(source_path,"generic models-binomial glmer 1-RF + 4-FF-noD-interactions.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m24, m27, test="Chisq") # Adding A does not improve fit
anova(m22, m26, test="Chisq") # Adding B does not improve fit
```

```{r}
morph_male_glmer2 <- glmer(flew_b ~ thorax_c * wing_c + days_from_start_c + (1 | ID) , family=binomial, data=data_male) 
tidy_regression(morph_male_glmer2, is_color=output_col) 
```

<a id="morph_glmer3"></a>

#### glmer() A=wing2body, B=thorax, C=days_from_start, X=ID, Y=trial_type

```{r message=FALSE}
# left off here
R = data_male$flew_b
A = data_male$wing2body_c 
B = data_male$thorax_c
C = data_male$days_from_start_c 
X = data_male$ID 
Y = data_male$trial_type 

data<-data.frame(R, A, B, C, X, Y)

model_script = paste0(source_path,"generic models-binomial glmer 2 RF + 3-FF-noCinteractions.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m5, m7, test="Chisq") # Adding B does not improve fit
anova(m5, m24, test="Chisq") # Adding Y does not improve fit
```

```{r}
morph_male_glmer3 <- glmer(flew_b ~ wing2body_c + days_from_start_c + (1 | ID), family=binomial, data=data_male) 
tidy_regression(morph_male_glmer3, is_color=output_col) 
```

* strong positive effect of wing2body, where if you have more wing to body, then more likely to fly
* negative effect of days from start, where farther out in days from the start day, the less likely the bug will fly.

### Summary

```{r}
tidy_regression(nomass_male, is_color=output_col)
tidy_regression(mass_male, is_color=output_col)
tidy_regression(mass_male_glmer, is_color=output_col)
tidy_regression(morph_male, is_color=output_col) 
tidy_regression(morph_male_glmer, is_color=output_col) 
tidy_regression(morph_male_glmer2, is_color=output_col)
tidy_regression(morph_male_glmer3, is_color=output_col) 
```












