---
title: "Modeling Mass"
author: "Anastasia Bernat"
date: "4/17/2020"
output: html_document
---

```{r setup, include=FALSE}
rm(list=ls())
setwd("~/Desktop/git_repositories/SBB-dispersal/avbernat/Dispersal/Winter_2020/stats/")

# For Modeling
library(lme4)

# For Data Summaries and Manipulation
library(lubridate)
library(dplyr)
library(mosaic)

# For Plotting
library(ggplotify)
library(gridExtra) # "grid" graphics

knitr::opts_chunk$set(echo = TRUE)
ggplot2::theme_set(ggplot2::theme_bw())
ggplot2::theme_update(panel.grid.major = ggplot2::element_line(colour = "grey75"))
```

## Winter 2020 Flight Trials: Mass Modeling {.tabset}

Flight Trials Winter 2020 Dataset was conducted from 2/17/2020 - 3/10/2020. Soapberry bugs were flight tested twice for multiple hours in the flight mill and observed from 8 AM to (5-8 PM) each day. Used multivariate (lm) and mixed effect modeling (lmer) to analyze the mass results.

### All Data

[Cleaning Data](#clean)

[lmer() host, sex, sym_dist | ID and trial_type](#lmer1)

[Morphology](#morph)

<a id="clean"></a>

#### Reading the data

```{r}
source_path = "~/Desktop/git_repositories/SBB-dispersal/avbernat/Rsrc/"

script_names = c("center_flight_data.R", # Re-centers data 
                 "clean_flight_data.R", # Loads and cleans data
                 "get_warnings.R", 
                 "compare_models.R",
                 "regression_output.R", # Cleans regression outputs; prints in color or B&W
                 "AICprobabilities.R",
                 "get_Akaike_weights.R",
                 "plotting-lm.R",
                 "plotting-lm2.R",
                 "plotting-polyreg.R")

for (script in script_names) { 
  path = paste0(source_path, script)
  source(path) 
}

output_col = FALSE # Change to TRUE if working in Base R or RStudio; FALSE if generating an HTML

data <- read_flight_data("data/all_flight_data-Winter2020.csv")
#data_all <- data[[1]]
#data_tested <- data[[2]]
```

<a id="test"></a>

#### Testing Models and Covariates

```{r}
####Check for convergence without splitting by trial type
test_model<-lmer(mass~host_c*sex_c*sym_dist+ (1|trial_type) + (1|ID), data=data_tested, REML=FALSE) 
tidy_regression(test_model, is_color=output_col)
getME(test_model, "lower")
```

**Experimental Set-Up Effects**

```{r echo=FALSE}
###### No effect of chamber
tidy_regression(lmer(mass_c~chamber + (1|ID) + (1|trial_type), data=data_tested, REML=FALSE), is_color=output_col) 

####### Effect of test date - meaning that mass fluctuates by day
tidy_regression(lmer(mass_c~days_from_start_c + (1|ID), data=data_tested, REML=FALSE), is_color=output_col)

####### No effect of test time
tidy_regression(lmer(mass_c~min_from_IncStart_c + (1|ID) + (1|trial_type), data=data_tested, REML=FALSE), is_color=output_col) 
```

**Biological Effects**

```{r, echo=FALSE}
####### Effect of number of eggs laid
tidy_regression(lmer(mass_c~total_eggs + (1|ID) + (1|trial_type), data=data_tested, REML=FALSE), is_color=output_col) 

####### Effect of whether eggs were laid or not
tidy_regression(lmer(mass_c~eggs_b + (1|ID), data=data_tested, REML=FALSE), is_color=output_col) 
```

**Morphology Effects**

```{r, echo=FALSE}
####### Effect of beak length
tidy_regression(lmer(mass_c~beak_c + (1|ID) + (1|trial_type), data=data_tested, REML=FALSE), is_color=output_col) # model didn't converge with (1|trial_type)

####### Effect of thorax length
tidy_regression(lmer(mass_c~thorax_c + (1|ID) + (1|trial_type), data=data_tested, REML=FALSE), is_color=output_col)

####### Effect of body length
tidy_regression(lmer(mass_c~body_c + (1|ID) + (1|trial_type), data=data_tested, REML=FALSE), is_color=output_col)

####### Effect of wing length
tidy_regression(lmer(mass_c~wing_c + (1|ID) + (1|trial_type), data=data_tested, REML=FALSE), is_color=output_col)
```

<a id="lmer1"></a>

#### All Data | lmer() A=host, B=sex, C=sym_dist, X=ID 

```{r}
data<-data.frame(R=data_tested$mass_c, 
                 A=data_tested$host_c, 
                 B=data_tested$sex_c, 
                 C=data_tested$sym_dist, 
                 X=data_tested$ID)

model_script = paste0(source_path,"generic models-gaussian glmer 1-RF + 3-FF REMLF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m13, m15, test="Chisq") # Adding A*B does not improve fit
```

```{r}
model_alldata <- lmer(mass_c ~ sex_c * sym_dist + host_c + (1|ID), data=data_tested, REML=FALSE)
tidy_regression(model_alldata, is_color=output_col) 

model_alldata <- lmer(mass_c ~ sex_c * sym_dist + host_c + (1|ID) + (1|trial_type), data=data_tested, REML=FALSE)
tidy_regression(model_alldata, is_color=output_col) # adding trial type improves fit
```

* positive effect of sex, where if Female then weigh more.
* negative weak effect of host where if from GRT then weigh less.
* negative weak effect of sym_dist

* negative weak effect of sex*sym_dist where if F and farther from Homestead, then weigh less

<a id="morph"></a>

#### Morphology Modeling

```{r}
data<-data.frame(R=data_tested$mass_c, 
                 A=data_tested$beak_c, 
                 B=data_tested$thorax_c, 
                 C=data_tested$body_c, 
                 D=data_tested$wing_c,
                 X=data_tested$ID)
data <- data %>%
  filter(!is.na(C))

model_script = paste0(source_path,"generic models-gaussian glmer 1-RF + 4-FF REMLF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m91, m107, test="Chisq") # Adding C*D does not improve fit
```

```{r}
allmorph_model <- lmer(mass_c ~ beak_c * thorax_c + beak_c * body_c + beak_c * wing_c + thorax_c * body_c + (1|ID), data=data_tested, REML=FALSE)
tidy_regression(allmorph_model, is_color=output_col) 

allmorph_model <- lmer(mass_c ~ beak_c * thorax_c + beak_c * body_c + beak_c * wing_c + thorax_c * body_c + (1|ID) + (1|trial_type), data=data_tested, REML=FALSE)
tidy_regression(allmorph_model, is_color=output_col) 
```

```{r, message=FALSE}
data<-data.frame(R=data_tested$mass_c, 
                 A=data_tested$beak_c, 
                 B=data_tested$thorax_c, 
                 C=data_tested$wing2body_c,
                 X=data_tested$ID)
data <- data %>%
  filter(!is.na(A))

model_script = paste0(source_path,"generic models-gaussian glmer 1-RF + 3-FF REMLF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m12, m14, test="Chisq") # Adding A*B does not improve fit
```

```{r}
allmorph_model <- lmer(mass_c ~  beak_c * wing2body_c + thorax_c + (1 | ID), data=data_tested, REML = FALSE)
tidy_regression(allmorph_model, is_color=output_col) 

allmorph_model <- lmer(mass_c ~  beak_c * wing2body_c + thorax_c + (1 | ID) + (1|trial_type), data=data_tested, REML = FALSE)
tidy_regression(allmorph_model, is_color=output_col) 
```

* positive weak effect of beak, where the longer beak the heavier the bug
* negative effect of wing2body, where the longer the wing to body the lighter the bug
* positive effect of thorax, where the longer the thorax the heavier the bug

* negative beak*wing2body interaction, where the longer the beak and longer the wing2body, the lighter the bug

### All Data Plotting 

#### Days from start by mass

```{r, warning=FALSE}
lm_plot(data_tested, "days_from_start_c","mass_c", "Days From Start", "Mass (g)", "Mass by Days From Start") 
tidy_regression(lmer(mass_c~days_from_start_c + (1|ID), data=data_tested, REML=FALSE), is_color=output_col)
```
#### Flight Duration by Mass

```{r fig.width=6, fig.height=2.5}

#pdf("massby-flightduration.pdf") # alternative way to save graph
flight_p1 <- as.grob(expression(
  plot(data_tested$mass,data_tested$minute_duration, 
          main = "All Trials",
          xlab = "Mass (g)", 
          ylab = "Flight duration (min)") %>% 
  # Seems like bugs with a mass between 0.025-0.075 g can hit longer flight durations
  abline(v=0.025, col="red") %>%
  abline(v=0.075, col="red")              
#dev.off() # alt way to save graph
))
yes_fly <- filter(data_tested, flew_b == 1)
flight_p2 <- as.grob(expression(
  plot(yes_fly$mass,yes_fly$minute_duration, 
                  main = "All Trials, only bugs that flew (yes_flew)",
                  xlab="Mass (g)",
                  ylab = "Flight duration(min)") %>%
  abline(v=0.025, col="red")%>%
  abline(v=0.075, col="red")
))

C_fly <- filter(yes_fly, flight_type =="C")
flight_p3 <- as.grob(expression(
  plot(C_fly$mass,C_fly$minute_duration,
                  main = "All Trials, only bugs that flew continuously",
                  xlab="Mass (g)",
                  ylab = "Flight duration(min)") %>%
  abline(v=0.025, col="red") %>%
  abline(v=0.075, col="red")
))

grob1 <- as.grob(expression(flight_p1))

grid.arrange(flight_p1, flight_p2, flight_p3, ncol=3)
```

#### Plotting mass vs. sym_dist * host * sex

```{r, warning=FALSE}
# filter out missing masses
data_tested <- data_tested %>%
  filter(!is.na(mass_c))

p <- gf_point(mass ~ sym_dist, data = data_tested,
                  color = ~host_plant, 
                  alpha = ~sex, 
                  ylab="Mass (g)", 
                  xlab="Distance From Sympatric Zone")
p
```

```{r}
#, fig.width=6, fig.height=2.5}
summary<-aggregate(mass~sex*host_plant*sym_dist, data=data_all, FUN=mean)

# Uncomment below to save graph

#pdf("lm-mass-model.pdf", width=9, height=6)

plot(summary$mass~summary$sym_dist, 
     col=c(1,2)[as.factor(summary$sex)], # Female = Black, Male = Red
     pch=c(19,22)[as.factor(summary$host_plant)],
     main="Observed Data",
     xlab = "Distance from Sympatric Zone (°)",
     ylab= "Mass (g)", # K. elegans = Squares C.corindum = circles
     ) 
legend("topright",
       legend = c("C.corindum and F","K.elegans and M"),
       col=c(1,2),
       pch = c(19,22))

#dev.off()
```

#### Plotting best-fit model (m38) | lmer(mass ~ sex)

```{r, message=FALSE, warning=FALSE}
#, fig.width=2.8, fig.height=2.3}
# Uncomment below to save graphs 
# For both graphs: need to add xaxt="n" inside plot function to make the customized x-axis

#pdf("lm-massbysex-model.pdf", width=6, height=6)
lm_plot2(data_tested, "sex_c","mass_c", "Sex", "Mass (g)", "Mass by Sex") 
axis(1, at = seq(-1,1,.5),labels=c("M", " ", " ", " ", "F"))
#dev.off()

#pdf("lm-massbyhost-model.pdf", width=6, height=6)
lm_plot2(data_tested, "host_c","mass_c", "Host Plant", "Mass (g)", "Mass by Host Plant") 
axis(1, at = seq(-1,1,.5),labels=c("BV", " ", " ", " ", "GRT"))
#dev.off()

lm_plot(data_tested, "sym_dist","mass_c", "Distance From Sympatric Zone", "Mass (g)", "Mass by Sym_Dist") 
```

#### Morphology Plots 

The following plots are nonlinear - it's worth exploring this nonlinear relationship with a polynomial regression.

```{r}
p <- gf_point(mass ~ wing, data = data_tested,
                  color = ~thorax, 
                  alpha = ~body, 
                  size = ~beak,
                  ylab="Mass (g)", 
                  xlab="Wing (mm)")
p
#ggsave("MassbyMorphology-alldata_gf", plot = p, device="jpeg", width = 9, height = 7)
```

**First-Order Regression**

```{r, warning=FALSE}
p1 <- gf_point(mass ~ wing2body, data = data_tested,
                  ylab="Mass (g)", 
                  xlab="Wing-to-Body Ratio") %>%
  gf_lm(alpha=0.2)
p2 <- gf_point(mass ~ wing2body, data = data_tested,
                  color = ~thorax, 
                  alpha = ~beak, 
                  ylab="Mass (g)", 
                  xlab="Wing-to-Body Ratio") %>%
  gf_lm(alpha=0.2)
p3 <- gf_point(mass ~ thorax, data = data_tested,
                  color = ~wing2body, 
                  alpha = ~beak, 
                  ylab="Mass (g)", 
                  xlab="Thorax (mm)") %>%
  gf_lm(alpha=0.2)
p4 <- gf_point(mass ~ beak, data = data_tested,
                  color = ~wing2body, 
                  alpha = ~thorax, 
                  ylab="Mass (g)", 
                  xlab="Beak (mm)") %>%
  gf_lm(alpha=0.2)
g <- grid.arrange(p1,p2,p3,p4, ncol=2, top="All Data")
#ggsave("MassbyMorphology-alldata2", plot = g, device="jpeg", width = 9, height = 7)
#ggsave("MassbyMorphology-alldata2-lm", plot = g, device="jpeg", width = 12, height = 7)
```

```{r, warning=FALSE, message=FALSE}
p5 <- as.grob(expression(
  plot(data_tested$body, data_tested$mass, ylab="Mass (g)", xlab="Body (mm)")))
p6 <- as.grob(expression(
  plot(data_tested$beak, data_tested$mass, ylab="Mass (g)", xlab="Beak (mm)")))
p7 <- as.grob(expression(
  plot(data_tested$thorax, data_tested$mass, ylab="Mass (g)", xlab="Throax (mm)")))
p8 <- as.grob(expression(
  plot(data_tested$wing, data_tested$mass, ylab="Mass (g)", xlab="Wing (mm)")))

g2 <- grid.arrange(p5,p6,p7,p8, ncol=2, top="All Data")

# Uncomment below to save graph:
#ggsave("MassbyMorphology-alldata", plot = g2, device="jpeg", width = 9, height = 7)

p9 <- plot(data_tested$wing2body, data_tested$mass, ylab="Mass (g)", xlab="Wing-to-Body Ratio", main="All Data")
```
```{r}
d <- data_tested %>%
  filter(!is.na(mass)) %>%
  filter(!is.na(body))
```

```{r warning=FALSE, results='hide'}
#, fig.width=5, fig.height=4}
lm_plot(d, "wing2body_c","mass_c", "Wing-to-Body Ratio", "Mass (g)", "Mass by Wing:Body") 

par(mfrow=c(2,2))
lm_plot(d, "body_c","mass_c", "Body Length (mm)", "Mass (g)", "Mass by Body Length") 
lm_plot(d, "wing_c","mass_c", "Wing Length (mm)", "Mass (g)", "Mass by Wing Length") 
lm_plot(d, "thorax_c","mass_c", "Thorax Length (mm)", "Mass (g)", "Mass by Thorax Length") 
lm_plot(d, "beak_c","mass_c", "Beak Length (mm)", "Mass (g)", "Mass by Beak Length") 
```

**Second-Order Polynomial Regression**

```{r}
#, fig.width=5, fig.height=4}
lm_polyplot(d, "wing2body_c","mass_c", "Wing-to-Body Ratio", "Mass (g)", "Mass by Wing:Body") 

#par(mfrow=c(2,2))
lm_polyplot(d, "body", "mass", "Body Length (mm)", "Mass (g)", "Mass by Body Length")
lm_polyplot(d, "thorax", "mass", "Thorax Length (mm)", "Mass (g)", "Mass by Thorax Length") # check some bugs 
lm_polyplot(d, "wing", "mass", "Wing Length (mm)", "Mass (g)", "Mass by Wing Length")
lm_polyplot(d, "beak", "mass", "Beak Length (mm)", "Mass (g)", "Mass by Beak Length")
```

More gf_point explanations: https://cran.r-project.org/web/packages/ggformula/vignettes/ggformula-blog.html

Try an exponential line next: https://www.theanalysisfactor.com/r-tutorial-5/


### Female vs. Male Histograms

```{r, fig.width=8, fig.height=2.5}
# Had twice as many males as females.
data_fem <- filter(data_tested, sex == "F") 
data_male <- filter(data_tested, sex == "M") 

data_fem1 <- filter(data_tested, sex == "F", trial_type=="T1") 
data_fem2 <- filter(data_tested, sex == "F", trial_type=="T2")  
data_male1 <- filter(data_tested, sex == "M", trial_type=="T1")  
data_male2 <- filter(data_tested, sex == "M", trial_type=="T2") 

h3 <- as.grob(expression(hist(data_fem1$mass, main="Females Trial 1", xlab= "Mass (g)")))
h4 <- as.grob(expression(hist(data_fem2$mass, main="Females Trial 2", xlab= "Mass (g)")))
h5 <- as.grob(expression(hist(data_male1$mass, main="Males Trial 1", xlab = "Mass (g)")))
h6 <- as.grob(expression(hist(data_male2$mass, main="Males Trial 2", xlab = "Mass (g)")))

h7 <- as.grob(expression(hist(data_fem$mass, main="Females", xlab= "Mass (g)")))
h8 <- as.grob(expression(hist(data_male$mass, main="Males", xlab= "Mass (g)")))

grid.arrange(h7,h8,ncol=2)
grid.arrange(h3,h4,h5,h6,ncol=4)
```


### Females

[Cleaning Data](#cleandata)

[Testing Covariates](#testf)

[lmer() A=host, B=sex, C=sym_dist, X=ID, Y=trial_type, Z=chamber](#lmerf)

[Morphology](#morphf)

[Egg and Morphology](#eggmorph)

<a id="cleandata"></a>

#### Reading the data

```{r}
data <- read_flight_data("data/all_flight_data-Winter2020.csv")
data_all <- data[[1]]
data_tested <- data[[2]]

# filter out missing masses
data_tested <- data_tested %>%
  filter(!is.na(mass))

data_fem <- data_tested[data_tested$sex=="F",]
data_fem <- center_data(data_fem)
```

<a id="testf"></a>

#### Testing Covariates

**Experimental Set-Up Effects**

```{r echo=FALSE}
###### No effect of chamber
tidy_regression(lmer(mass_c~chamber + (1|ID) + (1|trial_type), data=data_fem, REML=FALSE), is_color=output_col) 

####### Marginal effect of test date
tidy_regression(lmer(mass_c~days_from_start_c + (1|ID), data=data_fem, REML=FALSE), is_color=output_col) # t val = 2.29 - makes sense because female mass would fluctuate the most

####### No effect of test time
tidy_regression(lmer(mass_c~min_from_IncStart_c + (1|ID) + (1|trial_type), data=data_fem, REML=FALSE), is_color=output_col) 
```

**Biological Effects**

```{r, echo=FALSE}
####### Effect of number of eggs laid
tidy_regression(lmer(mass_c~total_eggs + (1|ID) + (1|trial_type), data=data_fem, REML=FALSE), is_color=output_col) 

####### Effect of whether eggs were laid or not
tidy_regression(lmer(mass_c~eggs_b + (1|ID), data=data_fem, REML=FALSE), is_color=output_col)
```

**Morphology Effects**

```{r, echo=FALSE}
####### Effect of beak length
tidy_regression(lmer(mass_c~beak_c + (1|ID) + (1|trial_type), data=data_fem, REML=FALSE), is_color=output_col)

####### Effect of thorax length
tidy_regression(lmer(mass_c~thorax_c + (1|ID) + (1|trial_type), data=data_fem, REML=FALSE), is_color=output_col)

####### Effect of body length
tidy_regression(lmer(mass_c~body_c + (1|ID), data=data_fem, REML=FALSE), is_color=output_col)

####### Effect of wing length
tidy_regression(lmer(mass_c~wing_c + (1|ID), data=data_fem, REML=FALSE), is_color=output_col) 
```

<a id="lmerf"></a>

#### Female Data | lmer() A=host, B=sex, C=sym_dist, X=ID 

```{r, message=FALSE}
data<-data.frame(R=data_fem$mass_c, 
                 A=data_fem$host_c, 
                 B=data_fem$sym_dist,
                 X=data_fem$ID)

model_script = paste0(source_path,"generic models-gaussian glmer 1-RF + 2-FF REMLF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m2, m3, test="Chisq") # Adding A does not improve fit
```

```{r}
femm_model <- lmer(mass_c ~ sym_dist + (1|ID), data=data_fem, REML=FALSE)
tidy_regression(femm_model, is_color=output_col)

femm_model <- lmer(mass_c ~ sym_dist + (1|ID) + (1|trial_type), data=data_fem, REML=FALSE)
tidy_regression(femm_model, is_color=output_col)
```

* marginal negative weak effect of sym_dist where the farther from homestead the less GRT female bugs weigh

Let's add eggs:

```{r message=FALSE}
data<-data.frame(R=data_fem$mass_c, 
                 A=data_fem$host_c, 
                 B=data_fem$sym_dist,
                 C=data_fem$eggs_b,
                 X=data_fem$ID)

model_script = paste0(source_path,"generic models-gaussian glmer 1-RF + 3-FF REMLF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m6, m10, test="Chisq") # Adding B*C does not improve fit
```

```{r}
femm_model <- lmer(mass_c ~ sym_dist + eggs_b + (1|ID), data=data_fem, REML=FALSE)
tidy_regression(femm_model, is_color=output_col) # When addd trial type as RF, it has zero variance so not needed
```

* no efffect of sym dist
* Positive effect of eggs laid on test day where if a female bug laid eggs on her test day, she weighed more.

```{r, message=FALSE}
data<-data.frame(R=data_fem$mass_c, 
                 A=data_fem$host_c, 
                 B=data_fem$sym_dist,
                 C=data_fem$total_eggs_c,
                 X=data_fem$ID)
data <- data %>%
  filter(!is.na(D))

model_script = paste0(source_path,"generic models-gaussian glmer 1-RF + 3-FF REMLF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m2, m4, test="Chisq") # Adding A does not improve fit
```


```{r}
femm_model <- lmer(mass_c ~ sym_dist + (1|ID), data=data_fem, REML=FALSE)
tidy_regression(femm_model, is_color=output_col)
femm_model <- lmer(mass_c ~ sym_dist + (1|ID) + (1 | trial_type), data=data_fem, REML=FALSE)
tidy_regression(femm_model, is_color=output_col)
```

* negative marginal effect of sym_dist where the farther the bug is from Homestead, the less it weighed

<a id="morphf"></a>

#### Morphology Modeling

```{r, message=FALSE}
data<-data.frame(R=data_fem$mass_c, 
                 A=data_fem$beak_c, 
                 B=data_fem$thorax_c,
                 C=data_fem$wing_c,
                 D=data_fem$body_c,
                 X=data_fem$ID) 
data <- data %>%
  filter(!is.na(A))

model_script = paste0(source_path,"generic models-gaussian glmer 1-RF + 4-FF REMLF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
femm_morph_model <- lmer(mass_c ~ beak_c * thorax_c + thorax_c * body_c + wing_c  + (1|ID), data=data_fem, REML=FALSE)
tidy_regression(femm_morph_model, is_color=output_col)
femm_morph_model <- lmer(mass_c ~ beak_c * thorax_c + thorax_c * body_c + wing_c  + (1|ID) + (1|trial_type), data=data_fem, REML=FALSE)
tidy_regression(femm_morph_model, is_color=output_col)
```

* no effect of beak
* no effect of thorax
* positive effect of body
* negative effect of beak
* negative effect of wing 

* beak*thorax negative effect where the longer the beak and throax the less the bug weighs
* positive effect of thorax*body where the larger the thorax and body, more less the bug weighs


<a id="eggmorph"></a>

#### Eggs and Morphology 

```{r, message=FALSE}
data<-data.frame(R=data_fem$mass_c, 
                 A=data_fem$eggs_b, 
                 B=data_fem$thorax_c,
                 C=data_fem$wing_c,
                 D=data_fem$body_c,
                 X=data_fem$ID) 
data <- data %>%
  filter(!is.na(A), !is.na(B)) 

model_script = paste0(source_path,"generic models-gaussian glmer 1-RF + 4-FF REMLF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))

## model_comparisonsAIC("src/generic models-gaussian glmer 1-RF + 4-FF.R") ## AB: wide difference between model selection, uncomment to compare.
```

```{r}
femm_eggmorph <- lmer(mass_c ~ eggs_b * body_c + thorax_c * body_c + wing_c  + (1|ID), data=data_fem, REML=TRUE)
tidy_regression(femm_eggmorph, is_color=output_col) # Adding (1|trial_type) does not improve fit
```

* if female laid eggs that day, then mass is larger
* the larger the body, the heavier the female
* marginal? effect of thorax
* negative effect of wing where the larger the wing the lighter the female

* no effect of eggs*body
* no effect of body*thorax

### Plotting Female Data

#### Sym Distance

```{r}
p <- gf_point(mass ~ sym_dist, data = data_fem, 
                  color = ~host_plant, 
                  ylab="Mass (g)", 
                  xlab="Distance From Sympatric Zone")
p
```

#### Eggs

```{r, warning=FALSE}
#, fig.width=2.8, fig.height=2.3}
data_eggs <- data_fem %>%
  filter(!is.na(total_eggs))

lm_plot(data_eggs, "total_eggs_c", "mass_c", "Total Eggs Laid", "Mass (g)", "Mass by Total Eggs Laid")

lm_plot2(data_fem, "eggs_b", "mass_c", "Eggs Laid", "Mass (g)", "Mass by Egg Laid on Test Day")
axis(1, at = seq(-1,1,.5),labels=c(" ", " ", "M ", " ", "F"))

lm_plot(data_fem, "sym_dist", "mass_c", "Distance From Sympatric Zone", "Mass (g)", "Mass by Distance from Sympatric Zone")
```

#### Morphology

```{r, warning=FALSE}
#, fig.width=2.8, fig.height=2.3}

lm_plot(data_fem, "wing2body_c", "mass_c", "Wing:Body", "Mass (g)", "Mass by Wing:Body")
lm_plot(data_fem, "thorax_c", "mass_c", "Thorax Length (mm)", "Mass (g)", "Mass by Thorax Length")
lm_plot(data_fem, "wing_c", "mass_c", "Wing Length (mm)", "Mass (g)", "Mass by Wing Length")
lm_plot(data_fem, "body_c", "mass_c", "Body Length (mm)", "Mass (g)", "Mass by Body Length")
lm_plot(data_fem, "beak_c", "mass_c", "Beak Length (mm)", "Mass (g)", "Mass by Beak Length")
```

```{r}
lm_polyplot(data_fem, "wing2body_c", "mass_c", "Wing:Body", "Mass (g)", "Mass by Wing:Body")
lm_polyplot(data_fem, "thorax_c", "mass_c", "Thorax Length (mm)", "Mass (g)", "Mass by Thorax Length")
lm_polyplot(data_fem, "wing_c", "mass_c", "Wing Length (mm)", "Mass (g)", "Mass by Wing Length")
lm_polyplot(data_fem, "body_c", "mass_c", "Body Length (mm)", "Mass (g)", "Mass by Body Length")
lm_polyplot(data_fem, "beak_c", "mass_c", "Beak Length (mm)", "Mass (g)", "Mass by Beak Length")
```

### Males

[Cleaning Data](#cleandatam)

[Testing Covariates](#testm)

[lmer() A=host, B=sym_dist, X=ID, Y=trial_type](#lmerm)

[Morphology1](#morphm)

<a id="cleandatam"></a>

#### Reading the data

```{r}
data <- read_flight_data("data/all_flight_data-Winter2020.csv")
data_all <- data[[1]]
data_tested <- data[[2]]

data_male <- data_tested[data_tested$sex=="M",]
data_male <- center_data(data_male)
```

<a id="testm"></a>

#### Testing Covariates

**Experimental Set-Up Effects**

```{r echo=FALSE}
###### No effect of chamber
tidy_regression(lmer(mass_c~chamber + (1|ID), data=data_male, REML=FALSE), is_color=output_col) 

####### No effect of test date
tidy_regression(lmer(mass_c~days_from_start_c + (1|ID) + (1|trial_type), data=data_male, REML=FALSE), is_color=output_col)

####### No effect of test time
tidy_regression(lmer(mass_c~min_from_IncStart_c + (1|ID), data=data_male, REML=FALSE), is_color=output_col) 
```

**Biological Effects** - none b/c males don't lay eggs

**Morphology Effects**

```{r, echo=FALSE}
####### Effect of beak length
tidy_regression(lmer(mass_c~beak_c + (1|ID) + (1|trial_type), data=data_male, REML=FALSE), is_color=output_col) 

####### Effect of thorax length
tidy_regression(lmer(mass_c~thorax_c + (1|ID) + (1|trial_type), data=data_male, REML=FALSE), is_color=output_col)

####### Effect of body length
tidy_regression(lmer(mass_c~body_c + (1|ID) + (1|trial_type), data=data_male, REML=FALSE), is_color=output_col)

####### Effect of wing length
tidy_regression(lmer(mass_c~wing_c + (1|ID) + (1|trial_type), data=data_male, REML=FALSE), is_color=output_col)
```

<a id="lmerm"></a>

#### Male Data | lmer() A=host, B=sym_dist, X=ID, Y=trial_type

```{r, message= FALSE}
data<-data.frame(R=data_male$mass_c, 
                 A=data_male$host_c, 
                 B=data_male$sym_dist,
                 X=data_male$ID)

model_script = paste0(source_path,"generic models-gaussian glmer 1-RF + 2-FF REMLF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m3, m4, test="Chisq") # Adding A*B does not improve fit
```

```{r}
male_model <- lmer(mass_c ~ host_c + sym_dist + (1 | ID), data=data_male, REML=FALSE)
tidy_regression(male_model, is_color=output_col)

male_model <- lmer(mass_c ~ host_c + sym_dist + (1 | ID) + (1 | trial_type), data=data_male, REML=FALSE)
tidy_regression(male_model, is_color=output_col)
```

* significant but weak negative effect of host
* no effect of sym_dist

<a id="morphm"></a>

#### Morphology Modeling

```{r, message=FALSE}
data<-data.frame(R=data_male$mass_c, 
                 A=data_male$beak_c, 
                 B=data_male$thorax_c,
                 C=data_male$wing_c,
                 D=data_male$body_c,
                 X=data_male$ID) # multiple models failed if add Y=data_male$trialtype
data <- data %>%
  filter(!is.na(A))

model_script = paste0(source_path,"generic models-gaussian glmer 1-RF + 4-FF REMLF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
# this changed
anova(m64, m90, test="Chisq") # Adding C*D does not improve fit
anova(m64, m80, test="Chisq") # Adding A*C does not improve fit
```

```{r}
male_morph_model <- lmer(mass_c ~ thorax_c * wing_c + thorax_c * body_c + wing_c * body_c + beak_c  + (1|ID), data=data_male, REML=FALSE)
tidy_regression(male_morph_model, is_color=output_col)

male_morph_model <- lmer(mass_c ~ thorax_c * wing_c + thorax_c * body_c + wing_c * body_c + beak_c  + (1|ID) + (1|trial_type), data=data_male, REML=FALSE)
tidy_regression(male_morph_model, is_color=output_col)
```

### Plotting Male Data

```{r}
data_mmale <- data_male %>%
  filter(!is.na(thorax))
p1 <- gf_point(mass ~ sym_dist, data = data_mmale, 
                  color = ~host_plant, 
                  ylab="Mass (g)", 
                  xlab="Distance From Sympatric Zone")
p1
```

```{r}
#, fig.width=2.8, fig.height=2.3}
lm_plot(data_mmale, "wing2body_c", "mass_c", "Wing:Body", "Mass (g)", "Mass by Wing:Body")
lm_plot(data_mmale, "thorax_c", "mass_c", "Thorax Length (mm)", "Mass (g)", "Mass by Thorax Length")
lm_plot(data_mmale, "wing_c", "mass_c", "Wing Length (mm)", "Mass (g)", "Mass by Wing Length")
lm_plot(data_mmale, "body_c", "mass_c", "Body Length (mm)", "Mass (g)", "Mass by Body Length")
lm_plot(data_mmale, "beak_c", "mass_c", "Beak Length (mm)", "Mass (g)", "Mass by Beak Length")
```

```{r}
# , fig.width=2.8, fig.height=2.3}
p.morph <- gf_point(mass ~ thorax, data = data_mmale, 
                  alpha = ~beak, 
                  ylab="Mass (g)", 
                  xlab="Thoarx Length (mm)")
p.morph
```

### Trial 1

[Cleaning Data](#cleanT1)

[Testing Covariates](#testT1)

[Sex * Host * Sym_Dist Modeling](#modelT1)

[Morphology](#morphT1)

<a id="cleanT1"></a>

#### Reading the data

```{r}
data <- read_flight_data("data/all_flight_data-Winter2020.csv")
data_all <- data[[1]]
data_tested <- data[[2]]

# filter out missing masses
data_tested <- data_tested %>%
  filter(!is.na(mass_c))

data_T1 <- filter(data_tested, trial_type == "T1")
data_T2 <- filter(data_tested, trial_type == "T2")
data_T1 <- center_data(data_T1)
```

<a id="testT1"></a>

#### Testing Covariates

**Experimental Set-Up Effects**

```{r echo=FALSE}
###### No effect of chamber
tidy_regression(lmer(mass_c~chamber + (1|days_from_start_c), data=data_T1, REML=FALSE), is_color=output_col) # (1|population) variance near zero so can remove

####### No effect of test date
tidy_regression(lmer(mass_c~days_from_start_c + (1|chamber), data=data_T1, REML=FALSE), is_color=output_col)

####### No effect of test time
tidy_regression(lmer(mass_c~min_from_IncStart_c + (1|chamber), data=data_T1, REML=FALSE), is_color=output_col) 
```

**Morphology Effects**

```{r, echo=FALSE}
####### Effect of beak length
tidy_regression(lmer(mass_c~beak_c + (1|population), data=data_T1, REML=FALSE), is_color=output_col) 

####### Effect of thorax length
tidy_regression(lmer(mass_c~thorax_c + (1|population), data=data_T1, REML=FALSE), is_color=output_col)

####### Effect of body length
tidy_regression(lmer(mass_c~body_c + (1|population), data=data_T1, REML=FALSE), is_color=output_col)

####### Effect of wing length
tidy_regression(lmer(mass_c~wing_c + (1|population), data=data_T1, REML=FALSE), is_color=output_col)
```

<a id="modelT1"></a>

#### Sex * Host * Sym_Dist Modeling

```{r, message=FALSE}
data<-data.frame(R=data_T1$mass_c, 
                 A=data_T1$host_c, 
                 B=data_T1$sym_dist,
                 C=data_T1$sex_c,
                 X=data_T1$population)

model_script = paste0(source_path,"generic models-gaussian glmer 1-RF + 3-FF REMLF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m10, m13, test="Chisq") # Adding A does not improve fit
```

```{r}
T1_model <- lmer(mass_c ~ sex_c * sym_dist + (1|population), data=data_T1, REML=FALSE)
tidy_regression(T1_model, is_color=output_col)
```

* positive effect of sex, where if female then weigh more
* negative effect of sym_dist, where if from GRT then weigh less

* negative effect of sym_dist*sex where if from GRT and F then weigh less

<a id="morphT1"></a>

#### Morphology Modeling

```{r, message=FALSE}
data<-data.frame(R=data_T1$mass_c, 
                 A=data_T1$beak_c, 
                 B=data_T1$thorax_c,
                 C=data_T1$body_c,
                 D=data_T1$wing_c,
                 X=data_T1$population)
data <- data %>%
  filter(!is.na(A))

model_script = paste0(source_path,"generic models-gaussian glmer 1-RF + 4-FF REMLF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
T1_morph_model <- lmer(mass_c ~ beak_c * thorax_c + beak_c * body_c + beak_c * wing_c + thorax_c * body_c  + (1 | population), data=data_T1, REML=FALSE)
tidy_regression(T1_morph_model, is_color=output_col)
```

* positive effect of beak
* positive effect of thorax
* positive effect of wing
* negative effect of wing 

* marginal negative effect of beak*thorax
* positve effect of beak*body
* negative effect of beak*wing
* positive effect of thorax*body


```{r}
data<-data.frame(R=data_T1$mass_c, 
                 A=data_T1$beak_c, 
                 B=data_T1$thorax_c,
                 C=data_T1$wing2body_c,
                 X=data_T1$population)
data <- data %>%
  filter(!is.na(A))

model_script = paste0(source_path,"generic models-gaussian glmer 1-RF + 3-FF REMLF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m12, m16, test="Chisq") # Adding B*C interaction does not improve fit
anova(m12, m14, test="Chisq") # Adding A*B interaction does not improve fit
```

```{r}
T1_morph_model <- lmer(mass_c ~ beak_c * wing2body_c + thorax_c + (1 | population), data=data_T1, REML=FALSE)
tidy_regression(T1_morph_model, is_color=output_col)
```

* positive effect of beak where longer then weigh more
* negative effect of wing2body where the more wing to body then weigh less
* positive effect of thorax where the wider the thorax the more the bug weighs

* negative effect of beak*wing2body wher ethe more wing to body and longer beak then weigh less


### Trial 2

[Cleaning Data](#cleanT2)

[Testing Covariates](#testT1)

[Sex * Host * Sym_Dist Modeling](#modelT2)

[Morphology](#morphT1)

<a id="cleanT2"></a>

#### Reading the data

```{r}
data <- read_flight_data("data/all_flight_data-Winter2020.csv")
data_all <- data[[1]]
data_tested <- data[[2]]

# filter out missing masses
data_tested <- data_tested %>%
  filter(!is.na(mass_c))

data_T2 <- filter(data_tested, trial_type == "T2")
data_T2 <- center_data(data_T2)
```

<a id="testT2"></a>

**Testing Experimental Effects**

```{r echo=FALSE}
###### No effect of chamber
tidy_regression(lmer(mass_c~chamber + (1|site), data=data_T2, REML=FALSE), is_color=output_col)

####### No effect of test date
tidy_regression(lmer(mass_c~days_from_start_c + (1|site), data=data_T2, REML=FALSE), is_color=output_col)

####### No effect of test time
tidy_regression(lmer(mass_c~min_from_IncStart_c + (1|site), data=data_T2, REML=FALSE), is_color=output_col) 
```

**Morphology Effects**

```{r, echo=FALSE}
####### Effect of beak length
tidy_regression(lmer(mass_c~beak_c + (1|population), data=data_T2, REML=FALSE), is_color=output_col) 

####### Effect of thorax length
tidy_regression(lmer(mass_c~thorax_c + (1|population), data=data_T2, REML=FALSE), is_color=output_col)

####### Effect of body length
tidy_regression(lmer(mass_c~body_c + (1|population), data=data_T2, REML=FALSE), is_color=output_col)

####### Effect of wing length
tidy_regression(lmer(mass_c~wing_c + (1|population), data=data_T2, REML=FALSE), is_color=output_col)
```

<a id="modelT2"></a>

#### Sex * Host * Sym_Dist Modeling

```{r, message=FALSE}
data<-data.frame(R=data_T2$mass_c, 
                 A=data_T2$host_c, 
                 B=data_T2$sym_dist,
                 C=data_T2$sex_c,
                 X=data_T2$population)

model_script = paste0(source_path,"generic models-gaussian glmer 1-RF + 3-FF REMLF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m5, m9, test="Chisq") # Adding A*C does not improve fit
```

```{r}
T2_model <- lmer(mass_c ~ host_c + sex_c + (1|population), data=data_T2, REML=FALSE)
tidy_regression(T2_model, is_color=output_col)
```

* positive effect of sex, where if female then weigh more.
* negative effect of sex, where if from GRT then weigh less.

<a id="moprhT2"></a>

#### Morphology Modeling

```{r message=FALSE}
data<-data.frame(R=data_T2$mass_c, 
                 A=data_T2$beak_c, 
                 B=data_T2$thorax_c,
                 C=data_T2$body_c,
                 D=data_T2$wing_c,
                 X=data_T2$population)
data <- data %>%
  filter(!is.na(A))

model_script = paste0(source_path,"generic models-gaussian glmer 1-RF + 4-FF REMLF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m49, m77, test="Chisq") # Adding A*C does not improve fit
```

```{r}
T2_morph_model <- lmer(mass_c ~ beak_c * wing_c + thorax_c * body_c  + (1|population), data=data_T2, REML=FALSE)
tidy_regression(T2_morph_model, is_color=output_col)
```

```{r, message=FALSE}
data<-data.frame(R=data_T2$mass_c, 
                 A=data_T2$beak_c, 
                 B=data_T2$thorax_c,
                 C=data_T2$wing2body_c,
                 X=data_T2$population)
data <- data %>%
  filter(!is.na(A))

model_script = paste0(source_path,"generic models-gaussian glmer 1-RF + 3-FF REMLF.R")
errors = catch_warnings(model_comparisonsAIC(model_script))
cat("Number of models that failed to converge: ", length(errors$warnings))
```

```{r}
anova(m7, m12, test="Chisq") # Adding A*C does improve the fit
anova(m12, m16, test="Chisq") # Adding B*C does not improve fit
```

```{r}
T2_morph_model <- lmer(mass_c ~ beak_c*wing2body_c + thorax_c + (1|population), data=data_T2, REML=TRUE)
tidy_regression(T2_morph_model, is_color=output_col)
```

* positive effect of beak length, where the longer beak the heavier the bug
* negative effect of wing2body ratio, where the larger the ratio, the lighter the bug.
* positive effect of thorax, where the longer the thorax the heavier the bug
* negative effect of beak*wing2body ratio where the longer the beak and ratio, the lighter the bug.

### T1 vs. T2 Plotting

#### Histograms 

```{r} 
#, fig.width=6, fig.height=2.5}
data_T1 <- filter(data_tested, trial_type == "T1")
data_T2 <- filter(data_tested, trial_type == "T2")
h1 <- as.grob(expression(hist(data_T1$mass, main="Trial 1", xlab= "Mass (g)")))
h2 <- as.grob(expression(hist(data_T2$mass, main="Trial 2", xlab = "Mass (g)")))
hist(data_all$mass, xlab="Mass (g)", main="Soapberry Bug Mass Histogram")

grid.arrange(h1, h2, ncol=2)
```

```{r}
h <- gf_histogram(~mass, data=data_tested, xlab="Mass (g)", ylab="Count",
             title="Soapberry Bug Mass Histogram", fill=~trial_type) %>%
    gf_theme(theme = theme_minimal() )
h
# Uncomment below to save graph:
#ggsave("Mass-Hist-splitbyTrial", plot = h, device="jpeg",width = 5, height = 3)
```

#### Trial 1: Mass by Sex, Host Plant and Distance from Sympatric Zone 

```{r, fig.width=6, fig.height=2.5}
p1 <- as.grob(expression(
  plot(data_T1$sex_c, data_T1$mass, ylab="Mass (g)", xlab="Sex")))
p2 <- as.grob(expression(
  plot(data_T1$host_c, data_T1$mass, ylab="Mass (g)", xlab="Host")))
p3 <- as.grob(expression(
  plot(data_T1$sym_dist, data_T1$mass, ylab="Mass (g)", 
       xlab="Distance From Sympatric Zone"))) 
p.all <- gf_point(mass ~ sym_dist, data = data_T1,
                  color = ~host_c, ylab="Mass (g)", 
                  xlab="Distance From Sympatric Zone", 
                  alpha = ~sex_c) # GRT = blue, BV = black
p.all

g <- grid.arrange(p1,p2,p3, ncol=3, top= "Trial 1")

# Uncomment below to save graph:
# ggsave("Massbysex,host,symp_zone", plot = g, device="jpeg", width = 9, height = 4)
# If Female and From GRT and farther from HS then have more mass.
```

#### Trial 1: Mass by Morphology

```{r}
p4 <- as.grob(expression(
  plot(data_T1$body, data_T1$mass, ylab="Mass (g)", xlab="Body (mm)")))
p5 <- as.grob(expression(
  plot(data_T1$beak, data_T1$mass, ylab="Mass (g)", xlab="Beak (mm)")))
p6 <- as.grob(expression(
  plot(data_T1$thorax, data_T1$mass, ylab="Mass (g)", xlab="Throax (mm)")))
p7 <- as.grob(expression(
  plot(data_T1$wing, data_T1$mass, ylab="Mass (g)", xlab="Wing (mm)")))

g2 <- grid.arrange(p4,p5,p6,p7, ncol=2, top="Trial 1")

# Uncomment below to save graph:
#ggsave("MassbyMorphologyT2", plot = g2, device="jpeg", width = 9, height = 7)
```

#### Trial 2: Mass by Morphology

```{r}
p8 <- as.grob(expression(
  plot(data_T2$body, data_T2$mass, ylab="Mass (g)", xlab="Body (mm)")))
p9 <- as.grob(expression(
  plot(data_T2$beak, data_T2$mass, ylab="Mass (g)", xlab="Beak (mm)")))
p10 <- as.grob(expression(
  plot(data_T2$thorax, data_T2$mass, ylab="Mass (g)", xlab="Throax (mm)")))
p11 <- as.grob(expression(
  plot(data_T2$wing, data_T2$mass, ylab="Mass (g)", xlab="Wing (mm)")))

g3 <- grid.arrange(p8,p9,p10,p11, ncol=2, top="Trial 2")

# Uncomment below to save graph:
#ggsave("MassbyMorphologyT2", plot = g3, device="jpeg", width = 9, height = 7)
```

